---
title: "Clustering part 3"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true

---

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```

```{r call-libraries} 
date()
library(Giotto)
library(future); if(parallelly::availableCores() > 2 & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); library(leiden);    
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

```


```{r readRDS-harmony-excl-annot}
# sln.sub <- readRDS("sln.clustering3.dim5k10.res0.1.20250416.084539.RDS")
sln <- readRDS("sln.clustering2.res0.5.20250411.1611.RDS")
print(paste("Primary object size: ", length(sln@cell_ID$cell), "cells"))
md <- pDataDT(sln) 
# saveRDS(md, file="md.sln.clustering2.RDS")
# saveRDS(Giotto::getExpression(sln, values = "scaled", output = c("matrix"),  set_defaults = TRUE ), "exprs.sln.clustering2.RDS")
print("cluster membership")
print(table(md$leiden_harmony))

md <- readRDS(file = "md.exprs.combined.sln.clustering2.res0.5.20250411.1611.RDS")  # sabrina's metadata with L1_annotations
df <- md[,c("cell_ID","L1_annotations")]
sln <- addCellMetadata(gobject = sln, new_metadata = df, by_column=T, column_cell_ID = 'cell_ID')
rm(df)
plots <- list()
```



# plot downsampled umap
```{r plot-UMAP-Downsample-excl1, fig.show="hold", out.width="50%"}
md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$leiden_harmony)))
names(leiden_colors) <- unique(md$leiden_harmony)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.projection', cell_color = 'L1_annotations',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

data <- as.data.frame( sln.sm@dimension_reduction$cells$cell$rna$umap$umap.projection@coordinates)
data$cluster <- sln.sm@cell_metadata$cell$rna$leiden_harmony
data <- data[sample(1:nrow(data),3e4), ]
ggplot(data) + geom_point(aes(x=Dim.1, y=Dim.2, color=cluster), show.legend = FALSE, alpha=0.2, size=0.2) + 
  geom_point(aes(x = Inf, y = Inf, color = cluster), size = 3, show.legend = TRUE) + 
  geom_density2d(aes(x=Dim.1, y=Dim.2), linewidth=0.25, color="black") + theme_classic()  + 
  # scale_color_manual(values = leiden_colors)   + 
  scale_x_continuous(limits=c(min(data$Dim.1) - .10*diff(range(data$Dim.1)),max(data$Dim.1) + .10* diff(range(data$Dim.1)))) +
  scale_y_continuous(limits=c(min(data$Dim.2) - .10*diff(range(data$Dim.2)),max(data$Dim.2) + .10* diff(range(data$Dim.2)))) + 
  ggtitle("All cells - downsampled") +  guides(color = guide_legend(override.aes = list(shape = 16)))
rm(data)

```

# plot heatmap of features vs clusters
```{r cluster-feat-heatmap-excl2-Cells, fig.show="hold", fig.width=10, fig.height=5}
# markers_scran <- findMarkers_one_vs_all(gobject = sln, method = "scran", expression_values = "normalized",
#                                         cluster_column = "leiden_harmony", min_feats = 3)
# 
# topgenes_scran <- unique(markers_scran[, head(.SD, 5), by = "cluster"][["feats"]])
# 
# plotMetaDataHeatmap(sln, expression_values = "normalized",
#                     metadata_cols = "leiden_harmony", selected_feats = topgenes_scran,
#                     y_text_size = 8, show_values = "zscores_rescaled")
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")
plotMetaDataHeatmap(sln, expression_values = "normalized",
                    metadata_cols = "L1_annotations", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
```

# plot CD3, CD4, CD20, MART1, and TOX on umap
```{r spatDimPlot-CD3E, fig.width=12, fig.height=12}
md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
dimFeatPlot2D(sln.sm, dim_reduction_name = 'umap.projection', expression_values = "scaled", 
              feats = c("CD3E","CD4", "CD8", "CD20","MART1","TOX"), 
              point_alpha = 0.2, point_shape = "no_border", cow_n_col = 2)
```

# violin plot
```{r show-violin, fig.show="hold", fig.width=10, fig.height=30, warning=F}
# don't try to do violinplot on more than 100k cells! 
# md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 5e4))$cell_ID
# sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
# violinPlot(sln.sm, feats = clusterChannels, cluster_column = "L1_annotations")

```


# filter down to tumor only
```{r show-tumor5-only, fig.show="hold", fig.height=12, out.width="50%"}
md <- pDataDT(sln); keepCells <- subset(md, list_ID == "tumor.5" & y %in% c(10000:20000) )$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$L1_annotations)))
names(leiden_colors) <- unique(md$L1_annotations)
leiden_colors.focus <- leiden_colors 
leiden_colors.focus <- rep("gray90",length(leiden_colors)); leiden_colors.focus["1"] <- leiden_colors["1"]
spatPlot2D(sln.sm, cell_color = 'L1_annotations', cell_color_code = leiden_colors, 
                         point_size = 1, point_shape="no_border", point_alpha=0.3, axis_text=14, 
                         legend_text = 14)
spatPlot2D(sln.sm, cell_color = 'L1_annotations', cell_color_code = leiden_colors.focus, 
                         point_size = 1, point_shape="no_border", point_alpha=0.3, axis_text=14, 
                         legend_text = 14)

```


# Calculate clusters by Leiden
```{r calculate-Clusters-res0.7}
md <- pDataDT(sln)
print("Eliminating non-lymphocyte populations")
# goal is to make a subsetGiotto object consisting of TCells.1-2, BCells.1-21, TBMixed.1-31
toEliminate <- c("Artifact.1","Artifact.2", "Artifact.3","Tumor.1","Tumor.2","Tumor.3","Tumor.4","Tumor.5")
keepCells <- subset(md, !(L1_annotations %in% c(toEliminate)))$cell_ID
sln.sub <- subsetGiotto(gobject = sln, cell_ids = keepCells)
print(paste("New object size: ", length(sln.sub@cell_ID$cell), "cells"))
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "CD19", "CD20", "CD38", "CD45")

sln.sub <- runPCA(gobject = sln.sub, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(sln.sub, scale_unit = TRUE, scree_ylim = c(0,40))
sln.sub <- createNearestNetwork(gobject = sln.sub, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:5, k = 15)  # calculate nearest neighbors 
sln.sub <- doLeidenClusterIgraph(gobject = sln.sub, resolution = 0.7, n_iterations = 200, set_seed=T, seed_number = 43, 
                                network_name = 'NN.harmony', name = 'L1_annotations')  # make the clusters
filename <- paste0("sln.clustering3.dim3k10.res0.7.",  strftime(Sys.time() , "%Y%m%d.%H%M%S"), ".RDS")
saveRDS(sln.sub, filename)
```

# plot heatmap of features vs clusters
```{r cluster-feat-heatmap-res0.7, fig.show="hold", fig.width=10, fig.height=5}
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")
plotMetaDataHeatmap(sln.sub, expression_values = "normalized",
                    metadata_cols = "L1_annotations", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
```



# UMAP projection
```{r UMAP-projection-res0.7}
sln.sub <- runUMAPprojection(sln.sub, random_subset = 1e5, dimensions_to_use=1:8, n_threads = 8, name="umap.clust3.d8")
sln.sub <- runUMAPprojection(sln.sub, random_subset = 1e5, dimensions_to_use=1:5, n_threads = 8, name="umap.clust3.d5")

```

# plot downsampled umap
```{r plot-UMAP-Downsample-res0.7, fig.show="hold", out.width="50%"}
md <- pDataDT(sln.sub); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln.sub, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$leiden_harmony)))
names(leiden_colors) <- unique(md$leiden_harmony)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.clust3.d8', cell_color = 'L1_annotations',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.clust3.d5', cell_color = 'L1_annotations',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)
```


```{r save-object-res0.7}
filename <- paste0("sln.clustering3.dim5k15.res0.7.",  strftime(Sys.time() , "%Y%m%d.%H%M%S"), ".RDS")
saveRDS(sln.sub, filename)
md <- pDataDT(sln.sub)
saveRDS(md, paste0("md.",filename))
saveRDS(Giotto::getExpression(sln.sub, values = "scaled", output = c("matrix"),  set_defaults = TRUE ),
        paste0("exprs.",filename))

```





# Calculate clusters by Leiden
```{r calculate-Clusters-res1}
md <- pDataDT(sln)
print("Eliminating non-lymphocyte populations")
# goal is to make a subsetGiotto object consisting of TCells.1-2, BCells.1-21, TBMixed.1-31
toEliminate <- c("Artifact.1","Artifact.2", "Artifact.3","Tumor.1","Tumor.2","Tumor.3","Tumor.4","Tumor.5")
keepCells <- subset(md, !(L1_annotations %in% c(toEliminate)))$cell_ID
sln.sub <- subsetGiotto(gobject = sln, cell_ids = keepCells)
print(paste("New object size: ", length(sln.sub@cell_ID$cell), "cells"))
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "CD19", "CD20", "CD38", "CD45")

sln.sub <- runPCA(gobject = sln.sub, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(sln.sub, scale_unit = TRUE, scree_ylim = c(0,40))
sln.sub <- createNearestNetwork(gobject = sln.sub, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:5, k = 15)  # calculate nearest neighbors 
sln.sub <- doLeidenClusterIgraph(gobject = sln.sub, resolution = 1, n_iterations = 200, set_seed=T, seed_number = 43, 
                                network_name = 'NN.harmony', name = 'L1_annotations')  # make the clusters

```

# plot heatmap of features vs clusters
```{r cluster-feat-heatmap-res1, fig.show="hold", fig.width=10, fig.height=5}
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")
plotMetaDataHeatmap(sln.sub, expression_values = "normalized",
                    metadata_cols = "L1_annotations", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
```



# UMAP projection
```{r UMAP-projection-res1}
sln.sub <- runUMAPprojection(sln.sub, random_subset = 1e5, dimensions_to_use=1:8, n_threads = 8, name="umap.clust3.d8")
sln.sub <- runUMAPprojection(sln.sub, random_subset = 1e5, dimensions_to_use=1:5, n_threads = 8, name="umap.clust3.d5")

```

# plot downsampled umap
```{r plot-UMAP-Downsample-res1, fig.show="hold", out.width="50%"}
md <- pDataDT(sln.sub); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln.sub, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$leiden_harmony)))
names(leiden_colors) <- unique(md$leiden_harmony)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.clust3.d8', cell_color = 'L1_annotations',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.clust3.d5', cell_color = 'L1_annotations',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)
```


```{r save-object-res1}
filename <- paste0("sln.clustering3.dim5k15.res1.0.",  strftime(Sys.time() , "%Y%m%d.%H%M%S"), ".RDS")
saveRDS(sln.sub, filename)
md <- pDataDT(sln.sub)
saveRDS(md, paste0("md.",filename))
saveRDS(Giotto::getExpression(sln.sub, values = "scaled", output = c("matrix"),  set_defaults = TRUE ),
        paste0("exprs.",filename))

```




```{r call-date}
date()
```

Cluster numbers
        dim3k10   dim5k10     dim5k15     dim8k10           
0.1     >1000     >1000       14          >100
0.3     >1000     >1000       28          >100
0.5     >1000     >1000       48          >100


