---
title: "Clustering part 4"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```

```{r call-libraries} 
date();  timeStamp <- strftime(Sys.time() , "%Y%m%d.%H%M%S")
library(Giotto)
library(future); if(parallelly::availableCores() > 2 & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); library(leiden);    
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

```

```{r readRDS-clustering3}
sln <- readRDS("sln.clustering3.dim5k15.res0.7.20250420.092625.RDS")
md <- pDataDT(sln) 
print(paste("Primary object size: ", nrow(md), "cells"))
print("cluster membership")
print(table(md$leiden_harmony))
print(table(md$L1_annotations))
```


```{r assign_newLabels}
df <- data.frame(leiden.clust3.r0.7 = md$leiden.clust3.r0.7, L2_annotations = "Unknown")
df$cell_ID <- md$cell_ID
df$L2_annotations <- paste0("Clust.", df$leiden.clust3.r0.7)  # unknown clusters:
df[which(df$leiden.clust3.r0.7  == "1" ), 'L2_annotations'] <-  "TBMixed.1"
df[which(df$leiden.clust3.r0.7  == "2" ), 'L2_annotations'] <-  "TBMixed.2"
df[which(df$leiden.clust3.r0.7  == "3" ), 'L2_annotations'] <-  "TBMixed.3"
df[which(df$leiden.clust3.r0.7  == "4" ), 'L2_annotations'] <-  "TBMixed.4"
df[which(df$leiden.clust3.r0.7  == "5" ), 'L2_annotations'] <-  "BCells.1"
df[which(df$leiden.clust3.r0.7  == "6" ), 'L2_annotations'] <-  "TBMixed.5"
df[which(df$leiden.clust3.r0.7  == "7" ), 'L2_annotations'] <-  "BCells.2"
df[which(df$leiden.clust3.r0.7  == "8" ), 'L2_annotations'] <-  "TCells.1"
df[which(df$leiden.clust3.r0.7  == "9" ), 'L2_annotations'] <-  "BCells.3"
df[which(df$leiden.clust3.r0.7  == "10" ), 'L2_annotations'] <-  "TCells.2"
df[which(df$leiden.clust3.r0.7  == "11" ), 'L2_annotations'] <-  "TBMixed.6"
df[which(df$leiden.clust3.r0.7  == "12" ), 'L2_annotations'] <-  "TBMixed.7"
df[which(df$leiden.clust3.r0.7  == "13" ), 'L2_annotations'] <-  "TBMixed.8"
df[which(df$leiden.clust3.r0.7  == "14" ), 'L2_annotations'] <-  "TBMixed.9"
df[which(df$leiden.clust3.r0.7  == "15" ), 'L2_annotations'] <-  "TBMixed.10"
df[which(df$leiden.clust3.r0.7  == "16" ), 'L2_annotations'] <-  "TBMixed.11"
df[which(df$leiden.clust3.r0.7  == "17" ), 'L2_annotations'] <-  "TCells.3"
df[which(df$leiden.clust3.r0.7  == "18" ), 'L2_annotations'] <-  "TCells.4"
df[which(df$leiden.clust3.r0.7  == "19" ), 'L2_annotations'] <-  "TCells.5"
df[which(df$leiden.clust3.r0.7  == "20" ), 'L2_annotations'] <-  "TCells.6"
df[which(df$leiden.clust3.r0.7  == "21" ), 'L2_annotations'] <-  "TCells.7"
df[which(df$leiden.clust3.r0.7  == "22" ), 'L2_annotations'] <-  "TCells.8"
df[which(df$leiden.clust3.r0.7  == "23" ), 'L2_annotations'] <-  "TCells.9"
df[which(df$leiden.clust3.r0.7  == "24" ), 'L2_annotations'] <-  "TCells.10"
df[which(df$leiden.clust3.r0.7  == "25" ), 'L2_annotations'] <-  "TCells.11"
df[which(df$leiden.clust3.r0.7  == "26" ), 'L2_annotations'] <-  "BCells.4"
df[which(df$leiden.clust3.r0.7  == "27" ), 'L2_annotations'] <-  "TBMixed.12"
df[which(df$leiden.clust3.r0.7  == "28" ), 'L2_annotations'] <-  "TBMixed.13"
df[which(df$leiden.clust3.r0.7  == "29" ), 'L2_annotations'] <-  "BCells.5"
df[which(df$leiden.clust3.r0.7  == "30" ), 'L2_annotations'] <-  "TBMixed.14"
df[which(df$leiden.clust3.r0.7  == "31" ), 'L2_annotations'] <-  "TCells.12"
df[which(df$leiden.clust3.r0.7  == "32" ), 'L2_annotations'] <-  "BCells.6"
df[which(df$leiden.clust3.r0.7  == "33" ), 'L2_annotations'] <-  "TCells.13"
df[which(df$leiden.clust3.r0.7  == "34" ), 'L2_annotations'] <-  "TCells.14"
df[which(df$leiden.clust3.r0.7  == "35" ), 'L2_annotations'] <-  "BCells.7"
df[which(df$leiden.clust3.r0.7  == "36" ), 'L2_annotations'] <-  "BCells.8"
df[which(df$leiden.clust3.r0.7  == "37" ), 'L2_annotations'] <-  "TBMixed.15"
df[which(df$leiden.clust3.r0.7  == "38" ), 'L2_annotations'] <-  "TBMixed.16"
df[which(df$leiden.clust3.r0.7  == "39" ), 'L2_annotations'] <-  "TBMixed.17"
df[which(df$leiden.clust3.r0.7  == "40" ), 'L2_annotations'] <-  "TBMixed.18"
df[which(df$leiden.clust3.r0.7  == "41" ), 'L2_annotations'] <-  "TBMixed.19"
df[which(df$leiden.clust3.r0.7  == "42" ), 'L2_annotations'] <-  "BCells.9"
df[which(df$leiden.clust3.r0.7  == "43" ), 'L2_annotations'] <-  "TCells.15"
df[which(df$leiden.clust3.r0.7  == "44" ), 'L2_annotations'] <-  "TCells.16"
df[which(df$leiden.clust3.r0.7  == "45" ), 'L2_annotations'] <-  "TBMixed.20"
df[which(df$leiden.clust3.r0.7  == "46" ), 'L2_annotations'] <-  "BCells.10"
df[which(df$leiden.clust3.r0.7  == "47" ), 'L2_annotations'] <-  "TCells.17"
df[which(df$leiden.clust3.r0.7  == "48" ), 'L2_annotations'] <-  "TCells.18"
df[which(df$leiden.clust3.r0.7  == "49" ), 'L2_annotations'] <-  "BCells.11"
df[which(df$leiden.clust3.r0.7  == "50" ), 'L2_annotations'] <-  "TCells.19"
df[which(df$leiden.clust3.r0.7  == "51" ), 'L2_annotations'] <-  "BCells.12"
df[which(df$leiden.clust3.r0.7  == "52" ), 'L2_annotations'] <-  "BCells.13"
df[which(df$leiden.clust3.r0.7  == "53" ), 'L2_annotations'] <-  "TBMixed.21"
df[which(df$leiden.clust3.r0.7  == "54" ), 'L2_annotations'] <-  "BCells.14"
df[which(df$leiden.clust3.r0.7  == "55" ), 'L2_annotations'] <-  "BCells.15"
df[which(df$leiden.clust3.r0.7  == "56" ), 'L2_annotations'] <-  "TCells.20"
df[which(df$leiden.clust3.r0.7  == "57" ), 'L2_annotations'] <-  "TCells.21"
df[which(df$leiden.clust3.r0.7  == "58" ), 'L2_annotations'] <-  "TCells.22"
df[which(df$leiden.clust3.r0.7  == "59" ), 'L2_annotations'] <-  "TBMixed.22"
df[which(df$leiden.clust3.r0.7  == "60" ), 'L2_annotations'] <-  "TBMixed.23"
df[which(df$leiden.clust3.r0.7  == "61" ), 'L2_annotations'] <-  "TCells.23"
df[which(df$leiden.clust3.r0.7  == "62" ), 'L2_annotations'] <-  "TBMixed.24"
df[which(df$leiden.clust3.r0.7  == "63" ), 'L2_annotations'] <-  "TBMixed.25"
df[which(df$leiden.clust3.r0.7  == "64" ), 'L2_annotations'] <-  "BCells.16"
sln <- addCellMetadata(gobject = sln, new_metadata = df, by_column=T, column_cell_ID = 'cell_ID')
rm(df)

```



# TBMixed Cells: Cluster out TBMixed.1-25
TBMixed Cells Object Panel:
CD20
CD19
PAX5
HLA DR
CD38
IGD
KI67
CXCR5
CD21
CD11C
CD3E
CD4
CD8
TCF1
GZMB
CD39
PD1
TOX
FOXP3
ICOS
CD45RO

```{r subset-TBmixed-cells}
gc()
md <- pDataDT(sln); keepCells <- md$cell_ID[grep(pattern = "TBMixed", md$L2_annotations)]
TBmixcells <- subsetGiotto(sln, cell_ids = keepCells )
print("subcluster TBmixcells")
clusterChannels <- c("CD20", "CD19", "PAX5", "HLA.DR", "CD38", "IGD", "KI67", "CXCR5", "CD21", "CD11C", "CD3E", 
                     "CD4", "CD8", "TCF1", "GZMB", "CD39", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO")
TBmixcells <- runPCA(gobject = TBmixcells, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(TBmixcells, scale_unit = TRUE, scree_ylim = c(0,40))
TBmixcells = runGiottoHarmony(TBmixcells, vars_use = 'list_ID',  dim_reduction_to_use = "pca", dimensions_to_use = 1:5) 
TBmixcells <- createNearestNetwork(gobject = TBmixcells, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'TB.harm', dimensions_to_use = 1:5, k = 15)  # calculate nearest neighbors, focus on global differences 
TBmixcells <- doLeidenClusterIgraph(gobject = TBmixcells, resolution = 0.1, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'TB.harm', name = 'leiden_TBmixed.res0.1')   
TBmixcells <- doLeidenClusterIgraph(gobject = TBmixcells, resolution = 0.3, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'TB.harm', name = 'leiden_TBmixed.res0.3')  
TBmixcells <- runUMAPprojection(TBmixcells, random_subset = 1e5, n_threads = parallelly::availableCores(), 
                                name="umap.TBmixed")  # project UMAP coordinates to the full dataset based on a subsampling
print(table(TBmixcells$leiden_TBmixed.res0.1))

```

# save object 
```{r save-TBmixed-res0.5}
filename <- paste0("sln.clustering4.tbmixed.harm.",  timeStamp, ".RDS")
saveRDS(TBmixcells, filename)
md <- pDataDT(TBmixcells)
saveRDS(md, paste0("md.",filename))
saveRDS(Giotto::getExpression(TBmixcells, values = "scaled", output = c("matrix"),  set_defaults = TRUE ),
        paste0("exprs.",filename))

```


# heatmap and umap of TBmixed
```{r plot-UMAP-heatmap-TBmixed, fig.show="hold", fig.width=10, fig.height=15}
md <- pDataDT(TBmixcells); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = TBmixcells, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$leiden_TBmixed)))
names(leiden_colors) <- unique(md$leiden_TBmixed)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.TBmixed', cell_color = 'leiden_TBmixed',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)


dimFeatPlot2D(sln.sm, dim_reduction_name = 'umap.TBmixed', expression_values = "scaled", 
              feats = clusterChannels[1:8], 
              point_alpha = 0.2, point_shape = "no_border", cow_n_col = 2)

clusterChannels <- c("CD20", "CD19", "PAX5", "HLA.DR", "CD38", "IGD", "KI67", "CXCR5", "CD21", "CD11C", "CD3E", 
                     "CD4", "CD8", "TCF1", "GZMB", "CD39", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO")
plotMetaDataHeatmap(TBmixcells, expression_values = "normalized", title = "TBmixed res 0.1",
                    metadata_cols = "leiden_TBmixed", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(TBmixcells$leiden_TBmixed))

plotMetaDataHeatmap(TBmixcells, expression_values = "normalized", title = "TBmixed res 0.3",
                    metadata_cols = "leiden_TBmixed.res0.3", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(TBmixcells$leiden_TBmixed.res0.3))

```

# violin plot TBmixed
```{r show-violin-TBmixed, fig.show="hold", fig.width=10, fig.height=15, warning=F}
md <- pDataDT(TBmixcells); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 5e4))$cell_ID
sln.sm <- subsetGiotto(gobject = TBmixcells, cell_ids = keepCells)
violinPlot(sln.sm, feats = clusterChannels, cluster_column = "leiden_TBmixed")

print(paste("Primary object size: ", nrow(md), "cells"))
```



```{r call-date}
date()
```



seff 63157507
Job ID: 63157507
Cluster: slurm_cluster
User/Group: heratr01/heratr01
State: COMPLETED (exit code 0)
Nodes: 1
Cores per node: 10
CPU Utilized: 01:30:51
CPU Efficiency: 14.02% of 10:48:00 core-walltime
Job Wall-clock time: 01:04:48
Memory Utilized: 103.35 GB
Memory Efficiency: 51.67% of 200.00 GB












