---
title: "Clustering part 4"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```

```{r call-libraries} 
date()
library(Giotto)
library(future); if(parallelly::availableCores() > 2 & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); library(leiden);    
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

```

```{r readRDS-clustering3}
sln <- readRDS("sln.clustering3.dim5k15.res0.7.20250420.092625.RDS")
md <- pDataDT(sln) 
print(paste("Primary object size: ", nrow(md), "cells"))
print("cluster membership")
print(table(md$leiden_harmony))
print(table(md$L1_annotations))
```

# add annotation labels
```{r assign_newLabels}
df <- data.frame(leiden.clust3.r0.7 = md$leiden.clust3.r0.7, L2_annotations = "Unknown")
df$cell_ID <- md$cell_ID
df$L2_annotations <- paste0("Clust.", df$leiden.clust3.r0.7)  # unknown clusters:
df[which(df$leiden.clust3.r0.7  == "1" ), 'L2_annotations'] <-  "TBMixed.1"
df[which(df$leiden.clust3.r0.7  == "2" ), 'L2_annotations'] <-  "TBMixed.2"
df[which(df$leiden.clust3.r0.7  == "3" ), 'L2_annotations'] <-  "TBMixed.3"
df[which(df$leiden.clust3.r0.7  == "4" ), 'L2_annotations'] <-  "TBMixed.4"
df[which(df$leiden.clust3.r0.7  == "5" ), 'L2_annotations'] <-  "BCells.1"
df[which(df$leiden.clust3.r0.7  == "6" ), 'L2_annotations'] <-  "TBMixed.5"
df[which(df$leiden.clust3.r0.7  == "7" ), 'L2_annotations'] <-  "BCells.2"
df[which(df$leiden.clust3.r0.7  == "8" ), 'L2_annotations'] <-  "TCells.1"
df[which(df$leiden.clust3.r0.7  == "9" ), 'L2_annotations'] <-  "BCells.3"
df[which(df$leiden.clust3.r0.7  == "10" ), 'L2_annotations'] <-  "TCells.2"
df[which(df$leiden.clust3.r0.7  == "11" ), 'L2_annotations'] <-  "TBMixed.6"
df[which(df$leiden.clust3.r0.7  == "12" ), 'L2_annotations'] <-  "TBMixed.7"
df[which(df$leiden.clust3.r0.7  == "13" ), 'L2_annotations'] <-  "TBMixed.8"
df[which(df$leiden.clust3.r0.7  == "14" ), 'L2_annotations'] <-  "TBMixed.9"
df[which(df$leiden.clust3.r0.7  == "15" ), 'L2_annotations'] <-  "TBMixed.10"
df[which(df$leiden.clust3.r0.7  == "16" ), 'L2_annotations'] <-  "TBMixed.11"
df[which(df$leiden.clust3.r0.7  == "17" ), 'L2_annotations'] <-  "TCells.3"
df[which(df$leiden.clust3.r0.7  == "18" ), 'L2_annotations'] <-  "TCells.4"
df[which(df$leiden.clust3.r0.7  == "19" ), 'L2_annotations'] <-  "TCells.5"
df[which(df$leiden.clust3.r0.7  == "20" ), 'L2_annotations'] <-  "TCells.6"
df[which(df$leiden.clust3.r0.7  == "21" ), 'L2_annotations'] <-  "TCells.7"
df[which(df$leiden.clust3.r0.7  == "22" ), 'L2_annotations'] <-  "TCells.8"
df[which(df$leiden.clust3.r0.7  == "23" ), 'L2_annotations'] <-  "TCells.9"
df[which(df$leiden.clust3.r0.7  == "24" ), 'L2_annotations'] <-  "TCells.10"
df[which(df$leiden.clust3.r0.7  == "25" ), 'L2_annotations'] <-  "TCells.11"
df[which(df$leiden.clust3.r0.7  == "26" ), 'L2_annotations'] <-  "BCells.4"
df[which(df$leiden.clust3.r0.7  == "27" ), 'L2_annotations'] <-  "TBMixed.12"
df[which(df$leiden.clust3.r0.7  == "28" ), 'L2_annotations'] <-  "TBMixed.13"
df[which(df$leiden.clust3.r0.7  == "29" ), 'L2_annotations'] <-  "BCells.5"
df[which(df$leiden.clust3.r0.7  == "30" ), 'L2_annotations'] <-  "TBMixed.14"
df[which(df$leiden.clust3.r0.7  == "31" ), 'L2_annotations'] <-  "TCells.12"
df[which(df$leiden.clust3.r0.7  == "32" ), 'L2_annotations'] <-  "BCells.6"
df[which(df$leiden.clust3.r0.7  == "33" ), 'L2_annotations'] <-  "TCells.13"
df[which(df$leiden.clust3.r0.7  == "34" ), 'L2_annotations'] <-  "TCells.14"
df[which(df$leiden.clust3.r0.7  == "35" ), 'L2_annotations'] <-  "BCells.7"
df[which(df$leiden.clust3.r0.7  == "36" ), 'L2_annotations'] <-  "BCells.8"
df[which(df$leiden.clust3.r0.7  == "37" ), 'L2_annotations'] <-  "TBMixed.15"
df[which(df$leiden.clust3.r0.7  == "38" ), 'L2_annotations'] <-  "TBMixed.16"
df[which(df$leiden.clust3.r0.7  == "39" ), 'L2_annotations'] <-  "TBMixed.17"
df[which(df$leiden.clust3.r0.7  == "40" ), 'L2_annotations'] <-  "TBMixed.18"
df[which(df$leiden.clust3.r0.7  == "41" ), 'L2_annotations'] <-  "TBMixed.19"
df[which(df$leiden.clust3.r0.7  == "42" ), 'L2_annotations'] <-  "BCells.9"
df[which(df$leiden.clust3.r0.7  == "43" ), 'L2_annotations'] <-  "TCells.15"
df[which(df$leiden.clust3.r0.7  == "44" ), 'L2_annotations'] <-  "TCells.16"
df[which(df$leiden.clust3.r0.7  == "45" ), 'L2_annotations'] <-  "TBMixed.20"
df[which(df$leiden.clust3.r0.7  == "46" ), 'L2_annotations'] <-  "BCells.10"
df[which(df$leiden.clust3.r0.7  == "47" ), 'L2_annotations'] <-  "TCells.17"
df[which(df$leiden.clust3.r0.7  == "48" ), 'L2_annotations'] <-  "TCells.18"
df[which(df$leiden.clust3.r0.7  == "49" ), 'L2_annotations'] <-  "BCells.11"
df[which(df$leiden.clust3.r0.7  == "50" ), 'L2_annotations'] <-  "TCells.19"
df[which(df$leiden.clust3.r0.7  == "51" ), 'L2_annotations'] <-  "BCells.12"
df[which(df$leiden.clust3.r0.7  == "52" ), 'L2_annotations'] <-  "BCells.13"
df[which(df$leiden.clust3.r0.7  == "53" ), 'L2_annotations'] <-  "TBMixed.21"
df[which(df$leiden.clust3.r0.7  == "54" ), 'L2_annotations'] <-  "BCells.14"
df[which(df$leiden.clust3.r0.7  == "55" ), 'L2_annotations'] <-  "BCells.15"
df[which(df$leiden.clust3.r0.7  == "56" ), 'L2_annotations'] <-  "TCells.20"
df[which(df$leiden.clust3.r0.7  == "57" ), 'L2_annotations'] <-  "TCells.21"
df[which(df$leiden.clust3.r0.7  == "58" ), 'L2_annotations'] <-  "TCells.22"
df[which(df$leiden.clust3.r0.7  == "59" ), 'L2_annotations'] <-  "TBMixed.22"
df[which(df$leiden.clust3.r0.7  == "60" ), 'L2_annotations'] <-  "TBMixed.23"
df[which(df$leiden.clust3.r0.7  == "61" ), 'L2_annotations'] <-  "TCells.23"
df[which(df$leiden.clust3.r0.7  == "62" ), 'L2_annotations'] <-  "TBMixed.24"
df[which(df$leiden.clust3.r0.7  == "63" ), 'L2_annotations'] <-  "TBMixed.25"
df[which(df$leiden.clust3.r0.7  == "64" ), 'L2_annotations'] <-  "BCells.16"
sln <- addCellMetadata(gobject = sln, new_metadata = df, by_column=T, column_cell_ID = 'cell_ID')
rm(df)

```

select all cells in this object that have a CD8.Positive.Classification = 1 and cluster into CD8 T Cell Object, all other T Cells left will make up the CD4 Object.

# subcluster cd8 T cells
```{r subset-cd8T-cells}
gc()
md <- pDataDT(sln); keepCells <- md$cell_ID[grep(pattern = "TCells", md$L2_annotations)]
Tcells <- subsetGiotto(sln, cell_ids = keepCells )

md <- pDataDT(Tcells)
print(paste("Primary object size: ", nrow(md), "cells"))

keepCells <- md$cell_ID[which(md$CD8.Positive.Classification == 1)]
cd8.Tcells <- subsetGiotto(Tcells, cell_ids = keepCells )
print("subcluster CD8 T cells")
clusterChannels <- c("TCF1", "GZMB", "PD1", "TOX", "CD45RO", "HLA.DR")
cd8.Tcells <- runPCA(gobject = cd8.Tcells, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(cd8.Tcells, scale_unit = TRUE, scree_ylim = c(0,40))
cd8.Tcells <- createNearestNetwork(gobject = cd8.Tcells, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:5, k = 15)  # calculate nearest neighbors, focus on global differences 
cd8.Tcells <- doLeidenClusterIgraph(gobject = cd8.Tcells, resolution = 0.5, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd8Tcells.res0.5')  # make the clusters
cd8.Tcells <- doLeidenClusterIgraph(gobject = cd8.Tcells, resolution = 0.7, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd8Tcells.res0.7')  # make the clusters

cd8.Tcells <- runUMAPprojection(cd8.Tcells, random_subset = 1e5, n_threads = parallelly::availableCores(), 
                                name="umap.cd8", dimensions_to_use = 1:5)  # project UMAP coordinates to the full dataset 
```

# save cd8 objects 
```{r save-cd8Tcells-res0.5}
filename <- paste0("sln.clustering4.cd8tcells.dim5k15.",  strftime(Sys.time() , "%Y%m%d.%H%M%S"), ".RDS")
saveRDS(cd8.Tcells, filename)
md <- pDataDT(cd8.Tcells)
saveRDS(md, paste0("md.",filename))
saveRDS(Giotto::getExpression(cd8.Tcells, values = "scaled", output = c("matrix"),  set_defaults = TRUE ),
        paste0("exprs.",filename))

```

# cd8 heatmap & umap 
```{r plot-UMAP-heatmap-cd8cells }
md <- pDataDT(cd8.Tcells); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = cd8.Tcells, cell_ids = keepCells)

# ********************* res 0.5 ***********************
leiden_colors <- getDistinctColors(length(unique(md$leiden_cd8Tcells.res0.5)))
names(leiden_colors) <- unique(md$leiden_cd8Tcells.res0.5)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.cd8', cell_color = 'leiden_cd8Tcells.res0.5',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "GZMB", "CD39", "KI67", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO", "HLA.DR")

plotMetaDataHeatmap(cd8.Tcells, expression_values = "normalized",
                    metadata_cols = "leiden_cd8Tcells.res0.5", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd8.Tcells$leiden_cd8Tcells.res0.5))

# ********************* res 0.7 ***********************

leiden_colors <- getDistinctColors(length(unique(md$leiden_cd8Tcells.res0.7)))
names(leiden_colors) <- unique(md$leiden_cd8Tcells.res0.7)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.cd8', cell_color = 'leiden_cd8Tcells.res0.7',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "GZMB", "CD39", "KI67", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO", "HLA.DR")

plotMetaDataHeatmap(cd8.Tcells, expression_values = "normalized",
                    metadata_cols = "leiden_cd8Tcells.res0.7", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd8.Tcells$leiden_cd8Tcells.res0.7))

```

# cd8 violin plot 
```{r show-violin-cd8Tcells, fig.show="hold", fig.width=10, fig.height=15, warning=F}
md <- pDataDT(cd8.Tcells); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 5e4))$cell_ID
sln.sm <- subsetGiotto(gobject = cd8.Tcells, cell_ids = keepCells)
violinPlot(sln.sm, feats = clusterChannels, cluster_column = "leiden_cd8Tcells.res0.5")
print(paste("Primary object size: ", nrow(md), "cells"))

```

# remove outlier cd8 cluster 
```{r subset-cd8T-cells-again}
gc()
print("remove res0.5 cluster 32")
md <- pDataDT(cd8.Tcells)
finalLabel <- data.frame(cell_ID = subset(md, leiden_cd8Tcells.res0.5 == 32 )$cell_ID, L1_annotations = "CD8.outlier")
saveRDS( object = finalLabel, file = "finalAnnotation.cd8outlier.RDS")

keepCells <- subset(md, !(leiden_cd8Tcells.res0.5 == 32 ))$cell_ID
cd8.Tcells <- subsetGiotto(gobject = cd8.Tcells, cell_ids = keepCells)

print("subcluster CD8 T cells")
clusterChannels <- c("TCF1", "TOX", "PD1", "CD45RO", "HLA.DR")
cd8.Tcells <- runPCA(gobject = cd8.Tcells, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(cd8.Tcells, scale_unit = TRUE, scree_ylim = c(0,40))
cd8.Tcells <- createNearestNetwork(gobject = cd8.Tcells, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:5, k = 15)  # calculate nearest neighbors, focus on global differences 

cd8.Tcells <- doLeidenClusterIgraph(gobject = cd8.Tcells, resolution = 0.2, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd8Tcells.res0.2')  # make the clusters
cd8.Tcells <- doLeidenClusterIgraph(gobject = cd8.Tcells, resolution = 0.3, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd8Tcells.res0.3')  # make the clusters
cd8.Tcells <- doLeidenClusterIgraph(gobject = cd8.Tcells, resolution = 0.5, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd8Tcells.res0.5')  # make the clusters
cd8.Tcells <- doLeidenClusterIgraph(gobject = cd8.Tcells, resolution = 1, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd8Tcells.res1')  # make the clusters
cd8.Tcells <- runUMAPprojection(cd8.Tcells, random_subset = 1e5, n_threads = parallelly::availableCores(), 
                                name="umap.cd8", dimensions_to_use = 1:5)  # project UMAP coordinates to the full dataset  

md <- pDataDT(cd8.Tcells); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = cd8.Tcells, cell_ids = keepCells)

leiden_colors <- getDistinctColors(length(unique(md$leiden_cd8Tcells.res0.5)))
names(leiden_colors) <- unique(md$leiden_cd8Tcells.res0.5)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.cd8', cell_color = 'leiden_cd8Tcells.res0.5',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "GZMB", "CD39", "KI67", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO", "HLA.DR")


plotMetaDataHeatmap(cd8.Tcells, expression_values = "normalized", plot_title = "cd8 res 0.2",
                    metadata_cols = "leiden_cd8Tcells.res0.2", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd8.Tcells$leiden_cd8Tcells.res0.2))

plotMetaDataHeatmap(cd8.Tcells, expression_values = "normalized", plot_title = "cd8 res 0.3",
                    metadata_cols = "leiden_cd8Tcells.res0.3", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd8.Tcells$leiden_cd8Tcells.res0.3))

plotMetaDataHeatmap(cd8.Tcells, expression_values = "normalized", plot_title = "cd8 res 0.5",
                    metadata_cols = "leiden_cd8Tcells.res0.5", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd8.Tcells$leiden_cd8Tcells.res0.5))


plotMetaDataHeatmap(cd8.Tcells, expression_values = "normalized", plot_title = "cd8 res 1",
                    metadata_cols = "leiden_cd8Tcells.res1", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd8.Tcells$leiden_cd8Tcells.res1))

print(paste("Primary object size: ", nrow(md), "cells"))


```

# save cd8 objects 
```{r save-cd8Tcells-res0.5-revised}
# use prior filename so will overwrite with new object
# filename <- paste0("clustering4.cd8tcells.dim5k15.",  strftime(Sys.time() , "%Y%m%d.%H%M%S"), ".RDS")
saveRDS(cd8.Tcells, filename)
md <- pDataDT(cd8.Tcells)
saveRDS(md, paste0("md.",filename))
saveRDS(Giotto::getExpression(cd8.Tcells, values = "scaled", output = c("matrix"),  set_defaults = TRUE ),
        paste0("exprs.",filename))

```


# subcluster CD4 cells
```{r subset-cd4T-cells}
md <- pDataDT(Tcells)
keepCells <- md$cell_ID[which(md$CD8.Positive.Classification == 1)]
keepCells <- subset(md, !(cell_ID %in% keepCells))$cell_ID   # take everything from Tcells object that is not a CD8 T cell 
cd4.Tcells <- subsetGiotto(Tcells, cell_ids = keepCells )

print("subcluster CD4 T cells")
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "GZMB", "CD39", "KI67", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO", "HLA.DR")
cd4.Tcells <- runPCA(gobject = cd4.Tcells, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(cd4.Tcells, scale_unit = TRUE, scree_ylim = c(0,40))
cd4.Tcells <- createNearestNetwork(gobject = cd4.Tcells, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:5, k = 15)  # calculate nearest neighbors, focus on global differences 
cd4.Tcells <- doLeidenClusterIgraph(gobject = cd4.Tcells, resolution = 0.5, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd4Tcells.res0.5')  # make the clusters
cd4.Tcells <- doLeidenClusterIgraph(gobject = cd4.Tcells, resolution = 0.7, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd4Tcells.res0.7')  # make the clusters
cd4.Tcells <- runUMAPprojection(cd4.Tcells, random_subset = 1e5, n_threads = parallelly::availableCores(), 
                                name="umap.cd4", dimensions_to_use = 1:5)  # project UMAP coordinates to the full dataset  


```

# save CD4 objects 
```{r save-cd4Tcells}
filename <- paste0("sln.clustering4.cd4tcells.dim5k15.",  strftime(Sys.time() , "%Y%m%d.%H%M%S"), ".RDS")
saveRDS(cd4.Tcells, filename)
md <- pDataDT(cd4.Tcells)
saveRDS(md, paste0("md.",filename))
saveRDS(Giotto::getExpression(cd4.Tcells, values = "scaled", output = c("matrix"),  set_defaults = TRUE ),
        paste0("exprs.",filename))
print(paste("Primary object size: ", nrow(md), "cells"))

```

# heatmap & umap 
```{r plot-UMAP-heatmap-cd4cells }
md <- pDataDT(cd4.Tcells); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = cd4.Tcells, cell_ids = keepCells)

# ********************* res 0.5 ***********************
leiden_colors <- getDistinctColors(length(unique(md$leiden_cd4Tcells.res0.5)))
names(leiden_colors) <- unique(md$leiden_cd4Tcells.res0.5)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.cd4', cell_color = 'leiden_cd4Tcells.res0.5',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "GZMB", "CD39", "KI67", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO", "HLA.DR")

plotMetaDataHeatmap(cd4.Tcells, expression_values = "normalized",
                    metadata_cols = "leiden_cd4Tcells.res0.5", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd4.Tcells$leiden_cd4Tcells.res0.5))

# ********************* res 0.7 ***********************

leiden_colors <- getDistinctColors(length(unique(md$leiden_cd4Tcells.res0.7)))
names(leiden_colors) <- unique(md$leiden_cd4Tcells.res0.7)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.cd4', cell_color = 'leiden_cd4Tcells.res0.7',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "GZMB", "CD39", "KI67", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO", "HLA.DR")

plotMetaDataHeatmap(cd4.Tcells, expression_values = "normalized",
                    metadata_cols = "leiden_cd4Tcells.res0.7", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd4.Tcells$leiden_cd4Tcells.res0.7))


```


# remove outlier cd4 cluster 
```{r subset-cd4T-cells-again}
gc()
print("remove res0.5 cluster 26")
md <- pDataDT(cd4.Tcells)
finalLabel <- data.frame(cell_ID = subset(md, leiden_cd4Tcells.res0.5 == 26 )$cell_ID, L1_annotations = "CD4.outlier")
saveRDS( object = finalLabel, file = "finalAnnotation.cd4outlier.RDS")  # set aside for later 

keepCells <- subset(md, !(leiden_cd4Tcells.res0.5 == 26 ))$cell_ID
cd4.Tcells <- subsetGiotto(gobject = cd4.Tcells, cell_ids = keepCells)  # take away just those cluster 26 cells 

print("subcluster CD4 T cells")
clusterChannels <- c("TCF1", "TOX", "FOXP3", "CD45RO", "HLA.DR")
cd4.Tcells <- runPCA(gobject = cd4.Tcells, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(cd4.Tcells, scale_unit = TRUE, scree_ylim = c(0,40))
cd4.Tcells <- createNearestNetwork(gobject = cd4.Tcells, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:5, k = 15)  # calculate nearest neighbors, focus on global differences 

cd4.Tcells <- doLeidenClusterIgraph(gobject = cd4.Tcells, resolution = 0.2, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd4Tcells.res0.2')  # make the clusters
cd4.Tcells <- doLeidenClusterIgraph(gobject = cd4.Tcells, resolution = 0.3, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd4Tcells.res0.3')  # make the clusters
cd4.Tcells <- doLeidenClusterIgraph(gobject = cd4.Tcells, resolution = 0.5, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd4Tcells.res0.5')  # make the clusters
cd4.Tcells <- doLeidenClusterIgraph(gobject = cd4.Tcells, resolution = 1, n_iterations = 200, set_seed = T, seed_number = 43,
                                network_name = 'NN.harmony', name = 'leiden_cd4Tcells.res1')  # make the clusters
cd4.Tcells <- runUMAPprojection(cd4.Tcells, random_subset = 1e5, n_threads = parallelly::availableCores(), 
                                name="umap.cd4", dimensions_to_use = 1:5)  # project UMAP coordinates to the full dataset  


md <- pDataDT(cd4.Tcells); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = cd4.Tcells, cell_ids = keepCells)

leiden_colors <- getDistinctColors(length(unique(md$leiden_cd4Tcells.res0.5)))
names(leiden_colors) <- unique(md$leiden_cd4Tcells.res0.5)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.cd4', cell_color = 'leiden_cd4Tcells.res0.5',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "GZMB", "CD39", "KI67", "PD1", "TOX", "FOXP3", "ICOS", "CD45RO", "HLA.DR")


plotMetaDataHeatmap(cd4.Tcells, expression_values = "normalized", plot_title = "cd4 res 0.2",
                    metadata_cols = "leiden_cd4Tcells.res0.2", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd4.Tcells$leiden_cd4Tcells.res0.2))


plotMetaDataHeatmap(cd4.Tcells, expression_values = "normalized", plot_title = "cd4 res 0.3",
                    metadata_cols = "leiden_cd4Tcells.res0.3", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd4.Tcells$leiden_cd4Tcells.res0.3))


plotMetaDataHeatmap(cd4.Tcells, expression_values = "normalized", plot_title = "cd4 res 0.5",
                    metadata_cols = "leiden_cd4Tcells.res0.5", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd4.Tcells$leiden_cd4Tcells.res0.5))


plotMetaDataHeatmap(cd4.Tcells, expression_values = "normalized", plot_title = "cd4 res 1",
                    metadata_cols = "leiden_cd4Tcells.res1", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
print(table(cd4.Tcells$leiden_cd4Tcells.res1))



print(paste("Primary object size: ", nrow(md), "cells"))

```

# save cd4 objects 
```{r save-cd4Tcells-revised}
# use prior filename so will overwrite with new object
# filename <- paste0("clustering4.cd4tcells.dim5k15.",  strftime(Sys.time() , "%Y%m%d.%H%M%S"), ".RDS")
saveRDS(cd4.Tcells, filename)
md <- pDataDT(cd4.Tcells)
saveRDS(md, paste0("md.",filename))
saveRDS(Giotto::getExpression(cd4.Tcells, values = "scaled", output = c("matrix"),  set_defaults = TRUE ),
        paste0("exprs.",filename))

```


# cd4 violin plot
```{r show-violin-cd4Tcells, fig.show="hold", fig.width=10, fig.height=15, warning=F}
md <- pDataDT(cd4.Tcells); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 5e4))$cell_ID
sln.sm <- subsetGiotto(gobject = cd4.Tcells, cell_ids = keepCells)
violinPlot(sln.sm, feats = clusterChannels, cluster_column = "leiden_cd4Tcells.res0.5")

```




```{r call-date}
date()
```


seff 63157801
Job ID: 63157801
Cluster: slurm_cluster
User/Group: heratr01/heratr01
State: COMPLETED (exit code 0)
Nodes: 1
Cores per node: 10
CPU Utilized: 03:21:40
CPU Efficiency: 13.08% of 1-01:41:40 core-walltime
Job Wall-clock time: 02:34:10
Memory Utilized: 106.78 GB
Memory Efficiency: 53.39% of 200.00 GB













