---
title: "Clustering part 5 - Annotations"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```


```{r call-libraries} 
date(); timeStamp <- strftime(Sys.time() , "%Y%m%d.%H%M%S")
library(Giotto)
library(future); if(parallelly::availableCores() > 2 & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); #library(leiden);    
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

```

# read in CD8.Toxhi and CD8.Toxlo
```{r readRDS}
cd8.toxhi <- readRDS("sln.clustering4.cd8tcells.toxhi.dim5k15.20250507.104052.RDS")
cd8.toxlo <- readRDS("sln.clustering4.cd8tcells.toxlo.dim5k15.20250507.104052.RDS")
cd4.foxp3hi <- readRDS("sln.clustering4.cd4tcells.foxp3hi.dim5k15.20250507.104052.RDS")
cd4.foxp3lo <- readRDS("sln.clustering4.cd4tcells.foxp3lo.dim5k15.20250507.104052.RDS")
bcells <- readRDS("sln.clustering4.bcells.harm.dim5k15.20250502.112218.RDS")
tbmixed <- readRDS("sln.clustering4.tbmixed.harm.20250503.150210.RDS")
```


```{r makeList-of-annotations}
L2.annot <- list()
L2.annot[["CD8.toxhi"]] <- pDataDT(cd8.toxhi)[, .(cell_ID, L2_annotations, leiden_tox.hi.cd8)]
names(L2.annot[["CD8.toxhi"]]) <- c("cell_ID","L2_annotations", "L2.num")
L2.annot[["CD8.toxhi"]]$L2.char <- paste0("cd8.toxhi_", L2.annot[["CD8.toxhi"]]$L2.num)

L2.annot[["CD8.toxlo"]] <- pDataDT(cd8.toxlo)[, .(cell_ID, L2_annotations, leiden_tox.lo.cd8)]
names(L2.annot[["CD8.toxlo"]]) <- c("cell_ID","L2_annotations", "L2.num")
L2.annot[["CD8.toxlo"]]$L2.char <- paste0("cd8.toxlo_", L2.annot[["CD8.toxlo"]]$L2.num)

L2.annot[["cd4.foxp3hi"]] <- pDataDT(cd4.foxp3hi)[, .(cell_ID, L2_annotations, leiden_foxp3.hi.cd4)]
names(L2.annot[["cd4.foxp3hi"]]) <- c("cell_ID","L2_annotations", "L2.num")
L2.annot[["cd4.foxp3hi"]]$L2.char <- paste0("cd4.foxp3hi_", L2.annot[["cd4.foxp3hi"]]$L2.num)

L2.annot[["cd4.foxp3lo"]] <- pDataDT(cd4.foxp3lo)[, .(cell_ID, L2_annotations, leiden_foxp3.lo.cd4)]
names(L2.annot[["cd4.foxp3lo"]]) <- c("cell_ID","L2_annotations", "L2.num")
L2.annot[["cd4.foxp3lo"]]$L2.char <- paste0("cd4.foxp3lo_", L2.annot[["cd4.foxp3lo"]]$L2.num)

L2.annot[["bcells"]] <- pDataDT(bcells)[, .(cell_ID, L2_annotations, leiden_Bcells.res0.1)]
names(L2.annot[["bcells"]]) <- c("cell_ID","L2_annotations", "L2.num")
L2.annot[["bcells"]]$L2.char <- paste0("bcells_", L2.annot[["bcells"]]$L2.num)


L2.annot[["tbmixed"]] <- pDataDT(tbmixed)[, .(cell_ID, L2_annotations, leiden_TBmixed.res0.3)]
names(L2.annot[["tbmixed"]]) <- c("cell_ID","L2_annotations", "L2.num")
L2.annot[["tbmixed"]]$L2.char <- paste0("tbmixed_", L2.annot[["tbmixed"]]$L2.num)

L2.annot[["cd4outlier"]] <- readRDS("finalAnnotation.cd4outlier.RDS")
names(L2.annot[["cd4outlier"]]) <- c("cell_ID","L2_annotations")
L2.annot[["cd4outlier"]]$L2.num <- 42; L2.annot[["cd4outlier"]]$L2.char <- "CD4_OUTLIER_1"

L2.annot[["cd8outlier"]] <- readRDS("finalAnnotation.cd8outlier.RDS") 
names(L2.annot[["cd8outlier"]]) <- c("cell_ID","L2_annotations")
L2.annot[["cd8outlier"]]$L2.num <- 42; L2.annot[["cd8outlier"]]$L2.char <- "CD8_OUTLIER_1"


```


# assign L2 annotations
```{r assign_newLabels}
# start with CD8.toxhi
L2.annot[["CD8.toxhi"]]$L2_annotations <- paste0("Clust.", L2.annot[["CD8.toxhi"]]$L2_annotations)  # unknown clusters:
L2.annot[["CD8.toxhi"]][L2.num  == 1, 'L2_annotations':= "TCF1_MIXED_TOX_HI_CD8_1"]
L2.annot[["CD8.toxhi"]][L2.num  == 2, 'L2_annotations':= "TCF1_MIXED_TOX_HI_CD8_2"]
L2.annot[["CD8.toxhi"]][L2.num  == 3, 'L2_annotations':= "TCF1_HI_TOX_HI_CD8_1"]
L2.annot[["CD8.toxhi"]][L2.num  == 4, 'L2_annotations':= "TCF1_LO_TOX_HI_CD8_1"]
L2.annot[["CD8.toxhi"]][L2.num  == 5, 'L2_annotations':= "TCF1_LO_TOX_HI_CD8_2"]
L2.annot[["CD8.toxhi"]][L2.num  == 6, 'L2_annotations':= "TCF1_HI_TOX_HI_CD8_2"]
L2.annot[["CD8.toxhi"]][L2.num  == 7, 'L2_annotations':= "TCF1_HI_TOX_HI_CD8_3"]
L2.annot[["CD8.toxhi"]][L2.num  == 8, 'L2_annotations':= "TCF1_HI_TOX_HI_CD8_4"]
L2.annot[["CD8.toxhi"]][L2.num  == 9, 'L2_annotations':= "TCF1_HI_TOX_HI_CD8_5"]
L2.annot[["CD8.toxhi"]][L2.num  == 10, 'L2_annotations':= "TCF1_HI_TOX_HI_CD8_6"]

L2.annot[["CD8.toxlo"]]$L2_annotations <- paste0("Clust.", L2.annot[["CD8.toxlo"]]$L2_annotations)  # unknown clusters:
L2.annot[["CD8.toxlo"]][L2.num  == 1, 'L2_annotations':= "ACTIVATED_CD8_1"]
L2.annot[["CD8.toxlo"]][L2.num  == 2, 'L2_annotations':= "MEMORY_CD8_1"]
L2.annot[["CD8.toxlo"]][L2.num  == 3, 'L2_annotations':= "MEMORY_CD8_2"]
L2.annot[["CD8.toxlo"]][L2.num  == 4, 'L2_annotations':= "MEMORY_CD8_3"]
L2.annot[["CD8.toxlo"]][L2.num  == 5, 'L2_annotations':= "GZMB_CD8_1"]
L2.annot[["CD8.toxlo"]][L2.num  == 6, 'L2_annotations':= "MIXED_TOX_LO_CD8_1"]
L2.annot[["CD8.toxlo"]][L2.num  == 7, 'L2_annotations':= "MIXED_TOX_LO_CD8_2"]
L2.annot[["CD8.toxlo"]][L2.num  == 8, 'L2_annotations':= "MIXED_TOX_LO_CD8_3"]
L2.annot[["CD8.toxlo"]][L2.num  == 9, 'L2_annotations':= "MIXED_TOX_LO_CD8_4"]
L2.annot[["CD8.toxlo"]][L2.num  == 10, 'L2_annotations':= "GZMB_CD8_2"]
L2.annot[["CD8.toxlo"]][L2.num  == 11, 'L2_annotations':= "NAIVE_CD8_1"]
L2.annot[["CD8.toxlo"]][L2.num  == 12, 'L2_annotations':= "TERMINAL_EFFECTOR_CD8_1"]
L2.annot[["CD8.toxlo"]][L2.num  == 13, 'L2_annotations':= "GZMB_CD8_3"]
L2.annot[["CD8.toxlo"]][L2.num  == 14, 'L2_annotations':= "ACTIVATED_MEMORY_CD8_1"]
L2.annot[["CD8.toxlo"]][L2.num  == 15, 'L2_annotations':= "ACTIVATED_CD8_2"]
L2.annot[["CD8.toxlo"]][L2.num  == 16, 'L2_annotations':= "TERMINAL_EFFECTOR_CD8_2"]


L2.annot[["cd4.foxp3hi"]]$L2_annotations <- paste0("Clust.", L2.annot[["cd4.foxp3hi"]]$L2_annotations)  # unknown clusters:
L2.annot[["cd4.foxp3hi"]][L2.num  == 1, 'L2_annotations':= "ICOS_LO_TREG_1"]
L2.annot[["cd4.foxp3hi"]][L2.num  == 2, 'L2_annotations':= "ICOS_LO_TREG_2"]
L2.annot[["cd4.foxp3hi"]][L2.num  == 3, 'L2_annotations':= "MIXED_TREG_1"]
L2.annot[["cd4.foxp3hi"]][L2.num  == 4, 'L2_annotations':= "MIXED_TREG_2"]
L2.annot[["cd4.foxp3hi"]][L2.num  == 5, 'L2_annotations':= "MIXED_TREG_3"]
L2.annot[["cd4.foxp3hi"]][L2.num  == 6, 'L2_annotations':= "MIXED_TREG_4"]
L2.annot[["cd4.foxp3hi"]][L2.num  == 7, 'L2_annotations':= "ACTIVATED_TREG_1"]
L2.annot[["cd4.foxp3hi"]][L2.num  == 8, 'L2_annotations':= "MIXED_TREG_5"]
L2.annot[["cd4.foxp3hi"]][L2.num  == 9, 'L2_annotations':= "ACTIVATED_TREG_2"]


L2.annot[["cd4.foxp3lo"]]$L2_annotations <- paste0("Clust.", L2.annot[["cd4.foxp3lo"]]$L2_annotations)  # unknown clusters:
L2.annot[["cd4.foxp3lo"]][L2.num  == 1, 'L2_annotations':= "ACTIVATED_CD4_1"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 2, 'L2_annotations':= "ACTIVATED_CD4_2"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 3, 'L2_annotations':= "ACTIVATED_MEMORY_CD4_1"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 4, 'L2_annotations':= "ACTIVATED_CD4_4"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 5, 'L2_annotations':= "MEMORY_CD4_1"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 6, 'L2_annotations':= "ACTIVATED_MEMORY_CD4_2"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 7, 'L2_annotations':= "MEMORY_CD4_2"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 8, 'L2_annotations':= "MEMORY_CD4_3"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 9, 'L2_annotations':= "MEMORY_CD4_4"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 10, 'L2_annotations':= "CD4_MYELOID_NONIMMUNE_MIXED_1"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 11, 'L2_annotations':= "MEMORY_CD4_5"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 12, 'L2_annotations':= "CD4_MYELOID_NONIMMUNE_MIXED_2"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 13, 'L2_annotations':= "CD4_MYELOID_NONIMMUNE_MIXED_3"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 14, 'L2_annotations':= "CD4_MYELOID_NONIMMUNE_MIXED_4"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 15, 'L2_annotations':= "CD4_MYELOID_NONIMMUNE_MIXED_5"]
L2.annot[["cd4.foxp3lo"]][L2.num  == 16, 'L2_annotations':= "MEMORY_CD4_6"]


L2.annot[["bcells"]]$L2_annotations <- paste0("Clust.", L2.annot[["bcells"]]$L2_annotations)  # unknown clusters:
L2.annot[["bcells"]][L2.num  == 1, 'L2_annotations':= "BCELL_MYELOID_NONIMMUNE_MIXED_1"]
L2.annot[["bcells"]][L2.num  == 2, 'L2_annotations':= "PC_4"]
L2.annot[["bcells"]][L2.num  == 3, 'L2_annotations':= "MIXED_FOLLICULAR_BCELL_1"]
L2.annot[["bcells"]][L2.num  == 4, 'L2_annotations':= "MIXED_FOLLICULAR_AND_NONFOLLICULAR_BCELL_1"]
L2.annot[["bcells"]][L2.num  == 5, 'L2_annotations':= "NONIMMUNE_CELL_1"]
L2.annot[["bcells"]][L2.num  == 6, 'L2_annotations':= "KI67_HI_BCELL_1"]
L2.annot[["bcells"]][L2.num  == 7, 'L2_annotations':= "MIXED_FOLLICULAR_BCELL_2"]
L2.annot[["bcells"]][L2.num  == 8, 'L2_annotations':= "MIXED_FOLLICULAR_BCELL_3"]
L2.annot[["bcells"]][L2.num  == 9, 'L2_annotations':= "MANTLE_ZONE_BCELL_1"]
L2.annot[["bcells"]][L2.num  == 10, 'L2_annotations':= "BCELL_MYELOID_NONIMMUNE_MIXED_2"]
L2.annot[["bcells"]][L2.num  == 11, 'L2_annotations':= "NONIMMUNE_CELL_2"]
L2.annot[["bcells"]][L2.num  == 12, 'L2_annotations':= "NONIMMUNE_CELL_3"]
L2.annot[["bcells"]][L2.num  == 13, 'L2_annotations':= "BCELL_MYELOID_NONIMMUNE_MIXED_3"]
L2.annot[["bcells"]][L2.num  == 14, 'L2_annotations':= "MIXED_FOLLICULAR_AND_NONFOLLICULAR_BCELL_2"]
L2.annot[["bcells"]][L2.num  == 15, 'L2_annotations':= "MIXED_FOLLICULAR_AND_NONFOLLICULAR_BCELL_3"]
L2.annot[["bcells"]][L2.num  == 16, 'L2_annotations':= "MIXED_FOLLICULAR_BCELL_4"]
L2.annot[["bcells"]][L2.num  == 17, 'L2_annotations':= "MIXED_FOLLICULAR_AND_NONFOLLICULAR_BCELL_4"]

L2.annot[["tbmixed"]]$L2_annotations <- paste0("Clust.", L2.annot[["tbmixed"]]$L2_annotations)  # unknown clusters:
L2.annot[["tbmixed"]][L2.num  == 1, 'L2_annotations':= "CD4_CD20_MIXED_1"]
L2.annot[["tbmixed"]][L2.num  == 2, 'L2_annotations':= "PC_1"]
L2.annot[["tbmixed"]][L2.num  == 3, 'L2_annotations':= "CD4_CD8_CD20_MIXED_1"]
L2.annot[["tbmixed"]][L2.num  == 4, 'L2_annotations':= "CD8_PC_MIXED_1"]
L2.annot[["tbmixed"]][L2.num  == 5, 'L2_annotations':= "CD8_CD20_MIXED_1"]
L2.annot[["tbmixed"]][L2.num  == 6, 'L2_annotations':= "TCELL_BCELL_MYELOID_MIXED_1"]
L2.annot[["tbmixed"]][L2.num  == 7, 'L2_annotations':= "CD4_CD20_MIXED_2"]
L2.annot[["tbmixed"]][L2.num  == 8, 'L2_annotations':= "BCELL_MYELOID_MIXED_1"]
L2.annot[["tbmixed"]][L2.num  == 9, 'L2_annotations':= "CD4_CD20_MIXED_3"]
L2.annot[["tbmixed"]][L2.num  == 10, 'L2_annotations':= "CD4_CD20_MIXED_4"]
L2.annot[["tbmixed"]][L2.num  == 11, 'L2_annotations':= "CD8_CD20_MIXED_2"]
L2.annot[["tbmixed"]][L2.num  == 12, 'L2_annotations':= "PC_2"]
L2.annot[["tbmixed"]][L2.num  == 13, 'L2_annotations':= "CD4_CD20_MIXED_5"]
L2.annot[["tbmixed"]][L2.num  == 14, 'L2_annotations':= "PC_3"]
L2.annot[["tbmixed"]][L2.num  == 15, 'L2_annotations':= "CD4_CD20_MIXED_6"]
L2.annot[["tbmixed"]][L2.num  == 16, 'L2_annotations':= "CD8_CD20_MIXED_3"]
L2.annot[["tbmixed"]][L2.num  == 17, 'L2_annotations':= "BCELL_MYELOID_MIXED_2"]
L2.annot[["tbmixed"]][L2.num  == 18, 'L2_annotations':= "LZ_BCELL_1"]
L2.annot[["tbmixed"]][L2.num  == 19, 'L2_annotations':= "CD4_CD20_MIXED_7"]
L2.annot[["tbmixed"]][L2.num  == 20, 'L2_annotations':= "CD4_CD20_MIXED_8"]
L2.annot[["tbmixed"]][L2.num  == 21, 'L2_annotations':= "MIXED_GC_TCELL_1"]
L2.annot[["tbmixed"]][L2.num  == 22, 'L2_annotations':= "BCELL_MYELOID_MIXED_3"]
L2.annot[["tbmixed"]][L2.num  == 23, 'L2_annotations':= "CD4_CD8_CD20_MIXED_2"]
L2.annot[["tbmixed"]][L2.num  == 24, 'L2_annotations':= "CD4_CD20_MIXED_9"]
L2.annot[["tbmixed"]][L2.num  == 25, 'L2_annotations':= "ARTIFACT_1"]
L2.annot[["tbmixed"]][L2.num  == 26, 'L2_annotations':= "ARTIFACT_2"]
```


```{r assign_newLabels2}
L2.annot <- rbindlist(L2.annot)
```

# read in SLN 
```{r readRDS-full-sln-from-clustering2}
sln <- readRDS("sln.clustering2.res0.5.20250411.1611.RDS")
md <- pDataDT(sln) 
print(paste("Primary object size: ", nrow(md), "cells"))
print("cluster membership")
print(table(md$leiden_harmony))

df <- data.frame(leiden_harmony = md$leiden_harmony, L1_annotations = "Unknown")   # consider re-implementing as data.table
df$cell_ID <- md$cell_ID
df$L1_annotations <- paste0("Clust.", df$leiden_harmony)  # unknown clusters
df[which(df$leiden_harmony == "1" ), 'L1_annotations'] <-  "BCells.1"
df[which(df$leiden_harmony == "2" ), 'L1_annotations'] <-  "BCells.2"
df[which(df$leiden_harmony == "3" ), 'L1_annotations'] <-  "BCells.3"
df[which(df$leiden_harmony == "4" ), 'L1_annotations'] <-  "BCells.4"
df[which(df$leiden_harmony == "5" ), 'L1_annotations'] <-  "BCells.5"
df[which(df$leiden_harmony == "6" ), 'L1_annotations'] <-  "BCells.6"
df[which(df$leiden_harmony == "7" ), 'L1_annotations'] <-  "TBMixed.1"
df[which(df$leiden_harmony == "8" ), 'L1_annotations'] <-  "TBMixed.2"
df[which(df$leiden_harmony == "9" ), 'L1_annotations'] <-  "TBMixed.3"
df[which(df$leiden_harmony == "10" ), 'L1_annotations'] <-  "TBMixed.4"
df[which(df$leiden_harmony == "11" ), 'L1_annotations'] <-  "TBMixed.5"
df[which(df$leiden_harmony == "12" ), 'L1_annotations'] <-  "TBMixed.6"
df[which(df$leiden_harmony == "13" ), 'L1_annotations'] <-  "TBMixed.7"
df[which(df$leiden_harmony == "14" ), 'L1_annotations'] <-  "BCells.7"
df[which(df$leiden_harmony == "15" ), 'L1_annotations'] <-  "TBMixed.8"
df[which(df$leiden_harmony == "16" ), 'L1_annotations'] <-  "TBMixed.9"
df[which(df$leiden_harmony == "17" ), 'L1_annotations'] <-  "TBMixed.10"
df[which(df$leiden_harmony == "18" ), 'L1_annotations'] <-  "TBMixed.11"
df[which(df$leiden_harmony == "19" ), 'L1_annotations'] <-  "TBMixed.12"
df[which(df$leiden_harmony == "20" ), 'L1_annotations'] <-  "TBMixed.13"
df[which(df$leiden_harmony == "21" ), 'L1_annotations'] <-  "BCells.8"
df[which(df$leiden_harmony == "22" ), 'L1_annotations'] <-  "BCells.9"
df[which(df$leiden_harmony == "23" ), 'L1_annotations'] <-  "BCells.10"
df[which(df$leiden_harmony == "24" ), 'L1_annotations'] <-  "BCells.11"
df[which(df$leiden_harmony == "25" ), 'L1_annotations'] <-  "TBMixed.14"
df[which(df$leiden_harmony == "26" ), 'L1_annotations'] <-  "TBMixed.15"
df[which(df$leiden_harmony == "27" ), 'L1_annotations'] <-  "TBMixed.16"
df[which(df$leiden_harmony == "28" ), 'L1_annotations'] <-  "BCells.12"
df[which(df$leiden_harmony == "29" ), 'L1_annotations'] <-  "BCells.13"
df[which(df$leiden_harmony == "30" ), 'L1_annotations'] <-  "TBMixed.17"
df[which(df$leiden_harmony == "31" ), 'L1_annotations'] <-  "TBMixed.18"
df[which(df$leiden_harmony == "32" ), 'L1_annotations'] <-  "TBMixed.19"
df[which(df$leiden_harmony == "33" ), 'L1_annotations'] <-  "TBMixed.20"
df[which(df$leiden_harmony == "34" ), 'L1_annotations'] <-  "BCells.14"
df[which(df$leiden_harmony == "35" ), 'L1_annotations'] <-  "TBMixed.21"
df[which(df$leiden_harmony == "36" ), 'L1_annotations'] <-  "Tumor.1"
df[which(df$leiden_harmony == "37" ), 'L1_annotations'] <-  "Tumor.2"
df[which(df$leiden_harmony == "38" ), 'L1_annotations'] <-  "BCells.15"
df[which(df$leiden_harmony == "39" ), 'L1_annotations'] <-  "Tumor.3"
df[which(df$leiden_harmony == "40" ), 'L1_annotations'] <-  "Tumor.4"
df[which(df$leiden_harmony == "41" ), 'L1_annotations'] <-  "BCells.16"
df[which(df$leiden_harmony == "42" ), 'L1_annotations'] <-  "TBMixed.22"
df[which(df$leiden_harmony == "43" ), 'L1_annotations'] <-  "BCells.17"
df[which(df$leiden_harmony == "44" ), 'L1_annotations'] <-  "TBMixed.23"
df[which(df$leiden_harmony == "45" ), 'L1_annotations'] <-  "BCells.18"
df[which(df$leiden_harmony == "46" ), 'L1_annotations'] <-  "TBMixed.24"
df[which(df$leiden_harmony == "47" ), 'L1_annotations'] <-  "TBMixed.25"
df[which(df$leiden_harmony == "48" ), 'L1_annotations'] <-  "TBMixed.26"
df[which(df$leiden_harmony == "49" ), 'L1_annotations'] <-  "TCells.1"
df[which(df$leiden_harmony == "50" ), 'L1_annotations'] <-  "BCells.19"
df[which(df$leiden_harmony == "51" ), 'L1_annotations'] <-  "TBMixed.27"
df[which(df$leiden_harmony == "52" ), 'L1_annotations'] <-  "Artifact.1"
df[which(df$leiden_harmony == "53" ), 'L1_annotations'] <-  "TBMixed.28"
df[which(df$leiden_harmony == "54" ), 'L1_annotations'] <-  "Artifact.2"
df[which(df$leiden_harmony == "55" ), 'L1_annotations'] <-  "TCells.2"
df[which(df$leiden_harmony == "56" ), 'L1_annotations'] <-  "TBMixed.29"
df[which(df$leiden_harmony == "57" ), 'L1_annotations'] <-  "TBMixed.30"
df[which(df$leiden_harmony == "58" ), 'L1_annotations'] <-  "BCells.20"
df[which(df$leiden_harmony == "59" ), 'L1_annotations'] <-  "Tumor.5"
df[which(df$leiden_harmony == "60" ), 'L1_annotations'] <-  "BCells.21"
df[which(df$leiden_harmony == "61" ), 'L1_annotations'] <-  "Artifact.3"
df[which(df$leiden_harmony == "62" ), 'L1_annotations'] <-  "TBMixed.31"
      
sln <- addCellMetadata(sln, new_metadata = df, by_column = T, column_cell_ID="cell_ID")  # add in the annotations to the main object


md <- pDataDT(sln) 
print(table(md$L1_annotations))

toEliminate <- c("Artifact.1","Artifact.2", "Artifact.3")  # only take away Artifact clusters now, leave Tumor
keepCells <- subset(md, !(L1_annotations %in% c(toEliminate)))$cell_ID
sln <- subsetGiotto(gobject = sln, cell_ids = keepCells) 

md <- pDataDT(sln) 
print(table(md$L1_annotations))

missing <- setdiff(md$cell_ID, L2.annot$cell_ID)  # what have we left out  
md[cell_ID %in% missing, 'L1_annotations'] %>% table  # group them by prior annotation

```

# Format the metadata
```{r format-cellMetadata}
md <- pDataDT(sln)
df.new <- md[, .(cell_ID, L1_annotations)]
df.new$L2_annotations <- df.new$L3_annotations <- df.new$L1_annotations  # save the L1 annotations 
names(L2.annot) <- c("cell_ID","L3_annotations","L3.num","L3.char")  # changed my mind, let's call these L3_annotations
df.new <- merge(x = df.new, y=L2.annot, by='cell_ID', all.x = T)
df.new[!is.na(L3_annotations.y), L2_annotations := L3_annotations.y]  # replace L2_annot if L3_annot is not NA 
df.new[!is.na(L3.char), L1_annotations := L3.char]  # replace L2_annot if L3_annot is not NA 
df.new[, L1_annotations := sub("_.*", "", L1_annotations)]
df.new$L3_annotations.x <- NULL
names(df.new)[4] <- 'L3_annotations'
df.new$L2_annotations <- substr(df.new$L2_annotations, 1, nchar(df.new$L2_annotations) - 2)  # eliminate the last underscore & number
df.new[L2_annotations == 'CD4.outli', L2_annotations := 'CD4.outlier']
df.new[L2_annotations == 'CD8.outli', L2_annotations := 'CD8.outlier']
df.new[L1_annotations == 'CD4', L1_annotations := 'cd4.foxp3lo']
df.new[L1_annotations == 'CD8', L1_annotations := 'cd8.toxlo']
df.new %>% head
```

# addCellMetadata and check output
```{r add-CellMetadata}
sln <- addCellMetadata(gobject = sln, new_metadata = df.new, by_column=T, column_cell_ID = 'cell_ID')
# rm(df)

md <- pDataDT(sln)
md$L1_annotations %>% table
print(md$L1_annotations %>% table %>% length)

md$L2_annotations %>% table
print(md$L2_annotations %>% table %>% length)

md$L3_annotations %>% table
print(md$L3_annotations %>% table %>% length)

```

# calculate spatialKNN for later InteractionChangedFeats
```{r add-spatialKNN}
sln <- createSpatialKNNnetwork(sln, k=5, method = "dbscan",  name = "spatialknn.k5", verbose = T) 
```

# UMAP visualizations
```{r plotUMAP-by-annotations, fig.width=15}
md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 2e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)

print("L1 annotations")
leiden_colors <- getDistinctColors(length(unique(md$L1_annotations)))
names(leiden_colors) <- unique(md$L1_annotations)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.projection', cell_color = 'L1_annotations',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)


print("L2 annotations")
leiden_colors <- getDistinctColors(length(unique(md$L2_annotations)))
names(leiden_colors) <- unique(md$L2_annotations)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.projection', cell_color = 'L2_annotations',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5, legend_text=5)

# print("L3 annotations")    # *********** this block fails with "Viewport has zero dimensions" --> not clear why
# leiden_colors <- getDistinctColors(length(unique(md$L3_annotations)))
# names(leiden_colors) <- unique(md$L3_annotations)
# plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.clust3.d8', cell_color = 'L3_annotations',
#          cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
#          point_alpha=0.5)

md <- pDataDT(sln); keepCells <- subset(md, !(L2_annotations %in% c('CD4.outli','CD8.outli')))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)

clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")
plotMetaDataHeatmap(sln.sm, expression_values = "normalized", title = "L1 annotations", 
                    metadata_cols = "L1_annotations", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
plotMetaDataHeatmap(sln.sm, expression_values = "normalized", title = "L2 annotations",
                    metadata_cols = "L2_annotations", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
```

# tissue projection for SLN 5
```{r tissue-projection, fig.height=10}
md <- pDataDT(sln); keepCells <- subset(md, list_ID == "tumor.5" & y %in% c(10000:20000) )$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$L1_annotations)))
names(leiden_colors) <- unique(md$L1_annotations)
spatPlot2D(sln.sm, cell_color = 'L1_annotations', cell_color_code = leiden_colors, 
                         point_size = 0.75, point_shape="no_border", point_alpha=0.3, axis_text=14, 
                         legend_text = 14)
```


# final save object 
```{r save-SLN-object}
filename = 
saveRDS(object = sln,  file = paste0("sln.clustering5.annot.spatnet.", timeStamp, ".RDS"))
md <- pDataDT(sln)
saveRDS(object = md, file = paste0("md.sln.clustering5.annot.spatnet.", timeStamp, ".RDS"))
saveRDS(object = Giotto::getExpression(sln, values = "scaled", output = c("matrix"),  set_defaults = TRUE ),
        file = paste0("exprs.sln.clustering5.annot.spatnet.", timeStamp, ".RDS"))
```


```{r call-date}
date()
```


Job ID: 63754742
Cluster: slurm_cluster
User/Group: heratr01/heratr01
State: COMPLETED (exit code 0)
Nodes: 1
Cores per node: 10
CPU Utilized: 01:02:14
CPU Efficiency: 15.52% of 06:41:00 core-walltime
Job Wall-clock time: 00:40:06
Memory Utilized: 157.56 GB
Memory Efficiency: 78.78% of 200.00 GB









