---
title: "Data loading and harmony integration"
output:
  html_document: 
    df_print: default
    toc: true
---

Plan for 10 cores, 200 Gb RAM, and ~13 hours of time on command line on HPC

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```

```{r call-libraries}
date()
library(Giotto)
library(future); if(parallelly::availableCores() > 2 & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); library(leiden);     
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(harmony)
library(dbscan)
library(dplyr); library(leidenAlg)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

```


```{r set-up-and-read-objects}

results_folder <- "output"
instructions <- createGiottoInstructions(save_plot = TRUE, show_plot = TRUE, return_plot = FALSE, save_dir = results_folder)
files <- list.files("halo_export/")
files <- paste0("halo_export/",files)
names(files) <- c("ctrl.1","tumor.4", "tumor.5","ctrl.2","tumor.7","tumor.6","tumor.8","ctrl.3","tumor.9","tumor.10","ctrl.4","ctrl.5", "tumor.1","tumor.2","tumor.3")

# trimmed down version for testing 
# names(files) <- c("ctrl.1","tumor.4", "tumor.5","ctrl.2","tumor.7","tumor.8","tumor.9")

cleanHALO.local <- function(halo.obj, label=paste0(sample(letters, 5, T), collapse = ""), saveColumns)
{  
    # Ensure input is a data.table
  if (!data.table::is.data.table(halo.obj)) {   data.table::setDT(halo.obj)   }

  # Clean column names
  data.table::setnames(halo.obj, names(halo.obj), gsub(" ", ".", names(halo.obj)))

  # Extract Expression Data
  cellIntensity <- grep("Cell.Intensity", names(halo.obj), value = TRUE)
  expr <- setDF(halo.obj[, .SD, .SDcols = cellIntensity])
  names(expr) <- gsub("\\.Cell.Intensity$", "", names(expr))
  rownames(expr) <- paste0(label, "_", halo.obj$Object.Id)

  # Extract Locations
  locations <- data.table::data.table(
    x = rowMeans(halo.obj[, c('XMax', 'XMin'), with = FALSE]),
    y = rowMeans(halo.obj[, c('YMax', 'YMin'), with = FALSE])
  )

  # Extract Metadata
  saveColumns <-   gsub(" ", ".", saveColumns)
  metadata <- halo.obj[, .SD, .SDcols = saveColumns]
  metadata[, cellID := paste0(label, "_", Object.Id)]
  metadata[, x := locations$x]
  metadata[, y := locations$y]

  # Clean NaNs
  for (j in seq_len(ncol(metadata))) {
    set(metadata, which(is.nan(metadata[[j]])), j, 0)
  }

  return(list(expr = t(expr), metadata = metadata, locations = locations))
}

makeGiottoObject.local <- function(inputData, label, instructions)
{
  out <- inputData
  print(paste0("Making the Giotto object for ", label))
  sln <- createGiottoObject(expression = out$expr,  spatial_locs = out$locations, instructions = instructions)
  print("Adding cell metadata"); # gc()  
  sln <- addCellMetadata(sln, new_metadata = out$metadata, by_column = TRUE, column_cell_ID = "cellID" )
  cell_metadata <- pDataDT(sln) # pull metadata out into new object
  
  #filter out undesirable cells
  keepID <- cell_metadata[Classifier.Label != "Background" & Classifier.Label != "Artifact", ]$cell_ID
  sln <- subsetGiotto(sln, cell_ids = keepID)
  
  print("Starting normalization")
  sln <- normalizeGiotto(gobject = sln, scalefactor = 6000, norm_methods = "standard",
                                verbose = TRUE, log_norm = FALSE, library_size_norm = FALSE,
                                scale_feats = FALSE, scale_cells = TRUE)
  print("Finished normalization")
  sln <- addStatistics(gobject = sln, expression_values = "normalized")
  return(sln)
}


out <- list()
for(i in 1:length(files))
{
  a <- fread(files[i])
  cols <- names(a)[colnames(a) %like% paste(c('Object Id','Classifier Label','Positive Classification$', "Cell Area", "Tregs", "Activated Tregs", 
                                              "Tfh", "GZMB CD8", "Exhausted CD8", "Plasma Cells", "IFNy CD4 T Cells", "IFNy CD8 T cells", 
                                              "TOX CD8 T Cells"), collapse = '|')]
  temp <- cleanHALO.local(halo.obj = fread(files[i]),
            label=names(files)[i],
            saveColumns=cols)
  out[[i]] <- makeGiottoObject.local( inputData = temp,
                                label = names(files)[i], 
                                instructions= instructions)
  rm(temp)
  rm(a)
  rm(cols)
}


# sln <- adjustGiottoMatrix(gobject = sln, expression_values = "normalized",
#                                  batch_columns = "sample_Xtile_Ytile", covariate_columns = NULL,
#                                  return_gobject = TRUE, update_slot = "custom")

## visualize
spatPlot(gobject = out[[3]], point_size = 0.05, point_shape = "no_border", save_param = list(save_name="1_exampleSLN5"))
```

Joined objects check
```{r join-objects, fig.width=20, fig.height=25}
sln <- joinGiottoObjects(out, gobject_names = names(files), 
                         x_shift = c(0, 27000, 35000, 45000,70000,90000,129000,167000,
                                     214000,255000,310000,340000,390000,410000,450000), 
                         join_method = "shift", dry_run = F, verbose = TRUE)
spatPlot2D(sln, point_size=0.05, point_shape = "no_border", group_by = 'list_ID', show_legend = F )
rm(out)

```



```{r plot-ScreePlot}
# using channels that Sabrina used for downstream clustering
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")
sln <- runPCA(gobject = sln, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(sln, scale_unit = TRUE, scree_ylim = c(0,40))
```


```{r plot-PCA}
Giotto::plotPCA(gobject = sln, point_shape = "no_border", point_size = 0.2)
```



Calculate clusters by Leiden
```{r calculate-Clusters-nointegrate}
sln.sm <- subsetGiotto(gobject = sln, cell_ids = sample(sln$cell_ID, 1e6, replace=F)) # make a smaller dataset to work with 

sln.sm <- createNearestNetwork(gobject = sln.sm, dimensions_to_use = 1:10, k = 15)
sln.sm <- doLeidenClusterIgraph(gobject = sln.sm, resolution = 0.1, n_iterations = 100, name = 'trnsfr_leiden_harmony')

# sln.sm <- createNearestNetwork(gobject = sln.sm, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
#                             name = 'NN.harmony', dimensions_to_use = 1:10, k = 10)  # calculate nearest neighbors, focus on global differences 
# sln.sm <- doLeidenClusterIgraph(gobject = sln.sm, resolution = 0.3, n_iterations = 100, 
#                                 network_name = 'NN.harmony', name = 'leiden_harmony')  # make the clusters
# sln <- labelTransfer(x = sln, y = sln.sm, labels = "leiden_harmony")  # assign cluster labels in the full dataset based on the reduced data
```

UMAP projection
```{r UMAP-projections-nointegrate}
sln.sm <- runUMAPprojection(sln.sm, random_subset = 1e5)  # project UMAP coordinates to the full dataset based on a subsampling
```


Plot all of the tissues horizontally preintegration
```{r plot-tissues-horizontallys-nointegrate, fig.width=20, fig.height=15}
md <- pDataDT(sln.sm)
leiden_colors <- getDistinctColors(length(unique(md$trnsfr_leiden_harmony)))
names(leiden_colors) <- unique(md$trnsfr_leiden_harmony)
spatPlot(gobject = sln.sm, cell_color = "trnsfr_leiden_harmony", point_shape = "no_border", point_size = 0.1, 
         cell_color_code = leiden_colors, coord_fix_ratio = 1, label_size = 2)
```



UMAP plot of all the cells preintegration
```{r plot-UMAP-nointegrate}
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.projection', cell_color = 'trnsfr_leiden_harmony',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.5, point_shape="no_border",
         point_alpha=0.5)
```



Harmony integration
```{r harmony-integration}
sln = runGiottoHarmony(sln, vars_use = 'list_ID')
```

Calculate clusters by Leiden
```{r calculate-Clusters}
# sln.sm <- subsetGiotto(gobject = sln, cell_ids = sample(sln$cell_ID, 1e5, replace=F)) # make a smaller dataset to work with 

sln <- createNearestNetwork(gobject = sln, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony',
                            name = 'NN.harmony', dimensions_to_use = 1:10, k = 15)
sln <- doLeidenClusterIgraph(gobject = sln, resolution = 0.1, n_iterations = 100, network_name = 'NN.harmony',
                             name = 'trnsfr_leiden_harmony')

# sln.sm <- createNearestNetwork(gobject = sln.sm, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
#                             name = 'NN.harmony', dimensions_to_use = 1:10, k = 10)  # calculate nearest neighbors, focus on global differences 
# sln.sm <- doLeidenClusterIgraph(gobject = sln.sm, resolution = 0.3, n_iterations = 100, 
#                                 network_name = 'NN.harmony', name = 'leiden_harmony')  # make the clusters
# sln <- labelTransfer(x = sln, y = sln.sm, labels = "leiden_harmony")  # assign cluster labels in the full dataset based on the reduced data
```

UMAP projection
```{r UMAP-projection}
sln <- runUMAPprojection(sln, random_subset = 1e5)  # project UMAP coordinates to the full dataset based on a subsampling
```

Plot all of the integrated tissues horizontally
```{r plot-tissues-horizontally, fig.width=20, fig.height=15}
md <- pDataDT(sln)
leiden_colors <- getDistinctColors(length(unique(md$trnsfr_leiden_harmony)))
names(leiden_colors) <- unique(md$trnsfr_leiden_harmony)
spatPlot(gobject = sln, cell_color = "trnsfr_leiden_harmony", point_shape = "no_border", point_size = 0.1, 
         cell_color_code = leiden_colors, coord_fix_ratio = 1, label_size = 2)
```


UMAP plot of all the cells 
```{r plot-UMAP}
plotUMAP(gobject = sln, dim_reduction_name = 'umap.projection', cell_color = 'trnsfr_leiden_harmony',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.5, point_shape="no_border",
         point_alpha=0.5)
```

Plot the umap with a contour overlay
```{r plot-UMAP-customContour}
# data <- as.data.frame( sln@dimension_reduction$cells$cell$rna$umap$umap.projection@coordinates)
# data$cluster <- sln@cell_metadata$cell$rna$trnsfr_leiden_harmony
# leiden_colors <- getDistinctColors(length(unique(data$cluster)))
# names(leiden_colors) <- unique(data$cluster)
# data <- data[sample(1:nrow(data),3e5), ]
# ggplot(data) + geom_point(aes(x=Dim.1, y=Dim.2, color=cluster), show.legend = FALSE, alpha=0.2, size=0.2) + 
#   geom_point(aes(x = Inf, y = Inf, color = cluster), size = 3, show.legend = TRUE) + 
#   geom_density2d(aes(x=Dim.1, y=Dim.2), linewidth=0.25, color="black") + theme_classic()  + scale_color_manual(values = leiden_colors)   + 
#   scale_x_continuous(limits=c(min(data$Dim.1) - .10*diff(range(data$Dim.1)),max(data$Dim.1) + .10* diff(range(data$Dim.1)))) +
#   scale_y_continuous(limits=c(min(data$Dim.2) - .10*diff(range(data$Dim.2)),max(data$Dim.2) + .10* diff(range(data$Dim.2)))) + 
#   ggtitle("All cells - downsampled") +  guides(color = guide_legend(override.aes = list(shape = 16)))
# rm(data) 
```


```{r save-harmony-RDS}
filename <- paste0("sln.harmony.",  strftime(Sys.time() , "%Y%m%d.%H%M%S"), ".RDS")
saveRDS(sln, filename)
```























