---
title: "Full object subsetting"
output:
  html_document:
    df_print: paged    
    toc: TRUE

---

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```

```{r call-libraries} 
library(Giotto)
library(future); if(parallelly::availableCores() > 2 & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); library(leiden);    
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

```

```{r import-sln-meta}
sln <- readRDS("sln.harmony.meta.RDS")
print(paste("Primary object size: ", length(sln@cell_ID$cell), "cells"))
```


```{r plot-UMAP-AllCells, out.width="50%", fig.show="hold"}
md <- pDataDT(sln) 
leiden_colors <- getDistinctColors(length(unique(md$trnsfr_leiden_harmony)))
names(leiden_colors) <- unique(md$trnsfr_leiden_harmony)
if(.Platform$OS.type != "windows")  # only try to plot this if we're not in Windows (so presumably on HPC)
{ plotUMAP(gobject = sln, dim_reduction_name = 'umap.projection', cell_color = 'trnsfr_leiden_harmony',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)
  }

md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 2e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
data <- as.data.frame( sln.sm@dimension_reduction$cells$cell$rna$umap$umap.projection@coordinates)
data$cluster <- sln.sm@cell_metadata$cell$rna$trnsfr_leiden_harmony
data <- data[sample(1:nrow(data),1e5), ]
ggplot(data) + geom_point(aes(x=Dim.1, y=Dim.2, color=cluster), show.legend = FALSE, alpha=0.2, size=0.2) + 
  geom_point(aes(x = Inf, y = Inf, color = cluster), size = 3, show.legend = TRUE) + 
  geom_density2d(aes(x=Dim.1, y=Dim.2), linewidth=0.25, color="black") + theme_classic()  + scale_color_manual(values = leiden_colors)   + 
  scale_x_continuous(limits=c(min(data$Dim.1) - .10*diff(range(data$Dim.1)),max(data$Dim.1) + .10* diff(range(data$Dim.1)))) +
  scale_y_continuous(limits=c(min(data$Dim.2) - .10*diff(range(data$Dim.2)),max(data$Dim.2) + .10* diff(range(data$Dim.2)))) + 
  ggtitle("All cells - downsampled") +  guides(color = guide_legend(override.aes = list(shape = 16)))
rm(data) 

```


```{r plot-UMAP-Downsample}
md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$trnsfr_leiden_harmony)))
names(leiden_colors) <- unique(md$trnsfr_leiden_harmony)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.projection', cell_color = 'trnsfr_leiden_harmony',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

```


```{r cluster-feat-heatmap-AllCells, fig.show="hold", out.width="50%"}
markers_scran <- findMarkers_one_vs_all(gobject = sln, method = "scran", expression_values = "normalized",
                                        cluster_column = "trnsfr_leiden_harmony", min_feats = 3)

topgenes_scran <- unique(markers_scran[, head(.SD, 5), by = "cluster"][["feats"]])

plotMetaDataHeatmap(sln, expression_values = "normalized",
                    metadata_cols = "trnsfr_leiden_harmony", selected_feats = topgenes_scran,
                    y_text_size = 8, show_values = "zscores_rescaled")

clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")

plotMetaDataHeatmap(sln, expression_values = "normalized",
                    metadata_cols = "trnsfr_leiden_harmony", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")

```


```{r eliminate-most-negative-cluster-exl1}
df <- calculateMetaTable( sln,  expression_values = c("normalized"), metadata_cols = 'trnsfr_leiden_harmony' )
df <- df %>% group_by(trnsfr_leiden_harmony) %>% summarise(sum_value = mean(value, na.rm = TRUE), .groups = 'drop')
df <- df[order(df$sum_value, decreasing = F),]
toEliminate <- df[1:2,]$trnsfr_leiden_harmony
md <- pDataDT(sln); keepCells <- subset(md, !(trnsfr_leiden_harmony %in% c(toEliminate)))$cell_ID  # exclude clusters because pan-neg
sln <- subsetGiotto(gobject = sln, cell_ids = keepCells)
print(paste("Primary object size: ", length(sln@cell_ID$cell), "cells"))
```


```{r plot-ScreePlot-excl1}
# using channels that Sabrina used for downstream clustering
print("recalculate clusters post-exclusion")
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")
sln <- runPCA(gobject = sln, expression_values = "normalized", scale_unit = TRUE, 
              method = "factominer", feats_to_use = clusterChannels)
signPCA(sln, scale_unit = TRUE, scree_ylim = c(0,40))
```


Calculate clusters by Leiden
```{r calculate-Clusters-excl1}

sln.sm <- subsetGiotto(gobject = sln, cell_ids = sample(sln$cell_ID, 1e5, replace=F)) # make a smaller dataset to work with 
sln.sm <- createNearestNetwork(gobject = sln.sm, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:10, k = 10)  # calculate nearest neighbors, focus on global differences 
sln.sm <- doLeidenClusterIgraph(gobject = sln.sm, resolution = 0.3, n_iterations = 100, set.seed=T, seed_number = 43, 
                                network_name = 'NN.harmony', name = 'leiden_harmony')  # make the clusters
sln <- labelTransfer(x = sln, y = sln.sm, labels = "leiden_harmony")  # assign cluster labels in the full dataset based on the reduced data
```

UMAP projection
```{r UMAP-projection-excl1}
sln <- runUMAPprojection(sln, random_subset = 1e5)  # project UMAP coordinates to the full dataset based on a subsampling
```


```{r save-excl-1}
saveRDS(sln, "sln.harmony.excl-1.RDS")
```




```{r readRDS-excl-1}
sln <- readRDS("sln.harmony.excl-1.RDS")

```


```{r plot-UMAP-Downsample-excl1}
md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$trnsfr_leiden_harmony)))
names(leiden_colors) <- unique(md$trnsfr_leiden_harmony)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.projection', cell_color = 'trnsfr_leiden_harmony',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

```


```{r cluster-feat-heatmap-excl1-Cells, fig.show="hold", out.width="50%"}
markers_scran <- findMarkers_one_vs_all(gobject = sln, method = "scran", expression_values = "normalized",
                                        cluster_column = "trnsfr_leiden_harmony", min_feats = 3)

topgenes_scran <- unique(markers_scran[, head(.SD, 5), by = "cluster"][["feats"]])

plotMetaDataHeatmap(sln, expression_values = "normalized",
                    metadata_cols = "trnsfr_leiden_harmony", selected_feats = topgenes_scran,
                    y_text_size = 8, show_values = "zscores_rescaled")
clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")

plotMetaDataHeatmap(sln, expression_values = "normalized",
                    metadata_cols = "trnsfr_leiden_harmony", selected_feats = clusterChannels,
                    y_text_size = 8, show_values = "zscores_rescaled")
```



```{r eliminate-most-negative-cluster}
df <- calculateMetaTable( sln,  expression_values = c("normalized"), metadata_cols = 'trnsfr_leiden_harmony' )
df <- df %>% group_by(trnsfr_leiden_harmony) %>% summarise(sum_value = mean(value, na.rm = TRUE), .groups = 'drop')
toEliminate <- df[ which(df$sum_value == min(df$sum_value)), ]
```


```{r establish-L1annot}
md <- pDataDT(sln)
md$L1.annot <- paste0("Cluster ",md$trnsfr_leiden_harmony)
md <- md[, which(colnames(md) %in% c("cell_ID", "L1.annot"))]
sln <- addCellMetadata(gobject = sln, new_metadata = md, by_column=T, column_cell_ID = 'cell_ID')

```



```{r exclude-cluster-round2}
# 
# md <- pDataDT(sln); keepCells <- subset(md, !(trnsfr_leiden_harmony %in% c(13, 19, 30)))$cell_ID  # exclude clusters 21 and 23 because pan-neg
# sln <- subsetGiotto(gobject = sln, cell_ids = keepCells)
#  

```


```{r plot-ScreePlot-excl2}
# using channels that Sabrina used for downstream clustering
# clusterChannels <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
#                      "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")
# sln <- runPCA(gobject = sln, expression_values = "normalized", scale_unit = TRUE, 
#               method = "factominer", feats_to_use = clusterChannels)
# signPCA(sln, scale_unit = TRUE, scree_ylim = c(0,40))
```


Calculate clusters by Leiden
```{r calculate-Clusters-excl2}

# sln.sm <- subsetGiotto(gobject = sln, cell_ids = sample(sln$cell_ID, 1e5, replace=F)) # make a smaller dataset to work with 
# sln.sm <- createNearestNetwork(gobject = sln.sm, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
#                             name = 'NN.harmony', dimensions_to_use = 1:10, k = 10)  # calculate nearest neighbors, focus on global differences 
# sln.sm <- doLeidenClusterIgraph(gobject = sln.sm, resolution = 0.3, n_iterations = 100, set_seed = TRUE, seed_number = 43,
#                                 network_name = 'NN.harmony', name = 'leiden_harmony')  # make the clusters
# sln <- labelTransfer(x = sln, y = sln.sm, labels = "leiden_harmony")  # assign cluster labels in the full dataset based on the reduced data
```

UMAP projection
```{r UMAP-projection-excl2}
# sln <- runUMAPprojection(sln, random_subset = 1e5)  # project UMAP coordinates to the full dataset based on a subsampling
```


```{r save-excl-2}
# saveRDS(sln, "sln.harmony.excl-2.RDS")
```


```{r readRDS-exl2}
# sln <- readRDS("sln.harmony.excl-2.RDS")
```


```{r plot-UMAP-Downsample-excl2, fig.show="hold", out.width="50%"}
md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1.5e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$trnsfr_leiden_harmony)))
names(leiden_colors) <- unique(md$trnsfr_leiden_harmony)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.projection', cell_color = 'trnsfr_leiden_harmony',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.75, point_shape="no_border",
         point_alpha=0.5)

data <- as.data.frame( sln.sm@dimension_reduction$cells$cell$rna$umap$umap.projection@coordinates)
data$cluster <- sln.sm@cell_metadata$cell$rna$trnsfr_leiden_harmony
data <- data[sample(1:nrow(data),1e5), ]
ggplot(data) + geom_point(aes(x=Dim.1, y=Dim.2, color=cluster), show.legend = FALSE, alpha=0.2, size=0.2) + 
  geom_point(aes(x = Inf, y = Inf, color = cluster), size = 3, show.legend = TRUE) + 
  geom_density2d(aes(x=Dim.1, y=Dim.2), linewidth=0.25, color="black") + theme_classic()  + scale_color_manual(values = leiden_colors)   + 
  scale_x_continuous(limits=c(min(data$Dim.1) - .10*diff(range(data$Dim.1)),max(data$Dim.1) + .10* diff(range(data$Dim.1)))) +
  scale_y_continuous(limits=c(min(data$Dim.2) - .10*diff(range(data$Dim.2)),max(data$Dim.2) + .10* diff(range(data$Dim.2)))) + 
  ggtitle("All cells - downsampled") +  guides(color = guide_legend(override.aes = list(shape = 16)))
rm(data) 
```

```{r spatDimPlot-CD3E, fig.width=12, fig.height=12}
md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 1e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
dimFeatPlot2D(sln.sm, dim_reduction_name = 'umap.projection', expression_values = "scaled", 
              feats = c("CD3E","CD4", "CD8", "CD20","MART1","TOX"), 
              point_alpha = 0.2, point_shape = "no_border", cow_n_col = 2)
```


```{r calc-feat-heatmap-excl2-Cells}
markers_scran <- findMarkers_one_vs_all(gobject = sln, method = "scran", expression_values = "normalized",
                                        cluster_column = "trnsfr_leiden_harmony", min_feats = 3)

topgenes_scran <- unique(markers_scran[, head(.SD, 5), by = "cluster"][["feats"]])
```

```{r show-feat-heatmap-excl2-Cells, fig.show="hold", out.width="50%"}
features <- c("CD3E", "CD4", "CD8", "TCF1", "TOX", "CD45RO", "CD19", "CD20","HLA.DR", "CD38", "IGD", "CXCR5", 
                     "KI67", "CD45", "PAN.CYTO", "CD21", "MAC2/GAL3", "IFNG", "MART1", "GP100", "S100B")
plotMetaDataHeatmap(sln, expression_values = "normalized",
                    metadata_cols = "trnsfr_leiden_harmony", selected_feats = topgenes_scran,
                    y_text_size = 8, show_values = "zscores_rescaled")
plotMetaDataHeatmap(sln, expression_values = "normalized",
                    metadata_cols = "trnsfr_leiden_harmony", selected_feats = features,
                    y_text_size = 8, show_values = "zscores_rescaled")
```

df$L1.annot <- paste0("Unknown.", df$trnsfr_leiden_harmony)
df[which(df$trnsfr_leiden_harmony %in% c(28,15,20,18,26,33,6)), 'L1.annot'] <- "Tcells"
df[which(df$trnsfr_leiden_harmony %in% c(17,29,19,31,35,14,11,8)), 'L1.annot'] <- "Bcells"
df[which(df$trnsfr_leiden_harmony %in% c(3  )), 'L1.annot'] <- "GerminalCenter.1"
df[which(df$trnsfr_leiden_harmony %in% c(16)), 'L1.annot'] <- "Tumor.1"
df[which(df$trnsfr_leiden_harmony %in% c(21)), 'L1.annot'] <- "Tumor.2"
df[which(df$trnsfr_leiden_harmony %in% c(23)), 'L1.annot'] <- "Tumor.3"
df[which(df$trnsfr_leiden_harmony %in% c(36)), 'L1.annot'] <- "Tumor.4"


```{r show-tumor5-only, fig.show="hold",fig.height=12}
md <- pDataDT(sln); keepCells <- subset(md, list_ID == "tumor.5" & y %in% c(7500:22500) )$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$trnsfr_leiden_harmony)))
names(leiden_colors) <- unique(md$trnsfr_leiden_harmony)
spatPlot2D(sln.sm, cell_color = 'trnsfr_leiden_harmony', cell_color_code = leiden_colors, point_size = 1, point_shape="no_border",
         point_alpha=0.5, axis_text=14, legend_text = 14)
```


```{r show-violin, fig.show="hold", fig.width=10, fig.height=30}
# don't try to do violinplot on more than 100k cells! 
violinPlot(sln.sm, feats = features, cluster_column = "trnsfr_leiden_harmony")

```


```{r}
# x <- giottoToSeuratV4(gobject = sln)
giottoToAnnData(sln.sm)
```





```{r label-L1annotation}

df <- data.frame(trnsfr_leiden_harmony = md$trnsfr_leiden_harmony, L1.annot = "Unknown")
df$cell_ID <- md$cell_ID
df[which(df$trnsfr_leiden_harmony %in% c(12,19,20,27, 28)), 'L1.annot'] <- "Tcells"
df[which(df$trnsfr_leiden_harmony %in% c(6,8,18,19 )), 'L1.annot'] <- "Bcells"
df[which(df$trnsfr_leiden_harmony %in% c(17,22 )), 'L1.annot'] <- "Tumor"
sln <- addCellMetadata(sln, new_metadata = df, by_column = T, column_cell_ID="cell_ID")  # add in the annotations to the main object

md <- pDataDT(sln)
keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 2e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- getDistinctColors(length(unique(md$L1.annot)))
names(leiden_colors) <- unique(md$L1.annot)
plotUMAP(gobject = sln.sm, dim_reduction_name = 'umap.projection', cell_color = 'L1.annot',
         cell_color_code = leiden_colors, show_NN_network = F, point_size = 0.5, point_shape="no_border",
         point_alpha=0.35)

```






















```{r subset-B-and-T-cells}
gc()
md <- pDataDT(sln); keepCells <- md$cell_ID[which(md$L1.annot == "Tcells")]
Tcells <- subsetGiotto(sln, cell_ids = keepCells )

sln.sm <- subsetGiotto(gobject = Tcells, cell_ids = sample(sln$cell_ID, 1e5, replace=F)) # make a smaller dataset to work with 
sln.sm <- createNearestNetwork(gobject = sln.sm, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:10, k = 10)  # calculate nearest neighbors, focus on global differences 
sln.sm <- doLeidenClusterIgraph(gobject = sln.sm, resolution = 0.3, n_iterations = 100, 
                                network_name = 'NN.harmony', name = 'leiden_Tcells')  # make the clusters
Tcells <- labelTransfer(x = Tcells, y = sln.sm, labels = "leiden_Tcells")  # assign cluster labels in the full dataset based on the reduced data

saveRDS(Tcells, "Tcells.UMAP.RDS")

# ************************

keepCells <- md$cell_ID[which(md$L1.annot == "Bcells")]
Bcells <- subsetGiotto(sln, cell_ids = keepCells )
sln.sm <- subsetGiotto(gobject = Bcells, cell_ids = sample(sln$cell_ID, 1e5, replace=F)) # make a smaller dataset to work with 
sln.sm <- createNearestNetwork(gobject = sln.sm, dim_reduction_to_use = 'harmony', dim_reduction_name = 'harmony', 
                            name = 'NN.harmony', dimensions_to_use = 1:10, k = 10)  # calculate nearest neighbors, focus on global differences 
sln.sm <- doLeidenClusterIgraph(gobject = sln.sm, resolution = 0.3, n_iterations = 100, 
                                network_name = 'NN.harmony', name = 'leiden_Bcells')  # make the clusters
Bcells <- labelTransfer(x = Bcells, y = sln.sm, labels = "leiden_Bcells")  # assign cluster labels in the full dataset based on the reduced data

saveRDS(Bcells, "Bcells.UMAP.RDS")
# 
```

















