---
title: "Neighborhood analyses - Interaction Changed Features"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---
 

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```
 
 
```{r call-libraries} 
date()
library(Giotto)
library(future); if(parallelly::availableCores() > 2 )  # & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); # library(leiden);   # library(multinet);  
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
library(ComplexHeatmap)
library(circlize)
library(dplyr); library(tidyr)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

#  ****** libraries to install 
# library(factoextra)

```

# read in data object 
```{r read-annot-sln-spatnet}

sln <- readRDS("sln.clustering5.annot.spatnet.20250617.160512.RDS")
gc()
md <- pDataDT(sln)
stableColorsL2 <- getDistinctColors(length(unique(md$L2_annotations)));  names(stableColorsL2) <- unique(md$L2_annotations)
stage3.subgroup <- c("tumor.1","tumor.2","tumor.3","tumor.4","tumor.5","tumor.6","tumor.7", "tumor.8",
                                                 "tumor.9","tumor.10")
stage1.subgroup <- c("ctrl.1","ctrl.2","ctrl.3","ctrl.4","ctrl.5")
recur.subgroup <- c("tumor.3","tumor.4","tumor.5","tumor.7","tumor.8","tumor.10")
norecur.subgroup <- c("tumor.1","tumor.2", "tumor.6", "tumor.9")
saveRDS(md, file = "md.RDS")
```


# calculate ICF 
```{r ICF-perSample-calculateICF, fig.width=10}
list_ID <- unique(md$list_ID)
icf.persample <- list(); icf.persample.enrich <- list()
proteinsList <- c("BCA1", "CD3E", "CD4","CD8" ,"CD11C" ,"CD27","CD38","CD39", "CD45RO","CXCR3","CXCR5",
                  "FOXP3" , "GZMB" ,"HLA.DR", "ICOS","IFNG","IL21","IRF4","IRF7","KI67", "LAG3",
                  "PD1","PDL1","STAT1","STAT3","TCF1","TOX" ) # "TBET","PMTOR","CD19","CD20" ,"CD21","MAC2/GAL3","PAX5","VIMENTIN"
timeStamp <- format(Sys.time(), "%m%d.%H%M")
enrich.celltypes <- c("ACTIVATED_TREG", "TCF1_LO_TOX_HI_CD8", "ACTIVATED_CD8") # make a version that keeps all of these cells 

for (i in list_ID)
{
  if( .Platform$OS.type != "windows")    {    cat(paste("Starting on: ", i, " on ", date(),"\n"), 
                                                  file = paste0("progress.",timeStamp,".txt"), append = T)   }
  keepCells <- subset(md, list_ID %in% i )
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells$cell_ID)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=9, verbose=T, name = "spatialknn.k9.persample") 
  print("made spatialKNN")
  print(getSpatialNetwork(sln.sm))
  res <- findInteractionChangedFeats(gobject = sln.sm, selected_feats = proteinsList,
                                     spatial_network_name = "spatialknn.k9.persample", 
                                     cluster_column = "L2_annotations", diff_test = 'permutation', 
                                     adjust_method = 'fdr', nr_permutations = 1, do_parallel = T)  
  res$list_ID <- i
  icf.persample[i] <- res
  print(paste("Random downsample process for", i, "ends at", date()))
}

saveRDS(icf.persample, file = "neighborhood.ICFlist.perSample.sln.clustering5.spatnet.RDS")

# Job Wall-clock time: 01:51:30              Memory Utilized: 72.67 GB
```

# load ICF data 
```{r ICF-perSample-loadICF, fig.width=10}
icf.persample <- purrr::list_rbind(readRDS(file="neighborhood.ICFlist.perSample.sln.clustering5.spatnet.RDS"), 
                                    names_to = "list_ID")

 #    list_ID  feats       sel     other     log2fc        diff p.value p.adj  perm_sel perm_other perm_log2fc   perm_diff
 #     <char> <char>     <num>     <num>      <num>       <num>   <num> <num>     <num>      <num>       <num>       <num>
 # 1:  ctrl.1    CD4  6.678378  4.273293  0.6322200   2.4050848       0     0  4.443401   4.591068 -0.04614361 -0.14766649
 # 2:  ctrl.1   IRF7  4.335566  2.802213  0.6119648   1.5333530       0     0  2.833045   3.015845 -0.08722451 -0.18280052
 # 3:  ctrl.1 HLA.DR  7.155213  4.690167  0.5989422   2.4650466       0     0  5.143791   4.976156  0.04687391  0.16763547
 #                  cell_type             int_cell_type nr_select int_nr_select nr_other int_nr_other
 #                   <char>                    <char>     <int>         <int>    <int>        <int>
 # 1: TCF1_MIXED_TOX_HI_CD8        TCF1_HI_TOX_HI_CD8       271           309     1906         1054
 # 2: TCF1_MIXED_TOX_HI_CD8        TCF1_HI_TOX_HI_CD8       271           309     1906         1054
 # 3: TCF1_MIXED_TOX_HI_CD8        TCF1_HI_TOX_HI_CD8       271           309     1906         1054
 #                                          unif_int                                  spec_int type_int
 #                                       <char>                                    <char>   <char>
 # 1: TCF1_HI_TOX_HI_CD8--TCF1_MIXED_TOX_HI_CD8 TCF1_MIXED_TOX_HI_CD8--TCF1_HI_TOX_HI_CD8   hetero
 # 2: TCF1_HI_TOX_HI_CD8--TCF1_MIXED_TOX_HI_CD8 TCF1_MIXED_TOX_HI_CD8--TCF1_HI_TOX_HI_CD8   hetero
 # 3: TCF1_HI_TOX_HI_CD8--TCF1_MIXED_TOX_HI_CD8 TCF1_MIXED_TOX_HI_CD8--TCF1_HI_TOX_HI_CD8   hetero

```

# make volcano plots 
```{r ICF-perSample-calculate-volcano, fig.width=6}
# intxns <- findImportantInteractions( interaction.dt= interaction.byRecur, condition.exp="recur", condition.ctrl="no.recur", ef.threshold=1)[[1]]
plot.volcano.icf <- function( dt, receiver_cell, sender_cell, key.feat = NULL, plot = T)
{
  dt <- dt[int_cell_type == sender_cell & cell_type == receiver_cell]
  volcano.data <- dt[, .(mean_log2fc = mean(log2fc, na.rm = TRUE),
                                           p_value = t.test(log2fc, mu = 0)$p.value, 
                                           Welch_t = t.test(log2fc, mu = 0)$statistic ),
                                       by = .(feats)  ]
  volcano.data[, p_adj := p.adjust(p_value, method="fdr" ) ] [order(p_adj, decreasing = T)]
  volcano.data[, highlight.feat := "0"]
  volcano.data[feats %in% key.feat, highlight.feat := "1"]
  annotateX.incr <- max(volcano.data$mean_log2fc)-0.2*max(volcano.data$mean_log2fc)
  annotateX.decr <- min(volcano.data$mean_log2fc)+0.2*min(volcano.data$mean_log2fc)
  annotateY.ylim <- max(-log10(volcano.data$p_adj))*1.1
  if (plot == T)
  {
    print(
      ggplot(volcano.data, aes(y = -log10(p_adj), x=mean_log2fc, label = feats, color=highlight.feat)) + 
        geom_hline(yintercept = -log10(0.05), color="grey60", linetype="dashed", linewidth=0.1)+
        geom_point(aes(fill = highlight.feat), shape=21, color="white",  size=3) +   #fill="black",
        ggrepel::geom_text_repel(max.overlaps = 20) + 
        ggtitle(paste0(receiver_cell, " adjacent to ", sender_cell )) + 
        theme_classic()  + 
        scale_color_manual(values = c('black','red')) + scale_fill_manual(values = c('black','red'))+ 
        annotate("text", x = annotateX.incr, y = -0.15*annotateY.ylim, size = 3, label = "Increased expr") + 
        annotate("text", x = annotateX.decr, y = -0.15*annotateY.ylim, size = 3, label = "Reduced expr")  +
        coord_cartesian(clip = 'off', ylim= c(0, annotateY.ylim)) + theme(legend.position = 'none')
      )
  }
  return(volcano.data)
#  
#      feats mean_log2fc      p_value    Welch_t        p_adj
#    <char>       <num>        <num>      <num>        <num>
# 1: HLA.DR  0.67395872 1.071667e-07  9.8895908 2.893501e-06
# 2:   CD3E  0.02025484 5.118152e-01  0.6731441 5.757921e-01
# 3:   ICOS -0.05803399 7.299591e-02 -1.9384811 1.159347e-01
}

plot.volcano.icf(dt = icf.persample, receiver_cell = "ACTIVATED_TREG", sender_cell = "ACTIVATED_CD8")
plot.volcano.icf(dt = icf.persample, receiver_cell = "ACTIVATED_CD8", sender_cell = "ACTIVATED_TREG")

plot.volcano.icf(dt = icf.persample, receiver_cell = "ACTIVATED_TREG", sender_cell = "TCF1_HI_TOX_HI_CD8")
plot.volcano.icf(dt = icf.persample, receiver_cell = "TCF1_HI_TOX_HI_CD8", sender_cell = "ACTIVATED_TREG")
plot.volcano.icf(dt = icf.persample, receiver_cell = "MEMORY_CD4", sender_cell = "ACTIVATED_TREG")

```

This is curious. It's likely showing membrane spillover from the adjacent cell. All of the recipient cells have ICOS and Foxp3 which is likely coming from the ACTIVATED_TREG. 

# calculate statistics 
```{r plot-everything}

dtable <- icf.persample[,.(list_ID, feats, log2fc, cell_type, int_cell_type, nr_select, int_nr_select)]
volcano.dt <- dtable[cell_type == "ACTIVATED_TREG", .(mean_log2fc = mean(log2fc, na.rm = TRUE),
                                           p_value = t.test(log2fc, mu = 0)$p.value, 
                                           Welch_t = t.test(log2fc, mu = 0)$statistic ),
                                       by = .(feats, int_cell_type)  ]
volcano.dt <- volcano.dt[! int_cell_type %in% c("ARTIFACT", "CD4.outlier","CD8.outlier")]
volcano.dt[, p_adj := p.adjust(p_value, method="fdr" ) ] [order(p_adj, decreasing = T)]

```

# heatmap of the ICF 
```{r ICF-perSample-calculate-volcano-dupl, fig.width=8, eval=FALSE}
index.cell = "TCF1_LO_TOX_HI_CD8"
interacting.cell = "ACTIVATED_TREG"
df.for.plot <- icf.persample[int_cell_type == interacting.cell & cell_type == index.cell]
annotation.df <- data.frame(list_ID = c(stage1.subgroup, recur.subgroup, norecur.subgroup))
annotation.df$annot <- c(rep("stage1", 5), rep("stage3.recur", length(recur.subgroup)), 
                         rep("stage3.no.recur", length(norecur.subgroup))) 
rownames(annotation.df) <- annotation.df$list_ID
annotation.df$list_ID <- NULL
heatmap.feats <- proteinsList # c("IFNG", "GZMB", "HLADR")
heatmap.matrix <- df.for.plot[feats %in% heatmap.feats, .(log2fc, feats, list_ID)]
heatmap.matrix <- (heatmap.matrix %>% pivot_wider(names_from = feats, values_from=log2fc, id_cols = list_ID)) %>% 
   tibble::column_to_rownames("list_ID") 
heatmap.matrix <- heatmap.matrix[rownames(annotation.df), ] %>% as.matrix()
colors <- c("stage1" = "gray80", "stage3.no.recur" = "gray30", "stage3.recur" = "black")
ht.annot <- HeatmapAnnotation(df = annotation.df, col = list(annot = colors), which = "row")

Heatmap(  heatmap.matrix, name = "log2FC",
  row_split = annotation.df$annot, left_annotation = ht.annot, rect_gp = gpar(col = "white", lwd = 1),
  cluster_rows = TRUE, cluster_columns = TRUE, show_row_names = F, 
  row_names_gp = gpar(fontsize = 10),  column_names_gp = gpar(fontsize = 10), 
  column_title = (paste0(index.cell, " adjacent to ", interacting.cell)), row_title = " "
) 


```

# regroup Treg subgroups into one entity in L2_annotations
```{r regroup-Treg}
md.mod <- md
md.mod[, L2_annotations.mergeTreg:= L2_annotations]
md.mod[L2_annotations.mergeTreg %like% "TREG", L2_annotations.mergeTreg := "TREG"]
newdf <- data.frame(cell_ID=md.mod$cell_ID, L2_annotations.mergeTreg = md.mod$L2_annotations.mergeTreg)
sln <- addCellMetadata(gobject = sln, new_metadata = newdf, by_column=T, column_cell_ID = 'cell_ID')
md <- pDataDT(sln)
icf.persample.regroup <- list()
for (i in list_ID)
{
  if( .Platform$OS.type != "windows")    {    cat(paste("Starting on: ", i, " on ", date(),"\n"), 
                                                  file = paste0("progress.",timeStamp,".txt"), append = T)   }
  keepCells <- subset(md, list_ID %in% i )
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells$cell_ID)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=9, verbose=T, name = "spatialknn.k9.persample") 
  print("made spatialKNN")
  print(getSpatialNetwork(sln.sm))
  res <- findInteractionChangedFeats(gobject = sln.sm, selected_feats = proteinsList,
                                     spatial_network_name = "spatialknn.k9.persample", 
                                     cluster_column = "L2_annotations.mergeTreg", diff_test = 'permutation', 
                                     adjust_method = 'fdr', nr_permutations = 1, do_parallel = T)  
  res$list_ID <- i
  icf.persample.regroup[i] <- res
  print(paste("Random downsample process for", i, "ends at", date()))
}

saveRDS(icf.persample.regroup, file = "neighborhood.ICFlistmod.perSample.sln.clustering5.spatnet.RDS")

```

```{r load-ICF-groupedTreg}

icf.persample.regroup <- purrr::list_rbind(readRDS(file="neighborhood.ICFlistmod.perSample.sln.clustering5.spatnet.RDS"), 
                                    names_to = "list_ID")

```


# regroup the Treg 
What if we instead think about all the Treg together? Would that give us more Treg that aren't already in the terminal phenotype?  It might recover more cells that were possibly misclassified. Maybe it would show phenotypic change in the Treg that are near CD8, sort of like an orthogonal way to recreate the cellProximityEnrichment analysis 
```{r volcano-plots-regrouped, fig.width=8}
plot.volcano.icf(dt = icf.persample.regroup, receiver_cell = "TREG", sender_cell = "ACTIVATED_CD8")
plot.volcano.icf(dt = icf.persample.regroup, receiver_cell = "ACTIVATED_CD8", sender_cell = "TREG")
plot.volcano.icf(dt = icf.persample.regroup, receiver_cell = "TCF1_LO_TOX_HI_CD8", sender_cell = "TREG", key.feat = c("GZMB", "IFNG"))
  # ggsave(filename = 'output/icf.tcflo-near-treg.volcano.pdf', height=4, width=5)
plot.volcano.icf(dt = icf.persample.regroup, receiver_cell = "TREG", sender_cell = "TCF1_HI_TOX_HI_CD8")
plot.volcano.icf(dt = icf.persample.regroup, receiver_cell = "TCF1_HI_TOX_HI_CD8", sender_cell = "TREG", key.feat = c("GZMB", "IFNG"))
  # ggsave(filename = 'output/icf.tcfhi-near-treg.volcano.pdf', height=4, width=5)
plot.volcano.icf(dt = icf.persample.regroup, receiver_cell = "TCF1_HI_TOX_HI_CD8", sender_cell = "Tumor", key.feat = c("GZMB", "IFNG"))
  # ggsave(filename = 'output/icf.tcfhi-near-tumor.volcano.pdf', height=4, width=5)
plot.volcano.icf(dt = icf.persample.regroup, receiver_cell = "TCF1_LO_TOX_HI_CD8", sender_cell = "Tumor", key.feat = c("GZMB", "IFNG"))
  # ggsave(filename = 'output/icf.tcflo-near-tumor.volcano.pdf', height=4, width=5)
```

Still seeing membrane spillover effects, that didn't change. The CD8 adj to Treg again have lower GzmB and IFNg, esp in the Tcfhi CD8 subset. 

```{r heatmap-plots-regrouped, fig.width=8}
index.cell = "TCF1_LO_TOX_HI_CD8"
interacting.cell = "TREG"
df.for.plot <- plot.volcano.icf(dt = icf.persample.regroup, receiver_cell = index.cell, sender_cell = interacting.cell)

annotation.df <- data.frame(list_ID = c(stage1.subgroup, recur.subgroup, norecur.subgroup))
annotation.df$annot <- c(rep("stage1", 5), rep("stage3.recur", length(recur.subgroup)), 
                         rep("stage3.no.recur", length(norecur.subgroup))) 
rownames(annotation.df) <- annotation.df$list_ID
annotation.df$list_ID <- NULL
heatmap.feats <- proteinsList # c("IFNG", "GZMB", "HLADR")
heatmap.matrix <- df.for.plot[feats %in% heatmap.feats, .(log2fc, feats, list_ID)]
heatmap.matrix <- (heatmap.matrix %>% pivot_wider(names_from = feats, values_from=log2fc, id_cols = list_ID)) %>% 
   tibble::column_to_rownames("list_ID") 
heatmap.matrix <- heatmap.matrix[rownames(annotation.df), ] %>% as.matrix()
colors <- c("stage1" = "gray80", "stage3.no.recur" = "gray30", "stage3.recur" = "black")
ht.annot <- HeatmapAnnotation(df = annotation.df, col = list(annot = colors), which = "row")

Heatmap(  heatmap.matrix, name = "log2FC",
  row_split = annotation.df$annot, left_annotation = ht.annot, rect_gp = gpar(col = "white", lwd = 1),
  cluster_rows = TRUE, cluster_columns = TRUE, show_row_names = F, 
  row_names_gp = gpar(fontsize = 10),  column_names_gp = gpar(fontsize = 10), 
  column_title = (paste0(index.cell, " adjacent to ", interacting.cell)), row_title = " "
) 


```

```{r for-each-Cell-type}
# plot what is the effect of other sender celltypes on TCFlo for GzmB
probe.subset <- "TCF1_LO_TOX_HI_CD8";  probe.feature <- "GZMB"
dtable <- icf.persample.regroup[,.(list_ID, feats, log2fc, cell_type, int_cell_type, nr_select, int_nr_select)]  # simplify dtable
dtable <- dtable[cell_type == probe.subset, .(mean_log2fc = mean(log2fc, na.rm = TRUE), 
                                           p_value = t.test(log2fc, mu = 0)$p.value, 
                                           Welch_t = t.test(log2fc, mu = 0)$statistic ),
                                       by = .(feats, int_cell_type)  ] # calculate mean, pvalue, t stat
dtable <- dtable[! int_cell_type %in% c("ARTIFACT", "CD4.outlier","CD8.outlier")]
dtable[, p_adj := (p.adjust(p_value, method="fdr" )) ] # adjust p values using FDR method 
dtable[, log.p_adj := -log10(p.adjust(p_value, method="fdr" )) ]  
order <- dtable %>% arrange(mean_log2fc)
order <- order[feats == probe.feature, .(unique(int_cell_type))]  
dtable$int_cell_type <- factor(dtable$int_cell_type, levels=order$V1)  # factor to make graph display in order 
ggplot(dtable[feats == probe.feature], aes(y = int_cell_type, x=mean_log2fc, size=log.p_adj)) +  
  geom_point(pch=21, color='black', fill='black',stroke=0.5) + theme_classic() + 
  ggtitle(paste0(probe.feature, " in ", probe.subset)) + 
  theme(axis.title.y = element_blank())

plot.icf.affectCells <- function( dt, probe.subset, probe.feature)
{
  dtable <- dt[,.(list_ID, feats, log2fc, cell_type, int_cell_type, nr_select, int_nr_select)] # simplify
  dtable <- dtable[int_cell_type == probe.subset, .(mean_log2fc = mean(log2fc, na.rm = TRUE),
                                             p_value = t.test(log2fc, mu = 0)$p.value, 
                                             Welch_t = t.test(log2fc, mu = 0)$statistic ),
                                         by = .(feats, cell_type)  ]  # statistically summarize after subgrouping
  dtable <- dtable[! cell_type %in% c("ARTIFACT", "CD4.outlier","CD8.outlier")]
  dtable[, p_adj := (p.adjust(p_value, method="fdr" )) ]  
  dtable[, log.p_adj := -log10(p.adjust(p_value, method="fdr" )) ]  
  order <- dtable %>% arrange(mean_log2fc)
  order <- order[feats == probe.feature, .(unique(cell_type))]
  dtable$cell_type <- factor(dtable$cell_type, levels=order$V1)  # this is a problem 
  print(ggplot(dtable[feats == probe.feature], aes(y = cell_type, x=mean_log2fc, size=log.p_adj)) + 
    geom_point(pch=21, color='white', fill='black',stroke=0.5) + theme_classic() + 
    ggtitle(paste0(probe.subset, " affecting ", probe.feature)) + 
    theme(axis.title.y = element_blank())
  )
  return(dtable[feats==probe.feature])
  
#      feats                                cell_type  mean_log2fc     p_value     Welch_t       p_adj  log.p_adj
#    <char>                                   <fctr>        <num>       <num>       <num>       <num>      <num>
# 1:   IFNG                           CD4_CD20_MIXED -0.061328114 0.066587619 -1.98917806 0.121424481 0.91569374
# 2:   IFNG                                       PC -0.028835847 0.582465746 -0.56280971 0.682806484 0.16570236
# 3:   IFNG MIXED_FOLLICULAR_AND_NONFOLLICULAR_BCELL  0.048100229 0.495967493  0.69905601 0.602503326 0.22004055
}

# now seeking to make the reverse, where you see what Treg's effects on GzmB are in receiver cells
plot.icf.affectCells(dt = icf.persample.regroup, probe.subset = "TREG",  probe.feature = "GZMB")
  ggsave(filename = 'output/icf.treg-effects-on-GZMB.across-celltypes.pdf', height=6, width=6)
plot.icf.affectCells(dt = icf.persample.regroup, probe.subset = "TREG",  probe.feature = "IFNG")
  ggsave(filename = 'output/icf.treg-effects-on-IFNg.across-celltypes.pdf', height=6, width=6)
```



















# deprecated:  interaction changed genes
```{r Interaction-changed-genes-calculate, eval=FALSE}
# md <- pDataDT(sln)
keepCells <- subset(md, list_ID %in% stage3.subgroup)
keepCells <- subset(keepCells, cell_ID %in% sample(keepCells$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)
date()
## identify genes that are associated with proximity to other cell types
ICGscores.L2 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L2_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
saveRDS(object = ICGscores.L2, 
        file = "neighborhood.ICGlist.s3.sln.clustering5.annot.spatnet.RDS"); date()  # took 6 hours at 300 permutations each 


keepCells <- subset(md, list_ID %in% stage1.subgroup)
keepCells <- subset(keepCells, cell_ID %in% sample(keepCells$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)
date()
## identify genes that are associated with proximity to other cell types
ICGscores.L2 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L2_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
saveRDS(object = ICGscores.L2, 
        file = "neighborhood.ICGlist.s1.sln.clustering5.annot.spatnet.RDS"); date()



keepCells <- subset(md, list_ID %in% recur.subgroup)
keepCells <- subset(keepCells, cell_ID %in% sample(keepCells$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)
date()
## identify genes that are associated with proximity to other cell types
ICGscores.L2 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L2_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
saveRDS(object = ICGscores.L2, 
        file = "neighborhood.ICGlist.recur.sln.clustering5.annot.spatnet.RDS"); date()


keepCells <- subset(md, list_ID %in% norecur.subgroup)
keepCells <- subset(keepCells, cell_ID %in% sample(keepCells$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)
date()
## identify genes that are associated with proximity to other cell types
ICGscores.L2 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L2_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
saveRDS(object = ICGscores.L2, 
        file = "neighborhood.ICGlist.norecur.sln.clustering5.annot.spatnet.RDS"); date()

ICGscores.L2.recur = readRDS(file = "neighborhood.ICGlist.recur.sln.clustering5.annot.spatnet.RDS")
ICGscores.L2.norecur = readRDS(file = "neighborhood.ICGlist.norecur.sln.clustering5.annot.spatnet.RDS")


```


























