---
title: "Neighborhood analyses"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---
 

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```
 
 
```{r call-libraries} 
date()
library(Giotto)
library(future); if(parallelly::availableCores() > 2 & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); library(leiden);   # library(multinet);  
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
library(ComplexHeatmap)
library(circlize)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()
```

# read in data object 
```{r read-annot-sln-spatnet}

sln <- readRDS("sln.clustering5.annot.spatnet.20250515.121453.RDS")
md <- pDataDT(sln)
stableColorsL2 <- getDistinctColors(length(unique(md$L2_annotations)));  names(stableColorsL2) <- unique(md$L2_annotations)
```


# calculate proximity enrichment per sample for k=5
```{r custom-cellproxEnrich-function-k5}
cellProx.sim <- function(gobject, spat_unit = NULL, feat_type = NULL, spatial_network_name="spatialknn.k5", 
                         #spatial_network_name = "Delaunay_network",
    cluster_column, number_of_simulations = 1000, 
    adjust_method = c("none", "fdr", "bonferroni", "BH", "holm", "hochberg", "hommel", "BY"), 
    set_seed = TRUE, seed_number = 1234) 
{
    # Set feat_type and spat_unit
    spat_unit <- set_default_spat_unit(
        gobject = gobject,
        spat_unit = spat_unit
    )
    feat_type <- set_default_feat_type(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type
    )

    # p.adj test
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))

    spatial_network_annot <- annotateSpatialNetwork(
        gobject = gobject,
        feat_type = feat_type,
        spat_unit = spat_unit,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column
    )

    # remove double edges between same cells #
    # a simplified network does not have double edges between cells #

    # data.table variables
    unified_cells <- type_int <- N <- NULL

    spatial_network_annot <- GiottoUtils:::dt_sort_combine_two_columns(
        spatial_network_annot, "to", "from", "unified_cells"
    )
    spatial_network_annot <- spatial_network_annot[!duplicated(unified_cells)]

    sample_dt <- Giotto:::make_simulated_network(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column,
        number_of_simulations = number_of_simulations,
        set_seed = set_seed,
        seed_number = seed_number
    )

    # combine original and simulated network
    table_sim_results <- sample_dt[, .N, by = c(
        "unified_int", "type_int", "round"
    )]

    ## create complete simulations
    ## add 0 if no single interaction was found
    unique_ints <- unique(table_sim_results[, .(unified_int, type_int)])

    # data.table with 0's for all interactions
    minimum_simulations <- unique_ints[rep(
        seq_len(nrow(unique_ints)), number_of_simulations
    ), ]
    minimum_simulations[, round := rep(
        paste0("sim", seq_len(number_of_simulations)),
        each = nrow(unique_ints)
    )]
    minimum_simulations[, N := 0]

    table_sim_minimum_results <- rbind(table_sim_results, minimum_simulations)
    table_sim_minimum_results[, V1 := sum(N), by = c(
        "unified_int", "type_int", "round"
    )]
    table_sim_minimum_results <- unique(
        table_sim_minimum_results[, .(unified_int, type_int, round, V1)]
    )
    table_sim_results <- table_sim_minimum_results


    # data.table variables
    orig <- unified_int <- V1 <- original <- enrichm <- simulations <- NULL

    table_sim_results[, orig := "simulations"]
    spatial_network_annot[, round := "original"]

    table_orig_results <- spatial_network_annot[, .N, by = c(
        "unified_int", "type_int", "round"
    )]
    table_orig_results[, orig := "original"]
    data.table::setnames(table_orig_results, old = "N", new = "V1")

    table_results <- rbind(table_orig_results, table_sim_results)



    # add missing combinations from original or simulations
    # probably not needed anymore
    all_simulation_ints <- as.character(unique(table_results[
        orig == "simulations"
    ]$unified_int))
    all_original_ints <- as.character(unique(table_results[
        orig == "original"
    ]$unified_int))
    missing_in_original <- all_simulation_ints[
        !all_simulation_ints %in% all_original_ints
    ]
    missing_in_simulations <- all_original_ints[
        !all_original_ints %in% all_simulation_ints
    ]
    create_missing_for_original <- table_results[
        unified_int %in% missing_in_original
    ]
    create_missing_for_original <- unique(create_missing_for_original[
        , c("orig", "V1") := list("original", 0)
    ])
    create_missing_for_simulations <- table_results[
        unified_int %in% missing_in_simulations
    ]
    create_missing_for_simulations <- unique(
        create_missing_for_simulations[, c("orig", "V1") := list(
            "simulations", 0
        )]
    )
# stop **********************************************************************************************

    table_results <- do.call(
        "rbind",
        list(
            table_results, create_missing_for_original,
            create_missing_for_simulations
        )
    )
    return(table_results)
}

cellProx.calcP <- function(table_results, number_of_simulations = 1000, 
    adjust_method = c("none", "fdr", "bonferroni", "BH", "holm", "hochberg", "hommel", "BY"), 
    set_seed = TRUE, seed_number = 1234)
{
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))
    ## p-values
    combo_list <- rep(NA, length = length(unique(table_results$unified_int)))
    p_high <- rep(NA, length = length(unique(table_results$unified_int)))
    p_low <- rep(NA, length = length(unique(table_results$unified_int)))

    for (int_combo in seq_along(unique(table_results$unified_int))) {
        this_combo <- as.character(unique(table_results$unified_int)[int_combo])

        sub <- table_results[unified_int == this_combo]

        orig_value <- sub[orig == "original"]$V1
        sim_values <- sub[orig == "simulations"]$V1

        length_simulations <- length(sim_values)
        if (length_simulations != number_of_simulations) {
            additional_length_needed <- number_of_simulations -
                length_simulations
            sim_values <- c(sim_values, rep(0, additional_length_needed))
        }

        p_orig_higher <- 1 - (sum((orig_value + 1) > (sim_values + 1)) /
            number_of_simulations)
        p_orig_lower <- 1 - (sum((orig_value + 1) < (sim_values + 1)) /
            number_of_simulations)

        combo_list[[int_combo]] <- this_combo
        p_high[[int_combo]] <- p_orig_higher
        p_low[[int_combo]] <- p_orig_lower
    }
    res_pvalue_DT <- data.table::data.table(
        unified_int = as.vector(combo_list),
        p_higher_orig = p_high,
        p_lower_orig = p_low
    )


    # depletion or enrichment in barplot format
    table_mean_results <- table_results[, .(mean(V1)), by = c(
        "orig", "unified_int", "type_int"
    )]
    table_mean_results_dc <- data.table::dcast.data.table(
        data = table_mean_results, formula = type_int + unified_int ~ orig,
        value.var = "V1"
    )
    table_mean_results_dc[, original := ifelse(is.na(original), 0, original)]
    table_mean_results_dc[, enrichm := log2((original + 1) / (simulations + 1))]


    table_mean_results_dc <- merge(
        table_mean_results_dc, res_pvalue_DT,
        by = "unified_int"
    )
    data.table::setorder(table_mean_results_dc, enrichm)
    table_mean_results_dc[, unified_int := factor(unified_int, unified_int)]

    # adjust p-values for mht

    # data.table variables
    p.adj_higher <- p.adj_lower <- p_lower_orig <- p_higher_orig <-
        PI_value <- int_ranking <- NULL

    table_mean_results_dc[, p.adj_higher := stats::p.adjust(
        p_higher_orig,
        method = sel_adjust_method
    )]
    table_mean_results_dc[, p.adj_lower := stats::p.adjust(
        p_lower_orig,
        method = sel_adjust_method
    )]


    table_mean_results_dc[, PI_value := ifelse(p.adj_higher <= p.adj_lower,
        -log10(p.adj_higher + (1 / number_of_simulations)) * enrichm,
        -log10(p.adj_lower + (1 / number_of_simulations)) * enrichm
    )]
    data.table::setorder(table_mean_results_dc, PI_value)

    # order
    table_mean_results_dc <- table_mean_results_dc[order(-PI_value)]
    table_mean_results_dc[, int_ranking := seq_len(.N)]

    return(list(
        raw_sim_table = table_results,
        enrichm_res = table_mean_results_dc
    ))
}

list_ID <- unique(md$list_ID)
gc()

tableResults.k5 <- list()
for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=5, verbose=T, name = "spatialknn.k5.persample") 
  tableResults.k5[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L2_annotations', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "spatialknn.k5.persample")
}
dt <- purrr::list_rbind(tableResults.k5, names_to = "id")
saveRDS(dt, file="neighborhood.tableResults.k5.sln.RDS")  # all data merged
```

# calculate proximity enrichment per sample for k=9
```{r custom-cellproxEnrich-function-k10}
tableResults.k9 <- list()
annot.k9 <- list()
for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=9, verbose=T, name = "spatialknn.k9.persample") 
  print("made spatialKNN")
  annot.k9[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L2_annotations", spatial_network_name = 'spatialknn.k9.persample')
  tableResults.k9[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L2_annotations', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "spatialknn.k9.persample")
}
dt <- purrr::list_rbind(tableResults.k9, names_to = "id")
saveRDS(dt, file="neighborhood.tableResults.k9.sln.RDS")  # all data merged

dt <- readRDS(file="neighborhood.tableResults.k9.sln.RDS")  # all data merged
```

Fails on k=10 on tumor.6  (2 million rows) with long vectors error 
Works for k=9 for all samples 

# subgroup by specified metadata  
```{r custom-cellproxEnrich-function-subgroup}
# subset dt based on groupings of interest 
cellProx.subgroup <- function(dt)
{
  wide_dt <- dcast(dt,                # Your input data.table
                   unified_int ~ round, # Formula: rows ~ columns
                   value.var = "V1",  # Column with values for the new wide cells
                   fun.aggregate = mean, # Function to aggregate values (when multiple rows match a cell)
                   na.rm = TRUE,      # Tell mean() to ignore NA values during aggregation
                   fill = 0)    
  setkey(wide_dt, unified_int) # Good practice
  long_dt <- melt(wide_dt,
                  id.vars = "unified_int",       # Column(s) to keep as identifier variables
                  variable.name = "round",       # Name for the new column holding the former column names (1, 2, 3)
                  value.name = "V1",             # Name for the new column holding the values from the melted columns
                  variable.factor = FALSE)       # Keep the 'round' column as character initially (avoids factor issues)
  long_dt$orig <- "NA"
  long_dt$orig[grep("original", long_dt$round)] <- "original"
  long_dt$orig[grep("^sim", long_dt$round)] <- "simulations"
  long_dt$type_int <- "NA"
  long_dt$tosplit <- long_dt$unified_int
  long_dt <- as.data.table(tidyr::separate(long_dt, col = tosplit, into = c("left", "right"), sep = "--"))
  long_dt$type_int <- ifelse( long_dt$left == long_dt$right, "homo", "hetero")
  long_dt$left <- long_dt$right <- NULL
  out <- cellProx.calcP(long_dt, number_of_simulations = 200 )  # needs to match what was used to calculate sims
  return(out)
}
stage3.subgroup <- c("tumor.1","tumor.2","tumor.3","tumor.4","tumor.5","tumor.6","tumor.7", "tumor.8",
                                                 "tumor.9","tumor.10")
stage1.subgroup <- c("ctrl.1","ctrl.2","ctrl.3","ctrl.4","ctrl.5")
stage3 <- cellProx.subgroup(subset(dt, id %in% stage3.subgroup))
stage1 <- cellProx.subgroup(subset(dt, id %in% stage1.subgroup ))

x <- list(stage1 = stage1, stage3=stage3)
saveRDS(x, file="neighborhood.x.sln.RDS")  # all data merged 

recur <- cellProx.subgroup(subset(dt, id %in% c("tumor.3","tumor.4","tumor.5","tumor.7","tumor.8","tumor.10")))
no.recur <- cellProx.subgroup(subset(dt, id %in% c("tumor.1","tumor.2", "tumor.6") ))

sampleList <- unique(md$list_ID)
perSampleInteraction <- list()
for (i in seq_along(sampleList))
{
  print( paste("Interactions for sample", sampleList[i]) )
  perSampleInteraction[[i]] <- cellProx.subgroup(subset(dt, id %in% sampleList[i]))  # format in a per-sample fashion 
  perSampleInteraction[[i]]$name <- subset(dt, id %in% sampleList[i])$id[1]   # carry forward the sample name 
  # perSampleInteraction[[i]]$enrichm_res[, proportion := original/sum(original, na.rm=T) ]  
  names(perSampleInteraction)[i] <-  subset(dt, id %in% sampleList[i])$id[1]
}
saveRDS(perSampleInteraction, file="neighborhood.perSampleInteraction.RDS")  # all data merged 
perSampleInteraction <- readRDS(file="neighborhood.perSampleInteraction.RDS")

perSampleInteraction.enrichm_res <- list()
for (i in seq_along(perSampleInteraction))
{
  perSampleInteraction.enrichm_res[[i]] <- perSampleInteraction[[i]]$enrichm_res  # save the enrichm results table 
  names(perSampleInteraction.enrichm_res)[i] <- names(perSampleInteraction)[i]
}


perSampleInteraction.enrichm_res.s3 <- perSampleInteraction.enrichm_res[stage3.subgroup]
for (i in seq_along(perSampleInteraction.enrichm_res.s3))
{   perSampleInteraction.enrichm_res.s3[[i]][, list_ID := names(perSampleInteraction.enrichm_res.s3)[i]]   }
perSampleInteraction.enrichm_res.s3 <- rbindlist(perSampleInteraction.enrichm_res.s3)
perSampleInteraction.enrichm_res.s3$stage <- "stage3"

perSampleInteraction.enrichm_res.s1 <- perSampleInteraction.enrichm_res[stage1.subgroup]
for (i in seq_along(perSampleInteraction.enrichm_res.s1))
{   perSampleInteraction.enrichm_res.s1[[i]][, list_ID := names(perSampleInteraction.enrichm_res.s1)[i]]   }
perSampleInteraction.enrichm_res.s1 <- rbindlist(perSampleInteraction.enrichm_res.s1)
perSampleInteraction.enrichm_res.s1$stage <- "stage1"

interaction.byStage <- rbind(perSampleInteraction.enrichm_res.s1, perSampleInteraction.enrichm_res.s3)


```

# finish processing of subgroups at a summary level
```{r interact-with-pairwise-enrichment}
x <- readRDS(file="neighborhood.x.sln.RDS")  # all data merged 
stage1 <- x$stage1; stage3 <- x$stage3
md <- pDataDT(sln)
# x <- cellProximityBarplot(sln.sm, CPscore = out,  min_orig_ints = 3, min_sim_ints = 3, return_plot = T)
# ggsave(filename = "cellProxBarplot.pdf", x, height=20, width=7, scale = 2)

### md <- as.data.table(md)   # slated for deletion

addClusterLabels <- function(dt, md)
{
  df <- dt$enrichm_res
  df$tosplit <- df$unified_int
  df <- as.data.table(tidyr::separate(df, col = tosplit, into = c("left", "right"), sep = "--"))
  md$leiden_harmony <- as.character(md$leiden_harmony)
  setkey(md, leiden_harmony)
  df[md, on = .(left = leiden_harmony), left := i.L1_annotations]
  df[md, on = .(right = leiden_harmony), right := i.L1_annotations]
  return(df)
}

stage1$enrichm_res <- addClusterLabels(dt=stage1, md=md)
stage3$enrichm_res <- addClusterLabels(dt=stage3, md=md)
sub.stage1 <- stage1$enrichm_res
sub.stage1$subgroup <- "stage1"
sub.stage3 <- stage3$enrichm_res
sub.stage3$subgroup <- "stage3"
 # plot x=stage 1 PI_value, y= stage 3 PI_value
x <- rbind(sub.stage1, sub.stage3)
x$unified_int <- factor(x$unified_int, levels = sub.stage3$unified_int[order(sub.stage3$PI_value, decreasing = T)])

# ************** slated for deletion 
### df <- readRDS(file="neighborhood.recur.sln.clustering2.res0.5.20250411.1611.RDS")  # all data merged 
### recur <- df$recur; no.recur <- df$no.recur   
# **************  

# md <- as.data.table(md)
# recur$enrichm_res <- addClusterLabels(dt=recur, md=md)
# no.recur$enrichm_res <- addClusterLabels(dt=no.recur, md=md)
# sub.recur <- recur$enrichm_res
# sub.recur$subgroup <- "recur"
# sub.no.recur <- no.recur$enrichm_res
# sub.no.recur$subgroup <- "no.recur"
# df <- rbind(sub.recur, sub.no.recur)
# df$unified_int <- factor(df$unified_int, levels = sub.no.recur$unified_int[order(sub.no.recur$PI_value, decreasing = T)])
# saveRDS(df, file="neighborhood.recur.RDS")
# df <- readRDS(file = "neighborhood.recur.RDS")
```


# plots of pairwise enrichment
```{r plot-pairwise-enrichment, fig.width=10}
ggplot(x, aes(x=unified_int, y=enrichm, color=subgroup, label = unified_int)) + geom_point() + theme_classic() + 
  theme(axis.text.x = element_text(angle=45, hjust=1,vjust=1, size=3)) + ggrepel::geom_label_repel(max.overlaps = 30, size=2)

y <- merge(x=sub.stage1, y=sub.stage3, by='unified_int', suffixes = c(".stage1",".stage3"), all=T)
y$diff.PI_value <- y$PI_value.stage3 - y$PI_value.stage1
y$diff.enrich <- y$enrichm.stage3 - y$enrichm.stage1
y <- y[!is.na(y$diff.enrich),]  # remove NA values, which imply interactions not present in both settings
y$unified_int <- factor(y$unified_int, levels = y$unified_int[order(y$diff.enrich, decreasing = T)])
y$unified_char <- paste0(y$left.stage1,"--",y$right.stage1)

# y <- y[which(y$original.stage1 > 20),]

y[order(-diff.enrich)]

saveRDS(y, file="neighborhood.y.RDS")
y <- readRDS(file="neighborhood.y.RDS")

```


# show the average neighborhood for sln 5 
```{r show-niche-singleSample, fig.height=6}

showNiche.singleSample <- function(annot.table, indexCluster, title, filterPrefix='on', plots=TRUE)
{
  res <- annot.table[from_cell_type == indexCluster, to_cell_type] %>% table %>% prop.table  # calculate proportions 
  res <- as.data.table(res); colnames(res) <- c('subset','proportion')
  res <- res[order(res$proportion, decreasing=T)]
  stableColorsL2.fxn <- stableColorsL2
   if(filterPrefix == 'on')
  {  
    res[, subset := gsub(pattern = "CD8TOXHI_", replacement = "", x = .SD[[1]]), .SDcols = 1] 
    res[, subset := gsub(pattern = "CD8TOXLO_", replacement = "", x = .SD[[1]]), .SDcols = 1] 
    res[, subset := gsub(pattern = "TBMIXED_", replacement = "", x = .SD[[1]]), .SDcols = 1] 
    names(stableColorsL2.fxn) <- gsub(pattern = "CD8TOXHI_", replacement = "", x = names(stableColorsL2.fxn ))
    names(stableColorsL2.fxn) <- gsub(pattern = "CD8TOXLO_", replacement = "", x = names(stableColorsL2.fxn ))
    names(stableColorsL2.fxn) <- gsub(pattern = "TBMIXED_", replacement = "", x = names(stableColorsL2.fxn ))
    }
  res$subset <- factor(res$subset, levels = res$subset)   
  
  if(plots == T)
  {  
    # lollipop chart 
    print( 
      a <- ggplot(data = res, aes(y = subset, x = proportion, fill=subset, color = subset)) + 
      geom_bar(stat='identity', width=0.01) + 
      geom_point(size=5) + 
      theme_classic() + ggtitle(title) + 
        scale_fill_manual(values = stableColorsL2.fxn) +
        scale_color_manual(values = stableColorsL2.fxn) +
      theme(axis.text.y = element_text( size=6), axis.title.y = element_blank(), plot.title=element_text(size=8)) + 
        theme(legend.position = 'none')  
    )
    
    # stacked bar chart 
    print(
      ggplot(data = res, aes(y = proportion, x = 'blank',  fill = subset)) + 
      geom_col(width=1, position="fill") + 
      ggrepel::geom_text_repel( aes(label = subset ),
        position = position_stack(vjust = 0.5),           # specify stacked position to follow bar layer's default position
        size = 2,  color='black', direction = "y", xlim = c(1.5, NA), hjust = 0,
        segment.linetype = "dotted", box.padding = .4 ) +
      theme_classic() + ggtitle(paste("Neighbors of ", title)) +  
      theme(axis.text.x = element_blank()) +
      theme(axis.title.x = element_blank()) +
      # theme(legend.text = element_text(size=3), legend.title = element_text(size=3)) + 
      theme(legend.position = 'none') +
        # theme( plot.margin = margin(0,200,0,0,  unit="pt")) + 
      scale_fill_manual(values = stableColorsL2.fxn) + scale_color_manual(values = stableColorsL2.fxn) 
    )
  }
  if(plots==T)   return(a)
  else return(res)
}

a <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'TBMIXED_LZ_BCELL', title='TBMIXED_LZ_BCELL',
                            filterPrefix='on', plots=T)
b <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'TBMIXED_MIXED_GC_TCELL', title='TBMIXED_MIXED_GC_TCELL',
                            filterPrefix='on', plots=T)
gridExtra::grid.arrange(a,b, nrow=1)

a <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'CD8TOXHI_TCF1_HI_TOX_HI_CD8',
                            title='CD8TOXHI_TCF1_HI_TOX_HI_CD8', filterPrefix='on', plots=T)
b <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'CD8TOXHI_TCF1_LO_TOX_HI_CD8',
                            title='CD8TOXHI_TCF1_LO_TOX_HI_CD8', filterPrefix='on', plots=T)
gridExtra::grid.arrange(a,b, nrow=1) 
  
```



# show the average neighborhood for aggregated sln 
```{r show-niche-multipleSample}
showNiche.multiSample <- function(annot.table, indexCluster, title, filterPrefix='on', plots=T)
{
  res <- list()
  IDlist <- names(annot.table)
  for (i in IDlist)  
  {   
    res[[i]] <- showNiche.singleSample(annot.table = annot.k9[[i]], indexCluster = indexCluster,
                                       title=title, filterPrefix = filterPrefix, plots=F )  # return per-sample proportions
    res[[i]]$list_ID <- i  # name by list_ID
  }   
  res <- rbindlist(res)  # merge into one data table 
  ave <-  res %>% group_by(subset) %>% summarise(mean_proportion = mean(proportion, na.rm = TRUE))  # calculate mean over grouping
  ave <- ave[order(ave$mean_proportion, decreasing = T),]  # order by decreasing value 
  res$subset <- factor(res$subset, levels = ave$subset)  # set the res$subset factors based on the decreasing average
  stableColorsL2.fxn <- stableColorsL2
  if(filterPrefix == 'on')  # trim color names accordingly 
  {  
    names(stableColorsL2.fxn) <- gsub(pattern = "CD8TOXHI_", replacement = "", x = names(stableColorsL2.fxn ))
    names(stableColorsL2.fxn) <- gsub(pattern = "CD8TOXLO_", replacement = "", x = names(stableColorsL2.fxn ))
    names(stableColorsL2.fxn) <- gsub(pattern = "TBMIXED_", replacement = "", x = names(stableColorsL2.fxn ))
  }
  
  if(plots == T)
  { 
    # scatter plot of biological replicates
    print( 
      ggplot(res, aes(y = subset, x = proportion, fill=subset, label=list_ID)) +
      geom_point(size=2, position=position_dodge(width=1), pch=21, stroke=0.1, color='grey70') +
      stat_summary( fun.min = "mean", fun.max = "mean", geom = "errorbar",
        width = 0.75, color = "black", linewidth = 0.5 ) +
      theme_bw() + ggtitle(title) + 
      # ggrepel::geom_label_repel(size=2, box.padding = unit(0.1,"lines"), alpha=0.2) + 
      scale_fill_manual(values= stableColorsL2.fxn)  + 
      theme(axis.text = element_text(size=6), axis.title.y = element_blank()) + 
      theme(legend.position = 'none')
    )
  }
  return(res)
}  
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'TBMIXED_LZ_BCELL', title='TBMIXED_LZ_BCELL')
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'TBMIXED_MIXED_GC_TCELL', title='TBMIXED_MIXED_GC_TCELL')
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'CD8TOXHI_TCF1_HI_TOX_HI_CD8',
                      title='CD8TOXHI_TCF1_HI_TOX_HI_CD8')
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'CD8TOXHI_TCF1_LO_TOX_HI_CD8',
                      title='CD8TOXHI_TCF1_LO_TOX_HI_CD8')

```

```{r plot-average-niche}

library(dplyr)

plot.Average.Niche <- function( annot.table, indexCluster, title )
{
  # initialize the data table 
  grid_data <- data.table(  
    x = c(1,2,3,3,3,2,1,1,2),
    y = c(3,3,3,2,1,1,1,2,2),
    subset = c("red", "blue", "green", "yellow", "purple", "orange", "pink", "cyan", "brown"),
    proportion = c(0.8, 0.5, 0.7, 0.6, 0.9, 0.4, 0.7, 0.3, 0.8) # Arbitrary proportions for each square
  )
  
  # make the center the index cell 
  res <- showNiche.multiSample(annot.table = annot.table, indexCluster = indexCluster, title=indexCluster, plots = F)
  ave <- res %>% group_by(subset) %>% summarise(mean_proportion = mean(proportion, na.rm = TRUE))
  ave <- ave[1:9,]
  ave$x[1:9] <- grid_data$x; ave$y[1:9] <- grid_data$y
  grid_data <- as.data.table(ave)
  grid_data[x == 2 & y == 2, subset:= indexCluster]
  grid_data[x == 2 & y == 2, mean_proportion:= 0]
  stableColorsL2.fxn <- stableColorsL2
  if(filterPrefix == 'on')  # trim color names accordingly 
  { 
    res[, subset := gsub(pattern = "CD8TOXHI_", replacement = "", x = .SD[[1]]), .SDcols = 1] 
    res[, subset := gsub(pattern = "CD8TOXLO_", replacement = "", x = .SD[[1]]), .SDcols = 1] 
    res[, subset := gsub(pattern = "TBMIXED_", replacement = "", x = .SD[[1]]), .SDcols = 1] 
    names(stableColorsL2.fxn) <- gsub(pattern = "CD8TOXHI_", replacement = "", x = names(stableColorsL2.fxn ))
    names(stableColorsL2.fxn) <- gsub(pattern = "CD8TOXLO_", replacement = "", x = names(stableColorsL2.fxn ))
    names(stableColorsL2.fxn) <- gsub(pattern = "TBMIXED_", replacement = "", x = names(stableColorsL2.fxn ))
    }
  # grid_data$mean_proportion <- grid_data$mean_proportion 
  grid_data$label <- gsub(pattern = "_", replacement = " ", x = grid_data$subset)
  grid_data$label <- stringr::str_wrap(grid_data$label, 5)
  grid_data$label.annot.x <- c(0, 2, 4, 4, 4, 2, 0, 0, 2)
  grid_data$label.annot.y <- c(4, 4, 4, 2, 0, 0, 0, 2, 2)
  hex.coord <- data.frame(x = c(2.5, 2.25, 1.75, 1.5, 1.75, 2.25), y=c(2, 2.433, 2.433, 2, 1.567, 1.567))
  print(
    ggplot(grid_data[1:8,], aes( x=x, y=y, fill=subset, label=label, size=mean_proportion)) + 
      geom_polygon(data = hex.coord, inherit.aes = F, aes(x = x, y = y), fill = "white", color = "black", size=0.05) +
      geom_point(shape = 22, color = 'gray50') +   
      geom_text(aes(x = label.annot.x, y=label.annot.y, label = label), size=2) + 
      coord_equal() + # Ensure squares are square (base aspect ratio)
      scale_size(range = c(0,20)) + scale_fill_manual(values = stableColorsL2.fxn) + 
      ggtitle(title) + guides(fill = "none")  +#  scale_size_area() + 
      annotate("text", x = 2, y = 2, label = "index cell", colour = "black", fontface = 'bold', size=2) +  
      theme_void() +
      xlim(c(-0.5,4.5)) + ylim(c(-0.5,4.5))
    )
}
plot.Average.Niche(annot.k9, indexCluster = 'TBMIXED_LZ_BCELL', title='TBMIXED_LZ_BCELL')
plot.Average.Niche(annot.k9, indexCluster = 'TBMIXED_MIXED_GC_TCELL', title='TBMIXED_MIXED_GC_TCELL')
a <- plot.Average.Niche(annot.table = annot.k9, indexCluster = 'CD8TOXHI_TCF1_HI_TOX_HI_CD8',
                      title='CD8TOXHI_TCF1_HI_TOX_HI_CD8')
b <- plot.Average.Niche(annot.table = annot.k9, indexCluster = 'CD8TOXHI_TCF1_LO_TOX_HI_CD8',
                      title='CD8TOXHI_TCF1_LO_TOX_HI_CD8')

```


```{r plot-average-niche-comparisons, fig.width=10, fig.height=5}
gridExtra::grid.arrange(a,b, nrow=1) 
```


```{r plot-average-niche-by-stage, fig.width=10, fig.height=5}
a <- plot.Average.Niche(annot.table = annot.k9[stage1.subgroup], indexCluster = 'CD8TOXHI_TCF1_HI_TOX_HI_CD8',
                      title='CD8TOXHI_TCF1_HI_TOX_HI_CD8 stage 1')
b <- plot.Average.Niche(annot.table = annot.k9[stage3.subgroup], indexCluster = 'CD8TOXHI_TCF1_HI_TOX_HI_CD8',
                      title='CD8TOXHI_TCF1_HI_TOX_HI_CD8 stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 
a <- plot.Average.Niche(annot.table = annot.k9[stage1.subgroup], indexCluster = 'CD8TOXHI_TCF1_LO_TOX_HI_CD8',
                      title='CD8TOXHI_TCF1_LO_TOX_HI_CD8 stage 1')
b <- plot.Average.Niche(annot.table = annot.k9[stage3.subgroup], indexCluster = 'CD8TOXHI_TCF1_LO_TOX_HI_CD8',
                      title='CD8TOXHI_TCF1_LO_TOX_HI_CD8 stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 
```

# contrast stage 1 vs stage 3 for niche  
```{r contrast-niche-by-stage}

rgb_to_hex <- function(rgb) {
  sprintf("#%02X%02X%02X", rgb[1], rgb[2], rgb[3])
}

lighten_hex <- function(hex, factor = 0.3) { # You can adjust the factor
  rgb <- col2rgb(hex)
  new_rgb <- round(rgb + (255 - rgb) * factor)
  new_rgb <- pmax(0, pmin(255, new_rgb))
  return(rgb_to_hex(new_rgb))
}


niche.contrastStage <- function( annot.table, indexCluster, common.title, filterPrefix='on')
{
  stage1.res <- showNiche.multiSample(annot.table[stage1.subgroup], indexCluster, title = paste0(common.title,".stage1"),
                                      filterPrefix = 'on', plots=F)
  stage3.res <- showNiche.multiSample(annot.table[stage3.subgroup], indexCluster, title = paste0(common.title,".stage3"),
                                      filterPrefix = 'on', plots=F)
  stage1.res$stage <- 'stage1';  stage3.res$stage <- 'stage3'
  res <- rbind(stage1.res, stage3.res)
  res$stage <- factor(res$stage, levels = c("stage3", "stage1"))
  res$subset <- factor(res$subset, levels = levels(stage3.res$subset))
  stableColorsL2.fxn <- stableColorsL2
  if(filterPrefix == 'on')   # trim color names accordingly
  {
    names(stableColorsL2.fxn) <- gsub(pattern = "CD8TOXHI_", replacement = "", x = names(stableColorsL2.fxn ))
    names(stableColorsL2.fxn) <- gsub(pattern = "CD8TOXLO_", replacement = "", x = names(stableColorsL2.fxn ))
    names(stableColorsL2.fxn) <- gsub(pattern = "TBMIXED_", replacement = "", x = names(stableColorsL2.fxn ))
  }

  # Create a lighter version of the color palette
  lighter_stableColorsL2.fxn <- sapply(stableColorsL2.fxn, lighten_hex, factor = 0.6) # Adjust factor as needed

  # Create a combined color mapping key
  res$color_key <- interaction(res$subset, res$stage, sep = "_")

  # Create the combined color palette
  combined_palette <- c(stableColorsL2.fxn, lighter_stableColorsL2.fxn[names(stableColorsL2.fxn)])
  names(combined_palette) <- c(paste0(names(stableColorsL2.fxn), "_stage3"), paste0(names(lighter_stableColorsL2.fxn), "_stage1"))
  
  annotateX <- max(res$proportion)-0.1*max(res$proportion)
  print(
    ggplot(res, aes(y = subset, x = proportion, fill=color_key, label=list_ID, group=stage)) +
      geom_point(size=2, position=position_dodge(width=1), pch=21, stroke=0.1, color='grey70') +
      annotate('text',x= annotateX ,y=35, label="s1 in light shade\ns3 in dark shade", size=3)+
      stat_summary( fun.min = "mean", fun.max = "mean", geom = "errorbar", position = position_dodge(width=1),
                    width = 1, color = "black", linewidth = 0.5 ) +
      theme_bw() + ggtitle(common.title) +
      scale_fill_manual(name = 'subset & stage', values= combined_palette,  guide = guide_legend(ncol = 2))  +
      theme(axis.text = element_text(size=6), axis.title.y = element_blank()) +
      theme(legend.position = 'none')
  )
  return(res)
}

# ********* hold over for calculating adjusted and unadjusted statistics 
#   aggregated_DT <- data[, .(proportion), by = .(right, list_ID, stage)]
#   interacting.subsets <- unique(aggregated_DT$right)
#   res <- list()
#   for (i in seq_along(interacting.subsets))
#   {
#     data = aggregated_DT[ right %in% interacting.subsets[i], ]
#     res[[i]] <- wilcox.test(proportion ~ stage, data = data)
#     names(res)[i] <- data$right[1]
#   }
#   
#   print(paste("Unadjusted pvalues:")); 
#   print( round(sapply(res, function(x) { x$p.value }), digits = 2) )
#   print(paste("FDR-adjusted pvalues"));  
#   print( round(p.adjust( sapply(res, function(x) { x$p.value }), method="fdr"), digits = 2)) 
# }

niche.contrastStage( annot.k9, indexCluster = 'TBMIXED_LZ_BCELL', common.title='TBMIXED_LZ_BCELL')
niche.contrastStage( annot.k9, indexCluster = 'TBMIXED_MIXED_GC_TCELL', common.title='TBMIXED_MIXEDGC_TCELL')
niche.contrastStage( annot.k9, indexCluster = 'CD8TOXHI_TCF1_HI_TOX_HI_CD8', common.title='CD8TOXHI_TCF1_HI_TOX_HI_CD8')


```


```{r}

```


# average neighborhood representation - single entity for GC populations
```{r show-average-neighborhood}
# Filter for interactions involving the index cell type
intxn.index.s1 <- y[ , .(left.stage1, right.stage1, original.stage1) ]

# filter to the Toxhi CD8
filterIntxn <- function(simpleIntxnTable, title, filterprefix='on')
{
  simpleIntxnTable[, proportion := .SD[[1]]/sum(.SD[[1]], na.rm=T), .SDcols = 3 ] # .SD[[1]] is the first and only column selected by .SDcols
  simpleIntxnTable = simpleIntxnTable[order(-proportion)]
  target_col_name_for_lhs_sd <- names(simpleIntxnTable)[1]
  simpleIntxnTable[, (target_col_name_for_lhs_sd) := gsub(pattern = "CD8TOXHI_", replacement = "", x = .SD[[1]]), .SDcols = 1] 
  target_col_name_for_lhs_sd <- names(simpleIntxnTable)[2]
  if(filterprefix == 'on')
  {  
    simpleIntxnTable[, (target_col_name_for_lhs_sd) := gsub(pattern = "CD8TOXHI_", replacement = "", x = .SD[[1]]), .SDcols = 2] 
    simpleIntxnTable[, (target_col_name_for_lhs_sd) := gsub(pattern = "CD8TOXLO_", replacement = "", x = .SD[[1]]), .SDcols = 2] 
    simpleIntxnTable[, (target_col_name_for_lhs_sd) := gsub(pattern = "TBMIXED_", replacement = "", x = .SD[[1]]), .SDcols = 2] 
    }
  simpleIntxnTable[,  (target_col_name_for_lhs_sd) := factor(get(target_col_name_for_lhs_sd), levels= get(target_col_name_for_lhs_sd) )]

  print( 
    ggplot(data = simpleIntxnTable, aes(y = .data[[names(simpleIntxnTable)[2]]], x = proportion, fill=.data[[names(simpleIntxnTable)[2]]],
           color = .data[[names(simpleIntxnTable)[2]]])) + 
    geom_bar(stat='identity', width=0.01) + 
    geom_point(size=5) + 
    theme_classic() + ggtitle(paste("Neighbors of ", title)) +  scale_fill_manual(values = stableColorsL2) + 
    theme(axis.text.x = element_text(angle=45, hjust=1,vjust=1), legend.position = 'none')
  )
  print(
    ggplot(data = simpleIntxnTable, aes(x = .data[[names(simpleIntxnTable)[1]]], y =  .data[[names(simpleIntxnTable)[3]]],  
                                          fill=.data[[names(simpleIntxnTable)[2]]])) + 
    geom_bar(stat='identity', width=1, position="fill") + 
    theme_classic() + ggtitle(paste("Neighbors of ", title)) + scale_fill_manual(values = stableColorsL2) + 
    theme(axis.text.x = element_text(angle=45, hjust=1,vjust=1))
  )

  return(simpleIntxnTable)
}
filterIntxn(simpleIntxnTable = intxn.index.s1[left.stage1 == 'TBMIXED_LZ_BCELL'], title='TBMIXED_LZ_BCELL', filterprefix='off')
filterIntxn(simpleIntxnTable = intxn.index.s1[left.stage1 == 'TBMIXED_MIXED_GC_TCELL'], title='TBMIXED_MIXED_GC_TCELL', filterprefix='off')
```


```{r show-average-neighborhood-radialPlot}
grep("TOXHI", unique(y$left.stage1), value = T)
tcfhiCD8.s1 <- filterIntxn(simpleIntxnTable = intxn.index.s1[left.stage1 == 'CD8TOXHI_TCF1_HI_TOX_HI_CD8'], title='CD8+Tox+Tcf+')
tcfloCD8.s1 <- filterIntxn(simpleIntxnTable = intxn.index.s1[left.stage1 == 'CD8TOXHI_TCF1_LO_TOX_HI_CD8'], title='CD8+Tox+Tcf-')
tcfhiCD8.s1$left.stage1 <- paste0(tcfhiCD8.s1$left.stage1, ".s1")
tcfloCD8.s1$left.stage1 <- paste0(tcfloCD8.s1$left.stage1, ".s1")
comb <- rbind(tcfhiCD8.s1, tcfloCD8.s1)
ggplot(comb, aes(y = right.stage1, x = proportion, group=left.stage1, fill=left.stage1)) + 
  geom_bar( stat='identity', width=0.5, position=position_dodge(preserve='single')) + 
  # coord_radial("y", start = 0, direction = 1, rotate.angle=T , r_axis_inside = TRUE)+
  ggtitle(paste("Neighbors of CD8 T cells - Stage 1"))  +  
  scale_fill_manual(values= ( RColorBrewer::brewer.pal(n = 4, name = "Accent")) )  + 
  theme(   axis.title = element_blank(),axis.ticks = element_blank(), axis.text.y = element_text(color = "grey"),
           axis.text.x = element_text(color = "black", size = 4), legend.position = "bottom", legend.title = element_blank(), 
           plot.title = element_text(face = "bold", size = 18, hjust = 1), 
           panel.background = element_rect(fill = "white", color = "white"), panel.grid = element_blank(),  
           panel.grid.major.x = element_line(color="grey90", size=0.5))

ggplot(comb, aes(y = right.stage1, x = proportion, group=left.stage1, fill=left.stage1)) + 
  geom_bar( stat='identity', width=0.5, position=position_dodge(preserve='single')) + 
  coord_radial("y", start = 0, direction = 1, rotate.angle=T , r_axis_inside = TRUE)+
  ggtitle(paste("Neighbors of CD8 T cells - Stage 1"))  +  
  scale_fill_manual(values= ( RColorBrewer::brewer.pal(n = 4, name = "Accent")) )  + 
  theme(   axis.title = element_blank(),axis.ticks = element_blank(), axis.text.y = element_text(color = "grey"),
           axis.text.x = element_text(color = "black", size = 4), legend.position = "bottom", legend.title = element_blank(), 
           plot.title = element_text(face = "bold", size = 18, hjust = 1), 
           panel.background = element_rect(fill = "white", color = "white"), panel.grid = element_blank(),  
           panel.grid.major.x = element_line(color="grey90", size=0.5))
```

# average neighborhood representation - contrast stage 3 vs stage 1 
```{r show-average-neighborhood-contrast}
intxn.index.s3 <- y[ , .(left.stage3, right.stage3, original.stage3) ]
tcfhiCD8.s3 <- filterIntxn(simpleIntxnTable = intxn.index.s3[left.stage3 == 'CD8TOXHI_TCF1_HI_TOX_HI_CD8'], title='CD8+Tox+Tcf+')
tcfhiCD8.s3$left.stage3 <- paste0(tcfhiCD8.s3$left.stage3, ".s3")
names(tcfhiCD8.s1) <- names(tcfhiCD8.s3)

comb <- rbind(tcfhiCD8.s1, tcfhiCD8.s3)

ggplot(comb, aes(y = right.stage3, x = proportion, group=left.stage3, fill=left.stage3)) + 
  geom_bar( stat='identity', width=0.5, position=position_dodge(preserve='single')) + 
  coord_radial("y", start = 0, direction = 1, rotate.angle=T, r_axis_inside = TRUE)+
  ggtitle(paste("Tcfhi CD8 T cells - Stage 3 vs 1"))  +  
  scale_fill_manual(values= ( RColorBrewer::brewer.pal(n = 4, name = "Paired")) )  + 
  theme(   axis.title = element_blank(),axis.ticks = element_blank() , axis.text.y = element_text(color = "grey"),
           axis.text.x = element_text(color = "black", size = 4), legend.position = "bottom", legend.title = element_blank(), 
           plot.title = element_text(face = "bold", size = 18, hjust = 1), 
           panel.background = element_rect(fill = "white", color = "white"), panel.grid = element_blank(),  
           panel.grid.major.x = element_line(color="grey90", size=0.5))
```



# plot as individual biological replicates 
```{r show-average-neighborhood-contrast-biolReplicates}
significant.Neighbors.acrossStage <- function( data, interacting.partner)
{
  data <- as.data.table(tidyr::separate(data, col = unified_int, into = c("left", "right"), sep = "--"))
  data <- data[left == interacting.partner] 
  data[, proportion := original/sum(original), by=.(stage, list_ID)]
  
  print( 
    ggplot(data, aes(y = right, x = proportion, group = stage, fill=stage, label=list_ID)) +
      geom_point(size=3, position=position_dodge(width=1), pch=21, stroke=0.1, color='white') +
      theme_classic() + ggtitle(paste("Neighbors of ", interacting.partner)) + 
      ggrepel::geom_label_repel(size=2, box.padding = unit(0.1,"lines"), alpha=0.2) + 
      scale_fill_manual(values= ( RColorBrewer::brewer.pal(n = 4, name = "Paired")) )  + 
      theme(axis.text.x = element_text(angle=45, hjust=1,vjust=1))
  )
  aggregated_DT <- data[, .(proportion), by = .(right, list_ID, stage)]
  interacting.subsets <- unique(aggregated_DT$right)
  res <- list()
  for (i in seq_along(interacting.subsets))
  {
    data = aggregated_DT[ right %in% interacting.subsets[i], ]
    res[[i]] <- wilcox.test(proportion ~ stage, data = data)
    names(res)[i] <- data$right[1]
  }
  
  print(paste("Unadjusted pvalues:")); 
  print( round(sapply(res, function(x) { x$p.value }), digits = 2) )
  print(paste("FDR-adjusted pvalues"));  
  print( round(p.adjust( sapply(res, function(x) { x$p.value }), method="fdr"), digits = 2)) 
}

significant.Neighbors.acrossStage(data = interaction.byStage, interacting.partner = 'CD8TOXHI_TCF1_HI_TOX_HI_CD8')


```


# show a radial plot of contrast of neighborhoods 
```{r show-average-neighborhood-contrast-stage3vs1}
intxn.index.s3 <- y[ , .(left.stage3, right.stage3, original.stage3) ]
tcfloCD8.s3 <- filterIntxn(simpleIntxnTable = intxn.index.s3[left.stage3 == 'CD8TOXHI_TCF1_LO_TOX_HI_CD8'], title='CD8+Tox+Tcf-')
tcfloCD8.s3$left.stage3 <- paste0(tcfloCD8.s3$left.stage3, ".s3")
names(tcfloCD8.s1) <- names(tcfloCD8.s3)

comb <- rbind(tcfloCD8.s1, tcfloCD8.s3)

ggplot(comb, aes(y = right.stage3, x = proportion, group=left.stage3, fill=left.stage3)) + 
  geom_bar( stat='identity', width=0.5, position=position_dodge(preserve='single')) + 
  coord_radial("y", start = 0, direction = 1, rotate.angle=T, r_axis_inside = TRUE)+
  ggtitle(paste("Tcflo CD8 T cells - Stage 3 vs 1"))  +  
  scale_fill_manual(values= ( RColorBrewer::brewer.pal(n = 4, name = "Paired")) )  + 
  theme(   axis.title = element_blank(),axis.ticks = element_blank() , axis.text.y = element_text(color = "grey"),
           axis.text.x = element_text(color = "black", size = 4), legend.position = "bottom", legend.title = element_blank(), 
           plot.title = element_text(face = "bold", size = 18, hjust = 1), 
           panel.background = element_rect(fill = "white", color = "white"), panel.grid = element_blank(),  
           panel.grid.major.x = element_line(color="grey90", size=0.5)) 


```

# 
```{r plot-interacting-pair}
ggplot(y, aes(x=unified_int, y=diff.enrich, label=unified_int, size=original.stage3)) + 
  geom_point(pch=21, color='white', fill='black') + 
  ggrepel::geom_label_repel(max.overlaps = 20) + ggtitle("PI_value:  Stage3 - Stage1") + 
  geom_hline(yintercept = 0, color="grey50") + theme_classic() + 
  theme(axis.text.x = element_blank())  

```


```{r plot-pairwise-enrichment-recurrence, fig.width=10}
y <- merge(x=df[which(df$subgroup=='no.recur'),], y=df[which(df$subgroup == 'recur'),] , by='unified_int', suffixes = c(".no.recur",".recur"), all=T)
y$diff.PI_value <- y$PI_value.recur - y$PI_value.no.recur
y <- y[!is.na(y$diff.PI_value),]  # remove NA values, which imply interactions not present in both settings
y$unified_int <- factor(y$unified_int, levels = y$unified_int[order(y$diff.PI_value, decreasing = T)])
y$unified_char <- paste0(y$left.recur,"--",y$right.recur)

y <- y[which(y$original.recur > 20),]

ggplot(y, aes(x=unified_int, y=diff.PI_value, label=unified_char)) + geom_point() + theme_classic() + 
  ggrepel::geom_label_repel(max.overlaps = 20) + ggtitle("PI_value:  Recur - noRecur") + 
  geom_hline(yintercept = 0.001, color="grey50") + 
  theme(axis.text.x = element_blank())  

```

# show stage 1 interactions
```{r proximity-heatmap-stage1, fig.width=10}
cellProximityHeatmap( sln, CPscore = stage1, scale = TRUE, order_cell_types = TRUE, color_breaks = NULL, color_names = NULL, )
cellProximityNetwork(sln, CPscore = stage1)

cellProximityHeatmap.retMatrix <- function(gobject, CPscore, scale = TRUE, order_cell_types = TRUE, color_breaks = NULL,
                                           color_names = NULL, show_plot = NULL, return_plot = NULL, save_plot = NULL, 
                                           save_param = list(), default_save_name = "cellProximityHeatmap") 
{
    enrich_res <- CPscore$enrichm_res

    # data.table variables
    first_type <- second_type <- unified_int <- NULL

    enrich_res[, first_type := strsplit( x = as.character(unified_int), split = "--"    )[[1]][1],by = seq_len(nrow(enrich_res))]
    enrich_res[, second_type := strsplit( x = as.character(unified_int), split = "--" )[[1]][2],by = seq_len(nrow(enrich_res)) ]

    # create matrix
    enrich_mat <- data.table::dcast.data.table( data = enrich_res, formula = first_type ~ second_type, value.var = "enrichm")
    matrix_d <- as.matrix(enrich_mat[, -1])
    rownames(matrix_d) <- as.vector(enrich_mat[[1]])
    t_matrix_d <- t_flex(matrix_d)

    # fill in NAs based on values in upper and lower matrix triangle
    t_matrix_d[upper.tri(t_matrix_d)][is.na(t_matrix_d[
        upper.tri(t_matrix_d)
    ])] <- matrix_d[upper.tri(matrix_d)][
        is.na(t_matrix_d[upper.tri(t_matrix_d)])
    ]
    t_matrix_d[lower.tri(t_matrix_d)][is.na(t_matrix_d[
        lower.tri(t_matrix_d)
    ])] <- matrix_d[lower.tri(matrix_d)][
        is.na(t_matrix_d[lower.tri(t_matrix_d)])
    ]
    t_matrix_d[is.na(t_matrix_d)] <- 0
    final_matrix <- t_matrix_d

    # scale data
    if (scale == TRUE) {
        final_matrix <- t_flex(scale(t_flex(final_matrix)))
        final_matrix <- t_flex(final_matrix)
        final_matrix[lower.tri(final_matrix)] <- t_flex(final_matrix)[
            lower.tri(final_matrix)
        ]
    }

    # order cell types
    if (order_cell_types == TRUE) {
        cordist <- stats::as.dist(1 - cor_flex(final_matrix))
        clus <- stats::hclust(cordist)
        myorder <- clus$order
        mylabels <- clus$labels
        names(mylabels) <- seq_along(mylabels)
        sample_order <- mylabels[myorder]

        final_matrix <- final_matrix[sample_order, sample_order]
    }

    # create custom colors or not
    if (!is.null(color_breaks) & !is.null(color_names)) {
        if (length(color_breaks) != 3 | !is.numeric(color_breaks)) {
            stop("color_breaks needs to be a numerical vector of length 3")
        }

        if (length(color_names) != 3 | !is.character(color_names)) {
            stop("color_names needs to be a character vector of length 3")
        }

        heatm <- ComplexHeatmap::Heatmap(matrix = final_matrix, cluster_rows = F, cluster_columns = F,
                                         col = GiottoVisuals::colorRamp2( breaks = color_breaks, colors = color_names) )
    } else {
        heatm <- ComplexHeatmap::Heatmap( matrix = final_matrix, cluster_rows = FALSE, cluster_columns = FALSE )
    }

    return(  list(final_matrix=final_matrix, clus=clus) )
}


```

# show stage 3 interactions 
```{r proximity-heatmap-stage3, fig.width=10}
Giotto::cellProximityHeatmap( sln, CPscore = stage3, scale = F, order_cell_types = TRUE, return_plot = T)
final_matrix <- cellProximityHeatmap.retMatrix( sln, CPscore = stage3, scale = TRUE, order_cell_types = T)$final_matrix
clus <- cellProximityHeatmap.retMatrix( sln, CPscore = stage3, scale = TRUE, order_cell_types = T)$clus
plot(clus)
ComplexHeatmap::Heatmap(final_matrix)

cellProximityNetwork(sln, CPscore = stage3 )
write.csv(stage3$enrichm_res, file = "cellProximityHeatmap.stage3.csv")
```


# interaction changed genes
```{r Interaction-changed-genes}
gd = fDataDT(sln)

# high_expressed_genes = gd[mean_expr_det > quantile(gd$mean_expr_det)[4]]$feat_ID

## identify genes that are associated with proximity to other cell types
# ICGscoresHighGenes = findInteractionChangedFeats(gobject = sln, selected_feats = gd$feat_ID,
#                                                  spatial_network_name = "knn_network", cluster_column = 'L1_annotations',
#                                                  diff_test = 'permutation', adjust_method = 'fdr',
#                                                  nr_permutations = 200, do_parallel = T)
# ## visualize
# plotCellProximityGenes(sln, cpgObject = ICGscoresHighGenes, method = 'dotplot')

md <- pDataDT(sln); 
keepCells <- subset(md, list_ID %in% c("tumor.1","tumor.2","tumor.3","tumor.4","tumor.5","tumor.6","tumor.7", "tumor.8",
                                                 "tumor.9","tumor.10"))
keepCells <- subset(keepCells, cell_ID %in% sample(keepCells$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)
date()
## identify genes that are associated with proximity to other cell types
ICGscores.L1 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L1_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
ICGscores.L2 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L2_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
saveRDS(object = list(ICGscores.L1 = ICGscores.L1, ICGscores.L2 = ICGscores.L2), 
        file = "neighborhood.ICGlist.s3.sln.clustering5.annot.spatnet.RDS"); date()  # took 6 hours at 300 permutations each 


md <- pDataDT(sln); 
keepCells <- subset(md, list_ID %in% c("ctrl.1","ctrl.2","ctrl.3","ctrl.4","ctrl.5"))
keepCells <- subset(keepCells, cell_ID %in% sample(keepCells$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)
date()
## identify genes that are associated with proximity to other cell types
ICGscores.L1 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L1_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
ICGscores.L2 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L2_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
saveRDS(object = list(ICGscores.L1 = ICGscores.L1, ICGscores.L2 = ICGscores.L2), 
        file = "neighborhood.ICGlist.s1.sln.clustering5.annot.spatnet.RDS"); date()


df.icg <- readRDS("neighborhood.ICGlist.s3.sln.clustering5.annot.spatnet.RDS")[[2]][['ICFscores']] # ICGscores.L2
df.icg.fil <- df.icg[ p.adj<0.01 & cell_type == 'CD8TOXHI_TCF1_HI_TOX_HI_CD8']
```


```{r Interaction-changed-genes-byCircosPlot, fig.width=10, fig.retina=2, eval=FALSE}
pair_counts <- df.icg[p.adj < 0.1, .N, by = .(cell_type, int_cell_type)] # .N is a special data.table symbol that gives the number of rows in each group
pair_counts <- pair_counts[N > 1]
single.space <- c(rep(1, length(unique(pair_counts$cell_type)))); names(single.space) <- unique(pair_counts$cell_type)
grid.col = getDistinctColors(n = length(unique(pair_counts$cell_type))); names(grid.col) <- names(single.space)
# circos.par(gap.after = single.space)
circlize::chordDiagram(pair_counts[1:30], direction.type = c('arrows','diffHeight'), directional=-1, transparency = 0,
                       grid.col = grid.col, diffHeight = -mm_h(2), 
                       annotationTrack = "grid", preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(pair_counts))))),
                       # big.gap=20
                       # link.arr.type = "big.arrow"
                       ) # negative direction to show direction of the effect, e.g. target is modified for x many features because of source

# we go back to the first track and customize sector labels
circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
}, bg.border = NA) # here set bg.border to NA is important
title("interaction changed features - first 30")

```

# stop evaluation here 
```{r Interaction-changed-genes-multitrackCircos, eval=FALSE}
# what about showing a multi-track chordDiagram where the proteins changed the most are indicated 
# would have to filter the protein list to exclude things like DAPI which aren't meaningful 
chordDiagram(pair_counts[1:30], direction.type = c('arrows','diffHeight'), directional=-1, transparency = 0,
                       grid.col = grid.col, diffHeight = -mm_h(2), 
                       # link.arr.type = "big.arrow"
                       ) # negative direction to show direction of the effect, e.g. target is modified for x many features because of source
circos.track(track.index = 1, panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
              facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
  }, bg.border = NA) # here set bg.border to NA is important

prots <- data.table(fDataDT(sln)$feat_ID)
prots <- prots[ !V1 %in% c("DAPI-01", "IL21", "CXCR3", "CD45", "TBET", "MART1", "GP100", "PAN.CYTO","MAC2/GAL3","VIMENTIN") ]
df.icg.fil <- df.icg.fil[ feats %in% prots$V1]



date()

```


```{r, eval=FALSE}
set.seed(999)
mat = matrix(sample(18, 18), 3, 6) 
rownames(mat) = paste0("S", 1:3)
colnames(mat) = paste0("E", 1:6)
mat

adjac = data.frame(from = rep(rownames(mat), times = ncol(mat)),
    to = rep(colnames(mat), each = nrow(mat)),
    value = as.vector(mat),
    stringsAsFactors = FALSE)
adjac


```


Job ID: 63949935
Cluster: slurm_cluster
User/Group: heratr01/heratr01
State: COMPLETED (exit code 0)
Nodes: 1
Cores per node: 10
CPU Utilized: 02:56:58
CPU Efficiency: 12.36% of 23:51:20 core-walltime
Job Wall-clock time: 02:23:08
Memory Utilized: 329.74 GB
Memory Efficiency: 65.95% of 500.00 GB











# original function for cellProximityEnrichment from github
cellProxEnrich.orig <- function(gobject,
    spat_unit = NULL,
    feat_type = NULL,
    spatial_network_name = "Delaunay_network",
    cluster_column,
    number_of_simulations = 1000,
    adjust_method = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ),
    set_seed = TRUE,
    seed_number = 1234) 
{
    # Set feat_type and spat_unit
    spat_unit <- set_default_spat_unit(
        gobject = gobject,
        spat_unit = spat_unit
    )
    feat_type <- set_default_feat_type(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type
    )

    # p.adj test
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))

    spatial_network_annot <- annotateSpatialNetwork(
        gobject = gobject,
        feat_type = feat_type,
        spat_unit = spat_unit,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column
    )

    # remove double edges between same cells #
    # a simplified network does not have double edges between cells #

    # data.table variables
    unified_cells <- type_int <- N <- NULL

    spatial_network_annot <- GiottoUtils:::dt_sort_combine_two_columns(
        spatial_network_annot, "to", "from", "unified_cells"
    )
    spatial_network_annot <- spatial_network_annot[!duplicated(unified_cells)]

    sample_dt <- Giotto:::make_simulated_network(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column,
        number_of_simulations = number_of_simulations,
        set_seed = set_seed,
        seed_number = seed_number
    )

    # combine original and simulated network
    table_sim_results <- sample_dt[, .N, by = c(
        "unified_int", "type_int", "round"
    )]

    ## create complete simulations
    ## add 0 if no single interaction was found
    unique_ints <- unique(table_sim_results[, .(unified_int, type_int)])

    # data.table with 0's for all interactions
    minimum_simulations <- unique_ints[rep(
        seq_len(nrow(unique_ints)), number_of_simulations
    ), ]
    minimum_simulations[, round := rep(
        paste0("sim", seq_len(number_of_simulations)),
        each = nrow(unique_ints)
    )]
    minimum_simulations[, N := 0]

    table_sim_minimum_results <- rbind(table_sim_results, minimum_simulations)
    table_sim_minimum_results[, V1 := sum(N), by = c(
        "unified_int", "type_int", "round"
    )]
    table_sim_minimum_results <- unique(
        table_sim_minimum_results[, .(unified_int, type_int, round, V1)]
    )
    table_sim_results <- table_sim_minimum_results


    # data.table variables
    orig <- unified_int <- V1 <- original <- enrichm <- simulations <- NULL

    table_sim_results[, orig := "simulations"]
    spatial_network_annot[, round := "original"]

    table_orig_results <- spatial_network_annot[, .N, by = c(
        "unified_int", "type_int", "round"
    )]
    table_orig_results[, orig := "original"]
    data.table::setnames(table_orig_results, old = "N", new = "V1")

    table_results <- rbind(table_orig_results, table_sim_results)



    # add missing combinations from original or simulations
    # probably not needed anymore
    all_simulation_ints <- as.character(unique(table_results[
        orig == "simulations"
    ]$unified_int))
    all_original_ints <- as.character(unique(table_results[
        orig == "original"
    ]$unified_int))
    missing_in_original <- all_simulation_ints[
        !all_simulation_ints %in% all_original_ints
    ]
    missing_in_simulations <- all_original_ints[
        !all_original_ints %in% all_simulation_ints
    ]
    create_missing_for_original <- table_results[
        unified_int %in% missing_in_original
    ]
    create_missing_for_original <- unique(create_missing_for_original[
        , c("orig", "V1") := list("original", 0)
    ])
    create_missing_for_simulations <- table_results[
        unified_int %in% missing_in_simulations
    ]
    create_missing_for_simulations <- unique(
        create_missing_for_simulations[, c("orig", "V1") := list(
            "simulations", 0
        )]
    )

    table_results <- do.call(
        "rbind",
        list(
            table_results, create_missing_for_original,
            create_missing_for_simulations
        )
    )


    ## p-values
    combo_list <- rep(NA, length = length(unique(table_results$unified_int)))
    p_high <- rep(NA, length = length(unique(table_results$unified_int)))
    p_low <- rep(NA, length = length(unique(table_results$unified_int)))

    for (int_combo in seq_along(unique(table_results$unified_int))) {
        this_combo <- as.character(unique(table_results$unified_int)[int_combo])

        sub <- table_results[unified_int == this_combo]

        orig_value <- sub[orig == "original"]$V1
        sim_values <- sub[orig == "simulations"]$V1

        length_simulations <- length(sim_values)
        if (length_simulations != number_of_simulations) {
            additional_length_needed <- number_of_simulations -
                length_simulations
            sim_values <- c(sim_values, rep(0, additional_length_needed))
        }

        p_orig_higher <- 1 - (sum((orig_value + 1) > (sim_values + 1)) /
            number_of_simulations)
        p_orig_lower <- 1 - (sum((orig_value + 1) < (sim_values + 1)) /
            number_of_simulations)

        combo_list[[int_combo]] <- this_combo
        p_high[[int_combo]] <- p_orig_higher
        p_low[[int_combo]] <- p_orig_lower
    }
    res_pvalue_DT <- data.table::data.table(
        unified_int = as.vector(combo_list),
        p_higher_orig = p_high,
        p_lower_orig = p_low
    )


    # depletion or enrichment in barplot format
    table_mean_results <- table_results[, .(mean(V1)), by = c(
        "orig", "unified_int", "type_int"
    )]
    table_mean_results_dc <- data.table::dcast.data.table(
        data = table_mean_results, formula = type_int + unified_int ~ orig,
        value.var = "V1"
    )
    table_mean_results_dc[, original := ifelse(is.na(original), 0, original)]
    table_mean_results_dc[, enrichm := log2((original + 1) / (simulations + 1))]


    table_mean_results_dc <- merge(
        table_mean_results_dc, res_pvalue_DT,
        by = "unified_int"
    )
    data.table::setorder(table_mean_results_dc, enrichm)
    table_mean_results_dc[, unified_int := factor(unified_int, unified_int)]

    # adjust p-values for mht

    # data.table variables
    p.adj_higher <- p.adj_lower <- p_lower_orig <- p_higher_orig <-
        PI_value <- int_ranking <- NULL

    table_mean_results_dc[, p.adj_higher := stats::p.adjust(
        p_higher_orig,
        method = sel_adjust_method
    )]
    table_mean_results_dc[, p.adj_lower := stats::p.adjust(
        p_lower_orig,
        method = sel_adjust_method
    )]


    table_mean_results_dc[, PI_value := ifelse(p.adj_higher <= p.adj_lower,
        -log10(p.adj_higher + (1 / number_of_simulations)) * enrichm,
        -log10(p.adj_lower + (1 / number_of_simulations)) * enrichm
    )]
    data.table::setorder(table_mean_results_dc, PI_value)

    # order
    table_mean_results_dc <- table_mean_results_dc[order(-PI_value)]
    table_mean_results_dc[, int_ranking := seq_len(.N)]

    return(list(
        raw_sim_table = table_results,
        enrichm_res = table_mean_results_dc
    ))
}
