---
title: "Neighborhood analyses"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---
 

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```
 
 
```{r call-libraries} 
date();  timeStamp <- strftime(Sys.time() , "%Y%m%d.%H%M%S")
library(Giotto)
library(future); if(parallelly::availableCores() > 2 )  # & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); # library(leiden);   # library(multinet);  
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
library(ComplexHeatmap)
library(circlize)
library(dplyr); library(tidyr)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

#  ****** libraries to install 
# library(factoextra)
```

# read in data object 
```{r read-annot-sln-spatnet}

sln <- readRDS("sln.clustering5.annot.spatnet.20250617.160512.RDS")
gc()
md <- pDataDT(sln)
stableColorsL2 <- getDistinctColors(length(unique(md$L2_annotations)));  names(stableColorsL2) <- unique(md$L2_annotations)
classifier_label_colors <- c('B Cell Follicle'='#107BDA','GC'='#EE6710', 'T/B Mixed'='yellow', 'T Cell Zone'='#FF0096',
                             'Tumor'='limegreen')
name_mapping <- c('ctrl.1' = '101', 'ctrl.2' = '102', 'ctrl.3' = '103', 'ctrl.4' = '104', 'ctrl.5' = '105',
                  'tumor.1' = '301', 'tumor.2' = '302', 'tumor.3' = '303', 'tumor.4' = '304', 'tumor.5' = '305',
                  'tumor.6' = '306', 'tumor.7' = '307','tumor.8' = '308','tumor.9' = '309', 'tumor.10' = '310' )
stage3.subgroup <- c("tumor.1","tumor.2","tumor.3","tumor.4","tumor.5","tumor.6","tumor.7", "tumor.8",
                                                 "tumor.9","tumor.10")
stage1.subgroup <- c("ctrl.1","ctrl.2","ctrl.3","ctrl.4","ctrl.5")
recur.subgroup <- c("tumor.3","tumor.4","tumor.5","tumor.7","tumor.8","tumor.10")
norecur.subgroup <- c("tumor.1","tumor.2", "tumor.6", "tumor.9")
gc()
```


```{r grant-example-images}
keepCells <- subset(md, list_ID == "tumor.5" & y %in% c(10000:20000) )$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
leiden_colors <- stableColorsL2
names(leiden_colors) <- unique(md$L2_annotations)
leiden_colors.focus <- leiden_colors 
leiden_colors.focus <- rep("gray90",length(leiden_colors)); leiden_colors.focus["1"] <- leiden_colors["1"]
spatPlot2D(sln.sm, cell_color = 'L2_annotations', cell_color_code = leiden_colors, 
                         point_size = 0.1, point_shape="no_border", point_alpha=0.4, axis_text=14, 
                         legend_text = 4)
# ggsave(filename = 'output/tumor.5.L2annot.pdf')
```


# calculate proximity enrichment per sample for k=5
```{r cellproxEnrich-function-functions}
cellProx.sim <- function(gobject, spat_unit = NULL, feat_type = NULL, spatial_network_name="spatialknn.k5", 
                         #spatial_network_name = "Delaunay_network",
    cluster_column, number_of_simulations = 1000, 
    adjust_method = c("none", "fdr", "bonferroni", "BH", "holm", "hochberg", "hommel", "BY"), 
    set_seed = TRUE, seed_number = 1234) 
{
    # Set feat_type and spat_unit
    spat_unit <- set_default_spat_unit(
        gobject = gobject,
        spat_unit = spat_unit
    )
    feat_type <- set_default_feat_type(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type
    )

    # p.adj test
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))

    spatial_network_annot <- annotateSpatialNetwork(
        gobject = gobject,
        feat_type = feat_type,
        spat_unit = spat_unit,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column
    )

    # remove double edges between same cells #
    # a simplified network does not have double edges between cells #

    # data.table variables
    unified_cells <- type_int <- N <- NULL

    spatial_network_annot <- GiottoUtils:::dt_sort_combine_two_columns(
        spatial_network_annot, "to", "from", "unified_cells"
    )
    spatial_network_annot <- spatial_network_annot[!duplicated(unified_cells)]

    sample_dt <- Giotto:::make_simulated_network(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column,
        number_of_simulations = number_of_simulations,
        set_seed = set_seed,
        seed_number = seed_number
    )

    # combine original and simulated network
    table_sim_results <- sample_dt[, .N, by = c(
        "unified_int", "type_int", "round"
    )]

    ## create complete simulations
    ## add 0 if no single interaction was found
    unique_ints <- unique(table_sim_results[, .(unified_int, type_int)])

    # data.table with 0's for all interactions
    minimum_simulations <- unique_ints[rep(
        seq_len(nrow(unique_ints)), number_of_simulations
    ), ]
    minimum_simulations[, round := rep(
        paste0("sim", seq_len(number_of_simulations)),
        each = nrow(unique_ints)
    )]
    minimum_simulations[, N := 0]

    table_sim_minimum_results <- rbind(table_sim_results, minimum_simulations)
    table_sim_minimum_results[, V1 := sum(N), by = c(
        "unified_int", "type_int", "round"
    )]
    table_sim_minimum_results <- unique(
        table_sim_minimum_results[, .(unified_int, type_int, round, V1)]
    )
    table_sim_results <- table_sim_minimum_results


    # data.table variables
    orig <- unified_int <- V1 <- original <- enrichm <- simulations <- NULL

    table_sim_results[, orig := "simulations"]
    spatial_network_annot[, round := "original"]

    table_orig_results <- spatial_network_annot[, .N, by = c(
        "unified_int", "type_int", "round"
    )]
    table_orig_results[, orig := "original"]
    data.table::setnames(table_orig_results, old = "N", new = "V1")

    table_results <- rbind(table_orig_results, table_sim_results)



    # add missing combinations from original or simulations
    # probably not needed anymore
    all_simulation_ints <- as.character(unique(table_results[
        orig == "simulations"
    ]$unified_int))
    all_original_ints <- as.character(unique(table_results[
        orig == "original"
    ]$unified_int))
    missing_in_original <- all_simulation_ints[
        !all_simulation_ints %in% all_original_ints
    ]
    missing_in_simulations <- all_original_ints[
        !all_original_ints %in% all_simulation_ints
    ]
    create_missing_for_original <- table_results[
        unified_int %in% missing_in_original
    ]
    create_missing_for_original <- unique(create_missing_for_original[
        , c("orig", "V1") := list("original", 0)
    ])
    create_missing_for_simulations <- table_results[
        unified_int %in% missing_in_simulations
    ]
    create_missing_for_simulations <- unique(
        create_missing_for_simulations[, c("orig", "V1") := list(
            "simulations", 0
        )]
    )
# stop **********************************************************************************************

    table_results <- do.call(
        "rbind",
        list(
            table_results, create_missing_for_original,
            create_missing_for_simulations
        )
    )
    return(table_results)
}

cellProx.calcP <- function(table_results, number_of_simulations = 1000, 
    adjust_method = c("none", "fdr", "bonferroni", "BH", "holm", "hochberg", "hommel", "BY"), 
    set_seed = TRUE, seed_number = 1234)
{
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))
    ## p-values
    combo_list <- rep(NA, length = length(unique(table_results$unified_int)))
    p_high <- rep(NA, length = length(unique(table_results$unified_int)))
    p_low <- rep(NA, length = length(unique(table_results$unified_int)))

    for (int_combo in seq_along(unique(table_results$unified_int))) {
        this_combo <- as.character(unique(table_results$unified_int)[int_combo])

        sub <- table_results[unified_int == this_combo]

        orig_value <- sub[orig == "original"]$V1
        sim_values <- sub[orig == "simulations"]$V1

        length_simulations <- length(sim_values)
        if (length_simulations != number_of_simulations) {
            additional_length_needed <- number_of_simulations -
                length_simulations
            sim_values <- c(sim_values, rep(0, additional_length_needed))
        }

        p_orig_higher <- 1 - (sum((orig_value + 1) > (sim_values + 1)) /
            number_of_simulations)
        p_orig_lower <- 1 - (sum((orig_value + 1) < (sim_values + 1)) /
            number_of_simulations)

        combo_list[[int_combo]] <- this_combo
        p_high[[int_combo]] <- p_orig_higher
        p_low[[int_combo]] <- p_orig_lower
    }
    res_pvalue_DT <- data.table::data.table(
        unified_int = as.vector(combo_list),
        p_higher_orig = p_high,
        p_lower_orig = p_low
    )


    # depletion or enrichment in barplot format
    table_mean_results <- table_results[, .(mean(V1)), by = c(
        "orig", "unified_int", "type_int"
    )]
    table_mean_results_dc <- data.table::dcast.data.table(
        data = table_mean_results, formula = type_int + unified_int ~ orig,
        value.var = "V1"
    )
    table_mean_results_dc[, original := ifelse(is.na(original), 0, original)]
    table_mean_results_dc[, enrichm := log2((original + 1) / (simulations + 1))]


    table_mean_results_dc <- merge(
        table_mean_results_dc, res_pvalue_DT,
        by = "unified_int"
    )
    data.table::setorder(table_mean_results_dc, enrichm)
    table_mean_results_dc[, unified_int := factor(unified_int, unified_int)]

    # adjust p-values for mht

    # data.table variables
    p.adj_higher <- p.adj_lower <- p_lower_orig <- p_higher_orig <-
        PI_value <- int_ranking <- NULL

    table_mean_results_dc[, p.adj_higher := stats::p.adjust(
        p_higher_orig,
        method = sel_adjust_method
    )]
    table_mean_results_dc[, p.adj_lower := stats::p.adjust(
        p_lower_orig,
        method = sel_adjust_method
    )]


    table_mean_results_dc[, PI_value := ifelse(p.adj_higher <= p.adj_lower,
        -log10(p.adj_higher + (1 / number_of_simulations)) * enrichm,
        -log10(p.adj_lower + (1 / number_of_simulations)) * enrichm
    )]
    data.table::setorder(table_mean_results_dc, PI_value)

    # order
    table_mean_results_dc <- table_mean_results_dc[order(-PI_value)]
    table_mean_results_dc[, int_ranking := seq_len(.N)]

    return(list(
        raw_sim_table = table_results,
        enrichm_res = table_mean_results_dc
    ))
}

list_ID <- unique(md$list_ID)
gc()
```


```{r cellproxEnrich-function-k5}
tableResults.k5 <- list()
annot.k5 <- list()
for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=5, verbose=T, name = "spatialknn.k5.persample")
  annot.k5[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L2_annotations", spatial_network_name = 'spatialknn.k5.persample')
  tableResults.k5[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L2_annotations', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "spatialknn.k5.persample")
}
# saveRDS(list(annot.k5=annot.k5, tableResults.k5 = tableResults.k5), file="neighborhood.annot.k5.sln.RDS")  # all data merged
# annot.k5 <- readRDS(file="neighborhood.annot.k5.sln.RDS")[["annot.k5"]]  # all data merged
```


```{r cellproxEnrich-function-k9}
# calculate proximity enrichment per sample for k=9
tableResults.k9 <- list()
annot.k9 <- list()
for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=9, verbose=T, name = "spatialknn.k9.persample") 
  print("made spatialKNN")
  annot.k9[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L2_annotations", spatial_network_name = 'spatialknn.k9.persample')
  tableResults.k9[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L2_annotations', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "spatialknn.k9.persample")
}

# saveRDS(list(annot.k9 = annot.k9, tableResults.k9 = tableResults.k9), file="neighborhood.annot.k9.sln.RDS")  # all data merged


# ******* L3 annotations proximity enrichment ***********
# tableResults.k9 <- list()
# annot.k9 <- list()
# for (i in list_ID)
# {
#   print(i)
#   keepCells <- subset(md, list_ID %in% i )$cell_ID
#   sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
#   sln.sm <- createSpatialKNNnetwork(sln.sm, k=9, verbose=T, name = "spatialknn.k9.persample") 
#   print("made spatialKNN")
#   annot.k9[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L3_annotations", spatial_network_name = 'spatialknn.k9.persample')
#   tableResults.k9[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L3_annotations', adjust_method = "fdr",
#                                        number_of_simulations = 200, spatial_network_name = "spatialknn.k9.persample")
# }

# pull example cell for study
# annot.k9$tumor.5[from %like% 215265 | to %like% 215265] %>% write.csv("neighbors.of.tumor5.215265.csv")

# saveRDS(list(annot.k9 = annot.k9, tableResults.k9 = tableResults.k9), file="neighborhood.annot.k9.L3.sln.RDS")  # all data merged
#  
dt <- purrr::list_rbind(tableResults.k9, names_to = "id")
```

Fails on k=10 on tumor.6  (2 million rows) with long vectors error 
Works for k=9 for all samples 

```{r read-in-annotk9, eval=FALSE}
dt <- purrr::list_rbind(readRDS(file="neighborhood.annot.k9.sln.RDS")[['tableResults.k9']], names_to = "id")
annot.k9 <- readRDS(file="neighborhood.annot.k9.sln.RDS")[['annot.k9']]  #  read in L2 annotations
```


# subgroup by specified metadata  
```{r custom-cellproxEnrich-function-subgroup}
# subset dt based on groupings of interest 
cellProx.subgroup <- function(dt)
{
  wide_dt <- dcast(dt,                # Your input data.table
                   unified_int ~ round, # Formula: rows ~ columns
                   value.var = "V1",  # Column with values for the new wide cells
                   fun.aggregate = mean, # Function to aggregate values (when multiple rows match a cell)
                   na.rm = TRUE,      # Tell mean() to ignore NA values during aggregation
                   fill = 0)    
  setkey(wide_dt, unified_int) # Good practice
  long_dt <- melt(wide_dt,
                  id.vars = "unified_int",       # Column(s) to keep as identifier variables
                  variable.name = "round",       # Name for the new column holding the former column names (1, 2, 3)
                  value.name = "V1",             # Name for the new column holding the values from the melted columns
                  variable.factor = FALSE)       # Keep the 'round' column as character initially (avoids factor issues)
  long_dt$orig <- "NA"
  long_dt$orig[grep("original", long_dt$round)] <- "original"
  long_dt$orig[grep("^sim", long_dt$round)] <- "simulations"
  long_dt$type_int <- "NA"
  long_dt$tosplit <- long_dt$unified_int
  long_dt <- as.data.table(tidyr::separate(long_dt, col = tosplit, into = c("left", "right"), sep = "--"))
  long_dt$type_int <- ifelse( long_dt$left == long_dt$right, "homo", "hetero")
  long_dt$left <- long_dt$right <- NULL
  out <- cellProx.calcP(long_dt, number_of_simulations = 200 )  # needs to match what was used to calculate sims
  return(out)
}

stage3 <- cellProx.subgroup(subset(dt, id %in% stage3.subgroup))
stage1 <- cellProx.subgroup(subset(dt, id %in% stage1.subgroup ))

recur <- cellProx.subgroup(subset(dt, id %in% c("tumor.3","tumor.4","tumor.5","tumor.7","tumor.8","tumor.10")))
no.recur <- cellProx.subgroup(subset(dt, id %in% c("tumor.1","tumor.2", "tumor.6", "tumor.9") ))

sampleList <- unique(md$list_ID)
perSampleInteraction <- list()
for (i in seq_along(sampleList))
{
  print( paste("Interactions for sample", sampleList[i]) )
  perSampleInteraction[[i]] <- cellProx.subgroup(subset(dt, id %in% sampleList[i]))  # format in a per-sample fashion 
  perSampleInteraction[[i]]$name <- subset(dt, id %in% sampleList[i])$id[1]   # carry forward the sample name 
  names(perSampleInteraction)[i] <-  subset(dt, id %in% sampleList[i])$id[1]
}

# saveRDS(perSampleInteraction, file="neighborhood.perSampleInteraction.RDS")  # all data merged 
# perSampleInteraction <- readRDS(file="neighborhood.perSampleInteraction.RDS")

perSampleInteraction.enrichm_res <- list()
for (i in seq_along(perSampleInteraction))
{
  perSampleInteraction.enrichm_res[[i]] <- perSampleInteraction[[i]]$enrichm_res  # save the enrichm results table 
  names(perSampleInteraction.enrichm_res)[i] <- names(perSampleInteraction)[i]
}

addLabels.fxn <- function(dt, label)
{
  for (i in seq_along(dt))   {   dt[[i]][, list_ID := names(dt)[i]]   }  # attach the sample name to a new column
  dt <- rbindlist(dt)  # turn the list into data table
  dt$condition <- label  # add in a column for the stage 
  return(dt)
}

interaction.byStage <- rbind(addLabels.fxn(perSampleInteraction.enrichm_res[stage3.subgroup], label = "stage3"),
             addLabels.fxn(perSampleInteraction.enrichm_res[stage1.subgroup], label = "stage1") ) 
interaction.byRecur <- rbind(addLabels.fxn(perSampleInteraction.enrichm_res[recur.subgroup], label = "recur"),
             addLabels.fxn(perSampleInteraction.enrichm_res[norecur.subgroup], label = "no.recur") ) 

```


# find all neighborhoods within a fixed radius around the cell 
```{r neighborhoods-by-fixed-radius}

gc()

tableResults.rad <- list()
annot.rad <- list()  # annot.radius
for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=50, maximum_distance = 15, verbose=T, name = "spatialknn.radius.persample")
  annot.rad[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L2_annotations", spatial_network_name = 'spatialknn.radius.persample')
  tableResults.rad[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L2_annotations', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "spatialknn.radius.persample")
}
# saveRDS(list(annot.rad = annot.rad, tableResults.rad = tableResults.rad), 
#         file="neighborhood.annot.radius.sln.RDS")  # all data merged
# annot.rad <- readRDS(file="neighborhood.annot.radius.sln.RDS")[['annot.rad']]  #  read in L2 annotations 
dt <- purrr::list_rbind(tableResults.rad, names_to = "id")

```


```{r interaction-by-sample-radius-loadData, eval=FALSE}
dt <- purrr::list_rbind(readRDS(file="neighborhood.annot.radius.sln.RDS")[['tableResults.rad']], names_to = "id")
```


```{r interaction-by-sample-radius-calculate}
stage3 <- cellProx.subgroup(subset(dt, id %in% stage3.subgroup))
stage1 <- cellProx.subgroup(subset(dt, id %in% stage1.subgroup ))

recur <- cellProx.subgroup(subset(dt, id %in% c("tumor.3","tumor.4","tumor.5","tumor.7","tumor.8","tumor.10")))
no.recur <- cellProx.subgroup(subset(dt, id %in% c("tumor.1","tumor.2", "tumor.6", "tumor.9") ))

sampleList <- unique(md$list_ID)
perSampleInteraction <- list()
for (i in seq_along(sampleList))
{
  print( paste("Interactions for sample", sampleList[i]) )
  perSampleInteraction[[i]] <- cellProx.subgroup(subset(dt, id %in% sampleList[i]))  # format in a per-sample fashion 
  perSampleInteraction[[i]]$name <- subset(dt, id %in% sampleList[i])$id[1]   # carry forward the sample name 
  names(perSampleInteraction)[i] <-  subset(dt, id %in% sampleList[i])$id[1]
}

perSampleInteraction.enrichm_res <- list()
for (i in seq_along(perSampleInteraction))
{
  perSampleInteraction.enrichm_res[[i]] <- perSampleInteraction[[i]]$enrichm_res  # save the enrichm results table 
  names(perSampleInteraction.enrichm_res)[i] <- names(perSampleInteraction)[i]
}

interaction.byStage.rad <- rbind(addLabels.fxn(perSampleInteraction.enrichm_res[stage3.subgroup], label = "stage3"),
             addLabels.fxn(perSampleInteraction.enrichm_res[stage1.subgroup], label = "stage1") ) 
interaction.byRecur.rad <- rbind(addLabels.fxn(perSampleInteraction.enrichm_res[recur.subgroup], label = "recur"),
             addLabels.fxn(perSampleInteraction.enrichm_res[norecur.subgroup], label = "no.recur") ) 
rm(dt)
# gc()
```


# show the average neighborhood for sln 5 
```{r show-niche-singleSample, fig.height=6}

removeArtifacts <- function(dt)
{
  dt <- dt[!unified_int %like% "ARTIFACT"]
  dt <- dt[!unified_int %like% "outlier"]
  return(dt)
}

showNiche.singleSample <- function(annot.table, indexCluster, title, plots=TRUE)
{
  # example annot.table[['tumor.5']]
  #                        from                     to sdimx_begin sdimy_begin sdimx_end sdimy_end distance     weight
  #                    <char>                 <char>       <num>       <num>     <num>     <num>    <num>      <num>
  # 1: tumor.5-tumor.5_100000  tumor.5-tumor.5_99987     39429.5     14436.0   39446.5   14427.0 19.23538 0.04941838
  # 2: tumor.5-tumor.5_100000  tumor.5-tumor.5_99971     39429.5     14436.0   39428.5   14416.5 19.52562 0.04871959
  # 3: tumor.5-tumor.5_100000  tumor.5-tumor.5_99999     39429.5     14436.0   39408.0   14439.0 21.70829 0.04403677
  #                    to_cell_type              from_cell_type type_int                                                 from_to
  #                         <char>                      <char>   <char>                                                  <char>
  # 1:                  MEMORY_CD4                  MEMORY_CD8   hetero                                   MEMORY_CD8-MEMORY_CD4
  # 2:                  MEMORY_CD4                  MEMORY_CD8   hetero                                   MEMORY_CD8-MEMORY_CD4
  # 3:                ICOS_LO_TREG                  MEMORY_CD8   hetero                                 MEMORY_CD8-ICOS_LO_TREG
  #                                                       unified_int
  #                                                      <char>
  # 1:                                   MEMORY_CD4--MEMORY_CD8
  # 2:                                   MEMORY_CD4--MEMORY_CD8
  # 3:                                 ICOS_LO_TREG--MEMORY_CD8
  # 
  
  res <- removeArtifacts(annot.table)
  res <- rbind( res[from_cell_type == indexCluster, to_cell_type], 
                res[to_cell_type == indexCluster, from_cell_type] )  
  # because the from/to is not really directional, and we'd be missing some of the parings if we assume directionality
  res <- res %>% table %>% prop.table  # calculate proportions 
  res <- as.data.table(res); colnames(res) <- c('subset','proportion')
  res <- res[order(res$proportion, decreasing=T)]
  stableColorsL2.fxn <- stableColorsL2

  res$subset <- factor(res$subset, levels = res$subset)   
  
  if(plots == T)
  {  
    # lollipop chart 
    print( 
      a <- ggplot(data = res, aes(y = subset, x = proportion, fill=subset, color = subset)) + 
      geom_bar(stat='identity', width=0.01) + 
      geom_point(size=5) + 
      theme_classic() + ggtitle(title) + 
        scale_fill_manual(values = stableColorsL2.fxn) +
        scale_color_manual(values = stableColorsL2.fxn) +
      theme(axis.text.y = element_text( size=6), axis.title.y = element_blank(), plot.title=element_text(size=8)) + 
        theme(legend.position = 'none')  
    )
    
    # stacked bar chart 
    print(
      ggplot(data = res, aes(y = proportion, x = 'blank',  fill = subset)) + 
      geom_col(width=1, position="fill") + 
      ggrepel::geom_text_repel( aes(label = subset ),
        position = position_stack(vjust = 0.5),           # specify stacked position to follow bar layer's default position
        size = 2,  color='black', direction = "y", xlim = c(1.5, NA), hjust = 0,
        segment.linetype = "dotted", box.padding = .4 ) +
      theme_classic() + ggtitle(paste("Neighbors of ", title)) +  
      theme(axis.text.x = element_blank()) +
      theme(axis.title.x = element_blank()) +
      # theme(legend.text = element_text(size=3), legend.title = element_text(size=3)) + 
      theme(legend.position = 'none') +
        # theme( plot.margin = margin(0,200,0,0,  unit="pt")) + 
      scale_fill_manual(values = stableColorsL2.fxn) + scale_color_manual(values = stableColorsL2.fxn) 
    )
  }
  if(plots==T)   return(a)
  else return(res)
}

a <- removeArtifacts(annot.k9[['tumor.5']])
b <- a[to_cell_type == 'ACTIVATED_TREG']
a <- a[from_cell_type == 'ACTIVATED_TREG']
# write.csv(x = rbind(a, b), file = "neighborsTable.ACTIVATED_TREG.tumor5.csv")
# write.csv(x = annot.k9[['tumor.5']], file = "neighborsTable.allCells.tumor5.csv")
#   
  
a <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'LZ_BCELL', title='LZ_BCELL',
                            plots=T)
b <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'MIXED_GC_TCELL', title='MIXED_GC_TCELL',
                            plots=T)
gridExtra::grid.arrange(a,b, nrow=1)

a <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                            title='TCF1_HI_TOX_HI_CD8', plots=T)
b <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                            title='TCF1_LO_TOX_HI_CD8', plots=T)
gridExtra::grid.arrange(a,b, nrow=1) 
  

```



# show the average neighborhood for all sln 
```{r show-niche-multipleSample}
showNiche.multiSample <- function(annot.table, indexCluster, title,  plots=T, parallel=F)
{
  library(future.apply)
  res <- list()
  IDlist <- names(annot.table)
  if(parallel == T)
  {  
    res_list <- future_lapply(IDlist, function(i) {
      single_res <- showNiche.singleSample( annot.table = annot.table[[i]], indexCluster = indexCluster, title = title, plots = F )
      single_res$list_ID <- i    # name by list_ID
      return(single_res)
    }, future.seed = TRUE) # future.seed = TRUE is crucial for reproducible results!
  }  
  
  if (parallel == F)
  {  
    for (i in IDlist)
    {
      res[[i]] <- showNiche.singleSample(annot.table = annot.table[[i]], indexCluster = indexCluster,
                                         title=title, plots=F )  # return per-sample proportions
      res[[i]]$list_ID <- i  # name by list_ID
    }
  }
  res <- rbindlist(res)  # merge into one data table 
  ave <-  res %>% group_by(subset) %>% summarise(mean_proportion = mean(proportion, na.rm = TRUE))  # calculate mean over grouping
  ave <- ave[order(ave$mean_proportion, decreasing = T),]  # order by decreasing value 
  res$subset <- factor(res$subset, levels = ave$subset)  # set the res$subset factors based on the decreasing average
  stableColorsL2.fxn <- stableColorsL2

  if(plots == T)
  { 
    # scatter plot of biological replicates
    print( 
      ggplot(res, aes(y = subset, x = proportion, fill=subset, label=list_ID)) +
      geom_point(size=2, position=position_dodge(width=1), pch=21, stroke=0.2, color='grey50') +
      stat_summary( fun.min = "mean", fun.max = "mean", geom = "errorbar",
        width = 0.75, color = "black", linewidth = 0.5 ) + # geom_path(aes(group = list_ID), alpha=0.2, color='grey80') + 
      theme_bw() + ggtitle(title) + 
      # ggrepel::geom_label_repel(size=2, box.padding = unit(0.1,"lines"), alpha=0.2) + 
      scale_fill_manual(values= stableColorsL2.fxn)  + 
      theme(axis.text = element_text(size=6), axis.title.y = element_blank()) + 
      theme(legend.position = 'none')
    )
  }
  return(res)
}  
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'LZ_BCELL', title='LZ_BCELL', parallel=F)
 # ggsave("output/LZBCELL.scatterplot.allSamples.pdf", height=6,width=5)
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'TCF1_HI_TOX_HI_CD8', title='TCF1_HI_TOX_HI_CD8')
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'TCF1_LO_TOX_HI_CD8', title='TCF1_LO_TOX_HI_CD8')

```

```{r plot-average-niche}

plot.Average.Niche.squares <- function( annot.table, indexCluster, title )
{
  # initialize the data table 
  grid_data <- data.table(  
    x = c(1,2,3,3,3,2,1,1,2),
    y = c(3,3,3,2,1,1,1,2,2),
    subset = c("red", "blue", "green", "yellow", "purple", "orange", "pink", "cyan", "brown"),
    proportion = c(0.8, 0.5, 0.7, 0.6, 0.9, 0.4, 0.7, 0.3, 0.8) # Arbitrary proportions for each square
  )
  
  # make the center the index cell 
  res <- showNiche.multiSample(annot.table = annot.table, indexCluster = indexCluster, title=indexCluster, plots = F)
  stableColorsL2.fxn <- stableColorsL2

  ave <- res %>% group_by(subset) %>% summarise(mean_proportion = mean(proportion, na.rm = TRUE))
  ave <- ave[1:9,]
  ave$x[1:9] <- grid_data$x; ave$y[1:9] <- grid_data$y
  grid_data <- as.data.table(ave)
  grid_data[x == 2 & y == 2, subset:= indexCluster]
  grid_data[x == 2 & y == 2, mean_proportion:= 0]

  # grid_data$mean_proportion <- grid_data$mean_proportion 
  grid_data$label <- gsub(pattern = "_", replacement = " ", x = grid_data$subset)
  grid_data$label <- stringr::str_wrap(grid_data$label, 5)
  grid_data$label.annot.x <- c(0.5, 2, 3.5, 3.5, 3.5, 2, 0.5, 0.5, 2)
  grid_data$label.annot.y <- c(3.75, 3.75, 3.75, 2, 0.25, 0.25, 0.25, 2, 2)
  hex.coord <- data.frame(x = c(2.5, 2.25, 1.75, 1.5, 1.75, 2.25), y=c(2, 2.433, 2.433, 2, 1.567, 1.567))
  print(
    ggplot(grid_data[1:8,], aes( x=x, y=y, fill=subset, label=label, size=mean_proportion)) + 
      geom_polygon(data = hex.coord, inherit.aes = F, aes(x = x, y = y), fill = "white", color = "black", size=0.05) +
      geom_point(shape = 22, color = 'gray50') +   
      geom_text(aes(x = label.annot.x, y=label.annot.y, label = label), size=2) + 
      coord_equal() + # Ensure squares are square (base aspect ratio)
      scale_size(range = c(0,20)) + scale_fill_manual(values = stableColorsL2.fxn) + 
      ggtitle(title) + guides(fill = "none")  +#  scale_size_area() + 
      annotate("text", x = 2, y = 2, label = title, colour = "black", fontface = 'bold', size=2) +  
      theme_void() +
      xlim(c(-0.5,4.5)) + ylim(c(-0.5,4.5))
    )
}

plot.Average.Niche.igraph <- function( annot.table, indexCluster, title )
{
  library(ggraph)
  # make the center the index cell 
  res <- showNiche.multiSample(annot.table = annot.table, indexCluster = indexCluster, title=indexCluster, plots = F)
  stableColorsL2.fxn <- stableColorsL2

  ave <- res %>% group_by(subset) %>% summarise(mean_proportion = mean(proportion, na.rm = TRUE))
  ave$subset <- as.character(ave$subset)
  if (indexCluster %in% ave$subset)   {    
    ave$subset[which(ave$subset == indexCluster)] <- paste0(indexCluster, ".")  
    duplColor <- stableColorsL2.fxn[which(names(stableColorsL2.fxn)==indexCluster)]
    names(duplColor) <- paste0(names(duplColor), ".")
    stableColorsL2.fxn <- c(stableColorsL2.fxn, duplColor)
  }
  center_node_df <- data.frame(subset = indexCluster, mean_proportion = 0) 
  setorder(ave, -mean_proportion)  # order ave from biggest to smallest 
  ave <- ave[1:8,]  # keep the top 8 
  grid_data <- as.data.table(bind_rows(center_node_df, ave))  # add the index cell to the data frame 
  grid_data$label <- grid_data$subset  # wrap labels  
  grid_data$label <- gsub(pattern = "_",replacement = " ", x = grid_data$label)
  grid_data$label <- stringr::str_wrap(grid_data$label, 8)  # wrap labels  
  edges <- data.frame(from = indexCluster, to = grid_data$subset[grid_data$subset != indexCluster] )  # new df with directional proportions
  nodes <- data.frame(name = grid_data$subset, label = grid_data$label, size = grid_data$mean_proportion) # map to node size
  graph <- tidygraph::tbl_graph(nodes = nodes, edges = edges, directed = TRUE)  # create an igraph object with nodes and edges 
  print( 
      a <- ggraph(graph, layout = 'dendrogram', circular = TRUE) +
      geom_edge_fan(width = 0.5, color = "gray60", alpha = 0.8) +
      geom_node_point(aes(fill = name, size = size), shape = 21, color = "black", stroke = 0.2) +
      geom_node_text(aes(label = label, x = x * 1.3, y = y * 1.3), repel = F, size = 2, 
                     hjust='outward', vjust='outward') +
      scale_size_continuous( name = "Interaction\nProportion", limits = c(0, 0.5), range = c(0.1, 15), 
                             breaks=c(0, 0.01, 0.05,0.1,0.2,0.4) ) + 
      scale_fill_manual(values = stableColorsL2.fxn, guide = "none") +
      geom_node_label( aes(label = label, filter = name == indexCluster), size = 2, alpha=1, fontface = 'bold') +
      theme_graph(base_family = 'sans') +
      labs(title = title) +
          xlim(c(-3, 3)) + 
      ylim(c(-3, 3)) +
      coord_fixed()
  )
  return(a)
}
plot.Average.Niche.igraph(annot.table = annot.k9, indexCluster = 'LZ_BCELL', title='Light zone B cells')
 # ggsave("output/LZBCELL.igraph.allSamples.pdf", height=5,width=5)
plot.Average.Niche.igraph(annot.table = annot.k9, indexCluster = 'PC', title='Plasma cells')

# plot.Average.Niche.squares(annot.table = annot.k9, indexCluster = 'LZ_BCELL', title='LZ_BCELL')

a <- plot.Average.Niche.igraph(annot.table = annot.k9, indexCluster = 'TCF1_HI_TOX_HI_CD8', title='TCF1_HI_TOX_HI_CD8')
b <- plot.Average.Niche.igraph(annot.table = annot.k9, indexCluster = 'TCF1_LO_TOX_HI_CD8', title='TCF1_LO_TOX_HI_CD8')

```


```{r plot-average-niche-comparisons, fig.width=10, fig.height=5}
gridExtra::grid.arrange(a,b, nrow=1) 
```



# locate cells that exist in a given neighborhood 
```{r finding-example-cells-for-neighborhood}
defineNeighborhoods <- function( annot.list)
{
  list_ID <- names(annot.list)
  res <- list()
  copymatrix <- list()
  for (i in list_ID)
  { 
    #                    from                   to sdimx_begin sdimy_begin sdimx_end sdimy_end distance     weight
    #                  <char>               <char>       <num>       <num>     <num>     <num>    <num>      <num>
    # 1: ctrl.1-ctrl.1_100000  ctrl.1-ctrl.1_99994      9482.0      9320.0    9482.5    9305.0 15.00833 0.06246747
    # 2: ctrl.1-ctrl.1_100000 ctrl.1-ctrl.1_100002      9482.0      9320.0    9500.0    9323.5 18.33712 0.05171401
    # 3: ctrl.1-ctrl.1_100000 ctrl.1-ctrl.1_100004      9482.0      9320.0    9462.5    9333.0 23.43608 0.04092309
    # 
    #                                      to_cell_type                           from_cell_type type_int
    #                                      <char>                                   <char>   <char>
    # 1:                    TCF1_MIXED_TOX_HI_CD8                       TCF1_HI_TOX_HI_CD8   hetero
    # 2:                           CD4_CD20_MIXED                       TCF1_HI_TOX_HI_CD8   hetero
    # 3:                             CD8_PC_MIXED                       TCF1_HI_TOX_HI_CD8   hetero
    # 
    #                                                    from_to                                              unified_int
    #                                                     <char>                                                   <char>
    # 1:                TCF1_HI_TOX_HI_CD8-TCF1_MIXED_TOX_HI_CD8                TCF1_HI_TOX_HI_CD8--TCF1_MIXED_TOX_HI_CD8
    # 2:                       TCF1_HI_TOX_HI_CD8-CD4_CD20_MIXED                       CD4_CD20_MIXED--TCF1_HI_TOX_HI_CD8
    # 3:                         TCF1_HI_TOX_HI_CD8-CD8_PC_MIXED                         CD8_PC_MIXED--TCF1_HI_TOX_HI_CD8
    
    print(paste0("pass 1: ", i))  # because the pairings are NOT DIRECTIONAL, I need to swap from/to and rbind it in
    a <- annot.list[[i]][, c(1:2,9:10)]  # simplify to just the necessary columns 
    b <- copy(a); names(b) <- c("to","from","from_cell_type", "to_cell_type")  # swap order of column names 
    a <- rbind(a, b)
    neighborhood.matrix <- a[, .N, by = .(from, to_cell_type, from_cell_type)] %>% 
      dcast(from + from_cell_type ~ to_cell_type, value.var= 'N', fill=0)  # count the cell types and make a wide data frame 
    neighborhood.matrix[, c('ARTIFACT', 'CD4.outlier', 'CD8.outlier'):= NULL]
    copymatrix[[i]] <- neighborhood.matrix
    
   #                    from                           from_cell_type ACTIVATED_CD4 ACTIVATED_CD8 ACTIVATED_MEMORY_CD4 
   #                  <char>                                   <char>         <int>         <int>                <int>              
   # 1: ctrl.1-ctrl.1_100000                       TCF1_HI_TOX_HI_CD8             0             0                    0                  
   # 2: ctrl.1-ctrl.1_100001                       TCF1_HI_TOX_HI_CD8             0             0                    0                  
   # 3: ctrl.1-ctrl.1_100002                           CD4_CD20_MIXED             0             0                    0  
   # 
   #    ACTIVATED_TREG ARTIFACT BCELL_MYELOID_MIXED BCELL_MYELOID_NONIMMUNE_MIXED CD4.outlier CD4_CD20_MIXED CD4_CD8_CD20_MIXED
   #             <int>    <int>               <int>                         <int>       <int>          <int>              <int>
   # 1:              0        0                   0                             1           0              1                  2
   # 2:              0        0                   0                             0           0              1                  0
   # 3:              0        0                   0                             1           0              0                  2
  }
  for (i in list_ID)
  { 
    print(paste0("pass 2: ", i))
    working.dt <- copy(copymatrix[[i]])
    id_cols <- c("from", "from_cell_type")
    count_cols <- setdiff(names(working.dt), id_cols)
    
    # 2. Calculate the total for each row and add it as a new column
    working.dt[, row_total := rowSums(.SD), .SDcols = count_cols]
    
    # 3. Divide each count column by the row_total to get the proportion
    #    We loop through the count columns and update them "by reference"
    working.dt[, (count_cols) := lapply(.SD, function(x) x / row_total), .SDcols = count_cols]
    res[[i]] <- working.dt
  } 
  return( list(prop.table=res, raw.table=copymatrix, list_ID=list_ID)) 
}

neighborhoods <- defineNeighborhoods(annot.list = annot.k9)
gc()

exampleNeighborhoods.filter <- function(res.list, indexCluster, chooseNeighbor, threshold)
{  
  filterList <- list()
  list_ID <- res.list[['list_ID']]
  res.list <- res.list[['prop.table']]
  for (i in list_ID)    { filterList[[i]] <- res.list[[i]][from_cell_type == indexCluster & get(chooseNeighbor) > threshold]    }
  return( rbindlist(filterList)  )
}

a <- exampleNeighborhoods.filter(neighborhoods, indexCluster = "LZ_BCELL", chooseNeighbor = "LZ_BCELL", threshold=0.2)
# write.csv(a, file = "LZ_BCELL.filtered.by.LZ_BCELL_over0.2n.csv")
a <- exampleNeighborhoods.filter(neighborhoods, indexCluster = "LZ_BCELL", chooseNeighbor = "PC", threshold=0.2)
# write.csv(a, file = "LZ_BCELL.filtered.by.PC_over0.2.csv")
a <- exampleNeighborhoods.filter(neighborhoods, indexCluster = "TCF1_HI_TOX_HI_CD8", chooseNeighbor = "ACTIVATED_TREG", threshold=0.15)
# write.csv(a, file = "TCF1_HI_TOX_HI_CD8.filtered.by.ACTIVATED_TREG_over0.15.csv")
a <- exampleNeighborhoods.filter(neighborhoods, indexCluster = "ACTIVATED_TREG", chooseNeighbor = "TCF1_HI_TOX_HI_CD8", threshold=-0.1)
# write.csv(a, file = "ACTIVATED_TREG.unfiltered.csv")

# example output
#                      from     from_cell_type ACTIVATED_CD4 ACTIVATED_CD8 ACTIVATED_MEMORY_CD4 ACTIVATED_MEMORY_CD8 ACTIVATED_TREG
#                    <char>             <char>         <num>         <num>                <num>                <num>          <num>
# 1:   ctrl.1-ctrl.1_101134 TCF1_HI_TOX_HI_CD8          0.00          0.00                 0.25                    0      0.2500000
# 2:   ctrl.1-ctrl.1_132837 TCF1_HI_TOX_HI_CD8          0.00          0.00                 0.00                    0      0.1666667
# 3:   ctrl.1-ctrl.1_132840 TCF1_HI_TOX_HI_CD8          0.00          0.00                 0.00                    0      0.1250000
# 4:   ctrl.1-ctrl.1_157997 TCF1_HI_TOX_HI_CD8          0.00          0.00                 0.20                    0      0.2000000
# 5:   ctrl.1-ctrl.1_169239 TCF1_HI_TOX_HI_CD8          0.00          0.00                 0.00                    0      0.2000000

tmp <- purrr::list_rbind(neighborhoods$prop.table, names_to = "list_ID") 
ggplot(tmp, aes(x = row_total)) + geom_histogram(binwidth=0.25, position='identity', fill='grey70', color='black') + 
  theme_classic() + ggtitle("All cells") + xlab("Number of neighbors")
# ggsave(filename = 'output/hist.numbNeighbors.k9.pdf', width=2,height=3)
rm(tmp)
```

# plot average niche by stage for CD8 subsets 
```{r plot-average-niche-by-stage, fig.width=10, fig.height=5}
a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 
a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                      title='TCF1_LO_TOX_HI_CD8 stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                      title='TCF1_LO_TOX_HI_CD8 stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 
```

# plot average niche by recur
```{r plot-average-niche-by-recur, fig.width=10, fig.height=5}
a <- plot.Average.Niche.igraph(annot.table = annot.k9[recur.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 recur')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[norecur.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 no recur')
gridExtra::grid.arrange(a,b, nrow=1) 
a <- plot.Average.Niche.igraph(annot.table = annot.k9[recur.subgroup], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                      title='TCF1_LO_TOX_HI_CD8 recur')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[norecur.subgroup], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                      title='TCF1_LO_TOX_HI_CD8 no recur')
gridExtra::grid.arrange(a,b, nrow=1) 
```


# contrast stage 1 vs stage 3 for niche  
```{r contrast-niche-by-stage, plot.height=10}

rgb_to_hex <- function(rgb) {
  sprintf("#%02X%02X%02X", rgb[1], rgb[2], rgb[3])
}

lighten_hex <- function(hex, factor = 0.3) { # You can adjust the factor
  rgb <- col2rgb(hex)
  new_rgb <- round(rgb + (255 - rgb) * factor)
  new_rgb <- pmax(0, pmin(255, new_rgb))
  return(rgb_to_hex(new_rgb))
}

niche.contrast <- function( annot.table, indexCluster, common.title, grouping.exp, grouping.ctrl, condition.exp, condition.ctrl)
{
  ctrl <- showNiche.multiSample(annot.table[grouping.ctrl], indexCluster, title = paste0(common.title,".",condition.ctrl),
                                      plots=F)
  expt <- showNiche.multiSample(annot.table[grouping.exp], indexCluster, title = paste0(common.title,".",condition.exp),
                                      plots=F)
  ctrl$stage <- condition.ctrl;  expt$stage <- condition.exp
  res <- rbind(ctrl, expt)
  res$stage <- factor(res$stage, levels = c(condition.exp, condition.ctrl))
  res$subset <- factor(res$subset, levels = levels(expt$subset))
  stableColorsL2.fxn <- stableColorsL2

  # Create a lighter version of the color palette
  lighter_stableColorsL2.fxn <- sapply(stableColorsL2.fxn, lighten_hex, factor = 0.6) # Adjust factor as needed

  # Create a combined color mapping key
  res$color_key <- interaction(res$subset, res$stage, sep = "_")

  # Create the combined color palette
  combined_palette <- c(stableColorsL2.fxn, lighter_stableColorsL2.fxn[names(stableColorsL2.fxn)])
  names(combined_palette) <- c(paste0(names(stableColorsL2.fxn), "_", condition.exp), 
                               paste0(names(lighter_stableColorsL2.fxn), "_", condition.ctrl))
  
  annotateX <- max(res$proportion)-0.2*max(res$proportion)
  print(a <- ggplot(res, aes(y = subset, x = proportion, fill=color_key, label=list_ID, group=stage)) +
      geom_point(size=2, position=position_dodge(width=1), pch=21, stroke=0.2, color='grey50') +
      annotate('text',x= annotateX ,y=35, label=paste0(condition.ctrl," in light shade\n", 
                                                       condition.exp," in dark shade"), size=3)+
      stat_summary( fun.min = "mean", fun.max = "mean", geom = "errorbar", position = position_dodge(width=1),
                    width = 1, color = "black", linewidth = 0.5 ) +
      theme_bw() + ggtitle(common.title) +
      scale_fill_manual(name = 'subset', values= combined_palette,  guide = guide_legend(ncol = 2))  +
      theme(axis.text = element_text(size=6), axis.title.y = element_blank()) +
      theme(legend.position = 'none')
  )
  return(list(res=res, scatterplot = a, colorScheme = combined_palette))
}
niche.contrast(annot.table = annot.k9, indexCluster = 'LZ_BCELL', common.title='LZ_BCELL', grouping.exp = stage3.subgroup, 
               grouping.ctrl = stage1.subgroup, condition.exp = "stage3", condition.ctrl = 'stage1')
# niche.contrast( annot.k9, indexCluster = 'TCF1_HI_TOX_HI_CD8', common.title='TCF1_HI_TOX_HI_CD8', grouping.exp = stage3.subgroup,
#                grouping.ctrl = stage1.subgroup, condition.exp = "stage3", condition.ctrl = 'stage1')
# niche.contrast( annot.k9, indexCluster = 'TCF1_LO_TOX_HI_CD8', common.title='TCF1_LO_TOX_HI_CD8', grouping.exp = stage3.subgroup,
#                grouping.ctrl = stage1.subgroup, condition.exp = "stage3", condition.ctrl = 'stage1')

```

# contrast stage using cellProxEnrich
```{r contrast-stage-using-cellProxEnrich, fig.width=10, fig.height=5}
findImportantInteractions <- function( interaction.dt, condition.exp, condition.ctrl, ef.threshold, leaveOutTumor = T)
{
  working.dt <- copy(interaction.dt)
# example output
#                                                  unified_int type_int original simulations   enrichm p_higher_orig p_lower_orig
#                                                         <fctr>   <char>    <num>       <num>     <num>         <num>        <num>
# 1:                                                Tumor--Tumor     homo     5976     110.155  5.748777             0            1
# 2:    CD8TOXHI_TCF1_HI_TOX_HI_CD8--CD8TOXHI_TCF1_HI_TOX_HI_CD8     homo       61       0.500  5.369234             0            1
# 3:              TBMIXED_MIXED_GC_TCELL--TBMIXED_MIXED_GC_TCELL     homo     6627     227.565  4.857897             0            1
#        p.adj_higher p.adj_lower  PI_value int_ranking list_ID condition
#           <num>       <num>     <num>       <int>  <char>    <char>
# 1:            0           1  13.22811           1 tumor.1    stage3
# 2:            0           1  12.35477           2 tumor.1    stage3
# 3:            0           1  11.17817           3 tumor.1    stage3

# per discussion with Rami Vanguri (6/24), we should consider clipping the enrichm scores at 0. these scores represent deviations away from random but at some level, you cannot be more random than random. the negative enrichm values imply loss of interactions but maybe that has to be thought of distinctly from gain of interactions with respect to the baseline of random interaction 
  working.dt[, enrichm := ifelse(enrichm > 0, yes = enrichm, no=0)]
  delta.pairwise.enrichment<- working.dt %>% 
    group_by(unified_int, condition) %>% 
    summarise(mean = mean(enrichm, na.rm = TRUE), 
              sd = sd(enrichm, na.rm = TRUE), 
              n_replicates = n()) %>%  ungroup()
# example output
# A tibble: 1,406 × 5
 #   unified_int                                              condition  mean    sd n_replicates
 #   <fct>                                                    <chr>     <dbl> <dbl>        <int>
 # 1 BCELLS_MANTLE_ZONE_BCELL--TBMIXED_MIXED_GC_TCELL         stage1        0     0            5
 # 2 BCELLS_MANTLE_ZONE_BCELL--TBMIXED_MIXED_GC_TCELL         stage3        0     0           10
 # 3 BCELLS_MIXED_FOLLICULAR_BCELL--Tumor                     stage1        0     0            5
 # 4 BCELLS_MIXED_FOLLICULAR_BCELL--Tumor                     stage3        0     0           10
 # 5 BCELLS_MANTLE_ZONE_BCELL--Tumor                          stage1        0     0            5
 # 6 BCELLS_MANTLE_ZONE_BCELL--Tumor                          stage3        0     0           10


# For each group, subtract the 'mean' value of stage1 from stage3. Use `mean[condition == '...']` to select the correct value.
  mean.diff.enrichm <- delta.pairwise.enrichment %>%
    group_by(unified_int) %>%  
    summarise( mean.diff = mean[condition == condition.exp] - mean[condition == (condition.ctrl)], 
               effect.size = (mean[condition == (condition.exp)] - mean[condition == (condition.ctrl)]) / sqrt(0.5*(sd[condition==(condition.exp)]^2 + sd[condition==(condition.ctrl)]^2))) %>% 
    arrange(desc(effect.size))     
  print("mean.diff.enrichm")
  print(as_tibble(mean.diff.enrichm))
# example output 
# A tibble: 703 × 3
 #   unified_int                                                            mean.diff       effect.size
 #   <fct>                                                                      <dbl>             <dbl>
 # 1 CD4FOXP3HI_ACTIVATED_TREG--CD8TOXLO_MIXED_TOX_LO_CD8                       0.503              1.61
 # 2 CD4.outlier--CD4FOXP3HI_MIXED_TREG                                         0.885              1.51
 # 3 CD8TOXHI_TCF1_MIXED_TOX_HI_CD8--CD8TOXLO_MIXED_TOX_LO_CD8                  0.499              1.47
 # 4 BCELLS_PC--TBMIXED_BCELL_MYELOID_MIXED                                     0.449              1.47
 # 5 CD8TOXHI_TCF1_MIXED_TOX_HI_CD8--CD8TOXLO_TERMINAL_EFFECTOR_CD8             0.810              1.33
 
  a <- ggplot(mean.diff.enrichm, aes(x = 1:nrow(mean.diff.enrichm), y=effect.size, label=unified_int)) +
      geom_point(size=2, pch=21, color="white",  alpha=0.5, fill="black") +
      ggrepel::geom_text_repel(size=2.5, max.overlaps = 10, segment.color = 'gray60', force=5) +
      ggtitle(paste0("effect size: ", condition.exp, " - ", condition.ctrl))  + theme_classic() +
      theme(axis.text.x = element_blank())
  
  df <- as.data.table(mean.diff.enrichm)[ abs(effect.size) > ef.threshold]
  df <- as.data.table(df)[ !unified_int %like% 'outlier' & !unified_int %like% 'ARTIFACT'] 
  if(leaveOutTumor == T) { df <- as.data.table(df)[!unified_int %like% 'Tumor'] }  # leave out tumor interactions 
  df$unified_int <- factor(df$unified_int, levels = df$unified_int)
  print("effect size table, filtered for effect size threshold")
  print(as_tibble(df))
# example output
# # A tibble: 30 × 3
#    unified_int                                        mean.diff       effect.size
#    <fct>                                                  <dbl>             <dbl>
#  1 ACTIVATED_TREG--MIXED_TOX_LO_CD8                       0.503              1.61
#  2 MIXED_TOX_LO_CD8--TCF1_MIXED_TOX_HI_CD8                0.499              1.47
#  3 TCF1_MIXED_TOX_HI_CD8--TERMINAL_EFFECTOR_CD8           0.810              1.33
#  4 CD4_MYELOID_NONIMMUNE_MIXED--TCF1_MIXED_TOX_HI_CD8     0.221              1.31
#  5 GZMB_CD8--NONIMMUNE_CELL                               0.179              1.18
#  6 ACTIVATED_MEMORY_CD4--MIXED_TOX_LO_CD8                 0.441              1.18
#  ...
# 26 MEMORY_CD8--MIXED_TREG                               -0.505              -1.54
# 27 ARTIFACT--CD4_MYELOID_NONIMMUNE_MIXED                -0.916              -1.60
# 28 ARTIFACT--MIXED_TOX_LO_CD8                           -0.530              -1.67
# 29 CD8_PC_MIXED--GZMB_CD8                               -0.497              -1.68
# 30 ACTIVATED_TREG--MEMORY_CD8                           -0.778              -2.08
  
  # annotateX.exp <- max(df$effect.size)-0.2*max(df$effect.size)
  # annotateX.ctrl <- min(df$effect.size)+0.2*min(df$effect.size)

  # b <- ggplot(df, aes(x = effect.size, y = unified_int))  +   
  #   geom_bar(stat='identity', width=0.05) +  geom_point(size=4, pch=21, color='white', fill='black', stroke=0.2) + 
  #         theme_classic() + ggtitle( paste0("Effect size (filtered): ", condition.exp, " - ", condition.ctrl))  +       
  #   theme(axis.text.y = element_text( size=8), axis.title.y = element_blank(), plot.title=element_text(size=10), 
  #         axis.title.x = element_text(size=10), axis.text.x = element_text(size=10)) + 
  #   theme(legend.position = 'none') + annotate("text", x = annotateX.exp, y = -2.2, size = 5, label = condition.exp) + 
  #   annotate("text", x = annotateX.ctrl, y = -2.2, size = 5, label = condition.ctrl) + xlab("Effect size") + 
  #   coord_cartesian(clip = 'off', ylim= c(0, nrow(df)))
  b <- findImportantInteractions.plotFiltered(df, condition.exp, condition.ctrl)
  return(list(df=df, fullggplot=a, filteredggplot=b)) 
}

findImportantInteractions.plotFiltered <- function(df, condition.exp, condition.ctrl, labelPlot =T)
{
  annotateX.exp <- max(df$effect.size)-0.2*max(df$effect.size)
  annotateX.ctrl <- min(df$effect.size)+0.2*min(df$effect.size)

  b <- ggplot(df, aes(x = effect.size, y = unified_int))  +   
  geom_bar(stat='identity', width=0.05, color='grey70', fill='grey50') +  
  geom_point(size=4, pch=21, color='white', fill='grey50', stroke=0.2) + 
        theme_classic() + 
  ggtitle( paste0("Effect size (filtered): ", condition.exp, " - ", condition.ctrl))  +       
  theme(axis.text.y = element_text( size=8, color='black'), axis.title.y = element_blank(), 
        plot.title=element_text(size=10), 
        axis.title.x = element_text(size=10, color='black'), axis.text.x = element_text(size=10, color='black')) + 
  theme(legend.position = 'none')  
      
  if(labelPlot == T)
  (
    b <- ggplot(df, aes(x = effect.size, y = unified_int))  +   
      geom_bar(stat='identity', width=0.05, color='grey70', fill='grey50') +  
      geom_point(size=4, pch=21, color='white', fill='grey50', stroke=0.2) + 
            theme_classic() + 
      ggtitle( paste0("Effect size (filtered): ", condition.exp, " - ", condition.ctrl))  +       
      theme(axis.text.y = element_text( size=8, color='black'), axis.title.y = element_blank(), 
            plot.title=element_text(size=10), 
            axis.title.x = element_text(size=10, color='black'), axis.text.x = element_text(size=10, color='black')) + 
      theme(legend.position = 'none') + annotate("text", x = annotateX.exp, y = -2.2, size = 5, label = condition.exp) + 
      annotate("text", x = annotateX.ctrl, y = -2.2, size = 5, label = condition.ctrl) + xlab("Effect size")  +
      coord_cartesian(clip = 'off', ylim= c(0, nrow(df)))
  )
  return(b)
}

a <- findImportantInteractions( interaction.dt= interaction.byStage, condition.exp="stage3", condition.ctrl="stage1", ef.threshold=1,
                                leaveOutTumor = T)
# ggsave(plot = a$filteredggplot, file = "output/interaction.Stage.EffectSize.pdf", height=5, width=6)

a$filteredggplot.selective <- findImportantInteractions.plotFiltered(df = a$df[unified_int %like% "ACTIVATED_TREG"], 
                                            condition.exp="stage3", condition.ctrl="stage1", labelPlot = F)
# ggsave(plot = a$filteredggplot.selective, file = "output/interaction.Stage.EffectSize.ACTIV-TREG.pdf", height=1.75, width=4)


globalChanges.Effectsize.plot <- function(importIntxn.obj, colNames = names(stableColorsL2), title = "Stage 3 vs Stage 1")
{
  importIntxn.obj <- as.data.table(importIntxn.obj$fullggplot$data)
  res_list <- lapply(X = colNames, 
                   FUN = function(x) { importIntxn.obj[unified_int %like% x, effect.size]} )
  names(res_list) <- colNames
  res_list <- lapply(res_list, as.data.table)
  b <- purrr::list_rbind(res_list, names_to = 'subset')
  names(b)[2] <- "Effect.size"
  b <- b[!is.nan('Effect.size')]; b <- b[!subset %like% 'ARTIFACT']; b<- b[!subset %like% 'outlier']; b <- b[!subset %like% 'Tumor']
  b %>% rstatix::dunn_test(formula = Effect.size ~ subset) 
  order <- b %>% group_by(subset) %>% rstatix::get_summary_stats(type = 'mean') %>% arrange(mean)
  b$subset <- factor(b$subset, levels = order$subset) 
  a <- ggplot(b, aes(y = subset, x=Effect.size)) + geom_vline(xintercept = 0, color='grey50')+ 
    geom_violin() +# geom_point(pch=21, color='grey50',fill='grey90', size=0.5) +  
    theme_bw() + theme(axis.title.y = element_blank()) + theme(axis.text = element_text(color='black')) + 
    ggtitle(title)
  return(a)
}  

b <- globalChanges.Effectsize.plot(importIntxn.obj = a, title = "Stage 3 vs Stage 1")
# ggsave(b, filename = 'output/EffectSize.allSubsets.byStage.violin.pdf', height=6,width=6)



a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'ACTIVATED_TREG',
                      title='ACTIVATED_TREG stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'ACTIVATED_TREG',
                      title='ACTIVATED_TREG stage 3')
gridExtra::grid.arrange(a, b, nrow=1) 
# ggsave(a, file = "output/ACTIVATED_TREG.igraph.s1.pdf", height=5,width=5)
# ggsave(b, file = "output/ACTIVATED_TREG.igraph.s3.pdf", height=5,width=5)

a <- niche.contrast( annot.k9, indexCluster = 'ACTIVATED_TREG', common.title='ACTIVATED_TREG' , grouping.exp = stage3.subgroup, 
               grouping.ctrl = stage1.subgroup, condition.exp = "stage3", condition.ctrl = 'stage1')
# ggsave(a[['scatterplot']], file = "output/ACTIVATED_TREG.scatterplot.s1-v-s3.pdf", height=6,width=6)

a[['res']][subset == 'MIXED_TOX_LO_CD8'] %>% group_by(stage) %>% rstatix::get_summary_stats(show=c("n", "mean", "sd", "median"))
a[['res']][subset == 'MIXED_TOX_LO_CD8'] %>% rstatix::wilcox_test(formula = proportion ~ stage)
a[['res']][subset == 'MIXED_TOX_LO_CD8'] %>% group_by(stage) %>% rstatix::shapiro_test(proportion)
a[['res']][subset == 'MIXED_TOX_LO_CD8'] %>% rstatix::welch_anova_test(formula = proportion ~ stage)

plot.boxplot <- function(df, probeSubset, grouping.var, plot.var='proportion', nonparam = T, colorList, title)
{
  df <- df[subset == probeSubset]
  # pVal <- df %>% rstatix::welch_anova_test(formula = get(plot.var) ~ get(grouping.var)) 
  formula.str <- as.formula(paste(plot.var, "~", grouping.var))
  if(nonparam==F)   { pVal <- df %>% rstatix::welch_anova_test(formula = formula.str) }
  if(nonparam==T)   { pVal <- df %>% rstatix::wilcox_test(formula = formula.str)  }
  colorList <- colorList[names(colorList) %like% probeSubset]
  ggplot(df, aes(x = get(grouping.var), y=get(plot.var))) + geom_boxplot(aes(fill = color_key)) + 
    geom_point(size=4, pch=21, fill='white', color='grey50',) + theme_bw() + theme(axis.text = element_text(color='black')) + 
    scale_fill_manual(values = colorList)  + theme(legend.position = 'null', axis.title.x = element_blank()) + 
    annotate('text', x=2, y=max(df$proportion), label = paste("P=", round(pVal$p, 2)), color='black')+
    ylab(paste(probeSubset, plot.var)) + ggtitle(title) 
}
plot.boxplot( df = a$res, probeSubset = 'MIXED_TOX_LO_CD8', grouping.var = 'stage', plot.var='proportion', colorList = a$colorScheme,
              title = "ACTIVATED_TREG")
# ggsave(filename = 'output/ACTIVATED_TREG-for-MIXED_TOX_LO_CD8.stage.boxplot.pdf', width=3,height=3)
plot.boxplot( df = a$res, probeSubset = 'MEMORY_CD8', grouping.var = 'stage', plot.var='proportion', colorList = a$colorScheme,
              title = "ACTIVATED_TREG")
# ggsave(filename = 'output/ACTIVATED_TREG-for-MEMORY_CD8.stage.boxplot.pdf', width=3,height=3)


plot.enrichm.interaction <- function(interaction.dt, probeSubset, nonparam=T, plot.var='enrichm', colorList, title)
{
  interaction.dt <- interaction.dt[unified_int == probeSubset]
  formula.str <- as.formula(paste(plot.var, "~", 'condition'))
  if(nonparam==F)   { pVal <- interaction.dt %>% rstatix::welch_anova_test(formula = formula.str) }
  if(nonparam==T)   { pVal <- interaction.dt %>% rstatix::wilcox_test(formula = formula.str)   }
  colorList <- colorList[names(colorList) %like% probeSubset]
  ggplot(interaction.dt, aes(x = condition, y=get(plot.var))) + geom_boxplot(fill = 'grey80', alpha=0.5) + 
    geom_point(size=4, pch=21, fill='white', color='grey30',) + theme_bw() + theme(axis.text = element_text(color='black')) + 
    scale_fill_manual(values = colorList)  + theme(legend.position = 'null', axis.title.x = element_blank()) + 
    annotate('text', x=1, y=1.2*max(interaction.dt[,.(get(plot.var))]), 
             label = paste("P=", round(pVal$p, digits=2)), color='black')+
    ggtitle(title) + ylab("Log2FC enrichment relative to random") #+   scale_y_continuous(expand = expansion(mult = 0.1))
}

plot.enrichm.interaction(interaction.dt = interaction.byStage, probeSubset = "ACTIVATED_TREG--MIXED_TOX_LO_CD8", 
                         colorList = a$colorScheme, title="ACTIVATED_TREG--\nMIXED_TOX_LO_CD8")
# ggsave(filename = 'output/ACTIVATED_TREG--MIXED_TOX_LO_CD8.stage.boxplot.pdf', width=2.5,height=3)
plot.enrichm.interaction(interaction.dt = interaction.byStage, probeSubset = "ACTIVATED_TREG--MEMORY_CD8", 
                         colorList = a$colorScheme, title="ACTIVATED_TREG--\nMEMORY_CD8")
# ggsave(filename = 'output/ACTIVATED_TREG--MEMORY_CD8.stage.boxplot.pdf', width=2.5,height=3)

```


# contrast recurrence  
```{r contrast-recur-using-cellProxEnrich, fig.width=8, fig.height=8}
a <- findImportantInteractions( interaction.dt= interaction.byRecur, condition.exp="recur", condition.ctrl="no.recur", ef.threshold=1)
# ggsave(plot = a$filteredggplot, file = "output/interaction.Recur.EffectSize.pdf", height=11, width=8)
a$filteredggplot
# a$df[, colorMe:=ifelse(unified_int %like% "ACTIVATED_TREG", "T","F")]
a$filteredggplot.selective <- findImportantInteractions.plotFiltered(df = a$df[unified_int %like% "ACTIVATED_TREG"], 
                                            condition.exp="recur", condition.ctrl="no.recur", labelPlot = F)
a$filteredggplot.selective
# ggsave(plot = a$filteredggplot.selective, file = "output/interaction.Recur.EffectSize.justTREG.pdf", height=2, width=4)

b <- globalChanges.Effectsize.plot(importIntxn.obj = a, title = "Recur vs NoRecur")
# ggsave(b, filename = 'output/EffectSize.allSubsets.byRecur.violin.pdf', height=6,width=6)
b


a <- plot.Average.Niche.igraph(annot.table = annot.k9[norecur.subgroup], indexCluster = 'ACTIVATED_TREG',
                      title='ACTIVATED_TREG no recurrence')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[recur.subgroup], indexCluster = 'ACTIVATED_TREG',
                      title='ACTIVATED_TREG recurrence')
gridExtra::grid.arrange(a, b, nrow=1)
# ggsave(a, file = "output/ACTIVATED_TREG.igraph.norecur.pdf", height=5,width=5)
# ggsave(b, file = "output/ACTIVATED_TREG.igraph.recur.pdf", height=5,width=5)

a <- plot.Average.Niche.igraph(annot.table = annot.k9[norecur.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 no recurrence')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[recur.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 recurrence')
gridExtra::grid.arrange(a, b, nrow=1) 
# ggsave(a, file = "output/TCF1_HI_TOX_HI_CD8.igraph.norecur.pdf", height=5,width=5)
# ggsave(b, file = "output/TCF1_HI_TOX_HI_CD8.igraph.recur.pdf", height=5,width=5)


a <- niche.contrast(annot.table = annot.k9, indexCluster = 'ACTIVATED_TREG', common.title='ACTIVATED_TREG', grouping.exp = recur.subgroup, 
               grouping.ctrl = norecur.subgroup, condition.exp = 'recur', condition.ctrl = 'no.recur')
ggsave(a[['scatterplot']], file = "output/ACTIVATED_TREG.scatterplot.recur-v-norecur.pdf", height=6,width=6)

b <- niche.contrast(annot.table = annot.k9, indexCluster = 'TCF1_HI_TOX_HI_CD8', common.title='TCF1_HI_TOX_HI_CD8', 
                    grouping.exp = recur.subgroup, grouping.ctrl = norecur.subgroup, condition.exp = 'recur', condition.ctrl = 'no.recur')
a[['res']][subset == 'TCF1_HI_TOX_HI_CD8'] %>% group_by(stage) %>% rstatix::get_summary_stats(show=c("n", "mean", "sd", "median"))
a[['res']][subset == 'TCF1_HI_TOX_HI_CD8'] %>% group_by(stage) %>% rstatix::shapiro_test(proportion)
a[['res']][subset == 'TCF1_HI_TOX_HI_CD8'] %>% rstatix::wilcox_test(formula = proportion ~ stage)
b[['res']][subset == 'ACTIVATED_TREG'] %>% group_by(stage) %>% rstatix::get_summary_stats(show=c("n", "mean", "sd", "median"))
# b[['res']][subset == 'ACTIVATED_TREG'] %>% group_by(stage) %>% rstatix::shapiro_test(proportion)
b[['res']][subset == 'ACTIVATED_TREG'] %>% rstatix::wilcox_test(formula = proportion ~ stage)

plot.boxplot( df = a$res, probeSubset = 'TCF1_HI_TOX_HI_CD8', grouping.var = 'stage', plot.var='proportion', colorList = a$colorScheme,
              title = "ACTIVATED_TREG")
# ggsave(filename = 'output/ACTIVATED_TREG-for-TCF1_HI_TOX_HI_CD8.recur.boxplot.pdf', width=3,height=3)
plot.boxplot( df = b$res, probeSubset = 'ACTIVATED_TREG', grouping.var = 'stage', plot.var='proportion', colorList = a$colorScheme,
              title = "TCF1_HI_TOX_HI_CD8")
# ggsave(filename = 'output/TCF1_HI_TOX_HI_CD8-for-ACTIVATED_TREG.recur.boxplot.pdf', width=3,height=3)

plot.enrichm.interaction(interaction.dt = interaction.byRecur, probeSubset = "ACTIVATED_TREG--TCF1_HI_TOX_HI_CD8", colorList = a$colorScheme, title="ACTIVATED_TREG--\nTCF1_HI_TOX_HI_CD8")
# ggsave(filename = 'output/ACTIVATED_TREG--TCF1_HI_TOX_HI_CD8.recur.boxplot.pdf', width=2.5,height=3)

plot.enrichm.interaction(interaction.dt = interaction.byRecur, probeSubset = "ACTIVATED_CD8--ACTIVATED_TREG", colorList = a$colorScheme, title="ACTIVATED_CD8--\nACTIVATED_TREG")
# ggsave(filename = 'output/ACTIVATED_CD8--ACTIVATED_TREG.recur.boxplot.pdf', width=2.5,height=3)
```


```{r determine-LOD, fig.width=8, fig.height=8}
interaction.byRecur[,.(original)] %>% rstatix::get_summary_stats(type = 'common') 

plot.enrichm.interaction(interaction.dt = interaction.byStage, probeSubset = "ACTIVATED_TREG--TCF1_HI_TOX_HI_CD8", colorList = a$colorScheme, title="ACTIVATED_TREG--\nTCF1_HI_TOX_HI_CD8")

# define limit of detection
# what is background rate in stage 1 where no effect is present 

stats <- interaction.byStage[ unified_int == "ACTIVATED_TREG--TCF1_HI_TOX_HI_CD8" & condition == 'stage1' , .(original)] %>% rstatix::get_summary_stats(type = 'common')
print(stats)
print(paste0('2*SD + mean = ', 2*stats$sd + stats$mean) )

# need the negative results, first take all of the non-significant interactions 
# need to find the null distribution for specific A-B interaction pair

dt <- purrr::list_rbind(readRDS(file="neighborhood.annot.k9.sln.RDS")[['tableResults.k9']], names_to = "id")
dt.subs <- dt[unified_int == "ACTIVATED_TREG--TCF1_HI_TOX_HI_CD8"   & 
                id %in% norecur.subgroup   & 
                orig == 'simulations']
dt.subs[, .(V1)] %>% rstatix::get_summary_stats(type = 'quantile',probs = 0.95) 
dt.subs[order(V1, decreasing=T), by=id]  # sort highest to lowest, do it by sample, look for 20th value (given 1/200 sims = 0.005 and 95% percentile is #20 in list from largest to smallest) 
result <- dt.subs %>% group_by(id) %>% rstatix::get_summary_stats(type = 'quantile', probs=0.95)

result$`95%` %>% mean

# so effective limit of detection was 249 pairings 
num.row.norecur <- md[list_ID %in% norecur.subgroup] %>% nrow
print(paste0("LOD: ", result$`95%` %>% mean / num.row.norecur, " interactions per 1e6 cells"))
rm(c(result, num.row.norecur, dt.subs))

```


# evaluate neighborhoods by radius approach
```{r evaluate-neighborhoods-by-radius}

showNiche.singleSample(annot.table = annot.rad[['tumor.5']], indexCluster = 'LZ_BCELL', title='LZ_BCELL.rad',
                            plots=T)
showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'LZ_BCELL', title='LZ_BCELL.k9',
                            plots=T)
niche.contrast(annot.table = annot.rad, indexCluster = 'LZ_BCELL', common.title='LZ_BCELL.rad', grouping.exp = stage3.subgroup, 
               grouping.ctrl = stage1.subgroup, condition.exp = "stage3", condition.ctrl = 'stage1')
niche.contrast(annot.table = annot.k9, indexCluster = 'LZ_BCELL', common.title='LZ_BCELL.k9', grouping.exp = stage3.subgroup, 
               grouping.ctrl = stage1.subgroup, condition.exp = "stage3", condition.ctrl = 'stage1')

a <- findImportantInteractions( interaction.dt= interaction.byStage.rad, condition.exp="stage3", condition.ctrl="stage1", ef.threshold=1)
a <- findImportantInteractions( interaction.dt= interaction.byRecur.rad, 
                                condition.exp="recur", condition.ctrl="no.recur", ef.threshold=0)
b <- findImportantInteractions( interaction.dt= interaction.byRecur, 
                                condition.exp="recur", condition.ctrl="no.recur", ef.threshold=0)
print(a$filteredggplot)
gridExtra::grid.arrange(plot.Average.Niche.igraph(annot.table = annot.rad[norecur.subgroup], indexCluster = 'ACTIVATED_TREG',
                      title='ACTIVATED_TREG no recurrence'),
                      plot.Average.Niche.igraph(annot.table = annot.rad[recur.subgroup], indexCluster = 'ACTIVATED_TREG',
                      title='ACTIVATED_TREG recurrence'), 
                      nrow=1) 
gridExtra::grid.arrange(plot.Average.Niche.igraph(annot.table = annot.rad[norecur.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 no recurrence'),
                      plot.Average.Niche.igraph(annot.table = annot.rad[recur.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 recurrence'), 
                      nrow=1) 

df.a <- a$df; df.b <- b$df;  df.a[, source := 'radius'];  df.b[, source := 'nearest.9']
df.a <- rbindlist(list(df.a, df.b));  df.a$unified_int <- as.character(df.a$unified_int)
df.a <- df.a[!unified_int %like% 'ARTIFACT']
df.a <- df.a %>% select(unified_int, source, effect.size) %>%
  tidyr::pivot_wider(names_from = source, values_from = effect.size  ) %>%
  # Arrange by the average effect size for a cleaner plot
  mutate(avg.effect = (radius + nearest.9) / 2) %>%     arrange(avg.effect) %>%
  mutate(unified_int = factor(unified_int, levels = unified_int))    # Reorder the factor levels 

ggplot(df.a, aes(y = unified_int)) +
  geom_segment(aes(x = radius, xend = nearest.9, y = unified_int, yend = unified_int),
      color = "grey80", linewidth = 0.5 ) + xlab("effect size") + ylab("Pairwise interaction") + 
  geom_point(aes(x = radius), color = "#d6604d", size = 1) +  geom_point(aes(x = nearest.9), color = "#4393c3", size = 1) + 
  annotate( "text", x = 2.1, y = 26, size = 5, label = "Method", fontface = "bold", ) +
  annotate( "point", x = 1.6, y = 14, size = 5, color = "#d6604d" ) +
  annotate( "text", x = 1.9, y = 14, size = 4, label = "Radius", hjust = "left" ) +
  annotate( "point", x = 1.6, y = 6, size = 5, color = "#4393c3" ) +
  annotate( "text", x = 1.9, y = 6, size = 4, label = "Nearest 9", hjust = "left" ) +
  theme_bw() + ggtitle("Radius vs Nearest-9") + 
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.x = element_line(color = "grey90", linetype = "dashed"), 
        axis.text.y = element_blank(), axis.text.x = element_text(size=12), axis.title = element_text(size=12), 
        title = element_text(size=12))
# ggsave(filename = 'output/interaction.radius-vs-nearest9.pdf', height=8, width=4)


DT <- dcast.data.table( data = md, formula = Recurrence + list_ID + SLN_Stage ~ L2_annotations )
DT[, Recurrence:= ifelse(Recurrence == 0, "no", "yes")]
DT$TotalNumber <- rowSums(DT[,4:38])
DT <- melt(DT, id.vars = c("Recurrence","list_ID","TotalNumber", "SLN_Stage"))
DT[, Freq:= value/TotalNumber]
  #    Recurrence list_ID TotalNumber SLN_Stage      variable value         Freq
  #        <char>  <char>       <num>    <char>        <fctr> <int>        <num>
  # 1:         no  ctrl.1      506375    Stage1 ACTIVATED_CD4 26889 0.0531009627
  # 2:         no  ctrl.2      742999    Stage1 ACTIVATED_CD4 39215 0.0527793443
  # 3:         no  ctrl.3      924247    Stage1 ACTIVATED_CD4 34117 0.0369132927

pValue <- DT[variable == "ACTIVATED_TREG"] %>% rstatix::wilcox_test(formula = Freq ~ SLN_Stage) 
ggplot(DT[variable == "ACTIVATED_TREG"], aes(x = SLN_Stage, y=Freq )) + # , fill=list_ID)) + 
  geom_point(size=5, color='grey30', pch=21, stroke = 0.3, fill='grey70') + ggtitle("ACTIVATED_TREG") + 
  # scale_fill_manual(values = getDistinctColors(n = length(unique(DT$list_ID)), seed=3)) + 
  annotate('text', x=1, y=max(DT[variable == "ACTIVATED_TREG", Freq]), label = paste("P=", round(pValue$p, digits=2) ), color='black')+
  ylab("Frequency of all cells") + theme_bw()
# ggsave(filename = "output/SimpleFreq_ACTIVATED_TREG_byStage.pdf" , height=4, width=2)

DT <- DT[SLN_Stage == 'Stage3']
pValue <- DT[variable == "ACTIVATED_TREG"] %>% rstatix::wilcox_test(formula = Freq ~ Recurrence) 
ggplot(DT[variable == "ACTIVATED_TREG"], aes(x = Recurrence, y=Freq )) + # , fill=list_ID)) + 
  geom_point(size=5, color='grey30', pch=21, stroke = 0.3, fill='grey70') + ggtitle("ACTIVATED_TREG") + 
  annotate('text', x=1, y=max(DT[variable == "ACTIVATED_TREG", Freq]), label = paste("P=", round(pValue$p, digits=2) ), color='black')+
  ylab("Frequency of all cells") + theme_bw()
# ggsave(filename = "output/SimpleFreq_ACTIVATED_TREG_byRecur.pdf" , height=4, width=2)
```


# Histograms of feature expression for CD8 T cells given Treg
```{r histogram-of-gzmB-expression-tcflo, plot.width=6}
# neighborhoods <- defineNeighborhoods(annot.list = annot.k9)   # from chunk 15 or so  
# gc()

neighborhoods.grouped <- purrr::list_rbind(neighborhoods$raw.table, names_to = "list_ID")
neighborhoods.grouped[, TREG := ACTIVATED_TREG + ICOS_LO_TREG + MIXED_TREG]
neighborhoods.grouped

TCFlo.CD8 <- neighborhoods.grouped[from_cell_type == "TCF1_LO_TOX_HI_CD8"]
TCFhi.CD8 <- neighborhoods.grouped[from_cell_type == "TCF1_HI_TOX_HI_CD8"]


permutation.test.neighborhoods <- function(num.permutations = 10000, data.tbl, group.0, group.1)
{
  obs.diff <- mean(data.tbl[source %in% group.0, data]) - mean(data.tbl[source %in% group.1, data])  
  print(paste0("Observed difference: ", round(obs.diff, 2)))  
  permutation.diffs <- numeric(num.permutations)  # now for a ridiculous number of permutations 
  for (i in 1:num.permutations) 
  {
    shuffled_source <- sample(data.tbl$source)    # randomly shuffle the group labels.

    # Calculate the difference in means for this random permutation.
    # 1. mutate(): Replaces the original 'source' column with the 'shuffled_source' labels.
    # 2. group_by(): Groups the data by the new, random source labels.
    # 3. summarise(): Calculates the mean for each of the two groups.
    # 4. pull(): Extracts the mean values as a vector.
    # 5. diff(): Calculates the difference between the two means in the vector.
    permutation.diffs[i] <- data.tbl %>%
      mutate(source = shuffled_source) %>%
      group_by(source) %>%
      summarise(mean = mean(data)) %>%
      pull(mean) %>%
      diff()
  }
  hist(permutation.diffs, breaks=100)
  # The p-value is the proportion of permuted differences that were as extreme as (or more extreme than) our observed difference. 
  # The 'abs()' function performs a two-tailed test. The logical operator yields T/F and then mean calculates the proportion 
  pV <- mean(abs(permutation.diffs) >= abs(obs.diff))
  print(paste0("P value is ", pV, ". Lowest possible P value for ",num.permutations, " iterations is ", 1/num.permutations))
  return(pV)
}

neighbors.histogram <- function(target.cell, interacting.type, feature, title, xlab, binwidth=0.25, permutations=10000, 
                                annotate.location = c(12,5000))
{
  target.subset <- target.cell[get(interacting.type) >= 1]  # start with interacting.type neighbors
  val.1 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat[feature, target.subset$from]    # get specific expression data
  target.subset <- target.cell[get(interacting.type) == 0]  # now with no interacting.type neighbors
  val.0 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat[feature, target.subset$from]    # get specific expression data
  dt <- data.table(data = c(val.0, val.1), 
                   source = c(rep("No.Neighbors", length(val.0)), 
                              rep("Neighbors", length(val.1))) )  # make a new dt where the expression values of FEAT can be plotted 
  p.val <- permutation.test.neighborhoods(data.tbl = dt, group.0 = "No.Neighbors", group.1 = "Neighbors", num.perm = permutations)
  dt %>% group_by(source) %>% rstatix::get_summary_stats(data, type = 'common')
  num.NoNeighbors <- dt[source == 'No.Neighbors', .N]
  num.Neighbors <- dt[source == 'Neighbors', .N]
  if(num.NoNeighbors > num.Neighbors)        {   dt$source <- factor(dt$source, levels = c("No.Neighbors","Neighbors"))  } 
    else {     dt$source <- factor(dt$source, levels = c("Neighbors","No.Neighbors"))  }
  plot <- ggplot(dt, aes(x = data, fill = source)) + geom_histogram(binwidth = binwidth, alpha=0.6, position='identity') + 
    scale_fill_manual(values = c('grey50','firebrick')) + 
    annotate('text', x=annotate.location[1], y=annotate.location[2], color='black', 
             label= paste0("Mean Neighbors: ", round(mean(dt[source == 'Neighbors', data]), 1), 
                           "\n",
                           "Mean No.Neighbors: ", round(mean(dt[source == 'No.Neighbors', data]), 1),
                           "\nPermutation test:  P <= ", ifelse(p.val==0, 1/permutations, p.val)),
             hjust = 0, size=3) +
    ggtitle(title) + xlim(c(0,28 ))+ 
    xlab(xlab) + ylab("Number of cells") +
    theme_classic() + theme(legend.justification = c("right", "top"), legend.position = 'inside', 
                            legend.title=element_blank()) 
  return(list(plot, dt, p.val))
}

neighbors.histogram(target.cell = TCFlo.CD8, interacting.type = 'TREG', feature='GZMB', 
                    title = 'GZMB in TCF1_LO_TOX_HI_CD8 due to Treg', xlab='GZMB expression', binwidth=0.25, permutations=10000,
                    annotate.location = c(20,1000))
# ggsave(filename = 'output/hist.GzmB.TCF1loToxhiCD8.pdf', width=4,height=3)

neighbors.histogram(target.cell = TCFlo.CD8, interacting.type = 'TREG', feature='IFNG', 
                    title = 'IFNG in TCF1_LO_TOX_HI_CD8 due to Treg', xlab='IFNg expression', binwidth=0.25, permutations=10000,
                    annotate.location = c(20,1000))
# ggsave(filename = 'output/hist.IFNG.TCF1loToxhiCD8.pdf', width=4,height=3)
```


```{r histogram-of-gzmB-expression-tcfhi, plot.width=6}
neighbors.histogram(target.cell = TCFhi.CD8, interacting.type = 'TREG', feature='GZMB', 
                    title = 'GZMB in TCF1_HI_TOX_HI_CD8 due to Treg', xlab='GZMB expression', binwidth=0.25, permutations=10000,
                    annotate.location = c(20,3000))
# ggsave(filename = 'output/hist.GZMB.TCF1hiToxhiCD8.pdf', width=4,height=3)


neighbors.histogram(target.cell = TCFhi.CD8, interacting.type = 'TREG', feature='IFNG', 
                    title = 'IFNG in TCF1_HI_TOX_HI_CD8 due to Treg', xlab='IFNg expression', binwidth=0.25, permutations=10000,
                    annotate.location = c(20,3000))
# ggsave(filename = 'output/hist.IFNG.TCF1hiToxhiCD8.pdf', width=4,height=3)
```

```{r effects-of-progTex-adjacency-on-treg}
neighborhoods.grouped.TREG <- neighborhoods.grouped[from_cell_type %like% "TREG"]
a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_HI_TOX_HI_CD8', feature='FOXP3', 
                    title = 'FOXP3 in TREG due to TCF1_HI_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                    annotate.location = c(13,3000))
ICF.table <- data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                        "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                        feature = 'FOXP3', interactingType = 'TCF1_HI_TOX_HI_CD8', p.val.permutation = a[[3]]) 
ggsave(a[[1]], filename = 'output/hist.TREG.FOXP3.TCF1hiToxhiCD8.pdf', width=4,height=3)

a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_HI_TOX_HI_CD8', feature='ICOS', 
                         title = 'ICOS in TREG due to TCF1_HI_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                         annotate.location = c(13,3000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'ICOS', interactingType = 'TCF1_HI_TOX_HI_CD8', p.val.permutation = a[[3]]))) 
ggsave(a[[1]], filename = 'output/hist.TREG.ICOS.TCF1hiToxhiCD8.pdf', width=4,height=3)

a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_HI_TOX_HI_CD8', feature='LAG3', 
                         title = 'LAG3 in TREG due to TCF1_HI_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                         annotate.location = c(13,6000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'LAG3', interactingType = 'TCF1_HI_TOX_HI_CD8', p.val.permutation = a[[3]])))
ggsave(a[[1]], filename = 'output/hist.TREG.LAG3.TCF1hiToxhiCD8.pdf', width=4,height=3)

a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_HI_TOX_HI_CD8', feature='CD39', 
                         title = 'CD39 in TREG due to TCF1_HI_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                         annotate.location = c(13,2000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'CD39', interactingType = 'TCF1_HI_TOX_HI_CD8', p.val.permutation = a[[3]])) )
ggsave(a[[1]], filename = 'output/hist.TREG.CD39.TCF1hiToxhiCD8.pdf', width=4,height=3)

a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_HI_TOX_HI_CD8', feature='PDL1', 
                         title = 'PDL1 in TREG due to TCF1_HI_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                         annotate.location = c(13,6000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'PDL1', interactingType = 'TCF1_HI_TOX_HI_CD8', p.val.permutation = a[[3]])) )
ggsave(a[[1]], filename = 'output/hist.TREG.PDL1.TCF1hiToxhiCD8.pdf', width=4,height=3)
```


```{r effects-of-termTex-adjacency-on-treg}
a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_LO_TOX_HI_CD8', feature='FOXP3', 
                    title = 'FOXP3 in TREG due to TCF1_LO_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                    annotate.location = c(13,3000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'FOXP3', interactingType = 'TCF1_LO_TOX_HI_CD8', p.val.permutation = a[[3]]))) 
ggsave(a[[1]], filename = 'output/hist.TREG.FOXP3.TCF1loToxhiCD8.pdf', width=4,height=3)

a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_LO_TOX_HI_CD8', feature='ICOS', 
                    title = 'ICOS in TREG due to TCF1_LO_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                    annotate.location = c(13,3000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'ICOS', interactingType = 'TCF1_LO_TOX_HI_CD8', p.val.permutation = a[[3]])))
ggsave(a[[1]], filename = 'output/hist.TREG.ICOS.TCF1loToxhiCD8.pdf', width=4,height=3)

a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_LO_TOX_HI_CD8', feature='LAG3', 
                    title = 'LAG3 in TREG due to TCF1_LO_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                    annotate.location = c(13,5000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'LAG3', interactingType = 'TCF1_LO_TOX_HI_CD8', p.val.permutation = a[[3]])))
ggsave(a[[1]], filename = 'output/hist.TREG.LAG3.TCF1loToxhiCD8.pdf', width=4,height=3)

a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_LO_TOX_HI_CD8', feature='CD39', 
                    title = 'CD39 in TREG due to TCF1_LO_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                    annotate.location = c(13,2000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'CD39', interactingType = 'TCF1_LO_TOX_HI_CD8', p.val.permutation = a[[3]])))
ggsave(a[[1]], filename = 'output/hist.TREG.CD39.TCF1loToxhiCD8.pdf', width=4,height=3)

a <- neighbors.histogram(target.cell = neighborhoods.grouped.TREG, interacting.type = 'TCF1_LO_TOX_HI_CD8', feature='PDL1', 
                    title = 'PDL1 in TREG due to TCF1_LO_TOX_HI_CD8', xlab='z score', binwidth=0.25, permutations=1000,
                    annotate.location = c(13,4000))
ICF.table <- rbindlist(list(ICF.table, 
                            data.table("Neighbors" = mean(a[[2]][source == 'Neighbors', data]), 
                                       "No.Neighbors" = mean(a[[2]][source == 'No.Neighbors', data]), 
                                       feature = 'PDL1', interactingType = 'TCF1_LO_TOX_HI_CD8', p.val.permutation = a[[3]])))
ggsave(a[[1]], filename = 'output/hist.TREG.PDL1.TCF1loToxhiCD8.pdf', width=4,height=3)

```


```{r plot-of-Treg-ICFtable-given-Tex}
ICF.table.bk <- copy(ICF.table)
ICF.table <- ICF.table[, fold.change:= Neighbors/No.Neighbors]
ICF.table <- ICF.table[order(fold.change, decreasing = T)]
ICF.table$feature <- factor(ICF.table$feature, levels = ICF.table$feature[1:5])
ggplot(ICF.table, aes(y=feature, x=fold.change, fill=interactingType)) + 
  geom_bar(stat='identity', position=position_dodge(), width = 0.5) + 
  scale_fill_manual(values = c("#4393c3", "#d6604d")) + 
  xlab("Fold change in z-score") + ylab("Feature") + 
  theme_classic() + geom_vline(xintercept=1, linetype='dashed',  linewidth = 0.5, color='grey50') +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.x = element_blank(), 
        axis.text.y = element_text(size=12), axis.text.x = element_text(size=12), axis.title = element_text(size=12), 
        title = element_text(size=12))  +
  ggtitle('TREG ICF due to Tex subsets')
  # geom_text(aes(label=paste0("p=", ifelse(p.val.permutation < 0.001, "<0.001", round(p.val.permutation,3)))), 
  #           position=position_dodge(width=0.9), vjust=-0.5, size=3)  
# ggsave(filename = 'output/TREG_ICF_by_Tex.pdf', width=6,height=4)

```


# Dose response for Treg adjacency 
```{r treg-effect-gzmb-doseResponse}

target.cell <- TCFlo.CD8[TREG == 1]  # start with TREG neighbors
val.1 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat['GZMB', target.cell$from] 
target.cell <- TCFlo.CD8[TREG < 1]  # now with no TREG neighbors
val.0 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat['GZMB', target.cell$from]
val.2 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat['GZMB', TCFlo.CD8[TREG==2,from]] 
val.3 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat['GZMB', TCFlo.CD8[TREG==3,from]] 
val.4 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat['GZMB', TCFlo.CD8[TREG==4,from]] 
val.5 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat['GZMB', TCFlo.CD8[TREG==5,from]] 
val.6 <- Giotto::getExpression(gobject=sln, values='normalized')@exprMat['GZMB', TCFlo.CD8[TREG==6,from]] 


dt <- data.table(data = c(val.0, val.1, val.2, val.3, val.4, val.5, val.6), 
                 source = c(rep("No.Treg.Neighbors", length(val.0)), 
                            rep("One.Treg.Neighbors", length(val.1)),                            
                            rep("Two.Treg.Neighbors", length(val.2)),
                            rep("Three.Treg.Neighbors", length(val.3)),
                            rep("Four.Treg.Neighbors", length(val.4)),
                            rep("Five.Treg.Neighbors", length(val.5)),
                            rep("Six.Treg.Neighbors", length(val.6))
                            )
                 )  # make a new dt where the expression values of FEAT can be plotted 
dt$source <- factor(dt$source, levels = c("No.Treg.Neighbors","One.Treg.Neighbors","Two.Treg.Neighbors", "Three.Treg.Neighbors",
                                          "Four.Treg.Neighbors","Five.Treg.Neighbors", "Six.Treg.Neighbors"))
dt %>% group_by(source) %>% rstatix::get_summary_stats(data, type = 'common')
dt %>% rstatix::wilcox_test(formula = data ~ source)

```

Not clear if there is a dose-response effect. I think it gets complicated when there are <1000 cells. Probably this cannot be analyzed unless there are more data points to evaluate 

# Find cells expressing GzmB and show in tissue 
```{r draw-cell-interactions-in-silico, plot.width=6}

make.subset.for.spatPlot <- function(target, select.row, meta.data)
{
  target <- names(target[select.row])  # pick the target cell, collect the cell ID 
  print(target)
  target <- meta.data[cell_ID == target]  
  keepCells <- subset(meta.data, 
                      list_ID == target$list_ID & 
                        y %in% c((target$y-125):(target$y+125)) & 
                        x %in% c((target$x-125):(target$x+125)) )$cell_ID  
  return(keepCells)
}

plot.Feat.pairedVoroni <- function(target, select.row, feat.show, max.upper, subset.colors, meta.data, pass.title, 
                                   voronoi.only = F, g.object=sln)
{
  keepCells <- make.subset.for.spatPlot(target = target, select.row = select.row, meta.data = meta.data)
  sln.small <- subsetGiotto(g.object, cell_ids = keepCells)
  a <- spatPlot(sln.small, cell_color = 'L2_annotations', cell_color_code = subset.colors,  
             point_shape = 'voronoi', vor_max_radius = 30, point_size = 1, point_alpha=0, vor_border_color='grey40', 
             title = pass.title, legend_text = 6, legend_symbol_size=1, return_plot=T)
  if(voronoi.only==T) return(a)
  b <- spatFeatPlot2D(sln.small, feats= feat.show, point_shape='voronoi', vor_max_radius = 30, vor_border_color = 'grey40', 
                 gradient_midpoint = 0, expression_values = 'normalized', gradient_limits = c(-5, max.upper), 
                 return_plot = T)
  return(list(a,b))
}

expr.feat.1 <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat['GZMB', TCFlo.CD8[TREG >=1,from ]]  # get all expr 
upperhalf <- expr.feat.1[which(expr.feat.1 > median(expr.feat.1))]   # get the upper half of the expression values
sel.1 <-  R.utils::withSeed( expr = names(upperhalf[sample(upperhalf,1)]), seed = 100); print(expr.feat.1[sel.1])
plotting.cells <- make.subset.for.spatPlot(target=expr.feat.1, select.row= sel.1, meta.data = md)  # give me what will be in the view port
expr.feat.1.max <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat['GZMB', plotting.cells ] %>% max  # get max expr

# now with no TREG neighbors
expr.feat.0 <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat['GZMB', TCFlo.CD8[TREG ==0,from ]]  # get all expr 
upperhalf <- expr.feat.0[which(expr.feat.0 > median(expr.feat.0))]   # get the upper half of the expression values
sel.0 <-  R.utils::withSeed( expr = names(upperhalf[sample(upperhalf,1)]), seed = 100); print(expr.feat.0[sel.0])
plotting.cells <- make.subset.for.spatPlot(expr.feat.0, select.row= sel.0, meta.data = md)
expr.feat.0.max <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat['GZMB', plotting.cells ] %>% max

target.cell <- TCFlo.CD8[TREG >=1]  
a <- plot.Feat.pairedVoroni(target = expr.feat.1, select.row = sel.1, feat.show='GZMB', max.upper= min(expr.feat.0.max, expr.feat.1.max),
                            subset.colors = stableColorsL2, meta.data = md, # voronoi.only = T,
                            pass.title = paste0("GZMB if 1+ Treg (", sel.1, ")"))
target.cell <- TCFlo.CD8[TREG == 0]  
b <- plot.Feat.pairedVoroni(target = expr.feat.0, select.row = sel.0, feat.show='GZMB', max.upper= min(expr.feat.0.max, expr.feat.1.max),
                            subset.colors = stableColorsL2, meta.data = md, # voronoi.only = T,
                            pass.title = paste0("GZMB if 0 Treg (", sel.0, ")"))

```



Though these example images might come across as cherry-picked. It will be very hard to show a single cell-based example to compare unless it's only the mean being shown (but the mean is a poor descriptor for a skew population)


```{r region-given-pairings}

tmp <- neighborhoods$raw.table
tmp <- purrr::list_rbind(tmp, names_to='list_ID')
tmp <- merge(y=tmp, x=md, by.y='from', by.x='cell_ID', all.x=T)
# find all rows representing pairings of Activated Treg with CD8, then display graph of spatial region

pairings <- tmp[from_cell_type %like% 'TCF1' & ACTIVATED_TREG >0]

# table(pairings$Classifier.Label, pairings$list_ID.x)
#                 ctrl.1 ctrl.2 ctrl.3 ctrl.4 ctrl.5 tumor.1 tumor.10 tumor.2 tumor.3 tumor.4 tumor.5 tumor.6 tumor.7 tumor.8 tumor.9
# B Cell Follicle    294     32     87    145     21      19       91      19      40      78     317     229     295     123     187
# GC                   0      0      0      0      0      10        0       0       0       0       0       1       1       1       0
# T Cell Zone        122    505     97   1284     65       4       37      25     308     104    1401    2121     138     446    1105
# T/B Mixed          684    276    143    698     68       2      125      14      95      31    1491    1838      50     392     920
# Tumor                5      0      0      0      3       0        2       0       0       6      28      18       7       0       1

res.table <- as.data.table(table(pairings$Classifier.Label, pairings$list_ID.x, pairings$from_cell_type) )
res.table[, sample.labels := name_mapping[V2] ]
res.table[, prop.N := proportions(N), by = c('V2', 'V3')]
res.table[, ave := mean(prop.N), by=c('V1','V3')]
res.table[, sd := sd(prop.N), by=c('V1', 'V3')]
for.factoring <- res.table[V1 == 'T/B Mixed' & V3 == 'TCF1_HI_TOX_HI_CD8']; setorder(for.factoring, -prop.N)
res.table.plot <- res.table[V3 == 'TCF1_HI_TOX_HI_CD8']
res.table.plot[, V2 := factor(res.table.plot$V2, levels = for.factoring$V2)]
ggplot(data = res.table.plot, aes(y = prop.N, x=V2, fill=V1)) + geom_bar(position='fill', stat='identity') + theme_classic() +
  ggtitle("Location of Treg interactions - TCF1_HI_TOX_HI_CD8") + ylab("Proportion of all regions") + labs(fill = "Region") + 
  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) + 
  scale_fill_manual(values = classifier_label_colors) # + scale_x_discrete(labels = res.table.plot$sample.labels)
# ggsave('output/locate.Treg.TCFhiCD8.stackedbar.pdf', width=5,height=5)
for.factoring <- res.table[V1 == 'T/B Mixed' & V3 == 'TCF1_LO_TOX_HI_CD8']; setorder(for.factoring, -prop.N)
res.table.plot <- res.table[V3 == 'TCF1_LO_TOX_HI_CD8']
res.table.plot[, V2 := factor(res.table.plot$V2, levels = for.factoring$V2)]
ggplot(data = res.table.plot, aes(y = prop.N, x=V2, fill=V1)) + geom_bar(position='fill', stat='identity') + theme_classic() +
  ggtitle("Location of Treg interactions - TCF1_LO_TOX_HI_CD8") + ylab("Proportion of all regions") + labs(fill = "Region") + 
  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) + 
  scale_fill_manual(values = classifier_label_colors)
# ggsave('output/locate.Treg.TCFloCD8.stackedbar.pdf', width=5,height=5)
rm(pairings); rm(res.table); rm(tmp); rm(res.table.plot)

```

```{r explore-B-cell-interactions}
a <- findImportantInteractions( interaction.dt= interaction.byRecur, condition.exp="recur", condition.ctrl="no.recur", ef.threshold=1)
a$df[unified_int %like% 'BCELL']
proteinsList <- c("BCA1", "CD11C" ,"CD19","CD20","HLA.DR", "PD1")
                  # "CD3E", "CD4","CD8" ,,"CD27","CD38","CD39", "CD45RO","CXCR3","CXCR5",
                  # "FOXP3" , "GZMB" , "ICOS","IFNG","IRF4","IRF7","KI67", "LAG3",
                  # "IL21","PDL1","STAT1","STAT3","TCF1","TOX" ) # "TBET","PMTOR" ,"CD21","MAC2/GAL3","PAX5","VIMENTIN"
# TCELL_BCELL_MYELOID_MIXED
# BCELL_MYELOID_MIXED
# MIXED_FOLLICULAR_BCELL
retrieve.subset <- md[sample(11e6, 1e4,replace = F), cell_ID]  # L2_annotations == probe.subset
features <- as.data.table(t(as.matrix(Giotto::getExpression(sln, output="matrix")[, retrieve.subset] )), keep.rownames="cell_ID")
features <- inner_join(features, md, by='cell_ID')


probe.subset <- 'NAIVE_CD8'
retrieve.subset <- md[, cell_ID]  # L2_annotations == probe.subset
features <- as.data.table(t(as.matrix(Giotto::getExpression(sln, output="matrix")[, retrieve.subset] )), keep.rownames="cell_ID")
features <- melt(features, id.vars = 'cell_ID')
features <- features[variable %in% proteinsList]
a <- ggplot(features, aes(x = value, y= variable, fill = variable)) +  
  ggridges::geom_density_ridges(    stat = "binline",  bins = 2000, # Use the "binline" stat for a histogram appearance
    scale = 3,  alpha = 0.8, draw_baseline = FALSE, color='gray20', linewidth=0.2 ) + 
  theme_classic() + scale_x_sqrt( breaks = seq(0,100,10), limits=c(0,100))+ ggtitle(probe.subset)
probe.subset <- 'BCELL_MYELOID_MIXED'
retrieve.subset <- md[L2_annotations == probe.subset, cell_ID]
features <- as.data.table(t(as.matrix(Giotto::getExpression(sln, output="matrix")[, retrieve.subset] )), keep.rownames="cell_ID")
features <- melt(features, id.vars = 'cell_ID')
features <- features[variable %in% proteinsList]
b<- ggplot(features, aes(x = value, y= variable, fill = variable)) +  
  ggridges::geom_density_ridges(    stat = "binline",  bins = 2000, # Use the "binline" stat for a histogram appearance
    scale = 3,  alpha = 0.8, draw_baseline = FALSE, color='gray20', linewidth=0.2  ) + # Removes the horizontal line at the bottom of each plot
  theme_classic() + scale_x_sqrt( breaks = seq(0,100,10), limits=c(0,100)) + ggtitle(probe.subset)
gridExtra::grid.arrange(a,b, nrow=1) 

```



# identify the distance to the nearest Treg 

```{r cellproxEnrich-function-k50}
tableResults.k50 <- list()
annot.k50 <- list()
for (i in list_ID)
{
  if( .Platform$OS.type != "windows")  {cat(paste("Starting on: ", i, " on ", date(),"\n"), 
                                            file = paste0("progress.",timeStamp,".txt"), append = T)   }
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=50, verbose=T, name = "spatialknn.k50.persample")
  annot.k50[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L2_annotations", spatial_network_name = 'spatialknn.k50.persample')
  # tableResults.k50[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L2_annotations', adjust_method = "fdr",
  #                                      number_of_simulations = 200, spatial_network_name = "spatialknn.k50.persample")
}
# saveRDS(list(annot.k50=annot.k50), file="neighborhood.annot.k50.sln.RDS")  # all data merged
# saveRDS(list(annot.k50=annot.k50, tableResults.k50 = tableResults.k50), file="neighborhood.annot.k50.sln.RDS")  # all data merged


# Job ID: 6403353
# Cluster: slurm_cluster
# User/Group: heratr01/heratr01
# State: COMPLETED (exit code 0)
# Nodes: 1
# Cores per node: 5
# CPU Utilized: 01:19:20
# CPU Efficiency: 28.08% of 04:42:30 core-walltime
# Job Wall-clock time: 00:56:30
# Memory Utilized: 135.58 GB
# Memory Efficiency: 27.12% of 500.00 GB
```


```{r identify-distance-to-Treg-loadData, eval=FALSE}
annot.k50 <- readRDS(file="neighborhood.annot.k50.sln.RDS")[["annot.k50"]]  # all data merged
```


```{r identify-distance-to-Treg}
#                   from                   to sdimx_begin sdimy_begin sdimx_end sdimy_end distance     weight
       #                  <char>               <char>       <num>       <num>     <num>     <num>    <num>      <num>
       # 1: ctrl.1-ctrl.1_100000  ctrl.1-ctrl.1_99994      9482.0      9320.0    9482.5    9305.0 15.00833 0.06246747
       # 2: ctrl.1-ctrl.1_100000 ctrl.1-ctrl.1_100002      9482.0      9320.0    9500.0    9323.5 18.33712 0.05171401
       # 3: ctrl.1-ctrl.1_100000 ctrl.1-ctrl.1_100004      9482.0      9320.0    9462.5    9333.0 23.43608 0.04092309
       #                                to_cell_type                           from_cell_type type_int
       #                                      <char>                                   <char>   <char>
       # 1:                    TCF1_MIXED_TOX_HI_CD8                       TCF1_HI_TOX_HI_CD8   hetero
       # 2:                           CD4_CD20_MIXED                       TCF1_HI_TOX_HI_CD8   hetero
       # 3:                             CD8_PC_MIXED                       TCF1_HI_TOX_HI_CD8   hetero
       #                                                    from_to                                              unified_int
       #                                                     <char>                                                   <char>
       # 1:                TCF1_HI_TOX_HI_CD8-TCF1_MIXED_TOX_HI_CD8                TCF1_HI_TOX_HI_CD8--TCF1_MIXED_TOX_HI_CD8
       # 2:                       TCF1_HI_TOX_HI_CD8-CD4_CD20_MIXED                       CD4_CD20_MIXED--TCF1_HI_TOX_HI_CD8
       # 3:                         TCF1_HI_TOX_HI_CD8-CD8_PC_MIXED                         CD8_PC_MIXED--TCF1_HI_TOX_HI_CD8
# average of 20.5 million rows per tissue 

neighborhoods.k50 <- defineNeighborhoods(annot.list = annot.k50)
# then make histogram for # neighbors for clarity
tmp <- purrr::list_rbind(neighborhoods.k50$prop.table, names_to = "list_ID")
ggplot(tmp, aes(x = row_total)) + geom_histogram(binwidth=0.25, position='identity', fill='grey70', color='black') +
  theme_classic() + ggtitle("All cells (k=50)") + xlab("Number of neighbors") + 
  scale_x_continuous(limits = c(40,80), breaks = seq(40,80,10))
# ggsave(filename = 'output/hist.numbNeighbors.k50.pdf', width=2,height=3)
rm(tmp)

annot.k50.filtered <- list()
for(i in list_ID)
{
  # first filter out most records but save anything CD8 or Treg
  annot.k50.filtered[[i]] <- annot.k50[[i]][from_to %like% 'TREG' | from_to %like% 'TCF']
  print(paste('Finished ', i))
}  # down to 1.1 million rows per biological replicate

tmp <- purrr::list_rbind(annot.k50.filtered, names_to = "list_ID") 

   #    list_ID                   from                     to sdimx_begin sdimy_begin sdimx_end sdimy_end distance     weight
   #     <char>                 <char>                 <char>       <num>       <num>     <num>     <num>    <num>      <num>
   # 1:  ctrl.1   ctrl.1-ctrl.1_100000    ctrl.1-ctrl.1_99994      9482.0      9320.0    9482.5    9305.0 15.00833 0.06246747
   # 2:  ctrl.1   ctrl.1-ctrl.1_100000   ctrl.1-ctrl.1_100002      9482.0      9320.0    9500.0    9323.5 18.33712 0.05171401
   # 3:  ctrl.1   ctrl.1-ctrl.1_100000   ctrl.1-ctrl.1_100004      9482.0      9320.0    9462.5    9333.0 23.43608 0.04092309
   #                     to_cell_type        from_cell_type type_int                                          from_to
   #                           <char>                <char>   <char>                                           <char>
   # 1:         TCF1_MIXED_TOX_HI_CD8    TCF1_HI_TOX_HI_CD8   hetero         TCF1_HI_TOX_HI_CD8-TCF1_MIXED_TOX_HI_CD8
   # 2:                CD4_CD20_MIXED    TCF1_HI_TOX_HI_CD8   hetero                TCF1_HI_TOX_HI_CD8-CD4_CD20_MIXED
   # 3:                  CD8_PC_MIXED    TCF1_HI_TOX_HI_CD8   hetero                  TCF1_HI_TOX_HI_CD8-CD8_PC_MIXED
   #                                          unified_int
   #                                               <char>
   # 1:         TCF1_HI_TOX_HI_CD8--TCF1_MIXED_TOX_HI_CD8
   # 2:                CD4_CD20_MIXED--TCF1_HI_TOX_HI_CD8
   # 3:                  CD8_PC_MIXED--TCF1_HI_TOX_HI_CD8
   # 

tmp <- tmp[!from_cell_type %in% c('ARTIFACT', 'CD4.outlier', 'CD8.outlier') ]
filtered.k50 <- tmp[!to_cell_type %in% c('ARTIFACT', 'CD4.outlier', 'CD8.outlier') ]
rm(tmp)
ggplot(filtered.k50, aes(x = distance)) + geom_histogram(position='identity', binwidth=0.25) + 
  scale_x_continuous(limits = c(0,150)) + theme_classic() + ggtitle("Filtered - distances between cells")

# aveDist <- filtered.k50[, .(meanDist = mean(distance, na.rm=T), SDdist = sd(distance, na.rm=T), numb = length(distance)), 
#                by=.(list_ID, from_cell_type, to_cell_type)]

   #    list_ID                from_cell_type                  to_cell_type meanDist   SDdist  numb
   #     <char>                        <char>                        <char>    <num>    <num> <int>
   # 1:  ctrl.1            TCF1_HI_TOX_HI_CD8         TCF1_MIXED_TOX_HI_CD8 67.53681 82.40942   824
   # 2:  ctrl.1            TCF1_HI_TOX_HI_CD8                CD4_CD20_MIXED 53.38692 17.50482  5140
   # 3:  ctrl.1            TCF1_HI_TOX_HI_CD8                  CD8_PC_MIXED 60.57265 41.93290  1831
   # 4:  ctrl.1            TCF1_HI_TOX_HI_CD8            CD4_CD8_CD20_MIXED 53.12021 17.87240   690
   
# goal is to compare phenotype of CD8 based on distance to Treg 
# bin the CD8 subset based on proximity to Treg, then retrieve expression for GzmB for plot 

Treg.interactions <- filtered.k50[to_cell_type %like% "TREG"]
Tumor.interactions <- filtered.k50[to_cell_type %like% "Tumor"]
nearest.Treg <- Treg.interactions[, .(min.dist.Treg = min(distance)), by = .(list_ID, from, from_cell_type)]
nearest.Tumor <- Tumor.interactions[, .(min.dist.Tumor = min(distance)), by = .(list_ID, from, from_cell_type)]


#  ***** TCF1_LO *****
spec.prot = 'GZMB'; subset = "TCF1_LO_TOX_HI_CD8"; num.perm = 1000; radius = 15; 
lab.more = paste0(">",radius); lab.less = paste0("<=",radius)
nearest.Treg.test <- nearest.Treg[from_cell_type == subset]
max.dist <- max(nearest.Treg.test$min.dist.Treg)
nearest.Treg.test <- nearest.Treg.test[, dist.bin := cut(min.dist.Treg, breaks = c(0,radius,Inf), labels= c(lab.less,lab.more),
                                                           right=T, include_lowest=T)]  # break into bins 
prot.data <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat[spec.prot, nearest.Treg.test$from ]
nearest.Treg.test <- merge(nearest.Treg.test, data.frame(from = names(prot.data), data = prot.data), by='from')
nearest.Treg.test %>% group_by(dist.bin) %>% select(data) %>% rstatix::get_summary_stats(type = 'common')
dt <- nearest.Treg.test[,.(source = dist.bin, data=data)]
p.val <- permutation.test.neighborhoods(dt, group.0 = lab.more, group.1 = lab.less, num.permutations = num.perm)
ypos = max(dt$data) * 0.85
ggplot(dt, aes(x = source, y=data)) + 
  geom_violin(fill='grey80', draw_quantiles = c(0.5)) + ggtitle(subset) + 
  xlab("Distance to nearest Treg") + ylab(spec.prot) + 
  annotate('text', x=0.5, y=ypos, color='black', 
           label= paste0("Mean ", lab.less, ": ", round(mean(dt[source == lab.less, data]), 1), 
                       "\n",
                       "Mean ", lab.more,": ", round(mean(dt[source == lab.more, data]), 1),
                       "\nPermutation test: P=", ifelse(p.val==0, 1/num.perm, p.val)),
           hjust = 0, size=3) +
  theme_classic() + theme(legend.position = 'none', axis.text = element_text(size=12), 
                          axis.text.x = element_text(size=12, hjust=1, vjust=1, angle=45)) 
# ggsave(filename = 'output/dist.to.Treg.GZMB.TCFloCD8.pdf', width=3, height=3)


#  ***** TCF1_LO *****
spec.prot = 'IFNG'; subset = "TCF1_LO_TOX_HI_CD8"; num.perm = 1000; radius = 15; 
lab.more = paste0(">",radius); lab.less = paste0("<=",radius)
nearest.Treg.test <- nearest.Treg[from_cell_type == subset]
max.dist <- max(nearest.Treg.test$min.dist.Treg)
nearest.Treg.test <- nearest.Treg.test[, dist.bin := cut(min.dist.Treg, breaks = c(0,radius,Inf), labels= c(lab.less,lab.more),
                                                           right=T, include_lowest=T)]  # break into bins 
prot.data <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat[spec.prot, nearest.Treg.test$from ]
nearest.Treg.test <- merge(nearest.Treg.test, data.frame(from = names(prot.data), data = prot.data), by='from')
nearest.Treg.test %>% group_by(dist.bin) %>% select(data) %>% rstatix::get_summary_stats(type = 'common')
dt <- nearest.Treg.test[,.(source = dist.bin, data=data)]
p.val <- permutation.test.neighborhoods(dt, group.0 = lab.more, group.1 = lab.less, num.permutations = num.perm)
ypos = max(dt$data) * 0.85
ggplot(dt, aes(x = source, y=data)) + 
  geom_violin(fill='grey80', draw_quantiles = c(0.5)) + ggtitle(subset) + 
  xlab("Distance to nearest Treg") + ylab(spec.prot) + 
  annotate('text', x=0.5, y=ypos, color='black', 
           label= paste0("Mean ", lab.less, ": ", round(mean(dt[source == lab.less, data]), 1), 
                       "\n",
                       "Mean ", lab.more,": ", round(mean(dt[source == lab.more, data]), 1),
                       "\nPermutation test: P=", ifelse(p.val==0, 1/num.perm, p.val)),
           hjust = 0, size=3) +
  theme_classic() + theme(legend.position = 'none', axis.text = element_text(size=12), 
                          axis.text.x = element_text(size=12, hjust=1, vjust=1, angle=45)) 
ggsave(filename = 'output/dist.to.Treg.IFNG.TCFloCD8.pdf', width=3, height=3)


#  ***** TCF1_HI *****
spec.prot = 'GZMB'; subset = "TCF1_HI_TOX_HI_CD8"; num.perm = 1000; radius = 15; 
lab.more = paste0(">",radius); lab.less = paste0("<=",radius)
nearest.Treg.test <- nearest.Treg[from_cell_type == subset]
max.dist <- max(nearest.Treg.test$min.dist.Treg)
nearest.Treg.test <- nearest.Treg.test[, dist.bin := cut(min.dist.Treg, breaks = c(0,radius,Inf), labels= c(lab.less,lab.more),
                                                           right=T, include_lowest=T)]  # break into bins 
prot.data <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat[spec.prot, nearest.Treg.test$from ]
nearest.Treg.test <- merge(nearest.Treg.test, data.frame(from = names(prot.data), data = prot.data), by='from')
nearest.Treg.test %>% group_by(dist.bin) %>% select(data) %>% rstatix::get_summary_stats(type = 'common')
dt <- nearest.Treg.test[,.(source = dist.bin, data=data)]
p.val <- permutation.test.neighborhoods(dt, group.0 = lab.more, group.1 = lab.less, num.permutations = num.perm)
ypos = max(dt$data) * 0.85
ggplot(dt, aes(x = source, y=data)) + 
  geom_violin(fill='grey80', draw_quantiles = c(0.5)) + ggtitle(subset) + 
  xlab("Distance to nearest Treg") + ylab(spec.prot) + 
  annotate('text', x=0.5, y=ypos, color='black', 
           label= paste0("Mean ", lab.less, ": ", round(mean(dt[source == lab.less, data]), 1), 
                       "\n",
                       "Mean ", lab.more,": ", round(mean(dt[source == lab.more, data]), 1),
                       "\nPermutation test: P=", ifelse(p.val==0, 1/num.perm, p.val)),
           hjust = 0, size=3) +
  theme_classic() + theme(legend.position = 'none', axis.text = element_text(size=12), 
                          axis.text.x = element_text(size=12, hjust=1, vjust=1, angle=45)) 
# ggsave(filename = 'output/dist.to.Treg.GZMB.TCFhiCD8.pdf', width=3, height=3)




#  ***** Tumor *****
spec.prot = 'GZMB'; subset = "TCF1_LO_TOX_HI_CD8"; num.perm = 1000; radius = 15; 
lab.more = paste0(">",radius); lab.less = paste0("<=",radius)
nearest.Treg.test <- nearest.Tumor[from_cell_type == subset]
max.dist <- max(nearest.Treg.test$min.dist.Treg)
nearest.Treg.test <- nearest.Treg.test[, dist.bin := cut(min.dist.Tumor, breaks = c(0,radius,Inf), labels= c(lab.less,lab.more),
                                                           right=T, include_lowest=T)]  # break into bins 
prot.data <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat[spec.prot, nearest.Treg.test$from ]
nearest.Treg.test <- merge(nearest.Treg.test, data.frame(from = names(prot.data), data = prot.data), by='from')
nearest.Treg.test %>% group_by(dist.bin) %>% select(data) %>% rstatix::get_summary_stats(type = 'common')
dt <- nearest.Treg.test[,.(source = dist.bin, data=data)]
p.val <- permutation.test.neighborhoods(dt, group.0 = lab.more, group.1 = lab.less, num.permutations = num.perm)
ypos = max(dt$data) * 1.15
ggplot(dt, aes(x = source, y=data)) + 
  geom_violin(fill='grey80', draw_quantiles = c(0.5)) + ggtitle(subset) + 
  xlab("Distance to nearest Tumor cell") + ylab(spec.prot) + 
  annotate('text', x=0.5, y=ypos, color='black', 
           label= paste0("Mean ", lab.less, ": ", round(mean(dt[source == lab.less, data]), 1), 
                       "\n",
                       "Mean ", lab.more,": ", round(mean(dt[source == lab.more, data]), 1),
                       "\nPermutation test: P=", ifelse(p.val==0, 1/num.perm, p.val)),
           hjust = 0, size=3) +
  theme_classic() + theme(legend.position = 'none', axis.text = element_text(size=12), 
                          axis.text.x = element_text(size=12, hjust=1, vjust=1, angle=45)) + 
  scale_y_continuous(limits = c(0,34))
# ggsave(filename = 'output/dist.to.Tumor.GZMB.TCFloCD8.pdf', width=3, height=3)


spec.prot = 'IFNG'; subset = "TCF1_LO_TOX_HI_CD8"; num.perm = 1000; radius = 15; 
lab.more = paste0(">",radius); lab.less = paste0("<=",radius)
nearest.Treg.test <- nearest.Tumor[from_cell_type == subset]
max.dist <- max(nearest.Treg.test$min.dist.Treg)
nearest.Treg.test <- nearest.Treg.test[, dist.bin := cut(min.dist.Tumor, breaks = c(0,radius,Inf), labels= c(lab.less,lab.more),
                                                           right=T, include_lowest=T)]  # break into bins 
prot.data <- Giotto::getExpression(gobject = sln, values = "normalized")@exprMat[spec.prot, nearest.Treg.test$from ]
nearest.Treg.test <- merge(nearest.Treg.test, data.frame(from = names(prot.data), data = prot.data), by='from')
nearest.Treg.test %>% group_by(dist.bin) %>% select(data) %>% rstatix::get_summary_stats(type = 'common')
dt <- nearest.Treg.test[,.(source = dist.bin, data=data)]
p.val <- permutation.test.neighborhoods(dt, group.0 = lab.more, group.1 = lab.less, num.permutations = num.perm)
ypos = max(dt$data) * 1
ggplot(dt, aes(x = source, y=data)) + 
  geom_violin(fill='grey80', draw_quantiles = c(0.5)) + ggtitle(subset) + 
  xlab("Distance to nearest Tumor cell") + ylab(spec.prot) + 
  annotate('text', x=0.5, y=ypos, color='black', 
           label= paste0("Mean ", lab.less, ": ", round(mean(dt[source == lab.less, data]), 1), 
                       "\n",
                       "Mean ", lab.more,": ", round(mean(dt[source == lab.more, data]), 1),
                       "\nPermutation test: P=", ifelse(p.val==0, 1/num.perm, p.val)),
           hjust = 0, size=3) +
  theme_classic() + theme(legend.position = 'none', axis.text = element_text(size=12), 
                          axis.text.x = element_text(size=12, hjust=1, vjust=1, angle=45)) + 
  scale_y_continuous(limits = c(0,40))
# ggsave(filename = 'output/dist.to.Tumor.IFNG.TCFloCD8.pdf', width=3, height=3)

rm(annot.k50)
```


Job ID: 6784618
Cluster: slurm_cluster
User/Group: heratr01/heratr01
State: COMPLETED (exit code 0)
Nodes: 1
Cores per node: 5
CPU Utilized: 05:28:57
CPU Efficiency: 26.30% of 20:50:55 core-walltime
Job Wall-clock time: 04:10:11
Memory Utilized: 317.18 GB
Memory Efficiency: 63.44% of 500.00 GB



