---
title: "Neighborhood analyses"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---
 

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```
 
 
```{r call-libraries} 
date()
library(Giotto)
library(future); if(parallelly::availableCores() > 2 & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); library(leiden);   # library(multinet);  
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
library(ComplexHeatmap)
set.seed(43)
sessionInfo()
```

# read in data object 
```{r read-in-data}
# sln <- readRDS("sln.clustering2.res0.5.20250411.1611.RDS")
# md <- readRDS(file = "md.exprs.combined.sln.clustering2.res0.5.20250411.1611.RDS")  # sabrina's metadata with L1_annotations
# df <- md[,c("cell_ID","L1_annotations")]
# sln <- addCellMetadata(gobject = sln, new_metadata = df, by_column=T, column_cell_ID = 'cell_ID')
# rm(df)
# sln <- createSpatialKNNnetwork(sln) 
# saveRDS(sln, "neighborhood.spatialnetwork.sln.clustering2.res0.5.20250411.1611.RDS")
sln <- readRDS("neighborhood.spatialnetwork.sln.clustering2.res0.5.20250411.1611.RDS")

```


# try a customized version of cellProxEnrich
```{r custom-cellproxEnrich-function}
cellProx.sim <- function(gobject, spat_unit = NULL, feat_type = NULL, spatial_network_name = "Delaunay_network",
    cluster_column, number_of_simulations = 1000, 
    adjust_method = c("none", "fdr", "bonferroni", "BH", "holm", "hochberg", "hommel", "BY"), 
    set_seed = TRUE, seed_number = 1234) 
{
    # Set feat_type and spat_unit
    spat_unit <- set_default_spat_unit(
        gobject = gobject,
        spat_unit = spat_unit
    )
    feat_type <- set_default_feat_type(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type
    )

    # p.adj test
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))

    spatial_network_annot <- annotateSpatialNetwork(
        gobject = gobject,
        feat_type = feat_type,
        spat_unit = spat_unit,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column
    )

    # remove double edges between same cells #
    # a simplified network does not have double edges between cells #

    # data.table variables
    unified_cells <- type_int <- N <- NULL

    spatial_network_annot <- GiottoUtils:::dt_sort_combine_two_columns(
        spatial_network_annot, "to", "from", "unified_cells"
    )
    spatial_network_annot <- spatial_network_annot[!duplicated(unified_cells)]

    sample_dt <- Giotto:::make_simulated_network(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column,
        number_of_simulations = number_of_simulations,
        set_seed = set_seed,
        seed_number = seed_number
    )

    # combine original and simulated network
    table_sim_results <- sample_dt[, .N, by = c(
        "unified_int", "type_int", "round"
    )]

    ## create complete simulations
    ## add 0 if no single interaction was found
    unique_ints <- unique(table_sim_results[, .(unified_int, type_int)])

    # data.table with 0's for all interactions
    minimum_simulations <- unique_ints[rep(
        seq_len(nrow(unique_ints)), number_of_simulations
    ), ]
    minimum_simulations[, round := rep(
        paste0("sim", seq_len(number_of_simulations)),
        each = nrow(unique_ints)
    )]
    minimum_simulations[, N := 0]

    table_sim_minimum_results <- rbind(table_sim_results, minimum_simulations)
    table_sim_minimum_results[, V1 := sum(N), by = c(
        "unified_int", "type_int", "round"
    )]
    table_sim_minimum_results <- unique(
        table_sim_minimum_results[, .(unified_int, type_int, round, V1)]
    )
    table_sim_results <- table_sim_minimum_results


    # data.table variables
    orig <- unified_int <- V1 <- original <- enrichm <- simulations <- NULL

    table_sim_results[, orig := "simulations"]
    spatial_network_annot[, round := "original"]

    table_orig_results <- spatial_network_annot[, .N, by = c(
        "unified_int", "type_int", "round"
    )]
    table_orig_results[, orig := "original"]
    data.table::setnames(table_orig_results, old = "N", new = "V1")

    table_results <- rbind(table_orig_results, table_sim_results)



    # add missing combinations from original or simulations
    # probably not needed anymore
    all_simulation_ints <- as.character(unique(table_results[
        orig == "simulations"
    ]$unified_int))
    all_original_ints <- as.character(unique(table_results[
        orig == "original"
    ]$unified_int))
    missing_in_original <- all_simulation_ints[
        !all_simulation_ints %in% all_original_ints
    ]
    missing_in_simulations <- all_original_ints[
        !all_original_ints %in% all_simulation_ints
    ]
    create_missing_for_original <- table_results[
        unified_int %in% missing_in_original
    ]
    create_missing_for_original <- unique(create_missing_for_original[
        , c("orig", "V1") := list("original", 0)
    ])
    create_missing_for_simulations <- table_results[
        unified_int %in% missing_in_simulations
    ]
    create_missing_for_simulations <- unique(
        create_missing_for_simulations[, c("orig", "V1") := list(
            "simulations", 0
        )]
    )
# stop **********************************************************************************************

    table_results <- do.call(
        "rbind",
        list(
            table_results, create_missing_for_original,
            create_missing_for_simulations
        )
    )
    return(table_results)
}

cellProx.calcP <- function(table_results, number_of_simulations = 1000, 
    adjust_method = c("none", "fdr", "bonferroni", "BH", "holm", "hochberg", "hommel", "BY"), 
    set_seed = TRUE, seed_number = 1234)
{
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))
    ## p-values
    combo_list <- rep(NA, length = length(unique(table_results$unified_int)))
    p_high <- rep(NA, length = length(unique(table_results$unified_int)))
    p_low <- rep(NA, length = length(unique(table_results$unified_int)))

    for (int_combo in seq_along(unique(table_results$unified_int))) {
        this_combo <- as.character(unique(table_results$unified_int)[int_combo])

        sub <- table_results[unified_int == this_combo]

        orig_value <- sub[orig == "original"]$V1
        sim_values <- sub[orig == "simulations"]$V1

        length_simulations <- length(sim_values)
        if (length_simulations != number_of_simulations) {
            additional_length_needed <- number_of_simulations -
                length_simulations
            sim_values <- c(sim_values, rep(0, additional_length_needed))
        }

        p_orig_higher <- 1 - (sum((orig_value + 1) > (sim_values + 1)) /
            number_of_simulations)
        p_orig_lower <- 1 - (sum((orig_value + 1) < (sim_values + 1)) /
            number_of_simulations)

        combo_list[[int_combo]] <- this_combo
        p_high[[int_combo]] <- p_orig_higher
        p_low[[int_combo]] <- p_orig_lower
    }
    res_pvalue_DT <- data.table::data.table(
        unified_int = as.vector(combo_list),
        p_higher_orig = p_high,
        p_lower_orig = p_low
    )


    # depletion or enrichment in barplot format
    table_mean_results <- table_results[, .(mean(V1)), by = c(
        "orig", "unified_int", "type_int"
    )]
    table_mean_results_dc <- data.table::dcast.data.table(
        data = table_mean_results, formula = type_int + unified_int ~ orig,
        value.var = "V1"
    )
    table_mean_results_dc[, original := ifelse(is.na(original), 0, original)]
    table_mean_results_dc[, enrichm := log2((original + 1) / (simulations + 1))]


    table_mean_results_dc <- merge(
        table_mean_results_dc, res_pvalue_DT,
        by = "unified_int"
    )
    data.table::setorder(table_mean_results_dc, enrichm)
    table_mean_results_dc[, unified_int := factor(unified_int, unified_int)]

    # adjust p-values for mht

    # data.table variables
    p.adj_higher <- p.adj_lower <- p_lower_orig <- p_higher_orig <-
        PI_value <- int_ranking <- NULL

    table_mean_results_dc[, p.adj_higher := stats::p.adjust(
        p_higher_orig,
        method = sel_adjust_method
    )]
    table_mean_results_dc[, p.adj_lower := stats::p.adjust(
        p_lower_orig,
        method = sel_adjust_method
    )]


    table_mean_results_dc[, PI_value := ifelse(p.adj_higher <= p.adj_lower,
        -log10(p.adj_higher + (1 / number_of_simulations)) * enrichm,
        -log10(p.adj_lower + (1 / number_of_simulations)) * enrichm
    )]
    data.table::setorder(table_mean_results_dc, PI_value)

    # order
    table_mean_results_dc <- table_mean_results_dc[order(-PI_value)]
    table_mean_results_dc[, int_ranking := seq_len(.N)]

    return(list(
        raw_sim_table = table_results,
        enrichm_res = table_mean_results_dc
    ))
}

md <- pDataDT(sln)
list_ID <- unique(md$list_ID)
gc()

tableResults <- list()

for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm) 
  
  tableResults[[i]] <- cellProx.sim(sln.sm, cluster_column = 'leiden_harmony', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "knn_network")
}
dt <- purrr::list_rbind(tableResults, names_to = "id")
saveRDS(dt, file="neighborhood.tableResults.sln.clustering2.res0.5.20250411.1611.RDS")  # all data merged 
dt <- readRDS(file="neighborhood.tableResults.sln.clustering2.res0.5.20250411.1611.RDS")  # all data merged 

# subset dt based on groupings of interest 
cellProx.subgroup <- function(dt)
{
  wide_dt <- dcast(dt,                # Your input data.table
                   unified_int ~ round, # Formula: rows ~ columns
                   value.var = "V1",  # Column with values for the new wide cells
                   fun.aggregate = sum, # Function to aggregate values (when multiple rows match a cell)
                   na.rm = TRUE,      # Tell sum() to ignore NA values during aggregation
                   fill = 0)    
  setkey(wide_dt, unified_int) # Good practice
  long_dt <- melt(wide_dt,
                  id.vars = "unified_int",       # Column(s) to keep as identifier variables
                  variable.name = "round",       # Name for the new column holding the former column names (1, 2, 3)
                  value.name = "V1",             # Name for the new column holding the values from the melted columns
                  variable.factor = FALSE)       # Keep the 'round' column as character initially (avoids factor issues)
  long_dt$orig <- "NA"
  long_dt$orig[grep("original", long_dt$round)] <- "original"
  long_dt$orig[grep("^sim", long_dt$round)] <- "simulations"
  long_dt$type_int <- "NA"
  long_dt$tosplit <- long_dt$unified_int
  long_dt <- as.data.table(tidyr::separate(long_dt, col = tosplit, into = c("left", "right"), sep = "--"))
  long_dt$type_int <- ifelse( long_dt$left == long_dt$right, "homo", "hetero")
  long_dt$left <- long_dt$right <- NULL
  out <- cellProx.calcP(long_dt, number_of_simulations = 200 )
  return(out)
}

stage3 <- cellProx.subgroup(subset(dt, id %in% c("tumor.1","tumor.2","tumor.3","tumor.4","tumor.5","tumor.6","tumor.7")))
stage1 <- cellProx.subgroup(subset(dt, id %in% c("ctrl.1","ctrl.2","ctrl.3","ctrl.4","ctrl.5") ))
x <- list(stage1 = stage1, stage3=stage3)
saveRDS(x, file="neighborhood.x.sln.clustering2.res0.5.20250411.1611.RDS")  # all data merged 


```

# subgroup by Stage 1 vs Stage 3
```{r interact-with-pairwise-enrichment}
x <- readRDS(file="neighborhood.x.sln.clustering2.res0.5.20250411.1611.RDS")  # all data merged 
stage1 <- x$stage1; stage3 <- x$stage3
md <- pDataDT(sln)
# x <- cellProximityBarplot(sln.sm, CPscore = out,  min_orig_ints = 3, min_sim_ints = 3, return_plot = T)
# ggsave(filename = "cellProxBarplot.pdf", x, height=20, width=7, scale = 2)
md <- as.data.table(md)

addClusterLabels <- function(dt, md)
{
  df <- dt$enrichm_res
  df$tosplit <- df$unified_int
  df <- as.data.table(tidyr::separate(df, col = tosplit, into = c("left", "right"), sep = "--"))
  md$leiden_harmony <- as.character(md$leiden_harmony)
  setkey(md, leiden_harmony)
  df[md, on = .(left = leiden_harmony), left := i.L1_annotations]
  df[md, on = .(right = leiden_harmony), right := i.L1_annotations]
  return(df)
}

stage1$enrichm_res <- addClusterLabels(dt=stage1, md=md)
stage3$enrichm_res <- addClusterLabels(dt=stage3, md=md)
sub.stage1 <- stage1$enrichm_res
sub.stage1$subgroup <- "stage1"
sub.stage3 <- stage3$enrichm_res
sub.stage3$subgroup <- "stage3"
 # plot x=stage 1 PI_value, y= stage 3 PI_value
x <- rbind(sub.stage1, sub.stage3)
x$unified_int <- factor(x$unified_int, levels = sub.stage3$unified_int[order(sub.stage3$PI_value, decreasing = T)])
```


# plots of pairwise enrichment
```{r plot-pairwise-enrichment, fig.width=10}
ggplot(x, aes(x=unified_int, y=PI_value, color=subgroup, label = unified_int)) + geom_point() + theme_classic() + 
  theme(axis.text.x = element_text(angle=45, hjust=1,vjust=1))

y <- merge(x=sub.stage1, y=sub.stage3, by='unified_int', suffixes = c(".stage1",".stage3"), all=T)
y$diff.PI_value <- y$PI_value.stage3 - y$PI_value.stage1
y <- y[!is.na(y$diff.PI_value),]  # remove NA values, which imply interactions not present in both settings
y$unified_int <- factor(y$unified_int, levels = y$unified_int[order(y$diff.PI_value, decreasing = T)])
# y$label <- NA
# y$label[which(abs(y$diff.PI_value )>10)] <- y$unified_int[which(abs(y$diff.PI_value )>10)]
y$unified_char <- paste0(y$left.stage1,"--",y$right.stage1)

y <- y[which(y$original.stage1 > 20),]

ggplot(y, aes(x=unified_int, y=diff.PI_value, label=unified_int)) + geom_point() + theme_classic() + 
  ggrepel::geom_label_repel(max.overlaps = 20) + ggtitle("PI_value:  Stage3 - Stage1") + 
  geom_hline(yintercept = 0, color="grey50") + 
  theme(axis.text.x = element_blank())  

ggplot(y, aes(x=unified_int, y=diff.PI_value, label=unified_char)) + geom_point() + theme_classic() + 
  ggrepel::geom_label_repel(max.overlaps = 20) + ggtitle("PI_value:  Stage3 - Stage1") + 
  geom_hline(yintercept = 0, color="grey50") + 
  theme(axis.text.x = element_blank())  

saveRDS(y, file="neighborhood.y.sln.clustering2.res0.5.20250411.1611.RDS")
```

# show stage 1 interactions
```{r proximity-heatmap-stage1, fig.width=10}
cellProximityHeatmap( sln, CPscore = stage1, scale = TRUE, order_cell_types = TRUE, color_breaks = NULL, color_names = NULL )
cellProximityNetwork(sln, CPscore = stage1)
```

# show stage 3 interactions 
```{r proximity-heatmap-stage3, fig.width=10}
cellProximityHeatmap( sln, CPscore = stage3, scale = TRUE, order_cell_types = TRUE, return_plot = T)
cellProximityNetwork(sln, CPscore = stage3, )

```


# interaction changed genes
```{r Interaction-changed-genes}
gd = fDataDT(sln)

# high_expressed_genes = gd[mean_expr_det > quantile(gd$mean_expr_det)[4]]$feat_ID

## identify genes that are associated with proximity to other cell types
# ICGscoresHighGenes = findInteractionChangedFeats(gobject = sln, selected_feats = gd$feat_ID,
#                                                  spatial_network_name = "knn_network", cluster_column = 'L1_annotations',
#                                                  diff_test = 'permutation', adjust_method = 'fdr',
#                                                  nr_permutations = 200, do_parallel = T)
# ## visualize
# plotCellProximityGenes(sln, cpgObject = ICGscoresHighGenes, method = 'dotplot')


md <- pDataDT(sln); keepCells <- subset(md, cell_ID %in% sample(md$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)

## identify genes that are associated with proximity to other cell types
ICGscoresHighGenes = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "knn_network", cluster_column = 'L1_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 50, do_parallel = F)
df <- ICGscoresHighGenes$ICFscores

# df.2 has 1e5 cells at 100 permutations 
date()
```


Job ID: 63088869
Cluster: slurm_cluster
User/Group: heratr01/heratr01
State: COMPLETED (exit code 0)
Nodes: 1
Cores per node: 5
CPU Utilized: 01:17:41
CPU Efficiency: 23.32% of 05:33:05 core-walltime
Job Wall-clock time: 01:06:37
Memory Utilized: 168.35 GB
Memory Efficiency: 84.17% of 200.00 GB














# original function for cellProximityEnrichment from github
cellProxEnrich.orig <- function(gobject,
    spat_unit = NULL,
    feat_type = NULL,
    spatial_network_name = "Delaunay_network",
    cluster_column,
    number_of_simulations = 1000,
    adjust_method = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ),
    set_seed = TRUE,
    seed_number = 1234) 
{
    # Set feat_type and spat_unit
    spat_unit <- set_default_spat_unit(
        gobject = gobject,
        spat_unit = spat_unit
    )
    feat_type <- set_default_feat_type(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type
    )

    # p.adj test
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))

    spatial_network_annot <- annotateSpatialNetwork(
        gobject = gobject,
        feat_type = feat_type,
        spat_unit = spat_unit,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column
    )

    # remove double edges between same cells #
    # a simplified network does not have double edges between cells #

    # data.table variables
    unified_cells <- type_int <- N <- NULL

    spatial_network_annot <- GiottoUtils:::dt_sort_combine_two_columns(
        spatial_network_annot, "to", "from", "unified_cells"
    )
    spatial_network_annot <- spatial_network_annot[!duplicated(unified_cells)]

    sample_dt <- Giotto:::make_simulated_network(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column,
        number_of_simulations = number_of_simulations,
        set_seed = set_seed,
        seed_number = seed_number
    )

    # combine original and simulated network
    table_sim_results <- sample_dt[, .N, by = c(
        "unified_int", "type_int", "round"
    )]

    ## create complete simulations
    ## add 0 if no single interaction was found
    unique_ints <- unique(table_sim_results[, .(unified_int, type_int)])

    # data.table with 0's for all interactions
    minimum_simulations <- unique_ints[rep(
        seq_len(nrow(unique_ints)), number_of_simulations
    ), ]
    minimum_simulations[, round := rep(
        paste0("sim", seq_len(number_of_simulations)),
        each = nrow(unique_ints)
    )]
    minimum_simulations[, N := 0]

    table_sim_minimum_results <- rbind(table_sim_results, minimum_simulations)
    table_sim_minimum_results[, V1 := sum(N), by = c(
        "unified_int", "type_int", "round"
    )]
    table_sim_minimum_results <- unique(
        table_sim_minimum_results[, .(unified_int, type_int, round, V1)]
    )
    table_sim_results <- table_sim_minimum_results


    # data.table variables
    orig <- unified_int <- V1 <- original <- enrichm <- simulations <- NULL

    table_sim_results[, orig := "simulations"]
    spatial_network_annot[, round := "original"]

    table_orig_results <- spatial_network_annot[, .N, by = c(
        "unified_int", "type_int", "round"
    )]
    table_orig_results[, orig := "original"]
    data.table::setnames(table_orig_results, old = "N", new = "V1")

    table_results <- rbind(table_orig_results, table_sim_results)



    # add missing combinations from original or simulations
    # probably not needed anymore
    all_simulation_ints <- as.character(unique(table_results[
        orig == "simulations"
    ]$unified_int))
    all_original_ints <- as.character(unique(table_results[
        orig == "original"
    ]$unified_int))
    missing_in_original <- all_simulation_ints[
        !all_simulation_ints %in% all_original_ints
    ]
    missing_in_simulations <- all_original_ints[
        !all_original_ints %in% all_simulation_ints
    ]
    create_missing_for_original <- table_results[
        unified_int %in% missing_in_original
    ]
    create_missing_for_original <- unique(create_missing_for_original[
        , c("orig", "V1") := list("original", 0)
    ])
    create_missing_for_simulations <- table_results[
        unified_int %in% missing_in_simulations
    ]
    create_missing_for_simulations <- unique(
        create_missing_for_simulations[, c("orig", "V1") := list(
            "simulations", 0
        )]
    )

    table_results <- do.call(
        "rbind",
        list(
            table_results, create_missing_for_original,
            create_missing_for_simulations
        )
    )


    ## p-values
    combo_list <- rep(NA, length = length(unique(table_results$unified_int)))
    p_high <- rep(NA, length = length(unique(table_results$unified_int)))
    p_low <- rep(NA, length = length(unique(table_results$unified_int)))

    for (int_combo in seq_along(unique(table_results$unified_int))) {
        this_combo <- as.character(unique(table_results$unified_int)[int_combo])

        sub <- table_results[unified_int == this_combo]

        orig_value <- sub[orig == "original"]$V1
        sim_values <- sub[orig == "simulations"]$V1

        length_simulations <- length(sim_values)
        if (length_simulations != number_of_simulations) {
            additional_length_needed <- number_of_simulations -
                length_simulations
            sim_values <- c(sim_values, rep(0, additional_length_needed))
        }

        p_orig_higher <- 1 - (sum((orig_value + 1) > (sim_values + 1)) /
            number_of_simulations)
        p_orig_lower <- 1 - (sum((orig_value + 1) < (sim_values + 1)) /
            number_of_simulations)

        combo_list[[int_combo]] <- this_combo
        p_high[[int_combo]] <- p_orig_higher
        p_low[[int_combo]] <- p_orig_lower
    }
    res_pvalue_DT <- data.table::data.table(
        unified_int = as.vector(combo_list),
        p_higher_orig = p_high,
        p_lower_orig = p_low
    )


    # depletion or enrichment in barplot format
    table_mean_results <- table_results[, .(mean(V1)), by = c(
        "orig", "unified_int", "type_int"
    )]
    table_mean_results_dc <- data.table::dcast.data.table(
        data = table_mean_results, formula = type_int + unified_int ~ orig,
        value.var = "V1"
    )
    table_mean_results_dc[, original := ifelse(is.na(original), 0, original)]
    table_mean_results_dc[, enrichm := log2((original + 1) / (simulations + 1))]


    table_mean_results_dc <- merge(
        table_mean_results_dc, res_pvalue_DT,
        by = "unified_int"
    )
    data.table::setorder(table_mean_results_dc, enrichm)
    table_mean_results_dc[, unified_int := factor(unified_int, unified_int)]

    # adjust p-values for mht

    # data.table variables
    p.adj_higher <- p.adj_lower <- p_lower_orig <- p_higher_orig <-
        PI_value <- int_ranking <- NULL

    table_mean_results_dc[, p.adj_higher := stats::p.adjust(
        p_higher_orig,
        method = sel_adjust_method
    )]
    table_mean_results_dc[, p.adj_lower := stats::p.adjust(
        p_lower_orig,
        method = sel_adjust_method
    )]


    table_mean_results_dc[, PI_value := ifelse(p.adj_higher <= p.adj_lower,
        -log10(p.adj_higher + (1 / number_of_simulations)) * enrichm,
        -log10(p.adj_lower + (1 / number_of_simulations)) * enrichm
    )]
    data.table::setorder(table_mean_results_dc, PI_value)

    # order
    table_mean_results_dc <- table_mean_results_dc[order(-PI_value)]
    table_mean_results_dc[, int_ranking := seq_len(.N)]

    return(list(
        raw_sim_table = table_results,
        enrichm_res = table_mean_results_dc
    ))
}
