---
title: "Neighborhood analyses"
output:
  html_document:
    df_print: paged    
    toc: TRUE
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
---
 

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}
```
 
 
```{r call-libraries} 
date()
library(Giotto)
library(future); if(parallelly::availableCores() > 2 )  # & .Platform$OS.type == "windows")  
  {plan('multisession', workers = parallelly::availableCores()-2)}
options(future.globals.maxSize= 1e12)
library(scran); library(MAST); # library(leiden);   # library(multinet);  
library(data.table)
library(gplots); library(ggplot2); library(viridis)
library(dbscan)
library(dplyr); library(leidenAlg)
library(ComplexHeatmap)
library(circlize)
set.seed(43)
knitr::opts_chunk$set(fig.retina = 2)
sessionInfo()

#  ****** libraries to install 
library(factoextra)

```

# read in data object 
```{r read-annot-sln-spatnet}

sln <- readRDS("sln.clustering5.annot.spatnet.20250617.160512.RDS")
md <- pDataDT(sln)
stableColorsL2 <- getDistinctColors(length(unique(md$L2_annotations)));  names(stableColorsL2) <- unique(md$L2_annotations)
stage3.subgroup <- c("tumor.1","tumor.2","tumor.3","tumor.4","tumor.5","tumor.6","tumor.7", "tumor.8",
                                                 "tumor.9","tumor.10")
stage1.subgroup <- c("ctrl.1","ctrl.2","ctrl.3","ctrl.4","ctrl.5")
recur.subgroup <- c("tumor.3","tumor.4","tumor.5","tumor.7","tumor.8","tumor.10")
norecur.subgroup <- c("tumor.1","tumor.2", "tumor.6")
```


# interaction changed genes
```{r Interaction-changed-genes}
# md <- pDataDT(sln)
keepCells <- subset(md, list_ID %in% c("tumor.1","tumor.2","tumor.3","tumor.4","tumor.5","tumor.6","tumor.7", "tumor.8",
                                                 "tumor.9","tumor.10"))
keepCells <- subset(keepCells, cell_ID %in% sample(keepCells$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)
date()
## identify genes that are associated with proximity to other cell types
ICGscores.L1 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L1_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
ICGscores.L2 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L2_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
saveRDS(object = list(ICGscores.L1 = ICGscores.L1, ICGscores.L2 = ICGscores.L2), 
        file = "neighborhood.ICGlist.s3.sln.clustering5.annot.spatnet.RDS"); date()  # took 6 hours at 300 permutations each 


# md <- pDataDT(sln); 
keepCells <- subset(md, list_ID %in% c("ctrl.1","ctrl.2","ctrl.3","ctrl.4","ctrl.5"))
keepCells <- subset(keepCells, cell_ID %in% sample(keepCells$cell_ID, 3e5))$cell_ID
sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
gd = fDataDT(sln.sm)
date()
## identify genes that are associated with proximity to other cell types
ICGscores.L1 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L1_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
ICGscores.L2 = findInteractionChangedFeats(gobject = sln.sm, selected_feats = gd$feat_ID,
                                                 spatial_network_name = "spatialknn.k5", cluster_column = 'L2_annotations',
                                                 diff_test = 'permutation', adjust_method = 'fdr',
                                                 nr_permutations = 200, do_parallel = F)
saveRDS(object = list(ICGscores.L1 = ICGscores.L1, ICGscores.L2 = ICGscores.L2), 
        file = "neighborhood.ICGlist.s1.sln.clustering5.annot.spatnet.RDS"); date()

```


# calculate proximity enrichment per sample for k=5
```{r custom-cellproxEnrich-function-k5}
cellProx.sim <- function(gobject, spat_unit = NULL, feat_type = NULL, spatial_network_name="spatialknn.k5", 
                         #spatial_network_name = "Delaunay_network",
    cluster_column, number_of_simulations = 1000, 
    adjust_method = c("none", "fdr", "bonferroni", "BH", "holm", "hochberg", "hommel", "BY"), 
    set_seed = TRUE, seed_number = 1234) 
{
    # Set feat_type and spat_unit
    spat_unit <- set_default_spat_unit(
        gobject = gobject,
        spat_unit = spat_unit
    )
    feat_type <- set_default_feat_type(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type
    )

    # p.adj test
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))

    spatial_network_annot <- annotateSpatialNetwork(
        gobject = gobject,
        feat_type = feat_type,
        spat_unit = spat_unit,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column
    )

    # remove double edges between same cells #
    # a simplified network does not have double edges between cells #

    # data.table variables
    unified_cells <- type_int <- N <- NULL

    spatial_network_annot <- GiottoUtils:::dt_sort_combine_two_columns(
        spatial_network_annot, "to", "from", "unified_cells"
    )
    spatial_network_annot <- spatial_network_annot[!duplicated(unified_cells)]

    sample_dt <- Giotto:::make_simulated_network(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column,
        number_of_simulations = number_of_simulations,
        set_seed = set_seed,
        seed_number = seed_number
    )

    # combine original and simulated network
    table_sim_results <- sample_dt[, .N, by = c(
        "unified_int", "type_int", "round"
    )]

    ## create complete simulations
    ## add 0 if no single interaction was found
    unique_ints <- unique(table_sim_results[, .(unified_int, type_int)])

    # data.table with 0's for all interactions
    minimum_simulations <- unique_ints[rep(
        seq_len(nrow(unique_ints)), number_of_simulations
    ), ]
    minimum_simulations[, round := rep(
        paste0("sim", seq_len(number_of_simulations)),
        each = nrow(unique_ints)
    )]
    minimum_simulations[, N := 0]

    table_sim_minimum_results <- rbind(table_sim_results, minimum_simulations)
    table_sim_minimum_results[, V1 := sum(N), by = c(
        "unified_int", "type_int", "round"
    )]
    table_sim_minimum_results <- unique(
        table_sim_minimum_results[, .(unified_int, type_int, round, V1)]
    )
    table_sim_results <- table_sim_minimum_results


    # data.table variables
    orig <- unified_int <- V1 <- original <- enrichm <- simulations <- NULL

    table_sim_results[, orig := "simulations"]
    spatial_network_annot[, round := "original"]

    table_orig_results <- spatial_network_annot[, .N, by = c(
        "unified_int", "type_int", "round"
    )]
    table_orig_results[, orig := "original"]
    data.table::setnames(table_orig_results, old = "N", new = "V1")

    table_results <- rbind(table_orig_results, table_sim_results)



    # add missing combinations from original or simulations
    # probably not needed anymore
    all_simulation_ints <- as.character(unique(table_results[
        orig == "simulations"
    ]$unified_int))
    all_original_ints <- as.character(unique(table_results[
        orig == "original"
    ]$unified_int))
    missing_in_original <- all_simulation_ints[
        !all_simulation_ints %in% all_original_ints
    ]
    missing_in_simulations <- all_original_ints[
        !all_original_ints %in% all_simulation_ints
    ]
    create_missing_for_original <- table_results[
        unified_int %in% missing_in_original
    ]
    create_missing_for_original <- unique(create_missing_for_original[
        , c("orig", "V1") := list("original", 0)
    ])
    create_missing_for_simulations <- table_results[
        unified_int %in% missing_in_simulations
    ]
    create_missing_for_simulations <- unique(
        create_missing_for_simulations[, c("orig", "V1") := list(
            "simulations", 0
        )]
    )
# stop **********************************************************************************************

    table_results <- do.call(
        "rbind",
        list(
            table_results, create_missing_for_original,
            create_missing_for_simulations
        )
    )
    return(table_results)
}

cellProx.calcP <- function(table_results, number_of_simulations = 1000, 
    adjust_method = c("none", "fdr", "bonferroni", "BH", "holm", "hochberg", "hommel", "BY"), 
    set_seed = TRUE, seed_number = 1234)
{
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))
    ## p-values
    combo_list <- rep(NA, length = length(unique(table_results$unified_int)))
    p_high <- rep(NA, length = length(unique(table_results$unified_int)))
    p_low <- rep(NA, length = length(unique(table_results$unified_int)))

    for (int_combo in seq_along(unique(table_results$unified_int))) {
        this_combo <- as.character(unique(table_results$unified_int)[int_combo])

        sub <- table_results[unified_int == this_combo]

        orig_value <- sub[orig == "original"]$V1
        sim_values <- sub[orig == "simulations"]$V1

        length_simulations <- length(sim_values)
        if (length_simulations != number_of_simulations) {
            additional_length_needed <- number_of_simulations -
                length_simulations
            sim_values <- c(sim_values, rep(0, additional_length_needed))
        }

        p_orig_higher <- 1 - (sum((orig_value + 1) > (sim_values + 1)) /
            number_of_simulations)
        p_orig_lower <- 1 - (sum((orig_value + 1) < (sim_values + 1)) /
            number_of_simulations)

        combo_list[[int_combo]] <- this_combo
        p_high[[int_combo]] <- p_orig_higher
        p_low[[int_combo]] <- p_orig_lower
    }
    res_pvalue_DT <- data.table::data.table(
        unified_int = as.vector(combo_list),
        p_higher_orig = p_high,
        p_lower_orig = p_low
    )


    # depletion or enrichment in barplot format
    table_mean_results <- table_results[, .(mean(V1)), by = c(
        "orig", "unified_int", "type_int"
    )]
    table_mean_results_dc <- data.table::dcast.data.table(
        data = table_mean_results, formula = type_int + unified_int ~ orig,
        value.var = "V1"
    )
    table_mean_results_dc[, original := ifelse(is.na(original), 0, original)]
    table_mean_results_dc[, enrichm := log2((original + 1) / (simulations + 1))]


    table_mean_results_dc <- merge(
        table_mean_results_dc, res_pvalue_DT,
        by = "unified_int"
    )
    data.table::setorder(table_mean_results_dc, enrichm)
    table_mean_results_dc[, unified_int := factor(unified_int, unified_int)]

    # adjust p-values for mht

    # data.table variables
    p.adj_higher <- p.adj_lower <- p_lower_orig <- p_higher_orig <-
        PI_value <- int_ranking <- NULL

    table_mean_results_dc[, p.adj_higher := stats::p.adjust(
        p_higher_orig,
        method = sel_adjust_method
    )]
    table_mean_results_dc[, p.adj_lower := stats::p.adjust(
        p_lower_orig,
        method = sel_adjust_method
    )]


    table_mean_results_dc[, PI_value := ifelse(p.adj_higher <= p.adj_lower,
        -log10(p.adj_higher + (1 / number_of_simulations)) * enrichm,
        -log10(p.adj_lower + (1 / number_of_simulations)) * enrichm
    )]
    data.table::setorder(table_mean_results_dc, PI_value)

    # order
    table_mean_results_dc <- table_mean_results_dc[order(-PI_value)]
    table_mean_results_dc[, int_ranking := seq_len(.N)]

    return(list(
        raw_sim_table = table_results,
        enrichm_res = table_mean_results_dc
    ))
}

list_ID <- unique(md$list_ID)
gc()

tableResults.k5 <- list()
annot.k5 <- list()
for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=5, verbose=T, name = "spatialknn.k5.persample")
  annot.k5[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L2_annotations", spatial_network_name = 'spatialknn.k5.persample')
  tableResults.k5[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L2_annotations', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "spatialknn.k5.persample")
}
saveRDS(list(annot.k5=annot.k5, tableResults.k5 = tableResults.k5), file="neighborhood.annot.k5.sln.RDS")  # all data merged
# annot.k5 <- readRDS(file="neighborhood.annot.k5.sln.RDS")[["annot.k5"]]  # all data merged
```


```{r cellproxEnrich-function-k9}
# calculate proximity enrichment per sample for k=9
tableResults.k9 <- list()
annot.k9 <- list()
for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=9, verbose=T, name = "spatialknn.k9.persample") 
  print("made spatialKNN")
  annot.k9[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L2_annotations", spatial_network_name = 'spatialknn.k9.persample')
  tableResults.k9[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L2_annotations', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "spatialknn.k9.persample")
}

saveRDS(list(annot.k9 = annot.k9, tableResults.k9 = tableResults.k9), file="neighborhood.annot.k9.sln.RDS")  # all data merged

tableResults.k9 <- list()
annot.k9 <- list()
for (i in list_ID)
{
  print(i)
  keepCells <- subset(md, list_ID %in% i )$cell_ID
  sln.sm <- subsetGiotto(gobject = sln, cell_ids = keepCells)
  sln.sm <- createSpatialKNNnetwork(sln.sm, k=9, verbose=T, name = "spatialknn.k9.persample") 
  print("made spatialKNN")
  annot.k9[[i]] <- annotateSpatialNetwork(sln.sm, cluster_column = "L3_annotations", spatial_network_name = 'spatialknn.k9.persample')
  tableResults.k9[[i]] <- cellProx.sim(sln.sm, cluster_column = 'L3_annotations', adjust_method = "fdr",
                                       number_of_simulations = 200, spatial_network_name = "spatialknn.k9.persample")
}

saveRDS(list(annot.k9 = annot.k9, tableResults.k9 = tableResults.k9), file="neighborhood.annot.k9.l3.sln.RDS")  # all data merged

annot.k9 <- readRDS(file="neighborhood.annot.k9.sln.RDS")[['annot.k9']]  #  read in L2 annotations 

```

Fails on k=10 on tumor.6  (2 million rows) with long vectors error 
Works for k=9 for all samples 

# subgroup by specified metadata  
```{r custom-cellproxEnrich-function-subgroup}
# subset dt based on groupings of interest 
cellProx.subgroup <- function(dt)
{
  wide_dt <- dcast(dt,                # Your input data.table
                   unified_int ~ round, # Formula: rows ~ columns
                   value.var = "V1",  # Column with values for the new wide cells
                   fun.aggregate = mean, # Function to aggregate values (when multiple rows match a cell)
                   na.rm = TRUE,      # Tell mean() to ignore NA values during aggregation
                   fill = 0)    
  setkey(wide_dt, unified_int) # Good practice
  long_dt <- melt(wide_dt,
                  id.vars = "unified_int",       # Column(s) to keep as identifier variables
                  variable.name = "round",       # Name for the new column holding the former column names (1, 2, 3)
                  value.name = "V1",             # Name for the new column holding the values from the melted columns
                  variable.factor = FALSE)       # Keep the 'round' column as character initially (avoids factor issues)
  long_dt$orig <- "NA"
  long_dt$orig[grep("original", long_dt$round)] <- "original"
  long_dt$orig[grep("^sim", long_dt$round)] <- "simulations"
  long_dt$type_int <- "NA"
  long_dt$tosplit <- long_dt$unified_int
  long_dt <- as.data.table(tidyr::separate(long_dt, col = tosplit, into = c("left", "right"), sep = "--"))
  long_dt$type_int <- ifelse( long_dt$left == long_dt$right, "homo", "hetero")
  long_dt$left <- long_dt$right <- NULL
  out <- cellProx.calcP(long_dt, number_of_simulations = 200 )  # needs to match what was used to calculate sims
  return(out)
}

stage3 <- cellProx.subgroup(subset(dt, id %in% stage3.subgroup))
stage1 <- cellProx.subgroup(subset(dt, id %in% stage1.subgroup ))

x <- list(stage1 = stage1, stage3=stage3)
# saveRDS(x, file="neighborhood.x.sln.RDS")  # all data merged 

recur <- cellProx.subgroup(subset(dt, id %in% c("tumor.3","tumor.4","tumor.5","tumor.7","tumor.8","tumor.10")))
no.recur <- cellProx.subgroup(subset(dt, id %in% c("tumor.1","tumor.2", "tumor.6") ))

sampleList <- unique(md$list_ID)
perSampleInteraction <- list()
for (i in seq_along(sampleList))
{
  print( paste("Interactions for sample", sampleList[i]) )
  perSampleInteraction[[i]] <- cellProx.subgroup(subset(dt, id %in% sampleList[i]))  # format in a per-sample fashion 
  perSampleInteraction[[i]]$name <- subset(dt, id %in% sampleList[i])$id[1]   # carry forward the sample name 
  # perSampleInteraction[[i]]$enrichm_res[, proportion := original/sum(original, na.rm=T) ]  
  names(perSampleInteraction)[i] <-  subset(dt, id %in% sampleList[i])$id[1]
}
saveRDS(perSampleInteraction, file="neighborhood.perSampleInteraction.RDS")  # all data merged 

perSampleInteraction <- readRDS(file="neighborhood.perSampleInteraction.RDS")

perSampleInteraction.enrichm_res <- list()
for (i in seq_along(perSampleInteraction))
{
  perSampleInteraction.enrichm_res[[i]] <- perSampleInteraction[[i]]$enrichm_res  # save the enrichm results table 
  names(perSampleInteraction.enrichm_res)[i] <- names(perSampleInteraction)[i]
}

addLabels.fxn <- function(dt, label)
{
  for (i in seq_along(dt))   {   dt[[i]][, list_ID := names(dt)[i]]   }  # attach the sample name to a new column
  dt <- rbindlist(dt)  # turn the list into data table
  dt$condition <- label  # add in a column for the stage 
  return(dt)
}

interaction.byStage <- rbind(addLabels.fxn(perSampleInteraction.enrichm_res[stage3.subgroup], label = "stage3"),
             addLabels.fxn(perSampleInteraction.enrichm_res[stage1.subgroup], label = "stage1") ) 
interaction.byRecur <- rbind(addLabels.fxn(perSampleInteraction.enrichm_res[recur.subgroup], label = "recur"),
             addLabels.fxn(perSampleInteraction.enrichm_res[norecur.subgroup], label = "no.recur") ) 
```

# finish processing of subgroups at a summary level
```{r interact-with-pairwise-enrichment, eval=FALSE}
x <- readRDS(file="neighborhood.x.sln.RDS")  # all data merged 
stage1 <- x$stage1; stage3 <- x$stage3
md <- pDataDT(sln)
# x <- cellProximityBarplot(sln.sm, CPscore = out,  min_orig_ints = 3, min_sim_ints = 3, return_plot = T)
# ggsave(filename = "cellProxBarplot.pdf", x, height=20, width=7, scale = 2)

### md <- as.data.table(md)   # slated for deletion

addClusterLabels <- function(dt, md)
{
  df <- dt$enrichm_res
  df$tosplit <- df$unified_int
  df <- as.data.table(tidyr::separate(df, col = tosplit, into = c("left", "right"), sep = "--"))
  md$leiden_harmony <- as.character(md$leiden_harmony)
  setkey(md, leiden_harmony)
  df[md, on = .(left = leiden_harmony), left := i.L1_annotations]
  df[md, on = .(right = leiden_harmony), right := i.L1_annotations]
  return(df)
}

stage1$enrichm_res <- addClusterLabels(dt=stage1, md=md)
stage3$enrichm_res <- addClusterLabels(dt=stage3, md=md)
sub.stage1 <- stage1$enrichm_res
sub.stage1$subgroup <- "stage1"
sub.stage3 <- stage3$enrichm_res
sub.stage3$subgroup <- "stage3"
 # plot x=stage 1 PI_value, y= stage 3 PI_value
x <- rbind(sub.stage1, sub.stage3)
x$unified_int <- factor(x$unified_int, levels = sub.stage3$unified_int[order(sub.stage3$PI_value, decreasing = T)])
```


# show the average neighborhood for sln 5 
```{r show-niche-singleSample, fig.height=6}

removeArtifacts <- function(dt)
{
  dt <- dt[!unified_int %like% "ARTIFACT"]
  dt <- dt[!unified_int %like% "outlier"]
  return(dt)
}

showNiche.singleSample <- function(annot.table, indexCluster, title, plots=TRUE)
{
  res <- removeArtifacts(annot.table)
  res <- res[from_cell_type == indexCluster, to_cell_type] %>% table %>% prop.table  # calculate proportions 
  res <- as.data.table(res); colnames(res) <- c('subset','proportion')
  res <- res[order(res$proportion, decreasing=T)]
  stableColorsL2.fxn <- stableColorsL2

  res$subset <- factor(res$subset, levels = res$subset)   
  
  if(plots == T)
  {  
    # lollipop chart 
    print( 
      a <- ggplot(data = res, aes(y = subset, x = proportion, fill=subset, color = subset)) + 
      geom_bar(stat='identity', width=0.01) + 
      geom_point(size=5) + 
      theme_classic() + ggtitle(title) + 
        scale_fill_manual(values = stableColorsL2.fxn) +
        scale_color_manual(values = stableColorsL2.fxn) +
      theme(axis.text.y = element_text( size=6), axis.title.y = element_blank(), plot.title=element_text(size=8)) + 
        theme(legend.position = 'none')  
    )
    
    # stacked bar chart 
    print(
      ggplot(data = res, aes(y = proportion, x = 'blank',  fill = subset)) + 
      geom_col(width=1, position="fill") + 
      ggrepel::geom_text_repel( aes(label = subset ),
        position = position_stack(vjust = 0.5),           # specify stacked position to follow bar layer's default position
        size = 2,  color='black', direction = "y", xlim = c(1.5, NA), hjust = 0,
        segment.linetype = "dotted", box.padding = .4 ) +
      theme_classic() + ggtitle(paste("Neighbors of ", title)) +  
      theme(axis.text.x = element_blank()) +
      theme(axis.title.x = element_blank()) +
      # theme(legend.text = element_text(size=3), legend.title = element_text(size=3)) + 
      theme(legend.position = 'none') +
        # theme( plot.margin = margin(0,200,0,0,  unit="pt")) + 
      scale_fill_manual(values = stableColorsL2.fxn) + scale_color_manual(values = stableColorsL2.fxn) 
    )
  }
  if(plots==T)   return(a)
  else return(res)
}

a <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'LZ_BCELL', title='LZ_BCELL',
                            plots=T)
b <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'MIXED_GC_TCELL', title='MIXED_GC_TCELL',
                            plots=T)
gridExtra::grid.arrange(a,b, nrow=1)

a <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                            title='TCF1_HI_TOX_HI_CD8', plots=T)
b <- showNiche.singleSample(annot.table = annot.k9[['tumor.5']], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                            title='TCF1_LO_TOX_HI_CD8', plots=T)
gridExtra::grid.arrange(a,b, nrow=1) 
  
```



# show the average neighborhood for all sln 
```{r show-niche-multipleSample}
showNiche.multiSample <- function(annot.table, indexCluster, title,  plots=T, parallel=F)
{
  res <- list()
  IDlist <- names(annot.table)
  if(parallel == T)
  {  
    library(future.apply)
    res_list <- future_lapply(IDlist, function(i) {
      single_res <- showNiche.singleSample( annot.table = annot.k9[[i]], indexCluster = indexCluster, title = title, plots = F )
      single_res$list_ID <- i    # name by list_ID
      return(single_res)
    }, future.seed = TRUE) # future.seed = TRUE is crucial for reproducible results!
  }  
  
  if (parallel == F)
  {  
    for (i in IDlist)
    {
      res[[i]] <- showNiche.singleSample(annot.table = annot.k9[[i]], indexCluster = indexCluster,
                                         title=title, plots=F )  # return per-sample proportions
      res[[i]]$list_ID <- i  # name by list_ID
    }
  }
  res <- rbindlist(res)  # merge into one data table 
  ave <-  res %>% group_by(subset) %>% summarise(mean_proportion = mean(proportion, na.rm = TRUE))  # calculate mean over grouping
  ave <- ave[order(ave$mean_proportion, decreasing = T),]  # order by decreasing value 
  res$subset <- factor(res$subset, levels = ave$subset)  # set the res$subset factors based on the decreasing average
  stableColorsL2.fxn <- stableColorsL2

  if(plots == T)
  { 
    # scatter plot of biological replicates
    print( 
      ggplot(res, aes(y = subset, x = proportion, fill=subset, label=list_ID)) +
      geom_point(size=2, position=position_dodge(width=1), pch=21, stroke=0.1, color='grey70') +
      stat_summary( fun.min = "mean", fun.max = "mean", geom = "errorbar",
        width = 0.75, color = "black", linewidth = 0.5 ) + # geom_path(aes(group = list_ID), alpha=0.2, color='grey80') + 
      theme_bw() + ggtitle(title) + 
      # ggrepel::geom_label_repel(size=2, box.padding = unit(0.1,"lines"), alpha=0.2) + 
      scale_fill_manual(values= stableColorsL2.fxn)  + 
      theme(axis.text = element_text(size=6), axis.title.y = element_blank()) + 
      theme(legend.position = 'none')
    )
  }
  return(res)
}  
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'LZ_BCELL', title='LZ_BCELL', parallel=F)
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'MIXED_GC_TCELL', title='MIXED_GC_TCELL')
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'TCF1_HI_TOX_HI_CD8', title='TCF1_HI_TOX_HI_CD8')
showNiche.multiSample(annot.table = annot.k9, indexCluster = 'TCF1_LO_TOX_HI_CD8', title='TCF1_LO_TOX_HI_CD8')

```

```{r plot-average-niche}

plot.Average.Niche.squares <- function( annot.table, indexCluster, title )
{
  # initialize the data table 
  grid_data <- data.table(  
    x = c(1,2,3,3,3,2,1,1,2),
    y = c(3,3,3,2,1,1,1,2,2),
    subset = c("red", "blue", "green", "yellow", "purple", "orange", "pink", "cyan", "brown"),
    proportion = c(0.8, 0.5, 0.7, 0.6, 0.9, 0.4, 0.7, 0.3, 0.8) # Arbitrary proportions for each square
  )
  
  # make the center the index cell 
  res <- showNiche.multiSample(annot.table = annot.table, indexCluster = indexCluster, title=indexCluster, plots = F)
  stableColorsL2.fxn <- stableColorsL2

  ave <- res %>% group_by(subset) %>% summarise(mean_proportion = mean(proportion, na.rm = TRUE))
  ave <- ave[1:9,]
  ave$x[1:9] <- grid_data$x; ave$y[1:9] <- grid_data$y
  grid_data <- as.data.table(ave)
  grid_data[x == 2 & y == 2, subset:= indexCluster]
  grid_data[x == 2 & y == 2, mean_proportion:= 0]

  # grid_data$mean_proportion <- grid_data$mean_proportion 
  grid_data$label <- gsub(pattern = "_", replacement = " ", x = grid_data$subset)
  grid_data$label <- stringr::str_wrap(grid_data$label, 5)
  grid_data$label.annot.x <- c(0.5, 2, 3.5, 3.5, 3.5, 2, 0.5, 0.5, 2)
  grid_data$label.annot.y <- c(3.75, 3.75, 3.75, 2, 0.25, 0.25, 0.25, 2, 2)
  hex.coord <- data.frame(x = c(2.5, 2.25, 1.75, 1.5, 1.75, 2.25), y=c(2, 2.433, 2.433, 2, 1.567, 1.567))
  print(
    ggplot(grid_data[1:8,], aes( x=x, y=y, fill=subset, label=label, size=mean_proportion)) + 
      geom_polygon(data = hex.coord, inherit.aes = F, aes(x = x, y = y), fill = "white", color = "black", size=0.05) +
      geom_point(shape = 22, color = 'gray50') +   
      geom_text(aes(x = label.annot.x, y=label.annot.y, label = label), size=2) + 
      coord_equal() + # Ensure squares are square (base aspect ratio)
      scale_size(range = c(0,20)) + scale_fill_manual(values = stableColorsL2.fxn) + 
      ggtitle(title) + guides(fill = "none")  +#  scale_size_area() + 
      annotate("text", x = 2, y = 2, label = title, colour = "black", fontface = 'bold', size=2) +  
      theme_void() +
      xlim(c(-0.5,4.5)) + ylim(c(-0.5,4.5))
    )
}

plot.Average.Niche.igraph <- function( annot.table, indexCluster, title )
{
  library(ggraph)
  # make the center the index cell 
  res <- showNiche.multiSample(annot.table = annot.table, indexCluster = indexCluster, title=indexCluster, plots = F)
  stableColorsL2.fxn <- stableColorsL2

  ave <- res %>% group_by(subset) %>% summarise(mean_proportion = mean(proportion, na.rm = TRUE))
  ave$subset <- as.character(ave$subset)
  if (indexCluster %in% ave$subset)   {    
    ave$subset[which(ave$subset == indexCluster)] <- paste0(indexCluster, ".")  
    duplColor <- stableColorsL2.fxn[which(names(stableColorsL2.fxn)==indexCluster)]
    names(duplColor) <- paste0(names(duplColor), ".")
    stableColorsL2.fxn <- c(stableColorsL2.fxn, duplColor)
  }
  center_node_df <- data.frame(subset = indexCluster, mean_proportion = 0) 
  setorder(ave, -mean_proportion)  # order ave from biggest to smallest 
  ave <- ave[1:8,]  # keep the top 8 
  grid_data <- as.data.table(bind_rows(center_node_df, ave))  # add the index cell to the data frame 
  grid_data$label <- grid_data$subset  # wrap labels  
  grid_data$label <- gsub(pattern = "_",replacement = " ", x = grid_data$label)
  grid_data$label <- stringr::str_wrap(grid_data$label, 8)  # wrap labels  
  edges <- data.frame(from = indexCluster, to = grid_data$subset[grid_data$subset != indexCluster] )  # new df with directional proportions
  nodes <- data.frame(name = grid_data$subset, label = grid_data$label, size = grid_data$mean_proportion) # map to node size
  graph <- tidygraph::tbl_graph(nodes = nodes, edges = edges, directed = TRUE)  # create an igraph object with nodes and edges 
  print( 
      a <- ggraph(graph, layout = 'dendrogram', circular = TRUE) +
      geom_edge_fan(width = 0.5, color = "gray60", alpha = 0.8) +
      geom_node_point(aes(fill = name, size = size), shape = 21, color = "black", stroke = 0.2) +
      geom_node_text(aes(label = label, x = x * 1.3, y = y * 1.3), repel = F, size = 2, 
                     hjust='outward', vjust='outward') +
      scale_size_continuous( name = "Interaction\nProportion", limits = c(0, 0.5), range = c(0.1, 15), 
                             breaks=c(0, 0.01, 0.05,0.1,0.2,0.3,0.4) ) + 
      scale_fill_manual(values = stableColorsL2.fxn, guide = "none") +
      geom_node_label( aes(label = label, filter = name == indexCluster), size = 2, alpha=1, fontface = 'bold') +
      theme_graph(base_family = 'sans') +
      labs(title = title) +
          xlim(c(-3, 3)) + 
      ylim(c(-3, 3)) +
      coord_fixed()
  )
  return(a)
}
plot.Average.Niche.igraph(annot.table = annot.k9, indexCluster = 'LZ_BCELL', title='LZ_BCELL')
plot.Average.Niche.igraph(annot.table = annot.k9, indexCluster = 'MIXED_GC_TCELL', title='MIXED_GC_TCELL')

# plot.Average.Niche.squares(annot.table = annot.k9, indexCluster = 'LZ_BCELL', title='LZ_BCELL')
# plot.Average.Niche.squares(annot.table = annot.k9, indexCluster = 'MIXED_GC_TCELL', title='MIXED_GC_TCELL')
a <- plot.Average.Niche.igraph(annot.table = annot.k9, indexCluster = 'TCF1_HI_TOX_HI_CD8', title='TCF1_HI_TOX_HI_CD8')
b <- plot.Average.Niche.igraph(annot.table = annot.k9, indexCluster = 'TCF1_LO_TOX_HI_CD8', title='TCF1_LO_TOX_HI_CD8')

```


```{r plot-average-niche-comparisons, fig.width=10, fig.height=5}
gridExtra::grid.arrange(a,b, nrow=1) 
```


```{r plot-average-niche-by-stage, fig.width=10, fig.height=5}
a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 
a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                      title='TCF1_LO_TOX_HI_CD8 stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                      title='TCF1_LO_TOX_HI_CD8 stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 
```

# plot average niche by recur
```{r plot-average-niche-by-recur, fig.width=10, fig.height=5}
a <- plot.Average.Niche.igraph(annot.table = annot.k9[recur.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 recur')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[norecur.subgroup], indexCluster = 'TCF1_HI_TOX_HI_CD8',
                      title='TCF1_HI_TOX_HI_CD8 no recur')
gridExtra::grid.arrange(a,b, nrow=1) 
a <- plot.Average.Niche.igraph(annot.table = annot.k9[recur.subgroup], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                      title='TCF1_LO_TOX_HI_CD8 recur')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[norecur.subgroup], indexCluster = 'TCF1_LO_TOX_HI_CD8',
                      title='TCF1_LO_TOX_HI_CD8 no recur')
gridExtra::grid.arrange(a,b, nrow=1) 
```

# contrast stage 1 vs stage 3 for niche  
```{r contrast-niche-by-stage, plot.height=10}

rgb_to_hex <- function(rgb) {
  sprintf("#%02X%02X%02X", rgb[1], rgb[2], rgb[3])
}

lighten_hex <- function(hex, factor = 0.3) { # You can adjust the factor
  rgb <- col2rgb(hex)
  new_rgb <- round(rgb + (255 - rgb) * factor)
  new_rgb <- pmax(0, pmin(255, new_rgb))
  return(rgb_to_hex(new_rgb))
}


niche.contrastStage <- function( annot.table, indexCluster, common.title)
{
  stage1.res <- showNiche.multiSample(annot.table[stage1.subgroup], indexCluster, title = paste0(common.title,".stage1"),
                                      plots=F)
  stage3.res <- showNiche.multiSample(annot.table[stage3.subgroup], indexCluster, title = paste0(common.title,".stage3"),
                                      plots=F)
  stage1.res$stage <- 'stage1';  stage3.res$stage <- 'stage3'
  res <- rbind(stage1.res, stage3.res)
  res$stage <- factor(res$stage, levels = c("stage3", "stage1"))
  res$subset <- factor(res$subset, levels = levels(stage3.res$subset))
  stableColorsL2.fxn <- stableColorsL2

  # Create a lighter version of the color palette
  lighter_stableColorsL2.fxn <- sapply(stableColorsL2.fxn, lighten_hex, factor = 0.6) # Adjust factor as needed

  # Create a combined color mapping key
  res$color_key <- interaction(res$subset, res$stage, sep = "_")

  # Create the combined color palette
  combined_palette <- c(stableColorsL2.fxn, lighter_stableColorsL2.fxn[names(stableColorsL2.fxn)])
  names(combined_palette) <- c(paste0(names(stableColorsL2.fxn), "_stage3"), paste0(names(lighter_stableColorsL2.fxn), "_stage1"))
  
  annotateX <- max(res$proportion)-0.1*max(res$proportion)
  print(
    ggplot(res, aes(y = subset, x = proportion, fill=color_key, label=list_ID, group=stage)) +
      geom_point(size=2, position=position_dodge(width=1), pch=21, stroke=0.1, color='grey70') +
      annotate('text',x= annotateX ,y=35, label="s1 in light shade\ns3 in dark shade", size=3)+
      stat_summary( fun.min = "mean", fun.max = "mean", geom = "errorbar", position = position_dodge(width=1),
                    width = 1, color = "black", linewidth = 0.5 ) +
      theme_bw() + ggtitle(common.title) +
      scale_fill_manual(name = 'subset & stage', values= combined_palette,  guide = guide_legend(ncol = 2))  +
      theme(axis.text = element_text(size=6), axis.title.y = element_blank()) +
      theme(legend.position = 'none')
  )
  return(res)
}

# ********* hold over for calculating adjusted and unadjusted statistics 
#   aggregated_DT <- data[, .(proportion), by = .(right, list_ID, stage)]
#   interacting.subsets <- unique(aggregated_DT$right)
#   res <- list()
#   for (i in seq_along(interacting.subsets))
#   {
#     data = aggregated_DT[ right %in% interacting.subsets[i], ]
#     res[[i]] <- wilcox.test(proportion ~ stage, data = data)
#     names(res)[i] <- data$right[1]
#   }
#   
#   print(paste("Unadjusted pvalues:")); 
#   print( round(sapply(res, function(x) { x$p.value }), digits = 2) )
#   print(paste("FDR-adjusted pvalues"));  
#   print( round(p.adjust( sapply(res, function(x) { x$p.value }), method="fdr"), digits = 2)) 
# }

niche.contrastStage( annot.k9, indexCluster = 'LZ_BCELL', common.title='LZ_BCELL')
niche.contrastStage( annot.k9, indexCluster = 'MIXED_GC_TCELL', common.title='MIXED_GC_TCELL')
niche.contrastStage( annot.k9, indexCluster = 'TCF1_HI_TOX_HI_CD8', common.title='TCF1_HI_TOX_HI_CD8')
niche.contrastStage( annot.k9, indexCluster = 'TCF1_LO_TOX_HI_CD8', common.title='TCF1_LO_TOX_HI_CD8')

```

# contrast stage using cellProxEnrich
```{r contrast-stage-using-cellProxEnrich}
working.dt <- interaction.byStage

delta.pairwise.enrichment<- working.dt %>% group_by(unified_int, condition) %>% summarise(mean = mean(enrichm, na.rm = TRUE), 
                                                                                          sd = sd(enrichm, na.rm = TRUE), 
                                                                                          n_replicates = n()) %>%  ungroup()
range_of_means <- delta.pairwise.enrichment %>%
  group_by(unified_int) %>% summarise( range_mean = max(mean, na.rm = TRUE) - min(mean, na.rm = TRUE), n_stages = n_distinct(condition)) 
delta.pairwise.enrichment <- range_of_means[order(range_of_means$range_mean, decreasing = T),]
delta.pairwise.enrichment$unified_int <- with(delta.pairwise.enrichment, reorder(unified_int, range_mean))
delta.pairwise.enrichment$rownum <- seq(1:nrow(delta.pairwise.enrichment))
delta.pairwise.enrichment <- as.data.table(delta.pairwise.enrichment)
delta.pairwise.enrichment[ !unified_int %like% "Tumor"][1:100]
ggplot(delta.pairwise.enrichment, aes(x = rownum, y=range_mean, label=unified_int)) + 
  geom_point(size=2, pch=21, color="white", size=0.25, alpha=0.5, fill="black") + 
  ggrepel::geom_text_repel(size=2.5, max.overlaps = 10, segment.color = 'gray60', force=5) + 
  ggtitle("Changed enrichm between stage 1 and 3 - all results") + theme_classic() + 
  theme(axis.text.x = element_blank()) 
ggplot(delta.pairwise.enrichment[1:100,], aes(x = rownum, y=range_mean, label=unified_int)) + 
  geom_point(size=2, pch=21, color="white", size=0.25, alpha=0.5, fill="black") + 
  ggrepel::geom_text_repel(size=2.5, max.overlaps = 7, segment.color = 'gray60', force=20 ) + 
  ggtitle("Changed enrichm between stage 1 and 3 - first 100") + theme_classic() + 
  theme(axis.text.x = element_blank()) 
ggplot(delta.pairwise.enrichment[ !unified_int %like% "Tumor"][1:100], aes(x = rownum, y=range_mean, label=unified_int)) + 
  geom_point(size=2, pch=21, color="white", size=0.25, alpha=0.5, fill="black") + 
  ggrepel::geom_text_repel(size=2.5, max.overlaps = 7, segment.color = 'gray60', force=20 ) + 
  ggtitle("Changed enrichm between stage 1 and 3 - exclude tumor") + theme_classic() + 
  theme(axis.text.x = element_blank()) 


```


# contrast recurrence using cellProxEnrich
```{r contrast-recur-using-cellProxEnrich, fig.width=10, fig.height=6}

working.dt <- interaction.byRecur

delta.pairwise.enrichment<- working.dt %>% group_by(unified_int, condition) %>% summarise(mean = mean(enrichm, na.rm = TRUE), 
                                                                                          sd = sd(enrichm, na.rm = TRUE), 
                                                                                          n_replicates = n()) %>%  ungroup()
range_of_means <- delta.pairwise.enrichment %>%
  group_by(unified_int) %>% summarise( range_mean = max(mean, na.rm = TRUE) - min(mean, na.rm = TRUE), n_stages = n_distinct(condition)) 
delta.pairwise.enrichment <- range_of_means[order(range_of_means$range_mean, decreasing = T),]
delta.pairwise.enrichment$unified_int <- with(delta.pairwise.enrichment, reorder(unified_int, range_mean))
delta.pairwise.enrichment$rownum <- seq(1:nrow(delta.pairwise.enrichment))
delta.pairwise.enrichment <- as.data.table(delta.pairwise.enrichment)
delta.pairwise.enrichment <- delta.pairwise.enrichment[ !unified_int %like% "GC_TCELL"]  # leave out the GC-T cell population
delta.pairwise.enrichment <- removeArtifacts(delta.pairwise.enrichment)
delta.pairwise.enrichment  # leave out the GC-T cell population

ggplot(delta.pairwise.enrichment, aes(x = rownum, y=range_mean, label=unified_int)) + 
  geom_point(size=2, pch=21, color="white", size=0.25, alpha=0.5, fill="black") + 
  ggrepel::geom_text_repel(size=2.5, max.overlaps = 10, segment.color = 'gray60', force=5) + 
  ggtitle("Changed enrichm between recur vs norecur - all results") + theme_classic() + 
  theme(axis.text.x = element_blank()) 
ggplot(delta.pairwise.enrichment[1:100,], aes(x = rownum, y=range_mean, label=unified_int)) + 
  geom_point(size=2, pch=21, color="white", size=0.25, alpha=0.5, fill="black") + 
  ggrepel::geom_text_repel(size=2.5, max.overlaps = 7, segment.color = 'gray60', force=20 ) + 
  ggtitle("Changed enrichm between recur vs norecur - first 100") + theme_classic() + 
  theme(axis.text.x = element_blank()) 
ggplot(delta.pairwise.enrichment[ !unified_int %like% "Tumor"][1:100], aes(x = rownum, y=range_mean, label=unified_int)) + 
  geom_point(size=2, pch=21, color="white", size=0.25, alpha=0.5, fill="black") + 
  ggrepel::geom_text_repel(size=2.5, max.overlaps = 7, segment.color = 'gray60', force=20 ) + 
  ggtitle("Changed enrichm between recur vs norecur - exclude tumor") + theme_classic() + 
  theme(axis.text.x = element_blank()) 

gridExtra::grid.arrange(plot.Average.Niche.igraph(annot.table = annot.k9[norecur.subgroup], 
                                                  indexCluster = 'TCELL_BCELL_MYELOID_MIXED',
                      title='TCELL_BCELL_MYELOID_MIXED no.recur'),
                        plot.Average.Niche.igraph(annot.table = annot.k9[recur.subgroup], 
                                                  indexCluster = 'TCELL_BCELL_MYELOID_MIXED',
                      title='TCELL_BCELL_MYELOID_MIXED recur')
                      , nrow=1) 
```


# activated cd4 t cells by stage
```{r explore-activated-cd4, fig.width=10, fig.height=6}
niche.contrastStage( annot.k9, indexCluster = 'ACTIVATED_MEMORY_CD4', common.title='ACTIVATED_MEMORY_CD4')
delta.pairwise.enrichment[unified_int %like% "ACTIVATED_MEMORY_CD4"]

a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'ACTIVATED_MEMORY_CD4',
                      title='ACTIVATED_MEMORY_CD4 stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'ACTIVATED_MEMORY_CD4',
                      title='ACTIVATED_MEMORY_CD4 stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 


```

# activated cd8 t cells by stage
```{r explore-activated-cd8, fig.width=10, fig.height=6}
niche.contrastStage( annot.k9, indexCluster = 'ACTIVATED_MEMORY_CD8', common.title='ACTIVATED_MEMORY_CD8')
delta.pairwise.enrichment[unified_int %like% "ACTIVATED_MEMORY_CD8"]

a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'ACTIVATED_MEMORY_CD8',
                      title='ACTIVATED_MEMORY_CD8 stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'ACTIVATED_MEMORY_CD8',
                      title='ACTIVATED_MEMORY_CD8 stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 


```

# Treg-ICOSlo by stage
```{r explore-treg, fig.width=10, fig.height=6}
niche.contrastStage( annot.k9, indexCluster = 'ICOS_LO_TREG', common.title='ICOS_LO_TREG')
delta.pairwise.enrichment[unified_int %like% "ICOS_LO_TREG"]

a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'ICOS_LO_TREG',
                      title='ICOS_LO_TREG stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'ICOS_LO_TREG',
                      title='ICOS_LO_TREG stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 

```

# Treg-Activated by stage
```{r explore-tregActivated, fig.width=10, fig.height=6}
niche.contrastStage( annot.k9, indexCluster = 'ACTIVATED_TREG', common.title='ACTIVATED_TREG')
delta.pairwise.enrichment[unified_int %like% "ACTIVATED_TREG"]

a <- plot.Average.Niche.igraph(annot.table = annot.k9[stage1.subgroup], indexCluster = 'ACTIVATED_TREG',
                      title='ACTIVATED_TREG stage 1')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[stage3.subgroup], indexCluster = 'ACTIVATED_TREG',
                      title='ACTIVATED_TREG stage 3')
gridExtra::grid.arrange(a,b, nrow=1) 

```


 
# MIXED_GC_TCELL by recur
```{r explore-activated-cd4-by-recur, fig.width=10, fig.height=6}
niche.contrastStage( annot.k9, indexCluster = 'MIXED_GC_TCELL', common.title='MIXED_GC_TCELL')
delta.pairwise.enrichment[unified_int %like% "MIXED_GC_TCELL"]

a <- plot.Average.Niche.igraph(annot.table = annot.k9[recur.subgroup], indexCluster = 'MIXED_GC_TCELL',
                      title='MIXED_GC_TCELL recur')
b <- plot.Average.Niche.igraph(annot.table = annot.k9[norecur.subgroup], indexCluster = 'MIXED_GC_TCELL',
                      title='MIXED_GC_TCELL norecur')
gridExtra::grid.arrange(a,b, nrow=1) 


```

# locate cells that exist in a given neighborhood 
```{r finding-example-cells-for-neighborhood}
defineNeighborhoods <- function( annot.list)
{
  list_ID <- names(annot.list)
  res <- list()
  copymatrix <- list()
  for (i in list_ID)
  { 
    print(paste0("pass 1: ", i))
    neighborhood.matrix <- annot.list[[i]][, .N, by = .(from, to_cell_type, from_cell_type)] %>% 
      dcast(from + from_cell_type ~ to_cell_type, value.var= 'N', fill=0)  # count the cell types and make a wide data frame 
    neighborhood.matrix[, c('ARTIFACT', 'CD4.outlier', 'CD8.outlier'):= NULL]
    copymatrix[[i]] <- neighborhood.matrix
  }
  for (i in list_ID)
  { 
    print(paste0("pass 2: ", i))
    working.dt <- copy(copymatrix[[i]])
    id_cols <- c("from", "from_cell_type")
    count_cols <- setdiff(names(working.dt), id_cols)
    
    # 2. Calculate the total for each row and add it as a new column
    working.dt[, row_total := rowSums(.SD), .SDcols = count_cols]
    
    # 3. Divide each count column by the row_total to get the proportion
    #    We loop through the count columns and update them "by reference"
    working.dt[, (count_cols) := lapply(.SD, function(x) x / row_total), .SDcols = count_cols]
    
    # 4. (Optional) Remove the temporary 'row_total' column
    working.dt[, row_total := NULL]
    res[[i]] <- working.dt
  } 
  return( list(prop.table=res, raw.table=copymatrix, list_ID=list_ID)) 
}

neighborhoods <- defineNeighborhoods(annot.k9)
gc()

exampleNeighborhoods.filter <- function(res.list, indexCluster, chooseNeighbor, threshold)
{  
  filterList <- list()
  list_ID <- res.list[['list_ID']]
  res.list <- res.list[['prop.table']]
  for (i in list_ID)    { filterList[[i]] <- res.list[[i]][from_cell_type == indexCluster & get(chooseNeighbor) > threshold]    }
  return( rbindlist(filterList)  )
}
  
a <- exampleNeighborhoods.filter(neighborhoods, indexCluster = "LZ_BCELL", chooseNeighbor = "LZ_BCELL", threshold=0.2)
# write.csv(a, file = "LZ_BCELL.filtered.by.LZ_BCELL_over0.2.csv")
a <- exampleNeighborhoods.filter(neighborhoods, indexCluster = "LZ_BCELL", chooseNeighbor = "PC", threshold=0.2)
# write.csv(a, file = "LZ_BCELL.filtered.by.PC_over0.2.csv")


```

# cluster neighborhoods by kmeans

The goal is to see patterns in the neighborhoods around individual cells. Could consider this at two different levels. 
1. there might be neighborhood patterns that are irrespective of what cell is in the middle of the neighborhood
  though this is of unclear utility. it's a feature arising from the data but interpretation is hard. will take a long time to calculate
2. neighborhood patterns might be more interesting to consider at level of individual cell type.  
```{r neighborhoods-by-kmeans-global, eval=F}
# matrix.data = (neighborhoods[['raw.table']][['ctrl.1']])[, 3:35]
# kvalues <- 1:20
# wss_values <- sapply(kvalues, function(k) {  kmeans(matrix.data, centers = k)$tot.withinss   })
# 
# # Create a data frame for plotting
# elbow_data <- data.frame(k = kvalues, wss = wss_values )
# 
# # Plot the elbow curve using ggplot2
# ggplot(elbow_data, aes(x = k, y = wss)) +
#   geom_line(linewidth = 1) + geom_point(size = 3) + 
#   labs( title = "Elbow Method for Optimal k", x = "Number of Clusters (k)", y = "Total Within-Cluster Sum of Squares (WSS)") +
#   theme_minimal()


# kmeans.res.global <- kmeans( x = rbindlist(neighborhoods[['raw.table']])[, 3:35], centers=7, nstart=25)
matrix.data = rbindlist(neighborhoods[['raw.table']])[from_cell_type == 'LZ_BCELL', 3:35]

# factoextra::fviz_nbclust(matrix.data[sample(nrow(matrix.data), size = 1e4, replace = F)], kmeans, method = "wss") + 
#   labs(subtitle = "Elbow plot")  # doen't work for 300k cells
# kmeans.res.LZB <- kmeans( x = matrix.data, centers=7, nstart=25)

# factoextra::fviz_cluster(object = kmeans.res.LZB, data = matrix.data, geom = "point", ellipse.type = "convex", ggtheme = theme_bw() )  # looks crummy, not well separated at all on PCA plot 
umap.out <- uwot::umap(matrix.data[sample(.N, size = 3e4)], n_neighbors = 7)
# plot.df <- as.data.table(cbind(umap.out, as.factor(kmeans.res.LZB$cluster)))
plot.df <- as.data.table(umap.out)
names(plot.df)[1:2] <- c('UMAP1', 'UMAP2')
# ggplot(plot.df[sample(nrow(plot.df),size=5e4, replace=F)], aes(x = UMAP1, y = UMAP2, fill = kmeans.cluster)) + scale_fill_viridis_b() + 
#   geom_point(size = 1, alpha = 0.3, pch=21, color='grey70') + labs(title = "UMAP Plot of K-Means Clusters",color = "Cluster" ) + theme_bw()
# i guess kmeans is wrong for this kind of data 
 
plot.df$db.clusters <- dbscan(plot.df[,1:2], eps = 1.2, minPts = 20)$cluster
ggplot(plot.df[sample(.N, size=1e4)], aes(x = UMAP1, y = UMAP2, fill = db.clusters)) + scale_fill_viridis_b() + 
  geom_density_2d(color='grey60', alpha=0.5) + 
  geom_point(size = 1, alpha = 0.3, pch=21, color='grey70') + labs(title = "UMAP Plot of Clusters", color = "Cluster" ) + theme_bw()

nngraph <- dbscan::sNN(matrix.data, k = 15)
# Convert to an igraph object
graph_obj <- igraph::graph_from_edgelist(as.matrix( nngraph$id), directed = FALSE)

# 2. Run Leiden clustering
leiden_clusters <- leidenAlg::leiden.community(nngraph) # Tune resolution


```


# plot as individual biological replicates 
```{r show-average-neighborhood-contrast-biolReplicates, eval=FALSE}
significant.Neighbors.acrossStage <- function( data, interacting.partner)
{
  data <- as.data.table(tidyr::separate(data, col = unified_int, into = c("left", "right"), sep = "--"))
  data <- data[left == interacting.partner] 
  data[, proportion := original/sum(original), by=.(stage, list_ID)]
  
  print( 
    ggplot(data, aes(y = right, x = proportion, group = stage, fill=stage, label=list_ID)) +
      geom_point(size=3, position=position_dodge(width=1), pch=21, stroke=0.1, color='white') +
      theme_classic() + ggtitle(paste("Neighbors of ", interacting.partner)) + 
      ggrepel::geom_label_repel(size=2, box.padding = unit(0.1,"lines"), alpha=0.2) + 
      scale_fill_manual(values= ( RColorBrewer::brewer.pal(n = 4, name = "Paired")) )  + 
      theme(axis.text.x = element_text(angle=45, hjust=1,vjust=1))
  )
  aggregated_DT <- data[, .(proportion), by = .(right, list_ID, stage)]
  interacting.subsets <- unique(aggregated_DT$right)
  res <- list()
  for (i in seq_along(interacting.subsets))
  {
    data = aggregated_DT[ right %in% interacting.subsets[i], ]
    res[[i]] <- wilcox.test(proportion ~ stage, data = data)
    names(res)[i] <- data$right[1]
  }
  
  print(paste("Unadjusted pvalues:")); 
  print( round(sapply(res, function(x) { x$p.value }), digits = 2) )
  print(paste("FDR-adjusted pvalues"));  
  print( round(p.adjust( sapply(res, function(x) { x$p.value }), method="fdr"), digits = 2)) 
}

significant.Neighbors.acrossStage(data = interaction.byStage, interacting.partner = 'TCF1_HI_TOX_HI_CD8')


```


# show a radial plot of contrast of neighborhoods 
```{r show-average-neighborhood-contrast-stage3vs1, eval = FALSE}
intxn.index.s3 <- y[ , .(left.stage3, right.stage3, original.stage3) ]
tcfloCD8.s3 <- filterIntxn(simpleIntxnTable = intxn.index.s3[left.stage3 == 'TCF1_LO_TOX_HI_CD8'], title='CD8+Tox+Tcf-')
tcfloCD8.s3$left.stage3 <- paste0(tcfloCD8.s3$left.stage3, ".s3")
names(tcfloCD8.s1) <- names(tcfloCD8.s3)

comb <- rbind(tcfloCD8.s1, tcfloCD8.s3)

ggplot(comb, aes(y = right.stage3, x = proportion, group=left.stage3, fill=left.stage3)) + 
  geom_bar( stat='identity', width=0.5, position=position_dodge(preserve='single')) + 
  coord_radial("y", start = 0, direction = 1, rotate.angle=T, r_axis_inside = TRUE)+
  ggtitle(paste("Tcflo CD8 T cells - Stage 3 vs 1"))  +  
  scale_fill_manual(values= ( RColorBrewer::brewer.pal(n = 4, name = "Paired")) )  + 
  theme(   axis.title = element_blank(),axis.ticks = element_blank() , axis.text.y = element_text(color = "grey"),
           axis.text.x = element_text(color = "black", size = 4), legend.position = "bottom", legend.title = element_blank(), 
           plot.title = element_text(face = "bold", size = 18, hjust = 1), 
           panel.background = element_rect(fill = "white", color = "white"), panel.grid = element_blank(),  
           panel.grid.major.x = element_line(color="grey90", size=0.5)) 


```


```{r plot-pairwise-enrichment-recurrence, fig.width=10}
y <- merge(x=df[which(df$subgroup=='no.recur'),], y=df[which(df$subgroup == 'recur'),] , by='unified_int', suffixes = c(".no.recur",".recur"), all=T)
y$diff.PI_value <- y$PI_value.recur - y$PI_value.no.recur
y <- y[!is.na(y$diff.PI_value),]  # remove NA values, which imply interactions not present in both settings
y$unified_int <- factor(y$unified_int, levels = y$unified_int[order(y$diff.PI_value, decreasing = T)])
y$unified_char <- paste0(y$left.recur,"--",y$right.recur)

y <- y[which(y$original.recur > 20),]

ggplot(y, aes(x=unified_int, y=diff.PI_value, label=unified_char)) + geom_point() + theme_classic() + 
  ggrepel::geom_label_repel(max.overlaps = 20) + ggtitle("PI_value:  Recur - noRecur") + 
  geom_hline(yintercept = 0.001, color="grey50") + 
  theme(axis.text.x = element_blank())  

```

# show stage 1 interactions
```{r proximity-heatmap-stage1, fig.width=10}
cellProximityHeatmap( sln, CPscore = stage1, scale = TRUE, order_cell_types = TRUE, color_breaks = NULL, color_names = NULL, )
cellProximityNetwork(sln, CPscore = stage1)

cellProximityHeatmap.retMatrix <- function(gobject, CPscore, scale = TRUE, order_cell_types = TRUE, color_breaks = NULL,
                                           color_names = NULL, show_plot = NULL, return_plot = NULL, save_plot = NULL, 
                                           save_param = list(), default_save_name = "cellProximityHeatmap") 
{
    enrich_res <- CPscore$enrichm_res

    # data.table variables
    first_type <- second_type <- unified_int <- NULL

    enrich_res[, first_type := strsplit( x = as.character(unified_int), split = "--"    )[[1]][1],by = seq_len(nrow(enrich_res))]
    enrich_res[, second_type := strsplit( x = as.character(unified_int), split = "--" )[[1]][2],by = seq_len(nrow(enrich_res)) ]

    # create matrix
    enrich_mat <- data.table::dcast.data.table( data = enrich_res, formula = first_type ~ second_type, value.var = "enrichm")
    matrix_d <- as.matrix(enrich_mat[, -1])
    rownames(matrix_d) <- as.vector(enrich_mat[[1]])
    t_matrix_d <- t_flex(matrix_d)

    # fill in NAs based on values in upper and lower matrix triangle
    t_matrix_d[upper.tri(t_matrix_d)][is.na(t_matrix_d[
        upper.tri(t_matrix_d)
    ])] <- matrix_d[upper.tri(matrix_d)][
        is.na(t_matrix_d[upper.tri(t_matrix_d)])
    ]
    t_matrix_d[lower.tri(t_matrix_d)][is.na(t_matrix_d[
        lower.tri(t_matrix_d)
    ])] <- matrix_d[lower.tri(matrix_d)][
        is.na(t_matrix_d[lower.tri(t_matrix_d)])
    ]
    t_matrix_d[is.na(t_matrix_d)] <- 0
    final_matrix <- t_matrix_d

    # scale data
    if (scale == TRUE) {
        final_matrix <- t_flex(scale(t_flex(final_matrix)))
        final_matrix <- t_flex(final_matrix)
        final_matrix[lower.tri(final_matrix)] <- t_flex(final_matrix)[
            lower.tri(final_matrix)
        ]
    }

    # order cell types
    if (order_cell_types == TRUE) {
        cordist <- stats::as.dist(1 - cor_flex(final_matrix))
        clus <- stats::hclust(cordist)
        myorder <- clus$order
        mylabels <- clus$labels
        names(mylabels) <- seq_along(mylabels)
        sample_order <- mylabels[myorder]

        final_matrix <- final_matrix[sample_order, sample_order]
    }

    # create custom colors or not
    if (!is.null(color_breaks) & !is.null(color_names)) {
        if (length(color_breaks) != 3 | !is.numeric(color_breaks)) {
            stop("color_breaks needs to be a numerical vector of length 3")
        }

        if (length(color_names) != 3 | !is.character(color_names)) {
            stop("color_names needs to be a character vector of length 3")
        }

        heatm <- ComplexHeatmap::Heatmap(matrix = final_matrix, cluster_rows = F, cluster_columns = F,
                                         col = GiottoVisuals::colorRamp2( breaks = color_breaks, colors = color_names) )
    } else {
        heatm <- ComplexHeatmap::Heatmap( matrix = final_matrix, cluster_rows = FALSE, cluster_columns = FALSE )
    }

    return(  list(final_matrix=final_matrix, clus=clus) )
}


```

# show stage 3 interactions 
```{r proximity-heatmap-stage3, fig.width=10}
Giotto::cellProximityHeatmap( sln, CPscore = stage3, scale = F, order_cell_types = TRUE, return_plot = T)
final_matrix <- cellProximityHeatmap.retMatrix( sln, CPscore = stage3, scale = TRUE, order_cell_types = T)$final_matrix
clus <- cellProximityHeatmap.retMatrix( sln, CPscore = stage3, scale = TRUE, order_cell_types = T)$clus
plot(clus)
ComplexHeatmap::Heatmap(final_matrix)

cellProximityNetwork(sln, CPscore = stage3 )
write.csv(stage3$enrichm_res, file = "cellProximityHeatmap.stage3.csv")
```




```{r Interaction-changed-genes-byCircosPlot, fig.width=10, fig.retina=2, eval=FALSE}


df.icg <- readRDS("neighborhood.ICGlist.s3.sln.clustering5.annot.spatnet.RDS")[[2]][['ICFscores']] # ICGscores.L2
df.icg.fil <- df.icg[ p.adj<0.01 & cell_type == 'CD8TOXHI_TCF1_HI_TOX_HI_CD8']

pair_counts <- df.icg[p.adj < 0.1, .N, by = .(cell_type, int_cell_type)] # .N is a special data.table symbol that gives the number of rows in each group
pair_counts <- pair_counts[N > 1]
single.space <- c(rep(1, length(unique(pair_counts$cell_type)))); names(single.space) <- unique(pair_counts$cell_type)
grid.col = getDistinctColors(n = length(unique(pair_counts$cell_type))); names(grid.col) <- names(single.space)
# circos.par(gap.after = single.space)
circlize::chordDiagram(pair_counts[1:30], direction.type = c('arrows','diffHeight'), directional=-1, transparency = 0,
                       grid.col = grid.col, diffHeight = -mm_h(2), 
                       annotationTrack = "grid", preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(pair_counts))))),
                       # big.gap=20
                       # link.arr.type = "big.arrow"
                       ) # negative direction to show direction of the effect, e.g. target is modified for x many features because of source

# we go back to the first track and customize sector labels
circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
}, bg.border = NA) # here set bg.border to NA is important
title("interaction changed features - first 30")

```

# stop evaluation here 
```{r Interaction-changed-genes-multitrackCircos, eval=FALSE}
# what about showing a multi-track chordDiagram where the proteins changed the most are indicated 
# would have to filter the protein list to exclude things like DAPI which aren't meaningful 
chordDiagram(pair_counts[1:30], direction.type = c('arrows','diffHeight'), directional=-1, transparency = 0,
                       grid.col = grid.col, diffHeight = -mm_h(2), 
                       # link.arr.type = "big.arrow"
                       ) # negative direction to show direction of the effect, e.g. target is modified for x many features because of source
circos.track(track.index = 1, panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
              facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
  }, bg.border = NA) # here set bg.border to NA is important

prots <- data.table(fDataDT(sln)$feat_ID)
prots <- prots[ !V1 %in% c("DAPI-01", "IL21", "CXCR3", "CD45", "TBET", "MART1", "GP100", "PAN.CYTO","MAC2/GAL3","VIMENTIN") ]
df.icg.fil <- df.icg.fil[ feats %in% prots$V1]



date()

```




Job ID: 63949935
Cluster: slurm_cluster
User/Group: heratr01/heratr01
State: COMPLETED (exit code 0)
Nodes: 1
Cores per node: 10
CPU Utilized: 02:56:58
CPU Efficiency: 12.36% of 23:51:20 core-walltime
Job Wall-clock time: 02:23:08
Memory Utilized: 329.74 GB
Memory Efficiency: 65.95% of 500.00 GB











# original function for cellProximityEnrichment from github
cellProxEnrich.orig <- function(gobject,
    spat_unit = NULL,
    feat_type = NULL,
    spatial_network_name = "Delaunay_network",
    cluster_column,
    number_of_simulations = 1000,
    adjust_method = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ),
    set_seed = TRUE,
    seed_number = 1234) 
{
    # Set feat_type and spat_unit
    spat_unit <- set_default_spat_unit(
        gobject = gobject,
        spat_unit = spat_unit
    )
    feat_type <- set_default_feat_type(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type
    )

    # p.adj test
    sel_adjust_method <- match.arg(adjust_method, choices = c(
        "none", "fdr", "bonferroni", "BH",
        "holm", "hochberg", "hommel",
        "BY"
    ))

    spatial_network_annot <- annotateSpatialNetwork(
        gobject = gobject,
        feat_type = feat_type,
        spat_unit = spat_unit,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column
    )

    # remove double edges between same cells #
    # a simplified network does not have double edges between cells #

    # data.table variables
    unified_cells <- type_int <- N <- NULL

    spatial_network_annot <- GiottoUtils:::dt_sort_combine_two_columns(
        spatial_network_annot, "to", "from", "unified_cells"
    )
    spatial_network_annot <- spatial_network_annot[!duplicated(unified_cells)]

    sample_dt <- Giotto:::make_simulated_network(
        gobject = gobject,
        spat_unit = spat_unit,
        feat_type = feat_type,
        spatial_network_name = spatial_network_name,
        cluster_column = cluster_column,
        number_of_simulations = number_of_simulations,
        set_seed = set_seed,
        seed_number = seed_number
    )

    # combine original and simulated network
    table_sim_results <- sample_dt[, .N, by = c(
        "unified_int", "type_int", "round"
    )]

    ## create complete simulations
    ## add 0 if no single interaction was found
    unique_ints <- unique(table_sim_results[, .(unified_int, type_int)])

    # data.table with 0's for all interactions
    minimum_simulations <- unique_ints[rep(
        seq_len(nrow(unique_ints)), number_of_simulations
    ), ]
    minimum_simulations[, round := rep(
        paste0("sim", seq_len(number_of_simulations)),
        each = nrow(unique_ints)
    )]
    minimum_simulations[, N := 0]

    table_sim_minimum_results <- rbind(table_sim_results, minimum_simulations)
    table_sim_minimum_results[, V1 := sum(N), by = c(
        "unified_int", "type_int", "round"
    )]
    table_sim_minimum_results <- unique(
        table_sim_minimum_results[, .(unified_int, type_int, round, V1)]
    )
    table_sim_results <- table_sim_minimum_results


    # data.table variables
    orig <- unified_int <- V1 <- original <- enrichm <- simulations <- NULL

    table_sim_results[, orig := "simulations"]
    spatial_network_annot[, round := "original"]

    table_orig_results <- spatial_network_annot[, .N, by = c(
        "unified_int", "type_int", "round"
    )]
    table_orig_results[, orig := "original"]
    data.table::setnames(table_orig_results, old = "N", new = "V1")

    table_results <- rbind(table_orig_results, table_sim_results)



    # add missing combinations from original or simulations
    # probably not needed anymore
    all_simulation_ints <- as.character(unique(table_results[
        orig == "simulations"
    ]$unified_int))
    all_original_ints <- as.character(unique(table_results[
        orig == "original"
    ]$unified_int))
    missing_in_original <- all_simulation_ints[
        !all_simulation_ints %in% all_original_ints
    ]
    missing_in_simulations <- all_original_ints[
        !all_original_ints %in% all_simulation_ints
    ]
    create_missing_for_original <- table_results[
        unified_int %in% missing_in_original
    ]
    create_missing_for_original <- unique(create_missing_for_original[
        , c("orig", "V1") := list("original", 0)
    ])
    create_missing_for_simulations <- table_results[
        unified_int %in% missing_in_simulations
    ]
    create_missing_for_simulations <- unique(
        create_missing_for_simulations[, c("orig", "V1") := list(
            "simulations", 0
        )]
    )

    table_results <- do.call(
        "rbind",
        list(
            table_results, create_missing_for_original,
            create_missing_for_simulations
        )
    )


    ## p-values
    combo_list <- rep(NA, length = length(unique(table_results$unified_int)))
    p_high <- rep(NA, length = length(unique(table_results$unified_int)))
    p_low <- rep(NA, length = length(unique(table_results$unified_int)))

    for (int_combo in seq_along(unique(table_results$unified_int))) {
        this_combo <- as.character(unique(table_results$unified_int)[int_combo])

        sub <- table_results[unified_int == this_combo]

        orig_value <- sub[orig == "original"]$V1
        sim_values <- sub[orig == "simulations"]$V1

        length_simulations <- length(sim_values)
        if (length_simulations != number_of_simulations) {
            additional_length_needed <- number_of_simulations -
                length_simulations
            sim_values <- c(sim_values, rep(0, additional_length_needed))
        }

        p_orig_higher <- 1 - (sum((orig_value + 1) > (sim_values + 1)) /
            number_of_simulations)
        p_orig_lower <- 1 - (sum((orig_value + 1) < (sim_values + 1)) /
            number_of_simulations)

        combo_list[[int_combo]] <- this_combo
        p_high[[int_combo]] <- p_orig_higher
        p_low[[int_combo]] <- p_orig_lower
    }
    res_pvalue_DT <- data.table::data.table(
        unified_int = as.vector(combo_list),
        p_higher_orig = p_high,
        p_lower_orig = p_low
    )


    # depletion or enrichment in barplot format
    table_mean_results <- table_results[, .(mean(V1)), by = c(
        "orig", "unified_int", "type_int"
    )]
    table_mean_results_dc <- data.table::dcast.data.table(
        data = table_mean_results, formula = type_int + unified_int ~ orig,
        value.var = "V1"
    )
    table_mean_results_dc[, original := ifelse(is.na(original), 0, original)]
    table_mean_results_dc[, enrichm := log2((original + 1) / (simulations + 1))]


    table_mean_results_dc <- merge(
        table_mean_results_dc, res_pvalue_DT,
        by = "unified_int"
    )
    data.table::setorder(table_mean_results_dc, enrichm)
    table_mean_results_dc[, unified_int := factor(unified_int, unified_int)]

    # adjust p-values for mht

    # data.table variables
    p.adj_higher <- p.adj_lower <- p_lower_orig <- p_higher_orig <-
        PI_value <- int_ranking <- NULL

    table_mean_results_dc[, p.adj_higher := stats::p.adjust(
        p_higher_orig,
        method = sel_adjust_method
    )]
    table_mean_results_dc[, p.adj_lower := stats::p.adjust(
        p_lower_orig,
        method = sel_adjust_method
    )]


    table_mean_results_dc[, PI_value := ifelse(p.adj_higher <= p.adj_lower,
        -log10(p.adj_higher + (1 / number_of_simulations)) * enrichm,
        -log10(p.adj_lower + (1 / number_of_simulations)) * enrichm
    )]
    data.table::setorder(table_mean_results_dc, PI_value)

    # order
    table_mean_results_dc <- table_mean_results_dc[order(-PI_value)]
    table_mean_results_dc[, int_ranking := seq_len(.N)]

    return(list(
        raw_sim_table = table_results,
        enrichm_res = table_mean_results_dc
    ))
}
